---
title: "P2-Causal-Analysis-NB-total-v3"
author: "Jeffrey Wu"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    df-print: paged
---

```{r,warning=FALSE,echo=FALSE}
set.seed(2)

library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(ggExtra)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(huge)
library(INLA)
library(HMMpa)
library(invgamma)
library(brinla)
library(reshape2)
library(patchwork)
library(jsonlite)
library(geosphere)
library(urbnmapr)
library(RAQSAPI)
library(con2aqi)
library(pscl)
library(reshape2)
library(corrplot)
library(superheat)
library(shapes)
library(scales)
library(kableExtra)
library(splines)
library(fmesher)
library(Matrix)
library(tictoc)
```

```{r}
# load("p2-eda-workspace-2.27.RData")
# inla_df = readRDS("inla_df_3.7.rds")

load("~/work/p2-eda-workspace-2.27.RData")
inla_df = readRDS("~/work/inla_df_5.13.rds")

inla_df = st_drop_geometry(inla_df)
```

## Creating cumulative covariate columns (with memory property)

```{r,warning = FALSE}
create_rolling_agg_var <- function(data, variable_name, county_col = "NAME", 
                               lags = c(3, 6), n_periods = 72, 
                               agg_fun = sum, agg_name = "cum") {
  
  # Initialize result lists
  results <- list()
  
  # Create a column for each lag
  for (lag in lags) {
    results[[paste0(agg_name, "_", variable_name, lag)]] <- c()
  }
  
  # Loop through each county
  for (i in 1:length(counties)) {
    
    # Extract county data
    county_data <- data %>% 
      filter(!!sym(county_col) == counties[i]) %>% 
      dplyr::select(!!sym(variable_name)) %>% 
      st_drop_geometry()
    
    county_values <- county_data[[variable_name]]
    
    # Calculate rolling aggregations for each lag
    for (lag in lags) {
      col_name <- paste0(agg_name, "_", variable_name, "_", lag)
      
      # Initialize county aggregation vector
      county_agg <- numeric(n_periods)
      
      # Fill initial values (first 'lag' periods)
      county_agg[1:(lag-1)] <- county_values[1:(lag-1)]
      
      # Calculate rolling aggregations
      for (j in lag:n_periods) {
        county_agg[j] <- agg_fun(county_values[(j - lag + 1):j])
      }
      
      # Append to results
      results[[col_name]] <- c(results[[col_name]], county_agg)
    }
  }
  
  # Add results to original dataframe
  for (col_name in names(results)) {
    data[[col_name]] <- results[[col_name]]
  }
  
  return(data)
}


#Treatment (cumulative sum)
inla_df <- create_rolling_agg_var(
  data = inla_df,
  variable_name = "monthly_total_pm25",
  county_col = "NAME",
  lags = c(3, 6),
  n_periods = 72,
  agg_fun = sum,
  agg_name = "cum"
)

#Precipitation (cumulative sum)
inla_df <- create_rolling_agg_var(
  data = inla_df,
  variable_name = "total_precip",
  county_col = "NAME",
  lags = c(3, 6),
  n_periods = 72,
  agg_fun = sum,
  agg_name = "cum"
)

#Temperature (mean)
inla_df <- create_rolling_agg_var(
  data = inla_df,
  variable_name = "avg_air_temp",
  county_col = "NAME",
  lags = c(3, 6),
  n_periods = 72,
  agg_fun = mean,
  agg_name = "avg"
)

#Humidity (mean)
inla_df <- create_rolling_agg_var(
  data = inla_df,
  variable_name = "avg_dew_point",
  county_col = "NAME",
  lags = c(3, 6),
  n_periods = 72,
  agg_fun = mean,
  agg_name = "avg"
)

#AQI (max)
inla_df <- create_rolling_agg_var(
  data = inla_df,
  variable_name = "AQI",
  county_col = "NAME",
  lags = c(3,6),
  n_periods = 72,
  agg_fun = max,
  agg_name = "max"
)

#Wind speed (max)
inla_df <- create_rolling_agg_var(
  data = inla_df,
  variable_name = "avg_wind_speed",
  county_col = "NAME",
  lags = c(3,6),
  n_periods = 72,
  agg_fun = max,
  agg_name = "max"
)
```

```{r}
#Want to use the variation in AQI not explained by treatment i.e. residuals 
# lm_aqi = lm(max_AQI_6 ~ monthly_total_pm25, data = inla_df)
lm_aqi = lm(max_AQI_6 ~ pm10 + pm25, data = inla_df)

aqi_resids = lm_aqi$residuals

inla_df$aqi_resids = aqi_resids

# plot(AQI~monthly_total_pm25,data= inla_df)

# test = lm(death_rate ~ aqi_resids + total_precip + avg_air_temp + monthly_total_pm25 + monthly_total_pm25_lag1 + monthly_total_pm25_lag3,data = inla_df)
# 
# summary(test)
```


```{r}
#Add in lags of treatment e.g. lags 1-6
inla_df <- inla_df %>%
  group_by(ID) %>%
  mutate(
    cum_monthly_total_pm25_6_lag1 = lag(cum_monthly_total_pm25_6, 1),
    cum_monthly_total_pm25_6_lag2 = lag(cum_monthly_total_pm25_6, 2),
    cum_monthly_total_pm25_6_lag3 = lag(cum_monthly_total_pm25_6, 3),
    cum_monthly_total_pm25_6_lag4 = lag(cum_monthly_total_pm25_6, 4),
    cum_monthly_total_pm25_6_lag5 = lag(cum_monthly_total_pm25_6, 5),
    cum_monthly_total_pm25_6_lag6 = lag(cum_monthly_total_pm25_6, 6)
  ) %>% ungroup()
```

```{r}
#Attach cluster labels and SDI scores 
inla_df = inla_df %>% left_join(clusterlabels,by = c("NAME" = "counties"))

# Assign column names to your yearly df
colnames(SDI_df) <- CA_counties
rownames(SDI_df) <- 2010:2019

# Convert wide to long: one row per year-county pair
df_long <- as.data.frame(SDI_df) %>%
  mutate(year = as.integer(rownames(.))) %>%
  pivot_longer(-year, names_to = "county", values_to = "value")

# Add all 12 months to each row
df_monthly_expanded <- df_long %>%
  crossing(month = 1:12) %>%
  arrange(county, year, month) %>% 
  filter(year >= 2014)

inla_df$SDI = df_monthly_expanded$value
inla_df$Cluster = factor(inla_df$Cluster)
```

# Simplify clusters to quintiles of SDI score instead of SKATER

```{r}
# Calculate quintile breakpoints
SDI_groups = quantile(inla_df$SDI, c(0.2, 0.4, 0.6, 0.8))

# Calculate average SDI per county
avg_SDI = apply(SDI_df, 2, mean)
avg_SDI_df = data.frame(counties, avg_SDI)

# Assign group labels based on quintiles
grouplabels = rep(NA, nrow(avg_SDI_df))  # Initialize vector

grouplabels[avg_SDI <= SDI_groups[1]] = 1
grouplabels[avg_SDI > SDI_groups[1] & avg_SDI <= SDI_groups[2]] = 2
grouplabels[avg_SDI > SDI_groups[2] & avg_SDI <= SDI_groups[3]] = 3
grouplabels[avg_SDI > SDI_groups[3] & avg_SDI <= SDI_groups[4]] = 4
grouplabels[avg_SDI > SDI_groups[4]] = 5

# Add to dataframe
avg_SDI_df$Group = grouplabels
group_labels = avg_SDI_df[,c(1,3)]

inla_df = inla_df %>% left_join(group_labels,by = c("NAME" = "counties"))
inla_df$Group = factor(inla_df$Group)
```

```{r}
#Attach cluster labels to each county
group_labels$Group = factor(group_labels$Group)
group_labels$subregion = tolower(group_labels$counties)

merged_data <- merge(ca_map, group_labels, by = "subregion", all.x = TRUE)

# Define distinct colors for each category
#distinct_colors = qualitative_hcl(7,palette = "dynamic")
distinct_colors <- c("red", "blue", "green", "purple", "yellow")
# distinct_colors <- c("red", "blue")

#Plot
gg_with_grouping <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = Group), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  scale_fill_manual(values = distinct_colors) + 
  theme(
      title = element_text(size = 16),
      legend.text  = element_text(size = 20),  # legend labels
      legend.title = element_text(size = 20)   # legend title
    )

gg_with_grouping
```

# Plotting variables by county

```{r,eval = FALSE}
#SDI 
inla_df %>% group_by(NAME) %>% summarize(avg_SDI = mean(SDI)) %>% arrange(avg_SDI) %>% print(n=58)

#monthly_total_pm25
summary_df <- inla_df %>%
  group_by(NAME, month) %>%
  summarize(avg_treat = mean(monthly_total_pm25, na.rm = TRUE), .groups = "drop") %>%
  arrange(avg_treat)

ggplot(summary_df, aes(x = month, y = avg_treat)) +
  geom_point() + facet_wrap(~NAME) + 
  scale_x_discrete(breaks = 1:12) +
  labs(x = "Month", y = "Avg Treatment") +
  theme_bw()

#cum_monthly_total_pm25_6
summary_df2 <- inla_df %>%
  group_by(NAME, month) %>%
  summarize(avg_cumtreat = mean(cum_monthly_total_pm25_6, na.rm = TRUE), .groups = "drop") %>%
  arrange(avg_cumtreat)

ggplot(summary_df2, aes(x = month, y = avg_cumtreat)) +
  geom_point() + facet_wrap(~NAME) + 
  scale_x_discrete(breaks = 1:12) +
  labs(x = "Month", y = "Avg Cumulative Treatment") +
  theme_bw()

#death rate
summary_df3 <- inla_df %>%
  group_by(NAME, month) %>%
  summarize(max_deathrate = max(death_rate, na.rm = TRUE), .groups = "drop") %>%
  arrange(max_deathrate)

ggplot(summary_df3, aes(x = month, y = max_deathrate)) +
  geom_point() + facet_wrap(~NAME) + 
  scale_x_discrete(breaks = 1:12) +
  labs(x = "Month", y = "Avg Death Rate") +
  theme_bw()
```

# Modeling treatment

## Estimate a spatial dependence structure

```{r}
huge_input = as.matrix(score_scaled)
#huge_input = as.matrix(SDI_df)

#Graph learning w HUGE
out.glasso = huge(huge_input,lambda = seq(0.95,0.05,by=-0.05),method="glasso")

glasso.stars = huge.select(out.glasso,criterion = "stars",stars.thresh = 0.1)
glasso.ric = huge.select(out.glasso,criterion = "ric")
glasso.ebic = huge.select(out.glasso,criterion = "ebic")

plot(glasso.stars)
plot(glasso.ric)
plot(glasso.ebic)

huge.est = glasso.ebic$refit
#huge.est

degree_connectivity = data.frame(colSums(huge.est))
colnames(degree_connectivity) = "node_connections"
degree_connectivity = cbind(c(1:num_counties),degree_connectivity)
```

### Transforming estimated adjacency matrix to graph filter H

The code below takes the adjacency matrix estimated in the previous step and transforms it into a graph filter H. The steps are explained in Antonian et al 2019 (Gareth Peters' paper). The cutoff transformation is applied to the eigenvalues of the graph Laplacian.

```{r}
#A = as.matrix(huge.est)
A = as.matrix(adj)

p = nrow(A)

#obtain graph Laplacian L
D = diag(p)
for (i in 1:p){
  d = sum(A[,i])
  D[i,i] = d
}

L = D - A

#eigendecomposition of L
Ldecomp = eigen(L)
U = as.matrix(Ldecomp$vectors)
Lambdas = Ldecomp$values

#test
#U %*% (diag(p)*Lambdas) %*% t(U)

#Function implementing cutoff tranform for eigenvalues 
cutoff.transform = function(lambdas,q){
  transformed = c()
  cutoff = quantile(lambdas,q)
  for (i in lambdas){
    if(i <= cutoff){
      transformed = c(transformed,i)
    }
    else{
      transformed = c(transformed,0)
    }
  }
  
  return(transformed)
}

#quantile(Lambdas,2/3)
transformed.L = cutoff.transform(Lambdas,2/3)
eta.L = diag(p)*transformed.L

#obtain graph filter
H = U %*% eta.L %*% t(U)

gfilter_weight = norm((1/7)*H^2,type = "F")
# gfilter_weight = norm(H^2,type = "F")
```

```{r}
matrix_heatmap = function(
  matrix, 
  title = "", 
  legend_title = "Covariance", 
  xlab = "Entry j",
  ylab = "Entry i",
  gradient_zones = c(0, 0.5, 0.999),
  save_path = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/Paper1-images/plos1-figs",
  prefix = "",           
  filename = "heatmap.png", # base filename
  text_scale = 2){
  r = nrow(matrix)
  c = ncol(matrix)
  
  df = as_tibble(cbind(expand.grid(rev(seq_len(r)), seq_len(c)), c(matrix))) %>%
    setNames(c("row", "col", "value"))
  
  df$value[df$value == 1] = 0.999
  
  x_min = min(df$row) - 0.5
  x_max = max(df$row) + 0.5
  y_min = min(df$col) - 0.5
  y_max = max(df$col) + 0.5
  
  plot = ggplot(df, mapping = aes(x = row, y = col, fill = value)) +
    geom_tile() +
    geom_rect(aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), 
              fill = NA, color = "black", inherit.aes = FALSE) +
    scale_fill_gradientn(
      colors = c("white", "lightblue", "darkblue"),
      values = rescale(gradient_zones),
      limits = c(0, 0.99),
      oob = squish,
      name = legend_title
    ) +
    scale_x_continuous(
      breaks = c(1,c),
      labels = paste(c(1,c)),  # reversed because we used rev() earlier
      name = xlab
    ) +
    scale_y_continuous(
      breaks = c(r,1),
      labels = paste(c(1,r)),
      name = ylab
    ) +
    ggtitle(title) +
    theme_minimal(base_size = 11 * text_scale) +
    theme(
      plot.margin = margin(t = 10, r = 30, b = 10, l = 10),
      legend.key.size = unit(1, "cm"),
      axis.text.x = element_text(hjust = 0.5),  # center x labels
      panel.grid = element_blank()  # remove grid lines
    )

    # ggsave(
    #   filename = paste0(prefix, filename),
    #   plot = plot,
    #   path = save_path,
    #   width = 8,
    #   height = 6,
    #   dpi = 300
    # )
    # 
  return(plot)
}

matrix_heatmap(H)
```

## Estimate a temporal dependence structure

### Calculating covariate similarity kernel K_cov

```{r}
cov_similarity_kernel = function(cov_data,time_span,rho_rbf,rho_periodic,sigma2,additive = TRUE){
  K_cov = matrix(0,nrow=time_span,ncol=time_span)
  i = 1
  j = 1
  
  for(t1 in 1:time_span){
    for (t2 in 1:time_span){
      A = cov_data %>% filter(time == t1) 
      B = cov_data %>% filter(time == t2) 
      # AQIa = unique(A$AQI)
      # AQIb = unique(B$AQI)
      
      diff = (A$cum_monthly_total_pm25_6-B$cum_monthly_total_pm25_6)^2
      
      if(additive == TRUE){
        K_cov[i,j] = (exp(- (mean(diff)) 
           / (2*rho_rbf)) + exp(- (2*sin(sum(abs(diff))*pi/12)^2)
           / (rho_periodic))) * sigma2
      } else{
        K_cov[i,j] = exp(- (mean(diff))
                 / (2*rho_rbf)) * exp(- (2*sin(sum(abs(diff))*pi/12)^2)
                 / (rho_periodic)) * sigma2
      }
      
      j = j+1
    }
    
    j = 1
    i = i+1
  }
  
  return(K_cov)
}
```

### Calculating simple time kernel K_time

```{r}
time_rbf_kernel = function(time_span,rho_rbf,rho_periodic,sigma2){
  K_time = matrix(NA,nrow = time_span, ncol = time_span)

  for (i in 1:time_span){
    for (j in 1:time_span){
      K_time[i,j] = exp(- (abs(i-j)^2) / (2*rho_rbf)) * sigma2 #JUST RBF
      
      # K_time[i,j] = exp(- (2*sin(sum(abs(i-j))*pi/12)^2) #JUST PERIODIC
      #          / (rho_periodic)) * sigma2
    }
  }

  return(K_time)
}

time_kernel = function(time_span,rho_rbf,rho_periodic,sigma2,additive = TRUE){
  K_time = matrix(NA,nrow = time_span, ncol = time_span)

  for (i in 1:time_span){
    for (j in 1:time_span){
      
      if(additive == TRUE){
        K_time[i,j] = (exp(- (abs(i-j)^2) / (2*rho_rbf)) + exp(- (2*sin(sum(abs(i-j))*pi/12)^2)
               / (rho_periodic))) * sigma2
      } else{
        K_time[i,j] = (exp(- (abs(i-j)^2) / (2*rho_rbf)) * exp(- (2*sin(sum(abs(i-j))*pi/12)^2)
                 / (rho_periodic))) * sigma2
      }
        
    }
  }

  return(K_time)
}

Krbf = time_rbf_kernel(time_span = length(unique(inla_df$time)),rho_rbf = 10, rho_periodic = 1,sigma2 = 1)
Klp = time_kernel(time_span = length(unique(inla_df$time)),rho_rbf = 0.01, rho_periodic = 1,sigma2 = 1)

matrix_heatmap(Krbf)
matrix_heatmap(Klp)
```

### Ensuring precision matrix is not computationally singular, so we jitter eigenvalues, using reciprocal condition number as constraint

```{r}
desingularize = function(covmatrix,threshold = 1e-2,increment = 0.01){
  
  tracker = 0
  
  while (rcond(covmatrix) <= threshold){
    #Perform spectral decomposition
    ev = eigen(covmatrix)
    L = ev$values
    V = ev$vectors
    
    # #Add a little noise to eigenvalues to bring away from 0
    L = L + increment
    
    # #Add 0.01 to eigenvalues in bottom five percentile to bring away from 0
    # cutoff = quantile(abs(L),0.05)
    # L[which(abs(L) < cutoff)] = L[which(abs(L) < cutoff)] + 0.01
    
    #Calculate new precision matrix 
    covmatrix = V %*% diag(L) %*% t(V)
    
    tracker = tracker + 1
  }
  
  results_list = list(covmatrix,tracker)
  #sprintf("%s iterations of desingularizer applied",tracker)
  return(results_list)
}

# test = desingularize(K_time)
```

## Compute kernels and then spatiotemporal covariance matrices

```{r,cache = FALSE}
K_df = inla_df %>% dplyr::select(time,cum_monthly_total_pm25_6) %>% st_drop_geometry()

###Treatment (34.674340,93.99833,5.052856) WAIC = 42756.57, (0.0185147378, 0.0695663638, 4.6295870) WAIC = 42535.99
K_cov1 = cov_similarity_kernel(K_df,time_span = 72,rho_rbf = 34.674340,rho_periodic = 93.99833,
                               sigma2 = 5.052856, additive = FALSE) 

K_cov1.2 = cov_similarity_kernel(K_df,time_span = 72,rho_rbf = 0.0185147378,rho_periodic = 0.0695663638,
                                 sigma2 = 4.6295870, additive = TRUE) 

K_cov1_heatmap = matrix_heatmap(K_cov1.2,title = "")

#Calculate trace norm of gram matrix
K_cov1_weight = norm((1/72)*K_cov1.2,type = "F")

#Calculate proposed kernel
covGP1_treatment = kronecker((K_cov1.2/72),(H^2/58))

#Need to ensure precision matrix is not computationally singular i.e det > 0
covGP_jittered = desingularize(covGP1_treatment,threshold = 1e-5,increment = 0.2)
covGP1_treatment = covGP_jittered[[1]]
num_jitters = covGP_jittered[[2]]

inv_covGP1_treatment = solve(covGP1_treatment)

#Heatmap of resulting inv_covGP1
inv_covGP1_treatment_heatmap = matrix_heatmap(inv_covGP1_treatment,title = "")



##Response (58.998325,24.22638,37.477465) WAIC = 19948.19, (0.0568448990, 0.0120379678, 3.9571684) WAIC = 20046.18
K_cov2 = cov_similarity_kernel(K_df,time_span = 72,rho_rbf = 58.998325,rho_periodic = 24.22638,
                               sigma2 = 37.477465, additive = FALSE) 

K_cov2.2 = cov_similarity_kernel(K_df,time_span = 72,rho_rbf = 0.0568448990,rho_periodic = 0.0120379678,
                                 sigma2 = 3.9571684, additive = TRUE) 

covGP1_response = kronecker((K_cov2.2/72),(H^2/58))

#Need to ensure precision matrix is not computationally singular i.e det > 0
covGP_jittered = desingularize(covGP1_response,threshold = 1e-5,increment = 0.2)
covGP1_response = covGP_jittered[[1]]
num_jitters = covGP_jittered[[2]]

inv_covGP1_response = solve(covGP1_response)

#Heatmap of resulting inv_covGP2 
inv_covGP1_response_heatmap = matrix_heatmap(inv_covGP1_response,title = "")
```

```{r,cache = FALSE}
#Calculating gram matrix K_time

###Treatment (9.364325,32.45866,24.731415) WAIC = 42753.99, (0.072685962, 0.3132679, 1.6001799) WAIC = 42535.99
K_time1 = time_kernel(time_span = length(unique(inla_df$time)),rho_rbf = 9.364325, 
                      rho_periodic = 32.45866,sigma2 = 24.731415, additive = FALSE)

K_time1.2 = time_kernel(time_span = length(unique(inla_df$time)),rho_rbf = 0.072685962, 
                      rho_periodic = 0.3132679,sigma2 = 1.6001799, additive = TRUE)


#Heatmap of resulting K 
K_time1_heatmap = matrix_heatmap(K_time1.2,title = "K_time heatmap")

#Calculate trace norm of gram matrix
K_time1_weight = norm((1/72)*K_time1.2,type = "F")

#Calculate proposed kernel
covGP2_treatment = kronecker((K_time1.2/72),(H^2/58))
covGP2_treatment2 = kronecker((Krbf/72),(H^2/58))


#Need to ensure precision matrix is not computationally singular i.e det > 0
covGP_jittered = desingularize(covGP2_treatment,threshold = 1e-5,increment = 0.2)
covGP2_treatment = covGP_jittered[[1]]
num_jitters = covGP_jittered[[2]]

covGP_jittered = desingularize(covGP2_treatment2,threshold = 1e-5,increment = 0.2)
covGP2_treatment2 = covGP_jittered[[1]]
num_jitters = covGP_jittered[[2]]

inv_covGP2_treatment = solve(covGP2_treatment)
inv_covGP2_treatment2 = solve(covGP2_treatment2)

#Heatmap of resulting inv_covGP2 
inv_covGP2_treatment_heatmap = matrix_heatmap(covGP2_treatment,title = "")



###Response (21.611833,50.17043,54.861003) WAIC = 19806.30 , (0.008535596, 6.8016543, 4.6925669) WAIC = 20086.22
K_time2 = time_kernel(time_span = length(unique(inla_df$time)),rho_rbf = 21.611833, 
                      rho_periodic = 50.17043, sigma2 = 54.861003, additive = FALSE)

K_time2.2 = time_kernel(time_span = length(unique(inla_df$time)),rho_rbf = 100, 
                      rho_periodic = 0.2, sigma2 = 54.861003, additive = FALSE)

K_time2.3 = time_kernel(time_span = length(unique(inla_df$time)),rho_rbf = 0.072685962, 
                      rho_periodic = 0.3132679,sigma2 = 1.6001799, additive = TRUE)



covGP2_response = kronecker((K_time2.3/72),(H^2/58))

#Heatmap of resulting K 
K_time2_heatmap = matrix_heatmap(K_time2.3,title = "K_time heatmap")

#Calculate trace norm of gram matrix
K_time_weight = norm((1/72)*K_time2.3,type = "F")

#Calculate proposed kernel
covGP2_response = kronecker((K_time2.3/72),(H^2/58))

#Need to ensure precision matrix is not computationally singular i.e det > 0
covGP_jittered = desingularize(covGP2_response,threshold = 1e-5,increment = 0.2)
covGP2_response = covGP_jittered[[1]]
num_jitters = covGP_jittered[[2]]

inv_covGP2_response = solve(covGP2_response)
# inv_covGP2_response = as(kronecker(solve(K_time2.2/72), solve(H^2/58)), "sparseMatrix")


#Heatmap of resulting inv_covGP2 
inv_covGP2_response_heatmap = matrix_heatmap(inv_covGP2_response,title = "")
```

## Next fit two KGR models with an estimated graph and a temporal kernel

### Using K_cov

```{r,warning=FALSE,cache = TRUE}
kgr_model1_treatment = function(dataset,link=1){
  
  ###Fit INLA model 
  kgr_formula1 = (cum_monthly_total_pm25_6 + 0.0001) ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + month + aqi_resids + total_precip + avg_air_temp + avg_dew_point + avg_wind_speed + f(id2,model = "generic0",Cmatrix = inv_covGP1_treatment)
  
  model = inla(formula = kgr_formula1,family = "gamma",data = dataset, num.threads = 1,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "simplified.laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$NAME, dataset$time, dataset$ID, dataset$cum_monthly_total_pm25_6, preds_model)
  colnames(preds_model) <- c("county", "time", "ID", "cum_monthly_total_pm25_6", "mean", "sd", 
                             "0.025quant", "0.5quant", "0.975quant", "mode")
  
  # scale = 0.05

  ###Generate uncertainty bands from the posterior predictive distribution as opposed to posterior of the linear predictor
  for (i in 1:nrow(preds_model)){
    ###Sample 1000 mu for each row
    row = data.frame(model$marginals.fitted.values[[i]])

    sample_mu = inla.rmarginal(1000,inla.smarginal(row))
    lower = quantile(sample_mu,0.025)
    mean = mean(sample_mu)
    upper = quantile(sample_mu,0.975)
    
    row$density = row$y / sum(row$y)
    
    phi_density = data.frame(model$marginals.hyperpar[[1]])
    sample_phi = inla.rmarginal(1000,inla.smarginal(phi_density))

    ###Solve for a and b for each row
    a = sample_phi
    b = sample_phi / sample_mu #narrower

    ###Sample from Gamma with these a and b
    sample_y = rgamma(1000,a,b)
    
    ###Obtain 95% credible interval
    preds_model$y_ppmean[i] = mean(sample_y, na.rm = TRUE)
    preds_model$y_pplower[i] = quantile(sample_y, 0.025, na.rm = TRUE)
    preds_model$y_ppupper[i] = quantile(sample_y, 0.975, na.rm = TRUE)
  }
  
  preds_model$interval_width = preds_model$`0.975quant` - preds_model$`0.025quant` 
  # marginal_fvs <- model$marginals.fitted.values

  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  kgr_model1_results = list(
    covmatrix = covGP1_treatment,
    prec = inv_covGP1_treatment,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model
  )
  
  return(kgr_model1_results)
}

# #Fit kgr_model1
# kgr_model1_treatment_fit = kgr_model1_treatment(dataset = inla_df)
# 
# #Extract DIC and WAIC
# kgr_model1_DIC = kgr_model1_treatment_fit$model_DIC
# kgr_model1_WAIC = kgr_model1_treatment_fit$model_WAIC
# 
# #Get summaries of parameter estimates
# kgr_model1_treatment_fit$model_summary
# kgr_model1_treatment_fit$bri_hyperpar_summary
# kgr_model1_treatment_fit$exp_effects
# 
# #Show plots
# kgr_model1_treatment_fit$param_plot
# kgr_model1_treatment_fit$hyperparam_plot
# 
# for (i in 1:58){
#   df = inla_df %>% filter(ID == i) %>% dplyr::select(cum_monthly_total_pm25_6)
#   preds = kgr_model1_treatment_fit$fitted_values %>% filter(ID == i) 
#   df = cbind(df,preds)
#   df$mean = df$mean - 0.0001
#   
#   # title = sprintf("Posterior Predictive Fits for County %s",i)
#   title = sprintf("%s County",counties[i])
#   
#   post_pred_plot = preds %>% ggplot(aes(x=time,y=cum_monthly_total_pm25_6)) + geom_point() + 
#     geom_line(aes(y=mean),color = "red") + geom_ribbon(aes(ymin = `0.025quant`,ymax = `0.975quant`), fill = "orange",alpha = 0.7) + 
#   geom_ribbon(aes(ymin = y_pplower,ymax = y_ppupper),fill = "skyblue",alpha = 0.3) + ggtitle(title) + theme_classic()
#   print(post_pred_plot)
# }
```

### Using K_time

```{r,warning=FALSE,cache = TRUE}
kgr_model2_treatment = function(dataset,link=1){
  
  ###Fit INLA model 
  kgr_formula2 = (cum_monthly_total_pm25_6 + 0.0001) ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + month + aqi_resids + total_precip + avg_air_temp + avg_dew_point + avg_wind_speed + f(id2,model = "generic0",Cmatrix = inv_covGP2_treatment)
  
  model = inla(formula = kgr_formula2,family = "gamma",data = dataset, num.threads = 1,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "simplified.laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$NAME, dataset$time, dataset$ID, dataset$cum_monthly_total_pm25_6, preds_model)
  colnames(preds_model) <- c("county", "time", "ID", "cum_monthly_total_pm25_6", "mean", "sd", 
                             "0.025quant", "0.5quant", "0.975quant", "mode")
  
  ###Generate uncertainty bands from the posterior predictive distribution as opposed to posterior of the linear predictor
  for (i in 1:nrow(preds_model)){
    ###Sample 1000 mu for each row
    row = data.frame(model$marginals.fitted.values[[i]])

    sample_mu = inla.rmarginal(1000,inla.smarginal(row))
    lower = quantile(sample_mu,0.025)
    mean = mean(sample_mu)
    upper = quantile(sample_mu,0.975)
    
    row$density = row$y / sum(row$y)
    
    phi_density = data.frame(model$marginals.hyperpar[[1]])
    sample_phi = inla.rmarginal(1000,inla.smarginal(phi_density))

    ###Solve for a and b for each row
    a = sample_phi
    b = sample_phi / sample_mu #narrower

    ###Sample from Gamma with these a and b
    sample_y = rgamma(1000,a,b)
    
    ###Obtain 95% credible interval
    preds_model$y_ppmean[i] = mean(sample_y, na.rm = TRUE)
    preds_model$y_pplower[i] = quantile(sample_y, 0.025, na.rm = TRUE)
    preds_model$y_ppupper[i] = quantile(sample_y, 0.975, na.rm = TRUE)
  }
  
  preds_model$interval_width = preds_model$`0.975quant` - preds_model$`0.025quant` 
  # marginal_fvs <- model$marginals.fitted.values

  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  kgr_model2_results = list(
    covmatrix = covGP2_treatment,
    prec = inv_covGP2_treatment,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model
  )
  
  return(kgr_model2_results)
}

# #Fit kgr_model2
# kgr_model2_treatment_fit = kgr_model2_treatment(dataset = inla_df)
# 
# #Extract DIC and WAIC
# kgr_model2_DIC = kgr_model2_treatment_fit$model_DIC
# kgr_model2_WAIC = kgr_model2_treatment_fit$model_WAIC
# 
# #Get summaries of parameter estimates
# kgr_model2_treatment_fit$model_summary
# kgr_model2_treatment_fit$bri_hyperpar_summary
# kgr_model2_treatment_fit$exp_effects
# 
# #Show plots
# kgr_model2_treatment_fit$param_plot
# kgr_model2_treatment_fit$hyperparam_plot
# 
# 
# for (i in 1:58){
#   df = inla_df %>% filter(ID == i) %>% dplyr::select(cum_monthly_total_pm25_6)
#   preds = kgr_model2_treatment_fit$fitted_values %>% filter(ID == i) 
#   df = cbind(df,preds)
#   df$mean = df$mean - 0.0001
#   
#   # title = sprintf("Posterior Predictive Fits for County %s",i)
#   title = sprintf("%s County",counties[i])
#   
#   post_pred_plot = preds %>% ggplot(aes(x=time,y=cum_monthly_total_pm25_6)) + geom_point() + 
#     geom_line(aes(y=mean),color = "red") + geom_ribbon(aes(ymin = `0.025quant`,ymax = `0.975quant`), fill = "orange",alpha = 0.7) + 
#   geom_ribbon(aes(ymin = y_pplower,ymax = y_ppupper),fill = "skyblue",alpha = 0.3) + ggtitle(title) + theme_classic()
#   print(post_pred_plot)
# }
```

# Calculate propensity scores -> stabilized IPWs

## Function to extract density from fitted values 

```{r}
est_density_ps = function(model_fit,model_preds){
  ps = rep(NA,nrow(model_preds))
  
  for (i in 1:nrow(model_preds)){
    row = data.frame(inla.smarginal(model_fit$marginals.fitted.values[[i]]))
    row$density = row$y / sum(row$y)
    
    obs_density = inla.dmarginal(model_preds$cum_monthly_total_pm25_6[i],row) / sum(row$y)
    pred_density = inla.dmarginal(model_preds$mean[i],row) / sum(row$y)
    
    ps[i] = pred_density
  }
  
  return(ps)
}

# all_density_ps_denom = est_density_ps(all_ps_denom_model,preds_model)
```

## Estimate PS for all observations

```{r,cache = TRUE}
#Numerator model
ps_num_formula = (cum_monthly_total_pm25_6 + 0.1) ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + month + cum_monthly_total_pm25_6_lag1 + cum_monthly_total_pm25_6_lag2 + cum_monthly_total_pm25_6_lag3 + cum_monthly_total_pm25_6_lag4 + cum_monthly_total_pm25_6_lag5 + cum_monthly_total_pm25_6_lag6 + f(id2, model = "generic0", Cmatrix = inv_covGP2_treatment2)
  
all_ps_num_model = inla(ps_num_formula, family = "gamma",data = inla_df,
                control.compute = list(dic=TRUE,waic=TRUE,
                                       return.marginals.predictor=TRUE),
                control.inla = list(strategy = "laplace"),
                control.predictor = list(compute = TRUE, link = 1))

preds_model <- all_ps_num_model$summary.fitted.values
preds_model$mean = preds_model$mean - 0.1
preds_model <- cbind(inla_df$NAME, inla_df$time, inla_df$ID, inla_df$cum_monthly_total_pm25_6, preds_model)
colnames(preds_model) <- c("county", "time", "ID", "cum_monthly_total_pm25_6", "mean", "sd",
                           "0.025quant", "0.5quant", "0.975quant", "mode")

all_density_ps_num = est_density_ps(all_ps_num_model,preds_model)

# for (i in 1:58){
#   df = inla_df %>% filter(ID == i) %>% dplyr::select(cum_monthly_total_pm25_6)
#   preds = preds_model %>% filter(ID == i)
#   df = cbind(df,preds)
#   df$mean = df$mean
# 
#   # title = sprintf("Posterior Predictive Fits for County %s",i)
#   title = sprintf("%s County",counties[i])
# 
#   post_pred_plot = preds %>% ggplot(aes(x=time,y=cum_monthly_total_pm25_6)) + geom_point() +
#     geom_line(aes(y=mean),color = "red") + geom_ribbon(aes(ymin = `0.025quant`,ymax = `0.975quant`), fill = "orange",alpha = 0.7) +
#   geom_ribbon(aes(ymin = `0.025quant`,ymax = `0.975quant`),fill = "skyblue",alpha = 0.3) + ggtitle(title) + theme_classic()
#   print(post_pred_plot)
# }

#Denominator model
ps_denom_formula = (cum_monthly_total_pm25_6 + 0.1) ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + month + aqi_resids + total_precip + avg_air_temp + avg_dew_point + avg_wind_speed + cum_monthly_total_pm25_6_lag1 + cum_monthly_total_pm25_6_lag2 + cum_monthly_total_pm25_6_lag3 + cum_monthly_total_pm25_6_lag4 + cum_monthly_total_pm25_6_lag5 + cum_monthly_total_pm25_6_lag6 + f(id2, model = "generic0", Cmatrix = inv_covGP2_treatment2)

all_ps_denom_model = inla(formula = ps_denom_formula,family = "gamma",data = inla_df,
                control.compute = list(dic=TRUE,waic=TRUE,
                                       return.marginals.predictor=TRUE),
                control.inla = list(strategy = "laplace"),
                control.predictor = list(compute = TRUE, link = 1))

preds_model <- all_ps_denom_model$summary.fitted.values
preds_model$mean = preds_model$mean - 0.1
preds_model <- cbind(inla_df$NAME, inla_df$time, inla_df$ID, inla_df$cum_monthly_total_pm25_6, preds_model)
colnames(preds_model) <- c("county", "time", "ID", "cum_monthly_total_pm25_6", "mean", "sd",
                           "0.025quant", "0.5quant", "0.975quant", "mode")

all_density_ps_denom = est_density_ps(all_ps_denom_model,preds_model)

# for (i in 1:58){
#   df = inla_df %>% filter(ID == i) %>% dplyr::select(cum_monthly_total_pm25_6)
#   preds = preds_model %>% filter(ID == i)
#   df = cbind(df,preds)
#   df$mean = df$mean
# 
#   # title = sprintf("Posterior Predictive Fits for County %s",i)
#   title = sprintf("%s County",counties[i])
# 
#   post_pred_plot = preds %>% ggplot(aes(x=time,y=cum_monthly_total_pm25_6)) + geom_point() +
#     geom_line(aes(y=mean),color = "red") + geom_ribbon(aes(ymin = `0.025quant`,ymax = `0.975quant`), fill = "orange",alpha = 0.7) +
#   geom_ribbon(aes(ymin = `0.025quant`,ymax = `0.975quant`),fill = "skyblue",alpha = 0.3) + ggtitle(title) + theme_classic()
#   print(post_pred_plot)
# }

num_ps_allpreds = all_ps_num_model$summary.fitted.values$mean
denom_ps_allpreds = all_ps_denom_model$summary.fitted.values$mean

ev_sw_all_obs = num_ps_allpreds / denom_ps_allpreds
density_sw_all_obs = all_density_ps_num / all_density_ps_denom

summary(num_ps_allpreds)
summary(denom_ps_allpreds)
summary(ev_sw_all_obs)

summary(all_density_ps_num)
summary(all_density_ps_denom)
summary(density_sw_all_obs)

#Compare with estimating PS for all observations
inla_df$sw1 = ev_sw_all_obs
inla_df$sw2 = density_sw_all_obs
```

```{r}
#Multiplying densities up to time k for each unit
inla_df$sw_prod = 0

for (i in counties){
  county_idx = which(inla_df$NAME == i)
  county_sw = inla_df$sw2[county_idx]
  sw_prod = county_sw[1]
  
  for(j in 2:72){
    sw_prod[j] = county_sw[j]*sw_prod[j-1]
  }
  
  inla_df$sw_prod[county_idx] = sw_prod
}

summary(inla_df$sw_prod)
```

## Checking covariate balancing property

```{r,cache = TRUE}
###Without GPS
summary(inla_df$cum_monthly_total_pm25_6)

group1 = which(inla_df$cum_monthly_total_pm25_6 <= 50) #2102
group2 = which(inla_df$cum_monthly_total_pm25_6 > 50 & inla_df$cum_monthly_total_pm25_6 <= 185) #1032
group3 = which(inla_df$cum_monthly_total_pm25_6 > 185) #1042

# mu1 = mean(inla_df$avg_air_temp[group1],na.rm = TRUE)
# mu2 = mean(inla_df$avg_air_temp[-group1],na.rm = TRUE)

t.test(x = na.omit(inla_df$avg_dew_point[group3]), y = na.omit(inla_df$avg_dew_point[-group3]))


###GPS adjusted 

#median treatment = 48.844 -> set Z = 49 for all observations and recalculate SWs
inla_df49 = inla_df
inla_df49$cum_monthly_total_pm25_6 = 49

ps_num_formula = (cum_monthly_total_pm25_6 + 0.1) ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + month + cum_monthly_total_pm25_6_lag1 + cum_monthly_total_pm25_6_lag2 + cum_monthly_total_pm25_6_lag3 + cum_monthly_total_pm25_6_lag4 + cum_monthly_total_pm25_6_lag5 + cum_monthly_total_pm25_6_lag6 + f(id2, model = "generic0", Cmatrix = inv_covGP2_treatment2)

all_ps_num_model49 = inla(ps_num_formula, family = "gamma",data = inla_df49,
                control.compute = list(dic=TRUE,waic=TRUE,
                                       return.marginals.predictor=TRUE),
                control.inla = list(strategy = "laplace"),
                control.predictor = list(compute = TRUE, link = 1))

preds_model49 <- all_ps_num_model49$summary.fitted.values
preds_model49$mean = preds_model49$mean - 0.1
preds_model49 <- cbind(inla_df49$NAME, inla_df49$time, inla_df49$ID, inla_df49$cum_monthly_total_pm25_6, preds_model49)
colnames(preds_model49) <- c("county", "time", "ID", "cum_monthly_total_pm25_6", "mean", "sd",
                           "0.025quant", "0.5quant", "0.975quant", "mode")

all_density_ps_num49 = est_density_ps(all_ps_num_model49,preds_model49)




ps_denom_formula = (cum_monthly_total_pm25_6 + 0.1) ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + month + aqi_resids + total_precip + avg_air_temp + avg_dew_point + avg_wind_speed + cum_monthly_total_pm25_6_lag1 + cum_monthly_total_pm25_6_lag2 + cum_monthly_total_pm25_6_lag3 + cum_monthly_total_pm25_6_lag4 + cum_monthly_total_pm25_6_lag5 + cum_monthly_total_pm25_6_lag6 + f(id2, model = "generic0", Cmatrix = inv_covGP2_treatment2)

all_ps_denom_model49 = inla(formula = ps_denom_formula,family = "gamma",data = inla_df49,
                control.compute = list(dic=TRUE,waic=TRUE,
                                       return.marginals.predictor=TRUE),
                control.inla = list(strategy = "laplace"),
                control.predictor = list(compute = TRUE, link = 1))

preds_model49 <- all_ps_denom_model49$summary.fitted.values
preds_model49$mean = preds_model49$mean - 0.1
preds_model49 <- cbind(inla_df49$NAME, inla_df49$time, inla_df49$ID, inla_df49$cum_monthly_total_pm25_6, preds_model49)
colnames(preds_model49) <- c("county", "time", "ID", "cum_monthly_total_pm25_6", "mean", "sd",
                           "0.025quant", "0.5quant", "0.975quant", "mode")

all_density_ps_denom49 = est_density_ps(all_ps_denom_model49,preds_model49)

density_sw_all_obs49 = all_density_ps_num49 / all_density_ps_denom49

# summary(all_density_ps_num49)
# summary(all_density_ps_denom49)
summary(density_sw_all_obs49)

inla_df49$sw = density_sw_all_obs49

#Multiplying densities up to time k for each unit
inla_df49$sw_prod = 0

for (i in counties){
  county_idx = which(inla_df49$NAME == i)
  county_sw = inla_df49$sw[county_idx]
  sw_prod = county_sw[1]
  
  for(j in 2:72){
    sw_prod[j] = county_sw[j]*sw_prod[j-1]
  }
  
  inla_df49$sw_prod[county_idx] = sw_prod
}

summary(inla_df49$sw_prod)
```

```{r}
#Split GPS values into 4 groups
gps_group1 = which(inla_df49$sw_prod <= 0.0000085) #1044
gps_group2 = which(inla_df49$sw_prod > 0.0000085 & inla_df49$sw_prod <= 0.0003436) #1044
gps_group3 = which(inla_df49$sw_prod > 0.0003436 & inla_df49$sw_prod <= 0.0113204) #1044
gps_group4 = which(inla_df49$sw_prod > 0.0113204) #1044

# summary(inla_df$cum_monthly_total_pm25_6[gps_group1])

#For each GPS group, compare avg X between in treatment groups 
nested_group1 = which(inla_df$cum_monthly_total_pm25_6[gps_group1] <= 50) #424 / 1044
nested_group2 = which(inla_df$cum_monthly_total_pm25_6[gps_group2] <= 50) #399 / 1044
nested_group3 = which(inla_df$cum_monthly_total_pm25_6[gps_group3] <= 50) #529 /1044
nested_group4 = which(inla_df$cum_monthly_total_pm25_6[gps_group4] <= 50) #750 / 1044

#T-test for each covariate
test1 = t.test(x = na.omit(inla_df$avg_dew_point[nested_group1]), y = na.omit(inla_df$avg_dew_point[-nested_group1]))
test2 = t.test(x = na.omit(inla_df$avg_dew_point[nested_group2]), y = na.omit(inla_df$avg_dew_point[-nested_group2]))
test3 = t.test(x = na.omit(inla_df$avg_dew_point[nested_group3]), y = na.omit(inla_df$avg_dew_point[-nested_group3]))
test4 = t.test(x = na.omit(inla_df$avg_dew_point[nested_group4]), y = na.omit(inla_df$avg_dew_point[-nested_group4]))


d1 = test1$estimate[1] - test1$estimate[2]
d2 = test2$estimate[1] - test2$estimate[2]
d3 = test3$estimate[1] - test3$estimate[2]
d4 = test4$estimate[1] - test4$estimate[2]

se1 = test1$stderr
se2 = test2$stderr
se3 = test3$stderr
se4 = test4$stderr

n1 = length(nested_group1) #nested_group size
n2 = length(nested_group2)
n3 = length(nested_group3)
n4 = length(nested_group4)

num = n1*d1 + n2*d2 + n3*d3 + n4*d4
denom = sqrt(n1^2*se1^2 + n2^2*se2^2 + n3^2*se3^2 + n4^2*se4^2) 

weighted_mean_diff = num / (n1+n2+n3+n4)
se_diff = denom / (n1+n2+n3+n4)

tstat = num/denom 
```

```{r}
#Automate for all covariates -> make a table
covariate_balancing_df = matrix(NA,nrow = 13,ncol=3)
tracker = 1

for (i in 10:22){
  test1 = t.test(x = na.omit(inla_df[nested_group1,i]), y = na.omit(inla_df[-nested_group1,i]))
  test2 = t.test(x = na.omit(inla_df[nested_group2,i]), y = na.omit(inla_df[-nested_group2,i]))
  test3 = t.test(x = na.omit(inla_df[nested_group3,i]), y = na.omit(inla_df[-nested_group3,i]))
  test4 = t.test(x = na.omit(inla_df[nested_group4,i]), y = na.omit(inla_df[-nested_group4,i]))
  
  
  d1 = test1$estimate[1] - test1$estimate[2]
  d2 = test2$estimate[1] - test2$estimate[2]
  d3 = test3$estimate[1] - test3$estimate[2]
  d4 = test4$estimate[1] - test4$estimate[2]
  
  se1 = test1$stderr
  se2 = test2$stderr
  se3 = test3$stderr
  se4 = test4$stderr
  
  n1 = length(nested_group1) #nested_group size
  n2 = length(nested_group2)
  n3 = length(nested_group3)
  n4 = length(nested_group4)
  
  num = n1*d1 + n2*d2 + n3*d3 + n4*d4
  denom = sqrt(n1^2*se1^2 + n2^2*se2^2 + n3^2*se3^2 + n4^2*se4^2) 
  
  weighted_mean_diff = num / (n1+n2+n3+n4)
  se_diff = denom / (n1+n2+n3+n4)
  
  tstat = num/denom 
  
  covariate_balancing_df[tracker,] = c(weighted_mean_diff,se_diff,tstat)
  tracker = tracker+1
}

covariate_balancing_df = as.data.frame(covariate_balancing_df)
covariate_balancing_df = cbind(colnames(inla_df)[10:22],covariate_balancing_df)
colnames(covariate_balancing_df) = c("covariate","weighted_diff","se","t-statistic")
covariate_balancing_df = covariate_balancing_df[-9,]
```

# Modeling mortality

## KGR models

### Using K_cov

```{r,cache = TRUE}
kgr_model1_response_nb = function(dataset,link=1){
  
  ###Fit INLA model 
  kgr_formula1 = deaths ~ -1 + offset(log(Total_Pop)) + month + cum_monthly_total_pm25_6*Cluster + pm25 + total_precip + avg_air_temp + avg_dew_point + avg_wind_speed + f(id2,model = "generic0",Cmatrix = inv_covGP1_response)
  
  model = inla(formula = kgr_formula1,family = "nbinomial",data = dataset, num.threads = 1,
                  control.compute = list(dic=TRUE,waic=TRUE, config = TRUE,
                                         return.marginals.predictor=TRUE),
                  #control.family = list(variant = 1), E = Total_Pop,
                  control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
                  control.inla = list(strategy = "simplified.laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$NAME, dataset$time, dataset$ID, dataset$deaths, preds_model)
  colnames(preds_model) <- c("county", "time", "ID", "deaths", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")

  pp_ysamples = list()
  
  ###NBINOMIAL SAMPLING
  
  n_density = data.frame(model$marginals.hyperpar[[1]])
  sample_n = inla.rmarginal(1000,inla.smarginal(n_density))
  pp_nsamples = sample_n  
  
  ###Generate uncertainty bands from the posterior predictive distribution as opposed to posterior of the linear predictor
  for (i in 1:nrow(preds_model)){
    ###Sample 1000 mu for each row
    row = data.frame(model$marginals.fitted.values[[i]])

    sample_mu = inla.rmarginal(1000,inla.smarginal(row))
    lower = quantile(sample_mu,0.025)
    mean = mean(sample_mu)
    upper = quantile(sample_mu,0.975)
    
    ###Sample from nbinom with these a and b
    sample_y = rnbinom(1000, n = sample_n, mu = sample_mu)
    pp_ysamples[[i]] = sample_y
    
    ###Obtain 95% credible interval
    preds_model$y_ppmean[i] = mean(sample_y, na.rm = TRUE)
    preds_model$y_pplower[i] = quantile(sample_y, 0.025, na.rm = TRUE)
    preds_model$y_ppupper[i] = quantile(sample_y, 0.975, na.rm = TRUE)
  }
  
  # preds_model$interval_width = preds_model$`0.975quant` - preds_model$`0.025quant` 
  # marginal_fvs <- model$marginals.fitted.values

  # #Exponentiating parameter to get better interpretation of estimates 
  # multeff <- exp(model$summary.fixed$mean)
  # names(multeff) <- model$names.fixed
  # 
  # #Plot of each parameters' posterior density 
  # mf <- melt(model$marginals.fixed)
  # cf <- spread(mf,Var2,value)
  # names(cf)[2] <- 'parameter'
  # param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
  #            scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  # sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  # hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
  #            xlab("linear predictor")
  
  #Store the results in the list
  kgr_model1_results = list(
    covmatrix = covGP1_response,
    prec = inv_covGP1_response,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    #exp_effects = multeff,
    #param_plot = param_plot,
    #hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model,
    pp_ysamples = pp_ysamples,
    pp_nsamples = pp_nsamples
  )
  
  return(kgr_model1_results)
}

#Fit kgr_model1
kgr_model1_response_fit = kgr_model1_response_nb(dataset = inla_df)

#Extract DIC and WAIC
kgr_model1_DIC = kgr_model1_response_fit$model_DIC
kgr_model1_WAIC = kgr_model1_response_fit$model_WAIC

#Get summaries of parameter estimates
kgr_model1_response_fit$model_summary
kgr_model1_response_fit$bri_hyperpar_summary
kgr_model1_response_fit$exp_effects

#Show plots
kgr_model1_response_fit$param_plot
kgr_model1_response_fit$hyperparam_plot


for (i in 1:58){
df = inla_df %>% filter(ID == i) %>% dplyr::select(deaths)
preds = kgr_model1_response_fit$fitted_values %>% filter(ID == i) 
df = cbind(df,preds)

# title = sprintf("Posterior Predictive Fits for County %s",i)
title = sprintf("%s County",counties[i])

post_pred_plot = preds %>% ggplot(aes(x=time,y=deaths)) + geom_point() + 
  geom_line(aes(y=mean),color = "red") + geom_ribbon(aes(ymin = `0.025quant`,ymax = `0.975quant`), fill = "orange",alpha = 0.7) + 
  geom_ribbon(aes(ymin = y_pplower,ymax = y_ppupper),fill = "skyblue",alpha = 0.3) + ggtitle(title) + theme_classic()
print(post_pred_plot)
}

# kgr_model1_response_fit$fitted_values %>% group_by(county) %>% summarize(avg_width = mean(interval_width))
```

### Using K_time

```{r,cache = TRUE}
kgr_model2_response_nb = function(dataset,link=1){
  
  ###Fit INLA model 
  kgr_formula2 = deaths ~ -1 + offset(log(Total_Pop)) + month + cum_monthly_total_pm25_6*Cluster + pm25 + total_precip + avg_air_temp + avg_dew_point + avg_wind_speed + f(id2,model = "generic0",Cmatrix = inv_covGP2_response)

  
  model = inla(formula = kgr_formula2,family = "nbinomial",data = inla_df, num.threads = 20,
                  control.compute = list(dic=TRUE,waic=TRUE,return.marginals.predictor=TRUE),
                  control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
                  control.inla = list(strategy = "simplified.laplace"),
                control.predictor = list(compute = TRUE, link = 1))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$NAME, dataset$time, dataset$ID, dataset$deaths, preds_model)
  colnames(preds_model) <- c("county", "time", "ID", "deaths", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  
  pp_ysamples = list()
  
  ###NBINOMIAL SAMPLING
  
  n_density = data.frame(model$marginals.hyperpar[[1]])
  sample_n = inla.rmarginal(1000,inla.smarginal(n_density))
  pp_nsamples = sample_n
  
  ###Generate uncertainty bands from the posterior predictive distribution as opposed to posterior of the linear predictor
  for (i in 1:nrow(preds_model)){
    ###Sample 1000 mu for each row
    row = data.frame(model$marginals.fitted.values[[i]])

    sample_mu = inla.rmarginal(1000,inla.smarginal(row))
    lower = quantile(sample_mu,0.025)
    mean = mean(sample_mu)
    upper = quantile(sample_mu,0.975)
    
    # row$density = row$y / sum(row$y)
    # sample_mu = sample(row$x,size = 100000, replace = TRUE, prob = row$density)
    # lower = quantile(sample_mu,0.025)
    # mean = mean(sample_mu)
    # upper = quantile(sample_mu,0.975)
    
    ###For some reason these mean + lower/upper don't match summary.fitted.values?? Replace
    # preds_model$`0.025quant`[i] = lower
    # preds_model$mean[i] = mean
    # preds_model$`0.975quant`[i] = upper
    
    ###Sample 1000 n
    
    # n_density2 = data.frame(model$internal.marginals.hyperpar[[1]])
    # n_density$density = n_density$y / sum(n_density$y)
    
    # n_density = data.frame(model$marginals.hyperpar[[1]])
    # sample_n = inla.rmarginal(1000,inla.smarginal(n_density))
    # pp_nsamples = sample_n
  
    # n_density$density = n_density$y / sum(n_density$y)
    # sample_phi = sample(n_density$x,size = 100000, replace = TRUE, 
    #                     prob = n_density$density)

    ###Solve for p for each row
    # sample_p = sample_n / (sample_mu + sample_n)

    ###Sample from nbinom with these a and b
    sample_y = rnbinom(1000, n = sample_n, mu = sample_mu)
    pp_ysamples[[i]] = sample_y
    
    ###Obtain 95% credible interval
    preds_model$y_ppmean[i] = mean(sample_y, na.rm = TRUE)
    preds_model$y_pplower[i] = quantile(sample_y, 0.025, na.rm = TRUE)
    preds_model$y_ppupper[i] = quantile(sample_y, 0.975, na.rm = TRUE)
  }
  
  # preds_model$interval_width = preds_model$`0.975quant` - preds_model$`0.025quant` 
  # marginal_fvs <- model$marginals.fitted.values

  # #Exponentiating parameter to get better interpretation of estimates 
  # multeff <- exp(model$summary.fixed$mean)
  # names(multeff) <- model$names.fixed
  # 
  # #Plot of each parameters' posterior density 
  # mf <- melt(model$marginals.fixed)
  # cf <- spread(mf,Var2,value)
  # names(cf)[2] <- 'parameter'
  # param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
  #            scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  # sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  # hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
  #            xlab("linear predictor")
  
  #Store the results in the list
  kgr_model2_results = list(
    covmatrix = covGP2_response,
    prec = inv_covGP2_response,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    #exp_effects = multeff,
    #param_plot = param_plot,
    #hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model,
    pp_ysamples = pp_ysamples,
    pp_nsamples = pp_nsamples
  )
  
  return(kgr_model2_results)
}

#Fit kgr_model2
kgr_model2_response_fit = kgr_model2_response_nb(dataset = inla_df)

#Extract DIC and WAIC
kgr_model2_DIC = kgr_model2_response_fit$model_DIC
kgr_model2_WAIC = kgr_model2_response_fit$model_WAIC

#Get summaries of parameter estimates
kgr_model2_response_fit$model_summary
kgr_model2_response_fit$bri_hyperpar_summary
kgr_model2_response_fit$exp_effects

#Show plots
kgr_model2_response_fit$param_plot
kgr_model2_response_fit$hyperparam_plot


for (i in 1:58){
df = inla_df %>% filter(ID == i) %>% dplyr::select(deaths)
preds = kgr_model2_response_fit$fitted_values %>% filter(ID == i)
df = cbind(df,preds)

# title = sprintf("Posterior Predictive Fits for County %s",i)
title = sprintf("%s County",counties[i])

post_pred_plot = preds %>% ggplot(aes(x=time,y=deaths)) + geom_point() +
  geom_line(aes(y=mean),color = "red") + geom_ribbon(aes(ymin = `0.025quant`,ymax = `0.975quant`), fill = "orange",alpha = 0.7) +
  geom_ribbon(aes(ymin = y_pplower,ymax = y_ppupper),fill = "skyblue",alpha = 0.3) + ggtitle(title) + theme_classic()
print(post_pred_plot)
}

# kgr_model2_response_fit$fitted_values %>% group_by(county) %>% summarize(avg_width = mean(interval_width))
```

# Evaluate out of sample performance

## Attach identical dataset except with response omitted - estimate model then make predictions on second half of dataset

```{r}
inla_df2 = inla_df

# inla_df2$deaths = NA

out_idx = which(inla_df2$time >= 36 & inla_df2$time <= 48)
out_idx = c(out_idx,which(inla_df2$time >= 60))

#Switch to specific time period
inla_df2$deaths[out_idx] = NA

inla_df_out = rbind(inla_df,inla_df2)
```

## KGR models

### Using K_cov

```{r,cache = TRUE}
#Fit kgr_model1
kgr_model1_response_outfit = kgr_model1_response_nb(dataset = inla_df2)

#Extract DIC and WAIC
kgr_model1_DIC = kgr_model1_response_outfit$model_DIC
kgr_model1_WAIC = kgr_model1_response_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model1_response_outfit$model_summary
kgr_model1_response_outfit$bri_hyperpar_summary
kgr_model1_response_outfit$exp_effects

#Show plots
kgr_model1_response_outfit$param_plot
kgr_model1_response_outfit$hyperparam_plot


for (i in 1:58){
  df = inla_df %>% filter(ID == i) %>% dplyr::select(deaths)
  preds = kgr_model1_response_outfit$fitted_values %>% filter(ID == i) 
  preds$deaths = df$deaths
  
  # title = sprintf("Posterior Predictive Fits for County %s",i)
  title = sprintf("%s County",counties[i])
  
  post_pred_plot = preds %>% ggplot(aes(x = time, y = deaths)) + 
    geom_point() + 
    geom_line(aes(y = mean), color = "red") + 
    geom_ribbon(aes(ymin = `0.025quant`, ymax = `0.975quant`), fill = "green", alpha = 0.7) + 
    geom_ribbon(aes(ymin = y_pplower, ymax = y_ppupper), fill = "yellow", alpha = 0.5) + 
    
    # Shaded regions for out-of-sample predictions
    geom_rect(aes(xmin = 36, xmax = 48, ymin = -Inf, ymax = Inf), fill = "cyan", alpha = 0.005) +
    geom_rect(aes(xmin = 60, xmax = max(time), ymin = -Inf, ymax = Inf), fill = "cyan", alpha = 0.005) +
  
    # Vertical lines for reference
    geom_vline(xintercept = 36, linetype = "dashed", color = "blue", linewidth = 1.5) + 
    geom_vline(xintercept = 48, linetype = "dashed", color = "blue", linewidth = 1.5) +  
    geom_vline(xintercept = 60, linetype = "dashed", color = "yellow", linewidth = 1.5) + 
    ggtitle(title) + theme_classic()
  
  print(post_pred_plot)
}


# kgr_model1_response_outfit$fitted_values %>% group_by(county) %>% summarize(avg_width = mean(interval_width))
```

### Using K_time

```{r,cache = TRUE}
#Fit kgr_model2
kgr_model2_response_outfit = kgr_model2_response_nb(dataset = inla_df2)

#Extract DIC and WAIC
kgr_model2_DIC = kgr_model2_response_outfit$model_DIC
kgr_model2_WAIC = kgr_model2_response_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model2_response_outfit$model_summary
kgr_model2_response_outfit$bri_hyperpar_summary
kgr_model2_response_outfit$exp_effects

#Show plots
kgr_model2_response_outfit$param_plot
kgr_model2_response_outfit$hyperparam_plot

for (i in 1:58){
  df = inla_df %>% filter(ID == i) %>% dplyr::select(deaths)
  preds = kgr_model2_response_outfit$fitted_values %>% filter(ID == i) 
  preds$deaths = df$deaths
  
  # title = sprintf("Posterior Predictive Fits for County %s",i)
  title = sprintf("%s County",counties[i])
  
  post_pred_plot = preds %>% ggplot(aes(x = time, y = deaths)) + 
    geom_point() + 
    geom_line(aes(y = mean), color = "red") + 
    geom_ribbon(aes(ymin = `0.025quant`, ymax = `0.975quant`), fill = "green", alpha = 0.7) + 
    geom_ribbon(aes(ymin = y_pplower, ymax = y_ppupper), fill = "yellow", alpha = 0.5) + 
    
    # Shaded regions for out-of-sample predictions
    geom_rect(aes(xmin = 36, xmax = 48, ymin = -Inf, ymax = Inf), fill = "cyan", alpha = 0.005) +
    geom_rect(aes(xmin = 60, xmax = max(time), ymin = -Inf, ymax = Inf), fill = "cyan", alpha = 0.005) +
  
    # Vertical lines for reference
    geom_vline(xintercept = 36, linetype = "dashed", color = "blue", linewidth = 1.5) + 
    geom_vline(xintercept = 48, linetype = "dashed", color = "blue", linewidth = 1.5) +  
    geom_vline(xintercept = 60, linetype = "dashed", color = "yellow", linewidth = 1.5) + 
    ggtitle(title) + theme_classic()
  
  print(post_pred_plot)
}

# kgr_model2_response_outfit$fitted_values %>% group_by(county) %>% summarize(avg_width = mean(interval_width))
```

## Attach identical dataset except with treatment omitted - estimate model then make predictions on second half of dataset

```{r}
inla_df3 = inla_df

# inla_df3$cum_monthly_total_pm25_6 = NA

#Switch to specific time period
inla_df3$monthly_total_pm25[out_idx] = NA

inla_df_out = rbind(inla_df,inla_df3)
```

## KGR models

### Using K_cov

```{r,cache = TRUE}
#Fit kgr_model1
kgr_model1_treatment_outfit = kgr_model1_treatment(dataset = inla_df3)

#Extract DIC and WAIC
kgr_model1_DIC = kgr_model1_treatment_outfit$model_DIC
kgr_model1_WAIC = kgr_model1_treatment_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model1_treatment_outfit$model_summary
kgr_model1_treatment_outfit$bri_hyperpar_summary
kgr_model1_treatment_outfit$exp_effects

#Show plots
kgr_model1_treatment_outfit$param_plot
kgr_model1_treatment_outfit$hyperparam_plot


for (i in 1:58){
  df = inla_df %>% filter(ID == i) %>% dplyr::select(cum_monthly_total_pm25_6)
  preds = kgr_model1_treatment_outfit$fitted_values %>% filter(ID == i) 
  preds$cum_monthly_total_pm25_6 = df$cum_monthly_total_pm25_6
  
  # title = sprintf("Posterior Predictive Fits for County %s",i)
  title = sprintf("%s County",counties[i])
  
  post_pred_plot = preds %>% ggplot(aes(x = time, y = cum_monthly_total_pm25_6)) + 
    geom_point() + 
    geom_line(aes(y = mean), color = "red") + 
    geom_ribbon(aes(ymin = `0.025quant`, ymax = `0.975quant`), fill = "green", alpha = 0.7) + 
    geom_ribbon(aes(ymin = y_pplower, ymax = y_ppupper), fill = "yellow", alpha = 0.5) + 
    
    # Shaded regions for out-of-sample predictions
    geom_rect(aes(xmin = 36, xmax = 48, ymin = -Inf, ymax = Inf), fill = "cyan", alpha = 0.005) +
    geom_rect(aes(xmin = 60, xmax = max(time), ymin = -Inf, ymax = Inf), fill = "cyan", alpha = 0.005) +
  
    # Vertical lines for reference
    geom_vline(xintercept = 36, linetype = "dashed", color = "blue", linewidth = 1.5) + 
    geom_vline(xintercept = 48, linetype = "dashed", color = "blue", linewidth = 1.5) +  
    geom_vline(xintercept = 60, linetype = "dashed", color = "yellow", linewidth = 1.5) + 
    ggtitle(title) + theme_classic()
  
  print(post_pred_plot)
}
# kgr_model1_treatment_outfit$fitted_values %>% group_by(county) %>% summarize(avg_width = mean(interval_width))
```

### Using K_time

```{r,cache = TRUE}
#Fit kgr_model2
kgr_model2_treatment_outfit = kgr_model2_treatment(dataset = inla_df3)

#Extract DIC and WAIC
kgr_model2_DIC = kgr_model2_treatment_outfit$model_DIC
kgr_model2_WAIC = kgr_model2_treatment_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model2_treatment_outfit$model_summary
kgr_model2_treatment_outfit$bri_hyperpar_summary
kgr_model2_treatment_outfit$exp_effects

#Show plots
kgr_model2_treatment_outfit$param_plot
kgr_model2_treatment_outfit$hyperparam_plot


for (i in 1:58){
  df = inla_df %>% filter(ID == i) %>% dplyr::select(cum_monthly_total_pm25_6)
  preds = kgr_model2_treatment_outfit$fitted_values %>% filter(ID == i) 
  preds$cum_monthly_total_pm25_6 = df$cum_monthly_total_pm25_6
  
  # title = sprintf("Posterior Predictive Fits for County %s",i)
  title = sprintf("%s County",counties[i])
  
  post_pred_plot = preds %>% ggplot(aes(x = time, y = cum_monthly_total_pm25_6)) + 
    geom_point() + 
    geom_line(aes(y = mean), color = "red") + 
    geom_ribbon(aes(ymin = `0.025quant`, ymax = `0.975quant`), fill = "green", alpha = 0.7) + 
    geom_ribbon(aes(ymin = y_pplower, ymax = y_ppupper), fill = "yellow", alpha = 0.5) + 
    
    # Shaded regions for out-of-sample predictions
    geom_rect(aes(xmin = 36, xmax = 48, ymin = -Inf, ymax = Inf), fill = "cyan", alpha = 0.005) +
    geom_rect(aes(xmin = 60, xmax = max(time), ymin = -Inf, ymax = Inf), fill = "cyan", alpha = 0.005) +
  
    # Vertical lines for reference
    geom_vline(xintercept = 36, linetype = "dashed", color = "blue", linewidth = 1.5) + 
    geom_vline(xintercept = 48, linetype = "dashed", color = "blue", linewidth = 1.5) +  
    geom_vline(xintercept = 60, linetype = "dashed", color = "yellow", linewidth = 1.5) + 
    ggtitle(title) + theme_classic()
  
  print(post_pred_plot)
}

# kgr_model2_treatment_outfit$fitted_values %>% group_by(county) %>% summarize(avg_width = mean(interval_width))
```

## Calculate coverage tables for treatment and outcome

### Treatment

In and out of sample:

```{r}
### County level
coverage_county = rep(0,58)

for (i in 1:58){
  county_idx = which(inla_df$NAME == CA_counties[i])
  true_treatment = inla_df$cum_monthly_total_pm25_6[county_idx]
  treatment_preds = kgr_model1_treatment_outfit$fitted_values[county_idx,] #change model if necessary
  
  lci = treatment_preds %>% dplyr::select(y_pplower)
  lci = lci$y_pplower - 0.0001
  
  uci = treatment_preds %>% dplyr::select(y_ppupper)
  uci = uci$y_ppupper - 0.0001
  
  captured = (true_treatment >= lci & true_treatment <= uci)
  coverage_county[i] = mean(captured)
}

#Overall 
coverage_overall = 0
true_treatment = inla_df$cum_monthly_total_pm25_6

all_lci = kgr_model1_treatment_outfit$fitted_values %>% dplyr::select(y_pplower)
all_lci = all_lci$y_pplower - 0.0001
  
all_uci = kgr_model1_treatment_outfit$fitted_values %>% dplyr::select(y_ppupper)
all_uci = all_uci$y_ppupper - 0.0001

captured = (true_treatment >= all_lci & true_treatment <= all_uci)
coverage_overall = mean(captured)


#Show results
coverage_county = cbind(CA_counties,coverage_county)
coverage_county

coverage_overall
```

Just out of sample:

```{r}
true_vals = inla_df[out_idx,]
inla_fits = kgr_model1_treatment_outfit$fitted_values[out_idx,] #change model if necessary

### County level
coverage_county2 = rep(0,58)

for (i in 1:58){
  county_idx = which(true_vals$NAME == CA_counties[i])
  true_treatment = true_vals$cum_monthly_total_pm25_6[county_idx]
  treatment_preds = inla_fits[county_idx,] 
  
  lci = treatment_preds %>% dplyr::select(y_pplower)
  lci = lci$y_pplower - 0.0001
  
  uci = treatment_preds %>% dplyr::select(y_ppupper)
  uci = uci$y_ppupper - 0.0001
  
  captured = (true_treatment >= lci & true_treatment <= uci)
  coverage_county2[i] = mean(captured)
}

#Overall 
coverage_overall2 = 0
true_treatment = true_vals$cum_monthly_total_pm25_6

all_lci = inla_fits %>% dplyr::select(y_pplower)
all_lci = all_lci$y_pplower - 0.0001
  
all_uci = inla_fits %>% dplyr::select(y_ppupper)
all_uci = all_uci$y_ppupper - 0.0001

captured = (true_treatment >= all_lci & true_treatment <= all_uci)
coverage_overall2 = mean(captured)


#Show results
coverage_county2 = cbind(CA_counties,coverage_county2)
coverage_county2

coverage_overall2
```

### Outcome

In and out of sample:

```{r}
### County level
coverage_county3 = rep(0,58)

for (i in 1:58){
  county_idx = which(inla_df$NAME == CA_counties[i])
  true_outcome = inla_df$deaths[county_idx]
  outcome_preds = kgr_model1_response_outfit$fitted_values[county_idx,] #change model if necessary
  
  lci = outcome_preds %>% dplyr::select(y_pplower)
  uci = outcome_preds %>% dplyr::select(y_ppupper)
  
  captured = (true_outcome >= lci$y_pplower & true_outcome <= uci$y_ppupper)
  coverage_county3[i] = mean(captured)
}

#Overall 
coverage_overall3 = 0
true_outcome = inla_df$deaths

all_lci = kgr_model1_response_outfit$fitted_values %>% dplyr::select(y_pplower)
all_uci = kgr_model1_response_outfit$fitted_values %>% dplyr::select(y_ppupper)

captured = (true_outcome >= all_lci$y_pplower & true_outcome <= all_uci$y_ppupper)
coverage_overall3 = mean(captured)


#Show results
coverage_county3 = cbind(CA_counties,coverage_county3)
coverage_county3

coverage_overall3
```

Just out of sample:

```{r}
true_vals = inla_df[out_idx,]
inla_fits = kgr_model1_response_outfit$fitted_values[out_idx,]

### County level
coverage_county4 = rep(0,58)

for (i in 1:58){
  county_idx = which(true_vals$NAME == CA_counties[i])
  true_outcome = true_vals$deaths[county_idx]
  outcome_preds = inla_fits[county_idx,] #change model if necessary
  
  lci = outcome_preds %>% dplyr::select(y_pplower)
  uci = outcome_preds %>% dplyr::select(y_ppupper)
  
  captured = (true_outcome >= lci$y_pplower & true_outcome <= uci$y_ppupper)
  coverage_county4[i] = sum(captured)/length(captured)
}

#Overall 
coverage_overall4 = 0
true_outcome = true_vals$deaths

all_lci = inla_fits %>% dplyr::select(y_pplower)
all_uci = inla_fits %>% dplyr::select(y_ppupper)

captured = (true_outcome >= all_lci$y_pplower & true_outcome <= all_uci$y_ppupper)
coverage_overall4 = mean(captured)


#Show results
coverage_county4 = cbind(CA_counties,coverage_county4)
coverage_county4

coverage_overall4
```

# Causal inference with MSMs

## Use covariate history instead of instantaneous values 

## GLM outcome model

```{r}
library(MASS) 

#Create scaled version of exposure
inla_df$cum_monthly_total_pm25_6_scaled = (inla_df$cum_monthly_total_pm25_6 - mean(inla_df$cum_monthly_total_pm25_6)) / sd(inla_df$cum_monthly_total_pm25_6)

wildfire_szn_df$cum_monthly_total_pm25_6_scaled = (wildfire_szn_df$cum_monthly_total_pm25_6 - mean(wildfire_szn_df$cum_monthly_total_pm25_6)) / sd(wildfire_szn_df$cum_monthly_total_pm25_6)

#Weighted NB regression
causal_glm1 = glm.nb(deaths ~ offset(log(Total_Pop)) + factor(month) + aqi_resids + cum_total_precip_6 + avg_avg_air_temp_6 + avg_avg_dew_point_6 + max_avg_wind_speed_6 +  cum_monthly_total_pm25_6_scaled + I(cum_monthly_total_pm25_6_scaled^2), weights = stabilized_IPW_kgr, data = inla_df)

summary(causal_glm1)

# NB regression with SWs included as covariate
causal_glm2 = glm.nb(deaths ~ offset(log(Total_Pop)) + factor(month) + aqi_resids + cum_total_precip_6 + avg_avg_air_temp_6 + avg_avg_dew_point_6 + max_avg_wind_speed_6 +  cum_monthly_total_pm25_6_scaled + I(cum_monthly_total_pm25_6_scaled^2) + stabilized_IPW_kgr, data = inla_df)

summary(causal_glm2)

detach("package:MASS", unload = TRUE) 
```

## KGR outcome model

```{r,cache = TRUE}
###Fit INLA model 
# kgr_formula2 = deaths ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + factor(month) + aqi_resids + cum_total_precip_6 + avg_avg_air_temp_6 + avg_avg_dew_point_6 + max_avg_wind_speed_6 +  cum_monthly_total_pm25_6_scaled + I(cum_monthly_total_pm25_6_scaled^2) + sw3 + f(id2,model = "generic0",Cmatrix = inv_covGP2)

kgr_formula2 = deaths ~ -1 + offset(log(Total_Pop)) + factor(month) + aqi_resids + cum_total_precip_6 + avg_avg_air_temp_6 + avg_avg_dew_point_6 + max_avg_wind_speed_6 +  Cluster*cum_monthly_total_pm25_6_scaled + sw_prod + f(id2,model = "generic0",Cmatrix = inv_covGP2_response)

clus_id = as.integer(inla_df$Cluster)

model = inla(formula = kgr_formula2,family = "nbinomial",data = inla_df, num.threads = 20,
                control.compute = list(dic=TRUE,waic=TRUE,
                                       return.marginals.predictor=TRUE),
                #control.family = list(variant = 1), E = Total_Pop,
                control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
                control.inla = list(strategy = "simplified.laplace"),
                control.predictor = list(compute = TRUE, link = 1))

###Extract relevant information and store in the list
model$summary.fixed

# model_summary <- model$summary.fixed
# bri_hyperpar_summary <- bri.hyperpar.summary(model)
# model_DIC <- model$dic$dic
# model_WAIC <- model$waic$waic
# preds_model <- model$summary.fitted.values
# preds_model <- cbind(inla_df$NAME, inla_df$time, inla_df$ID, inla_df$deaths, preds_model)
# colnames(preds_model) <- c("county", "time", "ID", "deaths", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
```

Interpretations:

cum_monthly_total_pm25_6 coefficient = 5.946441e-04 -\> exp(5.946441e-04) = 1.000595

-\>

Let's interpret for Santa Cruz, Los Angeles, Mariposa: An increase in 1ppm of wildfire PM2.5 over the last 6 months leads to a

SC death rate increases from 5.48 to 5.4832 -\> 15.013 to 15.022 total deaths

LA death rate increases from 8.40 to 8.405 -\> 848.881 to 849.386 total deaths

Mariposa death rate increases from 21.9 to 21.913 -\> 3.815 to 3.817 total deaths

cum_monthly_total_pm25_6_scaled coefficient = -\> exp() = ??

-\>

Let's interpret for Santa Cruz, Los Angeles, Mariposa: An increase in 1 sd of wildfire PM2.5 over the last 6 months leads to a

SC death rate increases from 5.48 to 5.4832 -\> 15.013 to 15.022 total deaths

LA death rate increases from 8.40 to 8.405 -\> 848.881 to 849.386 total deaths

Mariposa death rate increases from 21.9 to 21.913 -\> 3.815 to 3.817 total deaths


## Plotting monthly effects

```{r,eval = FALSE}
m_beta = model$summary.fixed[1:12,1]
month = c("January","February","March","April","May","June","July",
          "August","September","October","November","December")
monthly_fe = data.frame(month = factor(month, levels = month.name),m_beta = m_beta, value = exp(m_beta))


ggplot(monthly_fe, aes(x = month, y = value)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Monthly Fixed Effects", x = "Month", y = "Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Modeling a nonlinear function between treatment and outcome (DRF)

## FINAL OUTCOME MODEL

Setting up the spde: 

```{r}
inla_df$cum_monthly_total_pm25_6_binned = inla.group(inla_df$cum_monthly_total_pm25_6,n=25,method = "quantile")
inla_df$Cluster = as.numeric(inla_df$Cluster)
inla_df$cum_monthly_total_pm25_6_binned_rep = inla_df$cum_monthly_total_pm25_6_binned

delta = 0.15 * diff(range(inla_df$cum_monthly_total_pm25_6_binned))
K = 20
mesh = fm_mesh_1d(seq(min(inla_df$cum_monthly_total_pm25_6_binned) - delta, 
                      max(inla_df$cum_monthly_total_pm25_6_binned) + delta, length.out = K),
                  boundary = "neumann", degree = 2)

spde = inla.spde2.pcmatern(mesh, alpha = 2, prior.range = c(diff(range(inla_df$cum_monthly_total_pm25_6_binned))*2,0.5), 
                           prior.sigma = c(0.01,0.1)) #strong shrinkage: (0.001,0.1), no shrinkage: (1,0.1), vague: (5,0.01)

# inla_df$Z_global = inla_df$cum_monthly_total_pm25_6_binned
inla_df$Z_cluster = sapply(inla_df$cum_monthly_total_pm25_6_binned, function(z) which.min(abs(mesh$loc - z)))



inla_df$sw_binned = inla.group(inla_df$sw_prod,n=20,method = "quantile")

prior_fixed = list(mean = list(default = 0), 
                   prec = list(default = 0.001,`Group2:1` = 4,
                               `Group3:1` = 4,`Group4:1` = 4,
                               `Group5:1` = 4,`Group2:2` = 4,
                               `Group3:2` = 4,`Group4:2` = 4,
                               `Group5:2` = 4,`Group2:3` = 4,
                               `Group3:3` = 4,`Group4:3` = 4,
                               `Group5:3` = 4))
```

```{r,cache = TRUE}
# kgr_formula_hierarchical3 = deaths ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + 
#   Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + 
#   Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + 
#   Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + 
#   Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + 
#   Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + 
#   Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + 
#   Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + 
#   Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + factor(month) + 
#   aqi_resids +
#   cum_total_precip_6 +
#   avg_avg_air_temp_6 +
#   avg_avg_dew_point_6 +
#   max_avg_wind_speed_6 +
#   f(sw_binned,model = "rw2") +
#   # Global PM2.5 dose-response curve
#   ns(cum_monthly_total_pm25_6_binned,df=3) +
#   # Cluster-specific deviations from global curve (shrunken toward zero)
#   f(Z_cluster, model = spde, group = Cluster, control.group = list(model = "exchangeable")) +
#   f(ID,model = "generic0",Cmatrix = solve(H^2), group = time, control.group = list(model = "ar",order = 6))

# kgr_formula_hierarchical3 = deaths ~ offset(log(Total_Pop/100000)) + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + 
#   Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + 
#   Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + 
#   Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + 
#   Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + 
#   Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + 
#   Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + 
#   Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + 
#   Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + factor(month) + 
#   aqi_resids +
#   cum_total_precip_6 +
#   avg_avg_air_temp_6 +
#   avg_avg_dew_point_6 +
#   max_avg_wind_speed_6 +
#   f(sw_binned,model = "rw2") +
#   # Spline for treatment * SDI group 
#   Group*ns(cum_monthly_total_pm25_6_binned,df=3) +
#   f(ID,model = "generic0",Cmatrix = solve(H^2), group = time, control.group = list(model = "ar",order = 6))

kgr_formula_hierarchical3 = deaths ~ offset(log(Total_Pop/100000)) +
  f(NAME,model = "iid",hyper = list(prec = list(prior = "loggamma", param = c(10, 5))))+
  factor(month) + 
  aqi_resids +
  cum_total_precip_6 +
  avg_avg_air_temp_6 +
  avg_avg_dew_point_6 +
  max_avg_wind_speed_6 +
  f(sw_binned,model = "rw2") +
  # Spline for treatment * SDI group 
  Group*ns(cum_monthly_total_pm25_6_binned,df=3) +
  f(ID,model = "generic0",Cmatrix = solve(H^2), group = time, control.group = list(model = "ar",order = 6))

# kgr_formula_hierarchical3 = deaths ~ -1 +
#   offset(log(Total_Pop)) +
#   factor(month) +
#   aqi_resids +
#   cum_total_precip_6 +
#   avg_avg_air_temp_6 +
#   avg_avg_dew_point_6 +
#   max_avg_wind_speed_6 +
#   sw_prod + I(sw_prod^2) + I(sw_prod^3) +
#   # Global PM2.5 dose-response curve
#   ns(cum_monthly_total_pm25_6_binned,df=3) +
#   # Cluster-specific deviations from global curve (shrunken toward zero)
#   f(Z_cluster, model = spde, group = Cluster, control.group = list(model = "exchangeable")) +
#   f(ID,model = "generic0",Cmatrix = solve(H^2), group = time, control.group = list(model = "ar",order = 6))

tic()
model3 = inla(formula = kgr_formula_hierarchical3,family = "nbinomial",data = inla_df, num.threads = 20,
                control.compute = list(dic=TRUE,waic=TRUE,return.marginals.predictor=TRUE),
                control.fixed = prior_fixed,
                control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
                control.inla = list(strategy = "simplified.laplace"),
                control.predictor = list(compute = TRUE, link = 1))
toc()

model_summary <- model3$summary.fixed
bri_hyperpar_summary <- bri.hyperpar.summary(model3)
model_DIC <- model3$dic$dic
model_WAIC <- model3$waic$waic
preds_model <- model3$summary.fitted.values
preds_model <- cbind(inla_df$NAME, inla_df$time, inla_df$ID, inla_df$deaths, preds_model)
colnames(preds_model) <- c("county", "time", "ID", "deaths", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")

final_model_results = list(
  model_summary = model_summary,
  bri_hyperpar_summary = bri_hyperpar_summary,
  model_DIC = model_DIC,
  model_WAIC = model_WAIC,
  fitted_values = preds_model
)

model_summary
```


### Plotting DRF from coefficients

```{r}
base_spline_summary = model_summary[80:82,]
cluster2_summary = model_summary[c(83,87,91),]
cluster3_summary = model_summary[c(84,88,92),]
cluster4_summary = model_summary[c(85,89,93),]
cluster5_summary = model_summary[c(86,90,94),]

cluster1_coefs = base_spline_summary$mean
cluster2_coefs = cluster1_coefs + cluster2_summary$mean
cluster3_coefs = cluster1_coefs + cluster3_summary$mean
cluster4_coefs = cluster1_coefs + cluster4_summary$mean
cluster5_coefs = cluster1_coefs + cluster5_summary$mean

spline_basis = ns(unique(inla_df$cum_monthly_total_pm25_6_binned), df = 3)

cluster1_drf <- as.numeric(spline_basis %*% cluster1_coefs)
cluster2_drf <- as.numeric(spline_basis %*% cluster2_coefs)
cluster3_drf <- as.numeric(spline_basis %*% cluster3_coefs)
cluster4_drf <- as.numeric(spline_basis %*% cluster4_coefs)
cluster5_drf <- as.numeric(spline_basis %*% cluster5_coefs)

# For uncertainty bands (simplified approach)
# Extract standard errors
cluster1_se = base_spline_summary$sd
cluster2_se = sqrt(cluster1_se^2 + cluster2_summary$sd^2)
cluster3_se = sqrt(cluster1_se^2 + cluster3_summary$sd^2)
cluster4_se = sqrt(cluster1_se^2 + cluster4_summary$sd^2)
cluster5_se = sqrt(cluster1_se^2 + cluster5_summary$sd^2)

# Calculate uncertainty bands
cluster1_drf_se <- sqrt(rowSums((spline_basis %*% diag(cluster1_se^2)) * spline_basis))
cluster2_drf_se <- sqrt(rowSums((spline_basis %*% diag(cluster2_se^2)) * spline_basis))
cluster3_drf_se <- sqrt(rowSums((spline_basis %*% diag(cluster3_se^2)) * spline_basis))
cluster4_drf_se <- sqrt(rowSums((spline_basis %*% diag(cluster4_se^2)) * spline_basis))
cluster5_drf_se <- sqrt(rowSums((spline_basis %*% diag(cluster5_se^2)) * spline_basis))

# Combine into data frame
z_values = unique(inla_df$cum_monthly_total_pm25_6_binned)

cluster_curves <- rbind(
  data.frame(Z = z_values, DRF = cluster1_drf, 
             lower = cluster1_drf - 1.96 * cluster1_drf_se,
             upper = cluster1_drf + 1.96 * cluster1_drf_se,
             Cluster = "1"),
  data.frame(Z = z_values, DRF = cluster2_drf,
             lower = cluster2_drf - 1.96 * cluster2_drf_se,
             upper = cluster2_drf + 1.96 * cluster2_drf_se,
             Cluster = "2"),
  data.frame(Z = z_values, DRF = cluster3_drf,
             lower = cluster3_drf - 1.96 * cluster3_drf_se,
             upper = cluster3_drf + 1.96 * cluster3_drf_se,
             Cluster = "3"),
  data.frame(Z = z_values, DRF = cluster4_drf,
             lower = cluster4_drf - 1.96 * cluster4_drf_se,
             upper = cluster4_drf + 1.96 * cluster4_drf_se,
             Cluster = "4"),
  data.frame(Z = z_values, DRF = cluster5_drf,
             lower = cluster5_drf - 1.96 * cluster5_drf_se,
             upper = cluster5_drf + 1.96 * cluster5_drf_se,
             Cluster = "5")
)

# Define colors
my_colors <- c("darkblue", "purple", "magenta", "orange", "yellow")

# Plot
ggplot(cluster_curves, aes(x = Z, y = DRF, color = Cluster, fill = Cluster)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, color = NA) +
  geom_line(size = 1) +
  scale_color_manual(values = my_colors) +
  scale_fill_manual(values = my_colors) +
  labs(x = "Cumulative PM2.5 (binned)", 
       y = "DRF Effect",
       title = "Cluster-Specific Dose-Response Functions",
       subtitle = "With 95% Confidence Bands") +
  theme_minimal()
```

## Posterior predictive check for zero inflation

```{r,cache = TRUE}
zip_formula = deaths ~ -1 + offset(log(Total_Pop)) + 
  factor(month) + 
  aqi_resids +
  cum_total_precip_6 +
  avg_avg_air_temp_6 +
  avg_avg_dew_point_6 +
  max_avg_wind_speed_6 +  
  f(sw_binned,model = "rw2") +
  group*ns(cum_monthly_total_pm25_6_binned,df=3) +
  f(ID,model = "generic0",Cmatrix = solve(H^2), group = time, control.group = list(model = "ar",order = 6))


###Try zeroinflatedpoisson0, zeroinflatedpoisson1, zeroinflatednbinomial0, zeroinflatednbinomial1
zip_model = inla(formula = zip_formula,family = "zeroinflatedpoisson0",data = inla_df, num.threads = 1,
                control.compute = list(dic=TRUE,waic=TRUE,return.marginals.predictor=TRUE),
                #control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
                control.inla = list(strategy = "simplified.laplace"),
                control.predictor = list(compute = TRUE, link = 1))

preds_zip_model <- zip_model$summary.fitted.values
preds_zip_model <- cbind(inla_df$NAME, inla_df$time, inla_df$ID, inla_df$deaths, preds_zip_model)
colnames(preds_zip_model) <- c("county", "time", "ID", "deaths", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
```

```{r}
kgr_formula_hierarchical3 = deaths ~ -1 + offset(log(Total_Pop)) + 
  factor(month) + 
  aqi_resids +
  cum_total_precip_6 +
  avg_avg_air_temp_6 +
  avg_avg_dew_point_6 +
  max_avg_wind_speed_6 +
  f(sw_binned,model = "rw2") +
  group*ns(cum_monthly_total_pm25_6_binned,df=3) +
  f(ID,model = "generic0",Cmatrix = solve(H^2), group = time, control.group = list(model = "ar",order = 6))

model3 = inla(formula = kgr_formula_hierarchical3,family = "poisson",data = inla_df, num.threads = 20,
                control.compute = list(dic=TRUE,waic=TRUE,return.marginals.predictor=TRUE),
                #control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
                control.inla = list(strategy = "simplified.laplace"),
                control.predictor = list(compute = TRUE, link = 1))
```

```{r}
###ZIP
###Posterior predictive sampling
p_density = data.frame(zip_model$marginals.hyperpar$`zero-probability parameter for zero-inflated poisson_0`)
sample_p = inla.rmarginal(100000,inla.smarginal(p_density)) 
pp_psamples = sample_p  

pp_zip_samples = list()

###Generate uncertainty bands from the posterior predictive distribution as opposed to posterior of the linear predictor
for (i in 1:nrow(preds_zip_model)){ ###RENDER ERROR HERE
  ###Sample 1000 mu for each row
  row = data.frame(zip_model$marginals.fitted.values[[i]])

  sample_lambda = inla.rmarginal(100000,inla.smarginal(row))
  lower = quantile(sample_lambda,0.025)
  mean = mean(sample_lambda)
  upper = quantile(sample_lambda,0.975)

  sample_y = ifelse(rbinom(100000, size = 1, prob = sample_p) > 0, 0, rpois(100000, lambda = sample_lambda))
  pp_zip_samples[[i]] = sample_y
}

###Poisson
preds_pois_model <- model3$summary.fitted.values
preds_pois_model <- cbind(inla_df$NAME, inla_df$time, inla_df$ID, inla_df$deaths, preds_pois_model)
colnames(preds_pois_model) <- c("county", "time", "ID", "deaths", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")

###Posterior predictive sampling
pp_pois_samples = list()

###Generate uncertainty bands from the posterior predictive distribution as opposed to posterior of the linear predictor
for (i in 1:nrow(preds_pois_model)){
    ###Sample 1000 mu for each row
    row = data.frame(model3$marginals.fitted.values[[i]])
  
    sample_lambda = inla.rmarginal(100000,inla.smarginal(row))
    lower = quantile(sample_lambda,0.025)
    mean = mean(sample_lambda)
    upper = quantile(sample_lambda,0.975)
  
    sample_y = rpois(n=100000,lambda = sample_lambda)
    pp_pois_samples[[i]] = sample_y
  }

###PPC
true_prop0 = length(which(inla_df$deaths == 0)) / nrow(inla_df)
true_mv = mean(inla_df$deaths) / var(inla_df$deaths)

# ###ZIP
pp_zip_sample_df = do.call(rbind,pp_zip_samples)

zip_prop0 = c()
zip_mv = c()

for(i in 1:1000){
  deaths = pp_zip_sample_df[,i]
  zip_prop0[i] = length(which(deaths == 0)) / length(deaths)
  zip_mv[i] = mean(deaths) / var(deaths)
}


# ###POISSON
pp_pois_sample_df = do.call(rbind,pp_pois_samples)

pois_prop0 = c()
pois_mv = c()

for(i in 1:1000){
  deaths = pp_pois_sample_df[,i]
  pois_prop0[i] = length(which(deaths == 0)) / length(deaths)
  pois_mv[i] = mean(deaths) / var(deaths)
}


hist(zip_prop0,xlim = c(0.1,0.3),main = "ZIP Prop 0")
abline(v = true_prop0)

hist(pois_prop0,xlim = c(0.1,0.3),main = "Poisson Prop 0")
abline(v = true_prop0)


hist(zip_mv,xlim = c(0.006,0.008),main = "ZIP Mean/Var")
abline(v = true_mv)

hist(pois_mv,xlim = c(0.006,0.008),main = "Poisson Mean/Var")
abline(v = true_mv)


###Split into groups to parse signal
small_counties = na.omit(counties[which(CApops$POPESTIMATE2019 <= 100000)])
small_idx = which(inla_df$NAME %in% small_counties)
large_idx = which(!(inla_df$NAME %in% small_counties))

inla_df_small = inla_df[small_idx,]
inla_df_large = inla_df[large_idx,]

###Small
small_prop0 = length(which(inla_df_small$deaths == 0)) / nrow(inla_df_small)
small_mv = mean(inla_df_small$deaths) / var(inla_df_small$deaths)

zip_small_prop0 = c()
zip_small_mv = c()

for(i in 1:100000){
  deaths = pp_zip_sample_df[small_idx,i]
  zip_small_prop0[i] = length(which(deaths == 0)) / length(deaths)
  zip_small_mv[i] = mean(deaths) / var(deaths) 
}


pois_small_prop0 = c()
pois_small_mv = c()

for(i in 1:100000){
  deaths = pp_pois_sample_df[small_idx,i]
  pois_small_prop0[i] = length(which(deaths == 0)) / length(deaths)
  pois_small_mv[i] = mean(deaths) / var(deaths) 
}

hist(zip_small_prop0,xlim = c(0.1,0.3),main = "ZIP Prop 0 (Small Counties)")
abline(v = small_prop0)

hist(pois_small_prop0,xlim = c(0.1,0.3),main = "Poisson Prop 0 (Small Counties)")
abline(v = small_prop0)


hist(zip_small_mv,xlim = c(0.004,0.006),main = "ZIP Mean/Var (Small Counties)")
abline(v = small_mv)

hist(pois_small_mv,xlim = c(0.004,0.006),main = "Poisson Mean/Var (Small Counties)")
abline(v = small_mv)

###Large
large_prop0 = length(which(inla_df_large$deaths == 0)) / nrow(inla_df_large)
large_mv = mean(inla_df_large$deaths) / var(inla_df_large$deaths)

zip_large_prop0 = c()
zip_large_mv = c()

for(i in 1:100000){
  deaths = pp_zip_sample_df[large_idx,i]
  zip_large_prop0[i] = length(which(deaths == 0)) / length(deaths)
  zip_large_mv[i] = mean(deaths) / var(deaths) 
}


pois_large_prop0 = c()
pois_large_mv = c()

for(i in 1:100000){
  deaths = pp_pois_sample_df[large_idx,i]
  pois_large_prop0[i] = length(which(deaths == 0)) / length(deaths)
  pois_large_mv[i] = mean(deaths) / var(deaths) 
}

hist(zip_large_prop0,xlim = c(0.1,0.3),main = "ZIP Prop 0 (Large Counties)")
abline(v = large_prop0)

hist(pois_large_prop0,xlim = c(0.1,0.3),main = "Poisson Prop 0 (Large Counties)")
abline(v = large_prop0)


hist(zip_large_mv,xlim = c(0.02,0.025),main = "ZIP Mean/Var (Large Counties)")
abline(v = large_mv)

hist(pois_large_mv,xlim = c(0.02,0.025),main = "Poisson Mean/Var (Large Counties)")
abline(v = large_mv)
```

## Sensitivity analysis for unmeasured confounding

```{r,cache = TRUE}
#Create negative control proxies (Z = future exposure, W = past outcome)
inla_df <- inla_df %>%
  group_by(ID) %>%
  mutate(
    W1 = lag(deaths, 1),
    W4 = lag(deaths, 3),
    W8 = lag(deaths, 6),
    W12 = lag(deaths, 12),
    Zp1 = lead(cum_monthly_total_pm25_6,1),
    Zp4 = lead(cum_monthly_total_pm25_6,3),
    Zp8 = lead(cum_monthly_total_pm25_6,6),
    Zp12 = lead(cum_monthly_total_pm25_6,12)
  ) %>% ungroup()

#Fit model with proxies to see if nonexistent associations appear 
kgr_formula_hierarchical_wproxy = W1 ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + 
  Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + 
  Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + 
  Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + 
  Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + 
  Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + 
  Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + 
  Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + 
  Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + factor(month) + 
  aqi_resids +
  cum_total_precip_6 +
  avg_avg_air_temp_6 +
  avg_avg_dew_point_6 +
  max_avg_wind_speed_6 +
  f(sw_binned,model = "rw2") +
  # Global PM2.5 dose-response curve
  ns(cum_monthly_total_pm25_6_binned,df=3) + Zp1 +
  # Cluster-specific deviations from global curve (shrunken toward zero)
  f(Z_cluster, model = spde, group = Cluster, control.group = list(model = "exchangeable")) +
  f(ID,model = "generic0",Cmatrix = solve(H^2), group = time, control.group = list(model = "ar",order = 6))

proxy_model = inla(formula = kgr_formula_hierarchical_wproxy,family = "nbinomial",data = inla_df, num.threads = 5,
                control.compute = list(dic=TRUE,waic=TRUE,return.marginals.predictor=TRUE),
                control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
                control.inla = list(strategy = "simplified.laplace"),
                control.predictor = list(compute = TRUE, link = 1))

proxy_model$summary.fixed
proxy_model$summary.hyperpar
```

### FOR REFERENCE: 

```{r}
detach("package:MASS", unload = TRUE) 
plot((CA_sf %>% mutate(clus = clus5_min$groups))['clus'], main = "5 cluster example with population constraint")
```

```{r}
inla_df %>% group_by(NAME) %>% summarize(avg_z = mean(cum_monthly_total_pm25_6), max_z = max(cum_monthly_total_pm25_6), avg_y = mean(deaths), max_y = max(deaths), avg_yrate = mean(death_rate), max_yrate = max(death_rate)) %>% arrange(max_z)
```

# What is DRF? DRF models the expected value of potential outcomes Y as a function of the treatment level and covariates

## Example with z = 1000

```{r,eval = FALSE}
#Attach duplicate dataset with specified treatment/dose level 
# duplicate_df = inla_df
duplicate_df = inla_df %>% filter(NAME %in% counties[46:58])
# duplicate_df = inla_df %>% filter(NAME == "Kings")

duplicate_df$cum_monthly_total_pm25_6_binned = 1000
duplicate_df$deaths = NA 

inla_cf_df = rbind(inla_df,duplicate_df)
cf_idx = c((nrow(inla_df)+1):nrow(inla_cf_df))

prior_fixed = list(mean = list(default = 0), 
                   prec = list(default = 0.001,`group2:1` = 5,
                               `group3:1` = 5,`group4:1` = 5,
                               `group5:1` = 5,`group2:2` = 5,
                               `group3:2` = 5,`group4:2` = 5,
                               `group5:2` = 5,`group2:3` = 5,
                               `group3:3` = 5,`group4:3` = 5,
                               `group5:3` = 5))

# Same formula as above
kgr_formula = deaths ~ -1 + Intercept_1 + Intercept_2 + Intercept_3 + Intercept_4 + Intercept_5 + 
        Intercept_6 + Intercept_7 + Intercept_8 + Intercept_9 + Intercept_10 + Intercept_11 + Intercept_12 + 
        Intercept_13 + Intercept_14 + Intercept_15 + Intercept_16 + Intercept_17 + Intercept_18 + Intercept_19 + 
        Intercept_20 + Intercept_21 + Intercept_22 + Intercept_23 + Intercept_24 + Intercept_25 + Intercept_26 + 
        Intercept_27 + Intercept_28 + Intercept_29 + Intercept_30 + Intercept_31 + Intercept_32 + Intercept_33 + 
        Intercept_34 + Intercept_35 + Intercept_36 + Intercept_37 + Intercept_38 + Intercept_39 + Intercept_40 + 
        Intercept_41 + Intercept_42 + Intercept_43 + Intercept_44 + Intercept_45 + Intercept_46 + Intercept_47 + 
        Intercept_48 + Intercept_49 + Intercept_50 + Intercept_51 + Intercept_52 + Intercept_53 + Intercept_54 + 
        Intercept_55 + Intercept_56 + Intercept_57 + Intercept_58 + factor(month) + 
        aqi_resids +
        cum_total_precip_6 +
        avg_avg_air_temp_6 +
        avg_avg_dew_point_6 +
        max_avg_wind_speed_6 +
        f(sw_binned,model = "rw2") +
        # Spline for treatment * SDI group 
        group*ns(cum_monthly_total_pm25_6_binned,df=3) +
        f(ID,model = "generic0",Cmatrix = solve(H^2), group = time, control.group = list(model = "ar",order = 6))


model = inla(formula = kgr_formula, family = "nbinomial", data = inla_cf_df, num.threads = 1,
             control.compute = list(dic=TRUE,waic=TRUE,return.marginals.predictor=TRUE),
             control.fixed = prior_fixed,
             control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
             control.inla = list(strategy = "simplified.laplace"),
             control.predictor = list(compute = TRUE, link = 1))



###Extract relevant information and store in the list
model_summary <- model$summary.fixed
bri_hyperpar_summary <- bri.hyperpar.summary(model)
model_DIC <- model$dic$dic
model_WAIC <- model$waic$waic

preds_model <- model$summary.fitted.values
preds_model <- cbind(inla_cf_df$NAME, inla_cf_df$time, inla_cf_df$date, 
                     inla_cf_df$cum_monthly_total_pm25_6_binned,
                     inla_cf_df$deaths, preds_model)
colnames(preds_model) <- c("county", "time", "date", "exposure", "deaths", "mean", 
                           "sd", "0.025quant", "0.5quant", "0.975quant", "mode")

a = preds_model[cf_idx,]
mean(a$mean)

# a = preds_model[cf_idx,] %>% filter(county == "Kings")
# mean(a$mean)
```

## Function to evaluate DRFs for grid of treatment values (ran as background job with tmux)

```{r}
dose_grid = c(0,10,20,50,100,200,500,1000,2000,3000)

bins = list(c(1:15),c(16:30),c(31:45),c(46:58))

DRF_curves = function(dataset,dose_grid){
  cf_preds_list = list()
  
  for(i in 1:length(dose_grid)){
    cat("Calculating DRF for dose level:", dose_grid[i], "\n")
    
    # Initialize combined predictions dataframe for this dose
    combined_cf_preds = NULL
    
    for (j in 1:4){
      cat("  Processing batch", j, "of 4\n")
  
      # Get counties for this batch
      batch_counties = counties[bins[[j]]]
      
      # Create duplicate dataset for just these counties
      duplicate_df = dataset[dataset$NAME %in% batch_counties,]
      duplicate_df$cum_monthly_total_pm25_6_binned = dose_grid[i]
      duplicate_df$deaths = NA
      
      # Combine original dataset with counterfactual data for this batch
      inla_cf_df = rbind(dataset, duplicate_df)
      cf_idx = c((nrow(dataset)+1):nrow(inla_cf_df))

      kgr_formula_hierarchical = deaths ~ -1 +
        offset(log(Total_Pop)) +
        factor(month) +
        aqi_resids +
        cum_total_precip_6 +
        avg_avg_air_temp_6 +
        avg_avg_dew_point_6 +
        max_avg_wind_speed_6 +
        # Global PM2.5 dose-response curve
        ns(cum_monthly_total_pm25_6_binned,df = 3) +
        # Cluster-specific deviations from global curve (shrunken toward zero)
        f(Z_cluster, model = spde, group = Cluster, control.group = list(model = "exchangeable")) +
        f(id2, model = "generic0", Cmatrix = inv_covGP2_response)
      
      # Error handling for inla()
      kgr_model <- tryCatch({
        inla(formula = kgr_formula_hierarchical, family = "nbinomial", data = inla_cf_df, num.threads = 50,
             control.compute = list(dic=TRUE,waic=TRUE,return.marginals.predictor=TRUE),
             #control.fixed = prior_fixed,
             control.family = list(hyper = list(theta = list(prior = "pc.mgamma", param = 1))),
             control.inla = list(strategy = "simplified.laplace"),
             control.predictor = list(compute = TRUE, link = 1))
      }, error = function(e) {
        cat("Error fitting INLA model at dose", dose_grid[i], "batch", j, ":", conditionMessage(e), "\n")
        return(NULL)
      })
      
      if (is.null(kgr_model)) {
        cat("Skipping batch", j, "due to model fitting error\n")
        next
      }
      
      # Extract predictions for counterfactual data
      preds_model <- kgr_model$summary.fitted.values
      preds_model <- cbind(inla_cf_df$NAME, inla_cf_df$time, inla_cf_df$date, 
                           inla_cf_df$cum_monthly_total_pm25_6_binned, inla_cf_df$deaths, preds_model)
      colnames(preds_model) <- c("county", "time", "date", "exposure", "deaths", "mean", 
                                 "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
      
      # Get predictions for this batch's counterfactual data
      batch_cf_preds = preds_model[cf_idx,]
      
      # Append to combined predictions
      if (is.null(combined_cf_preds)) {
        combined_cf_preds = batch_cf_preds
      } else {
        combined_cf_preds = rbind(combined_cf_preds, batch_cf_preds)
      }
    }
    
    # Store combined predictions for this dose level
    cf_preds_list[[i]] = combined_cf_preds
  }
  
  return(cf_preds_list)
}

# full_data_DRFs = DRF_curves(inla_df,dose_grid)
# wildfire_szn_DRFs = DRF_curves(wildfire_szn_df,dose_grid)
```

## Bootstrapped version prototype

see estimate-drfs-v4.R or parallel-test.R

# ADRF: average of DRF over all individuals in the population

## Function to calculate ADRFs for grid of treatment values 

```{r}
# dose_grid = c(0,10,20,50,100,200,500,1000,2000,3000)
dose_grid = c(0,10,20,50,100,200,400,600,800,1000)

DRF_to_ADRF = function(DRF_list,dataset){
  ADRF = data.frame(Z = dose_grid, mu_hat = NA, weighted_mu_hat = NA, rate_hat = NA, weighted_rate_hat = NA)
  county_pops = dataset %>% select(NAME,date,Total_Pop) 

  for (i in 1:length(DRF_list)){
    dose_results = DRF_list[[i]]
    dose_results_pop = dose_results %>% left_join(county_pops,by = c("county" = "NAME","date" = "date"))
    
    # Calculate population-weighted mean
    total_pop = sum(dose_results_pop$Total_Pop)
    pop_weights = dose_results_pop$Total_Pop / total_pop
    
    ADRF$mu_hat[i] = mean(dose_results_pop$mean)
    #ADRF$weighted_mu_hat[i] = sum(dose_results_pop$mean * pop_weights)
    #ADRF$rate_hat[i] = mean((dose_results_pop$mean / dose_results_pop$Total_Pop * 100000)) # avg county level rate
    ADRF$rate_hat[i] = sum(dose_results_pop$mean)/total_pop*100000 # avg individual level rate
    #ADRF$weighted_rate_hat[i] = sum((dose_results_pop$mean / dose_results_pop$Total_Pop * 100000) * pop_weights)
  }

  return(ADRF)
}

# #Multiplicative spatiotemporal random effect: ar6 for time, graph filter for space + nonlinear SW + group*treatment interaction
# full_data_DRFs = readRDS("full_data_drfs_NB_totalcum6_swprodrw2_ar6xgfilter_groupint_9.30.rds") #strong shrinkage

collect_bootstrap_results = function(bootstrap_results){
  DRF_list = list()
  
  for (i in 1:length(dose_grid)){
    dose_results = bootstrap_results[[1]][[i]]
    
    for (j in 2:length(bootstrap_results)){
      dose_results = rbind(dose_results,bootstrap_results[[j]][[i]])
    }
    
    
    ###FILTER OUT BAD BOOTSTRAP SAMPLES (samples or just predictions???)
    bad_preds = which(dose_results$mean >= 10000)

    if(length(bad_preds) > 0){
      bad_sims = unique(dose_results$sample_id[bad_preds])
      good_idx = which(!(dose_results$sample_id %in% bad_sims))
      dose_results = dose_results[good_idx,]
    }
    
    DRF_list[[i]] = dose_results
  }
  
  return(DRF_list)
}

###stabilized_IPW_kgr estimated on treatment NOT cum_treatment
# full_data_DRFs = readRDS("full_data_bootstrapped_drfs_totalcum6_v2_kgr2_6.23.rds") #based on 5 samples
# wildfire_szn_DRFs = readRDS("wildfire_szn_bootstrapped_drfs_totalcum6_v2_kgr2_6.23.rds") #based on 5 sample

full_data_DRFs = readRDS("full_data_parallel_bootstrapped_drfs_totalcum6_kgr2_8.13.rds") #vague prior, 100 samples
#full_data_DRFs = readRDS("full_data_parallel_bootstrapped_drfs_totalcum6_kgr2_7.26.rds") #strong shrinkage, 100 samples
#full_data_DRFs = readRDS("full_data_parallel_bootstrapped_drfs_totalcum6_kgr2_7.27.rds") #no shrinkage, 100 samples

bootstrapped_DRFs = collect_bootstrap_results(full_data_DRFs)


DRF_to_ADRF2 = function(DRF_list, dataset) {
  ADRF = data.frame(
    Z = dose_grid, 
    # Unweighted estimates
    mu_hat = NA, mu_hat_lower = NA, mu_hat_upper = NA, 
    rate_hat = NA, rate_hat_lower = NA, rate_hat_upper = NA
    # Weighted estimates
    # weighted_mu_hat = NA, weighted_mu_hat_lower = NA, weighted_mu_hat_upper = NA,
    # weighted_rate_hat = NA, weighted_rate_hat_lower = NA, weighted_rate_hat_upper = NA
  )
  
  county_pops = dataset %>% select(NAME, date, Total_Pop) 
  
  for (i in 1:length(DRF_list)) {
    
    dose_results = DRF_list[[i]]
    dose_results_pop = dose_results %>% 
      left_join(county_pops, by = c("county" = "NAME", "date" = "date"))
    
    # Calculate both weighted and unweighted estimates for each sample
    dose_summary = dose_results_pop %>% 
      group_by(sample_id) %>% 
      summarize(
        # Unweighted estimates
        sample_mu_hat = mean(mean),
        sample_rate_hat = sum(mean)/sum(Total_Pop)*100000,
        # sample_rate_hat = mean(mean / Total_Pop * 100000),

                
        # # Weighted estimates  
        # sample_weighted_mu_hat = sum(mean * Total_Pop) / sum(Total_Pop),
        # sample_weighted_rate_hat = sum((mean / Total_Pop * 100000) * Total_Pop) / sum(Total_Pop),
        
        .groups = "drop"
      )
    
    # Calculate credible intervals for unweighted estimates
    ADRF$mu_hat[i] = mean(dose_summary$sample_mu_hat)
    ADRF$mu_hat_lower[i] = quantile(dose_summary$sample_mu_hat, 0.025)
    ADRF$mu_hat_upper[i] = quantile(dose_summary$sample_mu_hat, 0.975)
    
    ADRF$rate_hat[i] = mean(dose_summary$sample_rate_hat)
    ADRF$rate_hat_lower[i] = quantile(dose_summary$sample_rate_hat, 0.025)
    ADRF$rate_hat_upper[i] = quantile(dose_summary$sample_rate_hat, 0.975)
    
    # # Calculate credible intervals for weighted estimates
    # ADRF$weighted_mu_hat[i] = mean(dose_summary$sample_weighted_mu_hat)
    # ADRF$weighted_mu_hat_lower[i] = quantile(dose_summary$sample_weighted_mu_hat, 0.025)
    # ADRF$weighted_mu_hat_upper[i] = quantile(dose_summary$sample_weighted_mu_hat, 0.975)
    # 
    # ADRF$weighted_rate_hat[i] = mean(dose_summary$sample_weighted_rate_hat)
    # ADRF$weighted_rate_hat_lower[i] = quantile(dose_summary$sample_weighted_rate_hat, 0.025)
    # ADRF$weighted_rate_hat_upper[i] = quantile(dose_summary$sample_weighted_rate_hat, 0.975)
  }
  
  return(ADRF)
}

#full_data_ADRF = DRF_to_ADRF2(bootstrapped_DRFs,inla_df)
#wildfire_szn_ADRF = DRF_to_ADRF2(wildfire_szn_DRFs,wildfire_szn_df)
```

### NEWEST ITERATION

```{r}
# for (i in 1:length(bootstrap_sample_1)){
#   colnames(bootstrap_sample_10[[i]]) = c("sample_id","county", "time", "date", "exposure", "deaths", 
#                                         "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
# }

# bootstrap_results = lapply(1:10, function(i) {
#   readRDS(paste0("bootstrap_sample_", i, ".rds"))
# })

# parallel_bootstrap_test_results = list(bootstrap_sample_1,bootstrap_sample_2,bootstrap_sample_3,bootstrap_sample_4,
#             bootstrap_sample_5,bootstrap_sample_6,bootstrap_sample_7,bootstrap_sample_8,
#             bootstrap_sample_9,bootstrap_sample_10)

parallel_bootstrap_test_results = readRDS("~/work/cum6-bootstrap-samples/parallel_bootstrap_test_results-10.9.rds")

bootstrapped_DRFs = collect_bootstrap_results(parallel_bootstrap_test_results)
full_data_ADRF = DRF_to_ADRF2(bootstrapped_DRFs,inla_df)
```  


## Plot ADRF curves

```{r}
ggplot(data = full_data_ADRF, aes(x = Z)) + 
  geom_ribbon(aes(ymin = rate_hat_lower, ymax = rate_hat_upper), alpha = 0.3) +
  geom_line(aes(y = rate_hat), color = "red", size = 1.2) + 
  labs(x = "cum_monthly_total_pm25_6", y = "rate_hat", title = "Overall ADRF") + 
  theme_classic()

# ggplot(data = wildfire_szn_ADRF, aes(x = Z)) + 
#   geom_ribbon(aes(ymin = rate_hat_lower, ymax = rate_hat_upper), alpha = 0.3) +
#   geom_line(aes(y = rate_hat), color = "red", size = 1.2) + 
#   labs(x = "cum_monthly_total_pm25_6", y = "rate_hat", title = "Wildfire Season ADRF") + 
#   theme_classic()
```

## Calculate ADRF all counties -> make a plot: 

```{r}
county_DRF_to_ADRF = function(DRF_list, dataset) {
  county_ADRF_list = list()
  
  # Get county populations (single value per county)
  county_pops = dataset %>% select(NAME,date,Total_Pop) 
  
  for (i in counties) {
    ADRF = data.frame(Z = dose_grid, mu_hat = NA, weighted_mu_hat = NA, 
                      rate_hat = NA, weighted_rate_hat = NA)
    
    for (j in 1:length(DRF_list)) {
      county_preds = DRF_list[[j]] %>% filter(county == i)
      
      # Skip if no predictions for this county
      if (nrow(county_preds) == 0) next
      
      # Merge with time-specific populations
      county_preds_pop = county_preds %>%
        left_join(county_pops, by = c("county" = "NAME", "date" = "date"))
      
      # Population-weighted mean across time periods
      total_pop = sum(county_preds_pop$Total_Pop)
      pop_weights = county_preds_pop$Total_Pop / total_pop
      
      ADRF$mu_hat[j] = mean(county_preds_pop$mean)
      #ADRF$weighted_mu_hat[j] = sum(county_preds_pop$mean * pop_weights)
      #ADRF$rate_hat[j] = mean((county_preds_pop$mean / county_preds_pop$Total_Pop * 100000)) 
      ADRF$rate_hat[j] = sum(county_preds_pop$mean)/total_pop*100000 
      #ADRF$weighted_rate_hat[j] = sum((county_preds_pop$mean / county_preds_pop$Total_Pop * 100000) * pop_weights)
    }
    
    county_ADRF_list[[i]] = ADRF
  }
  
  return(county_ADRF_list)
}

plot_ADRFs = function(overall_ADRF, county_ADRFs, response,
                      hi.res = FALSE,input_title = "Estimated ADRFs",
                      prefix = "cum6-",filename = "county-adrfs.png") {
  
  county_data_combined <- bind_rows(county_ADRFs, .id = "county_id")
  
  if (hi.res) {
    county_data_combined <- county_data_combined %>% filter(Z <= 200)
    overall_ADRF <- overall_ADRF %>% filter(Z <= 200)
  }
  
  # Main ADRF plot
  p_main <- ggplot() + 
    geom_line(data = county_data_combined, 
              aes(x = Z, 
                  y = .data[[response]], 
                  group = county_id),
              color = "gray70", 
              alpha = 0.3, 
              size = 0.5) +
    geom_line(data = overall_ADRF, 
              aes(x = Z, y = .data[[response]]),
              color = "red", 
              size = 1.2) +
    labs(x = "cum_monthly_total_pm25_6", 
         y = response, 
         title = input_title) +
    theme_classic() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 12),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10)
    )
  
  # Marginal histogram/density
  p_top <- ggplot(inla_df, aes(x = cum_monthly_total_pm25_6)) +
    geom_density(fill = "gray80", alpha = 1.5) +  
    theme_void() +
    theme(
      plot.margin = margin(0, 5, 0, 5) # reduce gap
    )
  
  # Combine: p_top above p_main
  combined <- p_top / p_main + plot_layout(heights = c(1, 3))
  
#   ggsave(
#   filename = paste0(prefix,filename),
#   plot = combined,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
  
  return(combined)
}
```


```{r}
county_ADRFs = county_DRF_to_ADRF(bootstrapped_DRFs,inla_df)
plot_ADRFs(full_data_ADRF,county_ADRFs,response = "rate_hat",hi.res = FALSE)
plot_ADRFs(full_data_ADRF,county_ADRFs,response = "rate_hat",hi.res = TRUE,prefix = "cum6-hi.res-")
```

## Same analysis by SDI SKATER clusters and SDI groups/clusters

```{r}
cluster_DRF_to_ADRF = function(DRF_list, dataset, clusterlabels) {
  cluster_ADRF_list = list()
  
  # Get county populations and merge with cluster labels
  county_pops = dataset %>% 
    select(NAME, date, Total_Pop) %>%
    left_join(clusterlabels, by = c("NAME" = "counties"))

  for (cluster_id in 1:num_clus) {
    ADRF = data.frame(Z = dose_grid, mu_hat = NA, weighted_mu_hat = NA, rate_hat = NA, weighted_rate_hat = NA)
    
    # Get counties in this cluster
    cluster_counties = clusterlabels$counties[clusterlabels$Cluster == cluster_id]
    
    for (j in 1:length(DRF_list)) {
      # Get predictions for all counties in this cluster
      cluster_preds = DRF_list[[j]] %>% 
        filter(county %in% cluster_counties)
      
      # Skip if no predictions for this cluster
      if (nrow(cluster_preds) == 0) next
      
      # Merge with time-specific populations and cluster info
      cluster_preds_pop = cluster_preds %>%
        left_join(county_pops, by = c("county" = "NAME", "date" = "date"))
      
      # Population-weighted mean across time periods and counties in cluster
      total_pop = sum(cluster_preds_pop$Total_Pop, na.rm = TRUE)
      pop_weights = cluster_preds_pop$Total_Pop / total_pop
      
      ADRF$mu_hat[j] = mean(cluster_preds_pop$mean, na.rm = TRUE)
      #ADRF$weighted_mu_hat[j] = sum(cluster_preds_pop$mean * pop_weights, na.rm = TRUE)
      ADRF$rate_hat[j] = mean((cluster_preds_pop$mean / cluster_preds_pop$Total_Pop * 100000), na.rm = TRUE) 
      #ADRF$rate_hat[j] = sum(cluster_preds_pop$mean)/total_pop*100000 ??? not exactly the same
      #ADRF$weighted_rate_hat[j] = sum((cluster_preds_pop$mean / cluster_preds_pop$Total_Pop * 100000) * pop_weights, na.rm = TRUE)
    }
    
    cluster_ADRF_list[[as.character(cluster_id)]] = ADRF
  }
  
  return(cluster_ADRF_list)
}

SDIgroup_DRF_to_ADRF = function(DRF_list, dataset, group_labels) {
  cluster_ADRF_list = list()
  
  # Get county populations and merge with cluster labels
  county_pops = dataset %>% 
    select(NAME, date, Total_Pop) %>%
    left_join(group_labels, by = c("NAME" = "counties"))

  for (cluster_id in 1:num_clus) {
    ADRF = data.frame(Z = dose_grid, mu_hat = NA, weighted_mu_hat = NA, rate_hat = NA, weighted_rate_hat = NA)
    
    # Get counties in this cluster
    cluster_counties = group_labels$counties[group_labels$Group == cluster_id]
    
    for (j in 1:length(DRF_list)) {
      # Get predictions for all counties in this cluster
      cluster_preds = DRF_list[[j]] %>% 
        filter(county %in% cluster_counties)
      
      # Skip if no predictions for this cluster
      if (nrow(cluster_preds) == 0) next
      
      # Merge with time-specific populations and cluster info
      cluster_preds_pop = cluster_preds %>%
        left_join(county_pops, by = c("county" = "NAME", "date" = "date"))
      
      # Population-weighted mean across time periods and counties in cluster
      total_pop = sum(cluster_preds_pop$Total_Pop, na.rm = TRUE)
      pop_weights = cluster_preds_pop$Total_Pop / total_pop
      
      ADRF$mu_hat[j] = mean(cluster_preds_pop$mean, na.rm = TRUE)
      #ADRF$weighted_mu_hat[j] = sum(cluster_preds_pop$mean * pop_weights, na.rm = TRUE)
      #ADRF$rate_hat[j] = mean((cluster_preds_pop$mean / cluster_preds_pop$Total_Pop * 100000), na.rm = TRUE)
      ADRF$rate_hat[j] = sum(cluster_preds_pop$mean)/total_pop*100000
      #ADRF$weighted_rate_hat[j] = sum((cluster_preds_pop$mean / cluster_preds_pop$Total_Pop * 100000) * pop_weights, na.rm = TRUE)
    }
    
    cluster_ADRF_list[[as.character(cluster_id)]] = ADRF
  }
  
  return(cluster_ADRF_list)
}

plot_cluster_ADRFs = function(overall_ADRF, cluster_ADRFs, response, 
                              hi.res = FALSE, input_title = "Estimated Cluster ADRFs",
                              prefix = "cum6-",filename = "cluster-adrfs.png") {
  # Convert list of dataframes to single dataframe with cluster identifier
  cluster_data_combined <- bind_rows(cluster_ADRFs, .id = "cluster_id")
  
  if(hi.res == TRUE){
    cluster_data_combined = cluster_data_combined %>% filter(Z <= 200)
    overall_ADRF = overall_ADRF %>% filter(Z <= 200)
  }
  
  # Create a color palette for clusters
  cluster_colors <- c("darkblue", "purple", "magenta", "orange", "yellow")
  names(cluster_colors) = unique(cluster_data_combined$cluster_id)
  
  # Create the base plot with cluster curves
  p_main <- ggplot() + 
    # Cluster curves first
    geom_line(data = cluster_data_combined, 
              aes(x = Z, 
                  y = cluster_data_combined[,response], 
                  group = cluster_id,
                  color = cluster_id),
              alpha = 0.7, 
              size = 0.8) +
    # Then overall curve
    geom_line(data = overall_ADRF, 
              aes(x = Z, 
                  y = overall_ADRF[,response]),
              color = "black", 
              size = 1.2) +
    scale_color_manual(values = cluster_colors, name = "Cluster") +
    labs(x = "cum_monthly_total_pm25_6", 
         y = response, 
         title = input_title) +
    theme_classic() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 12),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10),
      legend.position = "right"
    )
  
  # Marginal histogram/density
  p_top <- ggplot(inla_df, aes(x = cum_monthly_total_pm25_6)) +
    geom_density(fill = "gray80", alpha = 1.5) +  
    theme_void() +
    theme(
      plot.margin = margin(0, 5, 0, 5) # reduce gap
    )

  # Combine: p_top above p_main
  combined <- p_top / p_main + plot_layout(heights = c(1, 3))
  
  # ggsave(
  #   filename = paste0(prefix,filename),
  #   plot = combined,
  #   path = "~/work/important-plots",
  #   width = 8,
  #   height = 10,
  #   dpi = 600
  # )
  
  return(combined)
}

# Example usage:
#cluster_ADRFs = cluster_DRF_to_ADRF(bootstrapped_DRFs, inla_df, clusterlabels)
cluster_ADRFs = SDIgroup_DRF_to_ADRF(bootstrapped_DRFs, inla_df, group_labels)

plot_cluster_ADRFs(full_data_ADRF, cluster_ADRFs, hi.res = FALSE,"rate_hat")
plot_cluster_ADRFs(full_data_ADRF, cluster_ADRFs, hi.res = TRUE,"rate_hat",prefix = "cum6-hi.res-")
```

# AGTE: evaluate ADRF at two levels of treatment for which we want to compare and take the difference

## Function to calculate 

```{r}
estimate_AGTE = function(z1,z0,ADRF_curve){
  # AGTE1 = (ADRF_curve$mu_hat[ADRF_curve$Z == z1]) - (ADRF_curve$mu_hat[ADRF_curve$Z == z0])
  # AGTE2 = (ADRF_curve$rate_hat[ADRF_curve$Z == z1]) - (ADRF_curve$rate_hat[ADRF_curve$Z == z0])
  AGTE3 = (ADRF_curve$rate_hat[ADRF_curve$Z == z1]) / (ADRF_curve$rate_hat[ADRF_curve$Z == z0])

  return(AGTE3)
}


#MAX VS MIN CONTRAST
estimate_AGTE(1000,0,full_data_ADRF)
# wildfire_szn_tau_hat = estimate_AGTE(3000,0,wildfire_szn_ADRF)

#2 SD CONTRAST
mean(inla_df$cum_monthly_total_pm25_6)
sd(inla_df$cum_monthly_total_pm25_6)
tau_hat = estimate_AGTE(1000,200,full_data_ADRF)

#Should we interpolate curve?
# f = splinefun(full_data_ADRF$Z,full_data_ADRF$weighted_rate_hat,method = "natural")
# 
# x = c(185,530,875)
# ls(envir = environment(f))
# splinecoef <- get("z", envir = environment(f))
# curve(f(x), 0, 1000, col = "green", lwd = 1.5)
# points(splinecoef, col = "purple",pch = 20, cex = 2)
# 
# new_z = c(185,530,875)
# interpolated_y = f(new_z)
# interpolated_y[3] / interpolated_y[1]


#THOMAS FIRE CASE STUDY (12/17-1/18)
thomas_fire_smoke = inla_df %>% filter(NAME %in% c("Santa Barbara","Ventura"),
  date >= as.Date("2017-11-01") & date <= as.Date("2018-02-01")) %>%
  select(NAME,date,monthly_total_pm25,cum_monthly_total_pm25_6,deaths)

###SB went from 0-541 in that month, Ventura went from 0-424
estimate_AGTE(600,0,full_data_ADRF)

tau_hat
# wildfire_szn_tau_hat
```

## Function to calculate with uncertainty quantification

```{r}
estimate_AGTE2 = function(DRF_list, dataset, z1, z0, type = "ratio") {
  # type: "difference" for Y(z1) - Y(z0), "ratio" for Y(z1) / Y(z0)
  
  # Find indices for the specified treatment levels
  z1_idx = which(dose_grid == z1)
  z0_idx = which(dose_grid == z0)
  
  if (length(z1_idx) == 0 | length(z0_idx) == 0) {
    stop("Specified treatment levels not found in dose_grid")
  }
  
  # Get the DRF results for both treatment levels
  z1_results = DRF_list[[z1_idx]]
  z0_results = DRF_list[[z0_idx]]
  
  county_pops = dataset %>% select(NAME, date, Total_Pop)
  
  # Join population data
  z1_results_pop = z1_results %>% 
    left_join(county_pops, by = c("county" = "NAME", "date" = "date"))
  z0_results_pop = z0_results %>% 
    left_join(county_pops, by = c("county" = "NAME", "date" = "date"))
  
  # Calculate estimates for each bootstrap sample
  z1_summary = z1_results_pop %>% 
    group_by(sample_id) %>% 
    summarize(
      # Unweighted estimates
      sample_mu_hat = mean(mean),
      sample_rate_hat = sum(mean)/sum(Total_Pop)*100000,
      
      # # Weighted estimates  
      # sample_weighted_mu_hat = sum(mean * Total_Pop) / sum(Total_Pop),
      # sample_weighted_rate_hat = sum((mean / Total_Pop * 100000) * Total_Pop) / sum(Total_Pop),
      
      .groups = "drop"
    )
  
  z0_summary = z0_results_pop %>% 
    group_by(sample_id) %>% 
    summarize(
      # Unweighted estimates
      sample_mu_hat = mean(mean),
      sample_rate_hat = sum(mean)/sum(Total_Pop)*100000,
      
      # # Weighted estimates  
      # sample_weighted_mu_hat = sum(mean * Total_Pop) / sum(Total_Pop),
      # sample_weighted_rate_hat = sum((mean / Total_Pop * 100000) * Total_Pop) / sum(Total_Pop),
      
      .groups = "drop"
    )
  
  # Merge by sample_id to ensure proper pairing
  agte_data = z1_summary %>% 
    inner_join(z0_summary, by = "sample_id", suffix = c("_z1", "_z0"))
  
  # Calculate AGTE for each bootstrap sample
  if (type == "difference") {
    agte_data = agte_data %>% 
      mutate(
        # Unweighted AGTE
        agte_mu = sample_mu_hat_z1 - sample_mu_hat_z0,
        agte_rate = sample_rate_hat_z1 - sample_rate_hat_z0
        
        # # Weighted AGTE
        # agte_weighted_mu = sample_weighted_mu_hat_z1 - sample_weighted_mu_hat_z0,
        # agte_weighted_rate = sample_weighted_rate_hat_z1 - sample_weighted_rate_hat_z0
      )
  } else if (type == "ratio") {
    agte_data = agte_data %>% 
      mutate(
        # Unweighted AGTE
        agte_mu = sample_mu_hat_z1 / sample_mu_hat_z0,
        agte_rate = sample_rate_hat_z1 / sample_rate_hat_z0
        
        # # Weighted AGTE
        # agte_weighted_mu = sample_weighted_mu_hat_z1 / sample_weighted_mu_hat_z0,
        # agte_weighted_rate = sample_weighted_rate_hat_z1 / sample_weighted_rate_hat_z0
      )
  } else {
    stop("type must be either 'difference' or 'ratio'")
  }
  
  # Calculate point estimates and confidence intervals
  agte_results = data.frame(
    z1 = z1,
    z0 = z0,
    type = type,
    
    # Unweighted estimates
    agte_mu_hat = mean(agte_data$agte_mu),
    agte_mu_lower = quantile(agte_data$agte_mu, 0.025),
    agte_mu_upper = quantile(agte_data$agte_mu, 0.975),
    
    agte_rate_hat = mean(agte_data$agte_rate),
    agte_rate_lower = quantile(agte_data$agte_rate, 0.025),
    agte_rate_upper = quantile(agte_data$agte_rate, 0.975)
    
    # # Weighted estimates
    # agte_weighted_mu_hat = mean(agte_data$agte_weighted_mu),
    # agte_weighted_mu_lower = quantile(agte_data$agte_weighted_mu, 0.025),
    # agte_weighted_mu_upper = quantile(agte_data$agte_weighted_mu, 0.975),
    # 
    # agte_weighted_rate_hat = mean(agte_data$agte_weighted_rate),
    # agte_weighted_rate_lower = quantile(agte_data$agte_weighted_rate, 0.025),
    # agte_weighted_rate_upper = quantile(agte_data$agte_weighted_rate, 0.975)
  )
  rownames(agte_results) = NULL
  
  return(agte_results)
}

estimate_AGTE2(bootstrapped_DRFs,inla_df,1000,0,"ratio")
estimate_AGTE2(bootstrapped_DRFs,inla_df,1000,0,"ratio")[7:9]
tau_hat = as.numeric(estimate_AGTE2(bootstrapped_DRFs,inla_df,1000,0,"ratio")[7])
```

3## Interpret + visualize 

```{r}
#LA 
inla_df %>% filter(NAME == "Los Angeles") %>% dplyr::select(Total_Pop,deaths,death_rate,cum_monthly_total_pm25_6) %>% summary()

4.2*10059182/100000 #Mean death rate -> deaths
(4.2*tau_hat)*10059182/100000 #Elevated death rate -> deaths

#Alameda
inla_df %>% filter(NAME == "Alameda") %>% dplyr::select(Total_Pop,deaths,death_rate,cum_monthly_total_pm25_6) %>% summary()

3.3*1613263/100000 #Mean death rate -> deaths
(3.3*tau_hat)*1613263/100000 #Elevated death rate -> deaths

#Alpine
inla_df %>% filter(NAME == "Alpine") %>% dplyr::select(Total_Pop,deaths,death_rate,cum_monthly_total_pm25_6) %>% summary()

0.08*33539/100000 #Mean death rate -> deaths
(0.08*tau_hat)*33539/100000 #Elevated death rate -> deaths

#SB
inla_df %>% filter(NAME == "Santa Barbara") %>% dplyr::select(Total_Pop,deaths,death_rate,cum_monthly_total_pm25_6) %>% summary()

3.123*439727/100000 #Mean death rate -> deaths
(3.123*tau_hat)*439727/100000 #Elevated death rate -> deaths
```

```{r}
visualize_causal_effect = function(ADRF_curve,z0 = 0,contrast = "ratio"){
  if(contrast == "difference"){
    causal_diff = data.frame(Z = dose_grid,diff = NA)
    
    for (i in 1:nrow(ADRF_curve)){
      ref_idx = which(ADRF_curve$Z == z0)
      causal_diff$diff[i] = ADRF_curve$rate_hat[i] - ADRF_curve$rate_hat[ref_idx]
    }
    
    p <- ggplot() + 
      geom_line(data = causal_diff, 
                aes(x = Z, y = diff),
                color = "red", 
                size = 1.2) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      labs(x = "cum_monthly_total_pm25_6", 
           y = "Difference", 
           title = "Causal Difference Compared to Z = 0") +
      theme_classic() +
      theme(
        plot.title = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10)
      )
    
    print(p)
    
  } else{
    risk_ratio = data.frame(Z = dose_grid,RR = NA)
    
    for (i in 1:nrow(ADRF_curve)){
      ref_idx = which(ADRF_curve$Z == z0)
      risk_ratio$RR[i] = ADRF_curve$rate_hat[i] / ADRF_curve$rate_hat[ref_idx]
    }
    
    p <- ggplot() + 
      geom_line(data = risk_ratio, 
                aes(x = Z, y = RR),
                color = "red", 
                size = 1.2) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "gray50") +
      labs(x = "cum_monthly_total_pm25_6", 
           y = "Risk Ratio", 
           title = "Causal Risk Ratio Compared to Z = 0") +
      theme_classic() +
      theme(
        plot.title = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10)
      )
    
    print(p)
  }
}

visualize_causal_effect(full_data_ADRF)
# visualize_causal_effect(county_ADRFs[[50]])
# visualize_causal_effect(cluster_ADRFs[[5]])
```

```{r}
visualize_causal_effect_with_ci = function(ADRF_curve, z0 = 0, contrast = "ratio"){
  
  if(contrast == "difference"){
    causal_diff = data.frame(
      Z = ADRF_curve$Z,
      diff = NA,
      diff_lower = NA,
      diff_upper = NA
    )
    
    ref_idx = which(ADRF_curve$Z == z0)
    
    for (i in 1:nrow(ADRF_curve)){
      
      result = estimate_AGTE2(bootstrapped_DRFs,inla_df,
                              dose_grid[i],dose_grid[ref_idx],"difference")
      
      causal_diff$diff[i] = result$agte_mu_hat
      causal_diff$diff_lower[i] = result$agte_mu_lower
      causal_diff$diff_upper[i] = result$agte_mu_upper
    }
    
    p <- ggplot(causal_diff, aes(x = Z, y = diff)) + 
      geom_ribbon(aes(ymin = diff_lower, ymax = diff_upper),
                  fill = "red", alpha = 0.2) +
      geom_line(color = "red", size = 1.2) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      labs(x = "cum_monthly_total_pm25_6", 
           y = "Difference", 
           title = paste0("Causal Difference Compared to Z = ", z0)) +
      theme_classic() +
      theme(
        plot.title = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10)
      )
    
    print(p)
    
  } else {
    risk_ratio = data.frame(
      Z = ADRF_curve$Z,
      RR = NA,
      RR_lower = NA,
      RR_upper = NA
    )
    
    ref_idx = which(ADRF_curve$Z == z0)
    
    for (i in 1:nrow(ADRF_curve)){
      
      result = estimate_AGTE2(bootstrapped_DRFs,inla_df,
                              dose_grid[i],dose_grid[ref_idx],"ratio")
      
      risk_ratio$RR[i] = result$agte_rate_hat
      risk_ratio$RR_lower[i] = result$agte_rate_lower
      risk_ratio$RR_upper[i] = result$agte_rate_upper
    }
    
    p <- ggplot(risk_ratio, aes(x = Z, y = RR)) + 
      geom_ribbon(aes(ymin = RR_lower, ymax = RR_upper),
                  fill = "red", alpha = 0.2) +
      geom_line(color = "red", size = 1.2) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "gray50") +
      labs(x = "cum_monthly_total_pm25_6", 
           y = "Risk Ratio", 
           title = paste0("Causal Risk Ratio Compared to Z = ", z0)) +
      theme_classic() +
      theme(
        plot.title = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10)
      )
    
    print(p)
  }
}

visualize_causal_effect_with_ci(full_data_ADRF)
# visualize_causal_effect_with_ci(county_ADRFs[[50]])
# visualize_causal_effect_with_ci(cluster_ADRFs[[5]])


###SOMETHING WRONG HERE
for (i in 1:5){
  visualize_causal_effect_with_ci(cluster_ADRFs[[i]])
}
```


### Interpret for each county (applies AGTE estimate to each county)

```{r}
interpret_AGTE_by_county = function(AGTE_estimate,dataset){
  
  AGTE_by_county = data.frame(county = counties,avg_pop = NA,avg_death_rate = NA,new_death_rate = NA,
                              avg_deaths = NA,new_deaths = NA)
  
  for(i in 1:length(counties)){
    county_data = dataset %>% filter(NAME == counties[i]) %>% dplyr::select(Total_Pop,death_rate)
    AGTE_by_county$avg_pop[i] = mean(county_data$Total_Pop)
    
    AGTE_by_county$avg_death_rate[i] = mean(county_data$death_rate)
    AGTE_by_county$new_death_rate[i] = mean(county_data$death_rate) * AGTE_estimate
    
    AGTE_by_county$avg_deaths[i] = mean(county_data$death_rate)*mean(county_data$Total_Pop)/100000
    AGTE_by_county$new_deaths[i] = (mean(county_data$death_rate) * AGTE_estimate)*mean(county_data$Total_Pop)/100000
  }
  
  return(AGTE_by_county)
}

AGTE_by_county = interpret_AGTE_by_county(AGTE_estimate = tau_hat, dataset = inla_df)
```

## Visualize effect on death rate

### Grouped barplot

```{r}
#Visualize
# Reshape data for grouped barplot
plot_data <- AGTE_by_county %>%
  dplyr::select(county, avg_death_rate, new_death_rate) %>%
  pivot_longer(cols = c(avg_death_rate, new_death_rate),
               names_to = "rate_type",
               values_to = "death_rate") %>%
  mutate(
    rate_type = case_when(
      rate_type == "avg_death_rate" ~ "Average Rate",
      rate_type == "new_death_rate" ~ "New Rate"
    )
  )

# Create grouped barplot
ggplot(plot_data, aes(x = county, y = death_rate, fill = rate_type)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = c("Average Rate" = "steelblue", "New Rate" = "coral"),
                    name = "Rate Type") +
  labs(title = "Death Rate Comparison by County",
       x = "County",
       y = "Death Rate (per 100k)") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10),
    legend.position = "top"
  )
```

### Stacked barplot

```{r}
#Visualize
# Calculate the increase and reshape for stacked barplot
plot_data <- AGTE_by_county %>%
  mutate(increase = new_death_rate - avg_death_rate) %>%
  dplyr::select(county, avg_death_rate, increase) %>%
  pivot_longer(cols = c(avg_death_rate, increase),
               names_to = "component",
               values_to = "value") %>%
  mutate(
    component = case_when(
      component == "avg_death_rate" ~ "Base Rate",
      component == "increase" ~ "Increase"
    ),
    component = factor(component, levels = c("Increase","Base Rate"))
  )

# Create stacked barplot
ggplot(plot_data, aes(x = county, y = value, fill = component)) +
  geom_col(width = 0.5) +
  scale_fill_manual(values = c("Base Rate" = "steelblue", "Increase" = "coral"),
                    name = "Component") +
  labs(title = "Death Rate: Base vs Increase by County",
       x = "County",
       y = "Death Rate (per 100k)") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10),
    legend.position = "top"
  )
```

## Visualize effect on death counts

### Grouped barplot

```{r}
#Visualize
# Reshape data for grouped barplot
plot_data <- AGTE_by_county %>%
  dplyr::select(county, avg_deaths, new_deaths) %>%
  pivot_longer(cols = c(avg_deaths, new_deaths),
               names_to = "count_type",
               values_to = "deaths") %>%
  mutate(
    count_type = case_when(
      count_type == "avg_deaths" ~ "Average Deaths",
      count_type == "new_deaths" ~ "New Deaths"
    )
  )

# Create grouped barplot
ggplot(plot_data, aes(x = county, y = deaths, fill = count_type)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = c("Average Deaths" = "steelblue", "New Deaths" = "coral"),
                    name = "County Type") +
  labs(title = "Death Counts Comparison by County",
       x = "County",
       y = "Respiratory Related Deaths") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10),
    legend.position = "top"
  )
```

### Stacked barplot

```{r}
#Visualize
# Calculate the increase and reshape for stacked barplot
plot_data <- AGTE_by_county %>%
  mutate(increase = new_deaths - avg_deaths) %>%
  dplyr::select(county, avg_deaths, increase) %>%
  pivot_longer(cols = c(avg_deaths, increase),
               names_to = "component",
               values_to = "value") %>%
  mutate(
    component = case_when(
      component == "avg_deaths" ~ "Baseline",
      component == "increase" ~ "Increase"
    ),
    component = factor(component, levels = c("Increase","Baseline"))
  )

# Create stacked barplot
ggplot(plot_data, aes(x = county, y = value, fill = component)) +
  geom_col(width = 0.5) +
  scale_fill_manual(values = c("Baseline" = "steelblue", "Increase" = "coral"),
                    name = "Component") +
  labs(title = "Deaths: Base vs Increase by County",
       x = "County",
       y = "Respiratory Related Deaths") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10),
    legend.position = "top"
  )
```

## Visualize relative risk for each county 

```{r}
county_pct_change = data.frame(county = counties, delta = NA)

for (i in 1:length(county_ADRFs)){
  county_pct_change$delta[i] = estimate_AGTE(1000,0,county_ADRFs[[i]])
}

county_pct_change %>% ggplot(aes(x=county, y=delta)) + 
  geom_bar(stat = "identity") +   labs(title = "Relative Risk in Death Rate by County",
       x = "County",
       y = "Relative Risk in Death Rate") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10),
    legend.position = "top"
  )
```


## Visualize relative risk for each cluster 

```{r}
cluster_pct_change = data.frame(Group = c("1","2","3","4","5"), delta = NA)

for (i in 1:length(cluster_ADRFs)){
  cluster_pct_change$delta[i] = estimate_AGTE(1000,0,cluster_ADRFs[[i]])
}

my_colors <- c("darkblue", "purple", "magenta", "orange", "yellow")

cluster_pct_change %>% ggplot(aes(x=Group, y=delta, fill=Group)) + 
  geom_bar(stat = "identity") + scale_fill_manual(values = my_colors) +
  labs(title = "Relative Risk in Death Rate by Cluster",
       x = "Cluster",
       y = "Relative Risk in Death Rate") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10),
    legend.position = "top"
  )
```