---
title: "Download-EPA"
format: html
editor: visual
---

## Downloading EPA data

```{r}
library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(huge)
library(INLA)
library(HMMpa)
library(invgamma)
library(brinla)
library(reshape2)
library(patchwork)
library(jsonlite)
library(maps)
library(geosphere)
library(urbnmapr)
library(RAQSAPI)
library(con2aqi)
library(pscl)
library(reshape2)
library(corrplot)
library(superheat)
library(shapes)
library(scales)
library(splines)
library(viridis)
library(RColorBrewer)
library(ggh4x)

aqs_credentials("jeffreywu@ucsb.edu","copperheron86")
```

# Get county and pollutant reference codes from EPA

```{r}
#Get county codes 
counties_url = "https://aqs.epa.gov/data/api/list/countiesByState?email=jeffreywu@ucsb.edu&key=copperheron86&state=06&"

countycodes = fromJSON(counties_url)
countycodes = countycodes[[2]]
california_counties = countycodes$code

#Get parameter codes
parameters_url = "https://aqs.epa.gov/data/api/list/parametersByClass?email=jeffreywu@ucsb.edu&key=copperheron86&pc=CRITERIA"

parametercodes = fromJSON(parameters_url)
parametercodes = parametercodes[[2]]
parametercodes = parametercodes[-7,]

pollutants = data.frame(parametercodes$code)
labels = c("lead","co","so2","no2","o3","pm10","pm25")
pollutants = cbind(pollutants,labels)
```

# Identifying a set of monitoring stations that are well distributed across CA

The goal here is to query all of the EPA measurements for all 7 pollutants and AQI from 2014-2019. The first step is to query all stations in California that collect measurements for at least one of the 7 pollutants.

FUNCTION THAT QUERIES STATION LOCATIONS FOR A GIVEN POLLUTANT

```{r}
query_aqs_station_data <- function(param,year){
  start_date <- paste0(year, "0101")
  end_date <- paste0(year, "1231")
  
  url <- paste0("https://aqs.epa.gov/data/api/monitors/byState?email=jeffreywu@ucsb.edu&key=copperheron86&param=", param, "&bdate=", start_date, "&edate=", end_date, "&state=06")

  myData <- fromJSON(url)
  station_data = myData[[2]]
  
  return(station_data)
}

test = query_aqs_station_data(pollutants$parametercodes.code[7],2014)
```

FOR EACH POLLUTANT, GRAB ALL MONITORING STATIONS FOR EACH YEAR

```{r,eval=FALSE}
stations_url = "https://aqs.epa.gov/data/api/monitors/byState?email=jeffreywu@ucsb.edu&key=copperheron86&param=88101&bdate=20140101&edate=20141231&state=06"

stations = fromJSON(stations_url)
stations = stations[[2]]

station_data2014_pm2.5 = stations %>% select(latitude,longitude,site_number, local_site_name,county_code,county_name)

#Get monitoring station locations for each pollutant for each year (takes approx 3 min)
all_pollutants_station_data <- list()
for (year in 2014:2019){
  year_data <- lapply(pollutants$parametercodes.code, 
                      query_aqs_station_data, year = year)
  all_pollutants_station_data[[as.character(year)]] <- year_data
}


# 1 - Lead, 2 - Carbon monoxide (CO), 3 - Sulfure dioxide (SO2), 4 - Nitrogen dioxide (NO2)
# 5 - Ozone (O3), 6 - Total PM10, 7 - PM2.5

all_pollutants_station_2014data = all_pollutants_station_data[[1]]
all_pollutants_station_2015data = all_pollutants_station_data[[2]]
all_pollutants_station_2016data = all_pollutants_station_data[[3]]
all_pollutants_station_2017data = all_pollutants_station_data[[4]]
all_pollutants_station_2018data = all_pollutants_station_data[[5]]
all_pollutants_station_2019data = all_pollutants_station_data[[6]]
```

# Identify subset of stations that has best spatial coverage wrt CA state

Starting off with a list of all the stations measuring each parameter/pollutant, we want to identify a subset of stations that measure each pollutant for each county. This overall set of stations (subset for each county combined together) should have a good spatial coverage of the state.

This is difficult because there is not a station in every county measuring each pollutant. So in order to identify a good set of stations to query data from, I first looked up the two largest cities in each county, based on population.

Lat/long for 2 biggest cities (based on population) in each county

```{r}
citylats = c(37.8044,37.5485,38.7743,38.8071,38.3527,38.3488,39.7285,39.7596,
             38.1231,38.0678,39.2143,39.1546,37.9780,38.0049,41.7558,41.7548,
             38.9399,38.6688,36.7378,36.8252,39.7474,39.5243,40.8021,40.8665,
             32.7920,32.6789,37.3614,37.3855,35.3733,35.7688,36.3275,36.3008,
             38.9582,38.8080,40.4163,40.2840,34.0522,33.7701,36.9613,37.1230,
             37.9735,38.1074,37.4849,37.4320,39.4457,39.4096,37.3022,37.0583,
             41.4871,41.4099,37.6485,38.5149,36.6777,36.6149,38.2975,38.1749,
             39.3280,39.2191,33.8366,33.7455,38.7521,38.7907,39.9341,40.3063,
             33.9806,33.9425,38.5816,38.4088,36.8525,36.8125,34.1083,34.0922,
             32.7157,32.6401,37.7749,37.9780,37.9577,37.7396,35.2828,35.6369,
             37.6879,37.5630,34.9530,34.4208,37.3387,37.3688,36.9741,36.9102,
             40.5865,40.4482,39.6763,39.5595,41.7354,41.3099,38.1041,38.2492,
             38.4404,38.2324,37.6393,37.4946,39.1404,39.1165,40.1785,39.9277,
             40.7310,40.4156,36.3301,36.2077,38.0297,37.9829,34.1975,34.1706,
             38.5449,38.6785,39.1277,39.0954)

citylongs = c(122.2712,121.9886,119.8219,119.7960,120.9327,120.7741,121.8375,121.6219,
              120.8509,120.5385,122.0094,122.1494,122.0311,121.8058,124.2026,124.1580,
              119.9772,120.9872,119.7871,119.7029,122.1964,122.1936,124.1637,124.0828,
              115.5631,115.4989,118.3997,118.4105,119.0187,119.2471,119.6457,119.7829,
              122.6264,122.5583,120.6530,120.5394,118.2437,118.1937,120.0607,120.2602,
              122.5311,122.5697,119.9663,120.0985,123.8053,123.3556,120.4830,120.8499,
              120.5425,120.6791,118.9721,119.4768,121.6555,121.8221,122.2869,122.2608,
              120.1833,121.0611,117.9143,117.8677,121.2880,121.2358,120.8980,121.2319,
              117.3755,117.2297,121.4944,121.3716,121.4016,121.3658,117.2898,117.4350,
              117.1611,117.0842,122.4194,122.0311,121.2908,121.4260,120.6596,120.6545,
              122.4702,122.3255,120.4357,119.6982,121.8853,122.0363,122.0308,121.7569,
              122.3917,122.2978,120.2410,120.8277,122.6345,122.3106,122.2566,122.0405,
              122.7141,122.6367,120.9970,120.8460,121.6169,121.6380,122.2358,122.1792,
              122.9420,123.2100,119.2966,119.3473,119.9741,120.3822,119.1771,118.8376,
              121.7405,121.7733,121.5508,121.5522)
citylongs = -1*citylongs
```

Alameda: Oakland (429082) and Fremont

Alpine: Alpine Village (225) and Mesa Vista

Amador: Ione (8363) and Jackson

Butte: Chico (94776) and Paradise

Calaveras: Rancho Calaveras (5324) and Angels Camp

Colusa: Colusa (5911) and Williams

Contra Costa: Concord (129688) and Antioch

Del Norte: Crescent City (6805) and Bertsch-Oceanview

El Dorado: South Lake Tahoe (22036) and Cameron Park

Fresno: Fresno (530093) and Clovis

Glenn: Orland (7644) and Willows

Humboldt: Eureka (26998) and Arcata

Imperial: El Centro (44120) and Calexico

Inyo: Bishop (3746) and Dixon Lane-Meadow Creek

Kern: Bakersfield (383579) and Delano

Kings: Hanford (56910) and Lemoore

Lake: Clearlake (15384) and Hidden Valley Lake

Lassen: Susanville (15165) and Janesville

Los Angeles: Los Angeles (3990000) and Long Beach

Madera: Madera (65706) and Chowchilla

Marin: San Rafael (58704) and Novato

Mariposa: Mariposa (1526) and Catheys Valley

Mendocino: Fort Bragg (7359) and Willits

Merced: Merced (83316) and Los Banos

Modoc: Alturas (2509) and California Pines

Mono: Mammoth Lakes (8127) and Walker

Monterey: Salinas (156259) and Seaside

Napa: Napa (79263) and American Canyon

Nevada: Truckee (16561) and Grass Valley

Orange: Anaheim (352005) and Santa Ana

Placer: Roseville (139117) and Rocklin

Plumas: East Quincy (2489) and Chester

Riverside: Riverside (330063) and Moreno Valley

Sacramento: Sacramento (508529) and Elk Grove

San Benito: Hollister (39749) and Ridgemark

San Bernandino: San Bernandino (215941) and Fontana

San Diego: San Diego (1426000) and Chula Vista

San Francisco: San Francisco (810000) and Concord

San Joaquin: Stockton (311178) and Tracy

San Luis Obispo: San Luis Obispo (47446) and Paso Robles

San Mateo: Daly City (107008) and San Mateo

Santa Barbara: Santa Maria (107408) and Santa Barbara

Santa Clara: San Jose (1030000) and Sunnyvale

Santa Cruz: Santa Cruz (64725) and Watsonville

Shasta: Redding (91772) and Anderson

Sierra: Loyalton (700) and Downieville

Siskiyou: Yreka (7556) and Mount Shasta

Solano: Vallejo (121913) and Fairfield

Sonoma: Santa Rosa (177586) and Petaluma

Stanislaus: Modesto (215030) and Turlock

Sutter: Yuba City and South Yuba City

Tehama: Red Bluff (14283) and Corning

Trinity: Weaverville (3667) and Post Mountain

Tulare: Visalia (133800) and Tulare

Tuolumne: Phoenix Lake-Cedar Ridge (5108) and Sonora

Ventura: Oxnard (209877) and Thousand Oaks

Yolo: Davis (69289) and Woodland

Yuba: Linda (17773) and Olivehurst

Then, I created the function below to choose a group of stations that are within a certain distance (Haversine distance from the latitude and longitude) of each city that I identified in the previous step. If there are less than 5 stations associated to a given city, the distance threshold (which starts at 100km) is increased by 50km.

FUNCTION THAT SELECTS SET OF STATIONS CLOSEST TO A GIVEN LAT/LONG

```{r}
# Function to filter stations based on spatial coverage
subset_stations_by_spatial_coverage <- function(station_data, reference_lat, reference_lon, max_distance_km=100) {
  # Calculate distances between stations and reference location
  distances <- distHaversine(
    cbind(station_data$longitude, station_data$latitude),
    c(reference_lon, reference_lat)
  )
  distances <- distances/1000
  
  # idx =  which(distances == min(distances))
  # #Identify station within min distance to centroid of county
  # station_data_subset <- station_data[idx, ]
  
  # Subset stations within the specified max_distance_km
  idx = which(distances <= max_distance_km)
  station_data_subset <- station_data[idx, ]
  station_data_subset <- cbind(station_data_subset,distances[idx])
  
  while (nrow(station_data_subset) < 5){
    max_distance_km = max_distance_km + 50
    station_data_subset = subset_stations_by_spatial_coverage(station_data, 
                              reference_lat, reference_lon, max_distance_km)
  }
  
  return(station_data_subset)
}

# # Construct the subset of stations based on spatial coverage criteria (test)
# reference_lat = citylats[3]
# reference_lon = citylongs[3]
# max_distance_km = 100
# 
# subset_stations <- subset_stations_by_spatial_coverage(station_data2014_pm2.5, reference_lat, reference_lon, max_distance_km)
# 
# # Print the subset of stations
# print(subset_stations)
```

```{r,eval=FALSE}
# Obtain centroid lat/longs for each county 
CA.counties2 = read.csv("counties.ca.data.csv")
ca.coordinates = data.frame(CA.counties2$county,CA.counties2$lat,CA.counties2$lng)
colnames(ca.coordinates) = c("county","lat","long")

ca.coordinates = ca.coordinates[order(ca.coordinates$county),]
row.names(ca.coordinates) = NULL
```

IMPORTANT FUNCTION: 

Given a dataset containing station locations/codes for a given pollutant and year, the function below selects 5-20 stations that are closest to the lat/longs for the two biggest cities in each county and puts the station information (code, lat, long, etc) into a dataframe.

```{r,eval=FALSE}
#Function that finds best monitoring station for each county for a specific pollutant for a specific year
# 1 - Lead, 2 - Carbon monoxide (CO), 3 - Sulfure dioxide (SO2), 4 - Nitrogen dioxide (NO2)
# 5 - Ozone (O3), 6 - Total PM10, 7 - PM2.5

best_stations = function(stationdata,pollutant){

  subset_list = list()
  
  #Load lat/longs for 58x2 cities into dataframe
  CA.coords = data.frame(rep(countycodes$value_represented,each = 2),citylats,citylongs)
  colnames(CA.coords) = c("County","Lat","Long")
  
  #Find closest station for each county centroid using subset_stations_by_spatial_coverage function
  for (i in 1:nrow(CA.coords)){
  reference_lat = CA.coords$Lat[i]
  reference_lon = CA.coords$Long[i]
  max_distance_km = 100
  
  subset_stations <- subset_stations_by_spatial_coverage(stationdata[[pollutant]], reference_lat, reference_lon, max_distance_km)
  subset_list[[i]] = subset_stations
  }
  
  #Combine pairs of city lists together 
  subset_list2 = list()
  sequence = seq(2,116,2)
  for(i in sequence){
    combine = rbind(subset_list[[i]],subset_list[[i-1]])
    subset_list2[[i-1]] = combine
  }
  subset_list2 =subset_list2[!sapply(subset_list2,is.null)]
  
  #Create a county label vector
  repnames = c()
  for(i in 1:58){
    repnames = c(repnames,nrow(subset_list2[[i]]))
  }  
  countylabels = rep(countycodes$value_represented,times = repnames)
  
  #Format the list into dataframe
  beststations = as.data.frame(do.call(rbind, subset_list2))
  beststations = cbind(countylabels,beststations$county_name,
                       beststations$`distances[idx]`,beststations)
  colnames(beststations)[c(1,2,3)] = c("measuring_county","station_county","distance_apart")
  rownames(beststations) = NULL
  
  return(beststations)
}

# #test cases
# pm2.5_stations_2014 = best_stations(all_pollutants_station_2014data,7)
# CO_stations_2016 = best_stations(all_pollutants_station_2016data,2)
```

CREATING BEST STATION LIST/DATAFRAME FOR EACH POLLUTANT, EACH ENTRY IS A YEAR

```{r,eval=FALSE}
#Generate list for best stations for each pollutant for each year
Lead_stations = list()

Lead_stations[[1]] = best_stations(all_pollutants_station_2014data,1)
Lead_stations[[2]] = best_stations(all_pollutants_station_2015data,1)
Lead_stations[[3]] = best_stations(all_pollutants_station_2016data,1)
Lead_stations[[4]] = best_stations(all_pollutants_station_2017data,1)
Lead_stations[[5]] = best_stations(all_pollutants_station_2018data,1)
Lead_stations[[6]] = best_stations(all_pollutants_station_2019data,1)



CO_stations = list()

CO_stations[[1]] = best_stations(all_pollutants_station_2014data,2)
CO_stations[[2]] = best_stations(all_pollutants_station_2015data,2)
CO_stations[[3]] = best_stations(all_pollutants_station_2016data,2)
CO_stations[[4]] = best_stations(all_pollutants_station_2017data,2)
CO_stations[[5]] = best_stations(all_pollutants_station_2018data,2)
CO_stations[[6]] = best_stations(all_pollutants_station_2019data,2)



SO2_stations = list()

SO2_stations[[1]] = best_stations(all_pollutants_station_2014data,3)
SO2_stations[[2]] = best_stations(all_pollutants_station_2015data,3)
SO2_stations[[3]] = best_stations(all_pollutants_station_2016data,3)
SO2_stations[[4]] = best_stations(all_pollutants_station_2017data,3)
SO2_stations[[5]] = best_stations(all_pollutants_station_2018data,3)
SO2_stations[[6]] = best_stations(all_pollutants_station_2019data,3)



NO2_stations = list()

NO2_stations[[1]] = best_stations(all_pollutants_station_2014data,4)
NO2_stations[[2]] = best_stations(all_pollutants_station_2015data,4)
NO2_stations[[3]] = best_stations(all_pollutants_station_2016data,4)
NO2_stations[[4]] = best_stations(all_pollutants_station_2017data,4)
NO2_stations[[5]] = best_stations(all_pollutants_station_2018data,4)
NO2_stations[[6]] = best_stations(all_pollutants_station_2019data,4)



O3_stations = list()

O3_stations[[1]] = best_stations(all_pollutants_station_2014data,5)
O3_stations[[2]] = best_stations(all_pollutants_station_2015data,5)
O3_stations[[3]] = best_stations(all_pollutants_station_2016data,5)
O3_stations[[4]] = best_stations(all_pollutants_station_2017data,5)
O3_stations[[5]] = best_stations(all_pollutants_station_2018data,5)
O3_stations[[6]] = best_stations(all_pollutants_station_2019data,5)



PM10_stations = list()

PM10_stations[[1]] = best_stations(all_pollutants_station_2014data,6)
PM10_stations[[2]] = best_stations(all_pollutants_station_2015data,6)
PM10_stations[[3]] = best_stations(all_pollutants_station_2016data,6)
PM10_stations[[4]] = best_stations(all_pollutants_station_2017data,6)
PM10_stations[[5]] = best_stations(all_pollutants_station_2018data,6)
PM10_stations[[6]] = best_stations(all_pollutants_station_2019data,6)



# Lead.PM10_stations = list()
# 
# Lead.PM10_stations[[1]] = best_stations(all_pollutants_station_2014data,7)
# Lead.PM10_stations[[2]] = best_stations(all_pollutants_station_2015data,7)
# Lead.PM10_stations[[3]] = best_stations(all_pollutants_station_2016data,7)
# Lead.PM10_stations[[4]] = best_stations(all_pollutants_station_2017data,7)
# Lead.PM10_stations[[5]] = best_stations(all_pollutants_station_2018data,7)
# Lead.PM10_stations[[6]] = best_stations(all_pollutants_station_2019data,7)



PM2.5_stations = list()

PM2.5_stations[[1]] = best_stations(all_pollutants_station_2014data,7)
PM2.5_stations[[2]] = best_stations(all_pollutants_station_2015data,7)
PM2.5_stations[[3]] = best_stations(all_pollutants_station_2016data,7)
PM2.5_stations[[4]] = best_stations(all_pollutants_station_2017data,7)
PM2.5_stations[[5]] = best_stations(all_pollutants_station_2018data,7)
PM2.5_stations[[6]] = best_stations(all_pollutants_station_2019data,7)
```

# Heatmap of population with station locations marked (2015)

Do the stations provide a good spatial coverage of California? To me, the coverage is reasonable especially because most of the northern and eastern counties are where most of the sparsely populated counties are located. There are probably not that many EPA stations there as a result. 

```{r,cache=TRUE}
pollutants1_2015 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Raw/pollutants1_2015_8.17.RData")
pollutants2_2015 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Raw/pollutants2_2015_8.17.RData")

stationlats = c(unique(pollutants1_2015$latitude),unique(pollutants2_2015$latitude))
stationlongs = c(unique(pollutants1_2015$longitude),unique(pollutants2_2015$longitude))

station_points = data.frame(stationlats,stationlongs)

#Plot
gg_pop_stations <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = pop), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  labs(title = "Heatmap of County Populations with Station Locations for 2015") +
  scale_fill_gradient(low = "lightblue", high = "darkblue")

# Add points
gg_pop_stations <- gg_pop_stations +
  geom_point(data = station_points, aes(x = stationlongs, y = stationlats), 
             color = "red", size = 1.5)

print(gg_pop_stations)
```

# Downloading and aggregating air quality data using direct API calls

Given a set of 5-20 monitoring stations for each county, we loop through its station codes (for each year 2014-2019) and query using the EPA's AQS function. This function only allows you to query a maximum of 4 parameters at once for a single year, so two calls to the function have to be made for each year (4 and 3).

(BELOW SHOWS IT BEING DONE FOR 2019)

END GOAL FINAL FORM: ONE BIG DATAFRAME (ALL POLLUTANTS ALL YEARS TOGETHER, USE FILTER TO SEPARATE)

```{r,eval=FALSE}
# 1 - Lead, 2 - Carbon monoxide (CO), 3 - Sulfure dioxide (SO2), 4 - Nitrogen dioxide (NO2)
# 5 - Ozone (O3), 6 - Total PM10, 7 - PM2.5

stations2019x = rbind(Lead_stations[[6]],CO_stations[[6]],
             SO2_stations[[6]],NO2_stations[[6]])
stations2019y = rbind(O3_stations[[6]],PM10_stations[[6]],PM2.5_stations[[6]])

sitenums2019x = stations2019x %>% select(county_code,site_number) %>% unique() #198 stations
sitenums2019y = stations2019y %>% select(county_code,site_number) %>% unique() #178 stations


#Trying EPA R Package query (took 15 + 20 min!) gives us a dataframe 
ccodes = sitenums2019y$county_code
snums = sitenums2019y$site_number
str1 = "2019-01-01"
str2 = "2019-12-31"

pollutants1_2019 = aqs_dailysummary_by_site(parameter = c("14129","42101","42401","42602"),bdate = as.Date(str1),edate = as.Date(str2),stateFIPS = "06",countycode = ccodes,sitenum = snums)

pollutants2_2019 = aqs_dailysummary_by_site(parameter = c("44201","81102","88101"),bdate = as.Date(str1),edate = as.Date(str2),stateFIPS = "06",countycode = ccodes,sitenum = snums)

###SAVE LIST LOCALLY
saveRDS(pollutants2_2019,file = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Raw/pollutants2_2019_8.18.RData")
```

After querying the raw data for each pollutant and each year, we want to go through it and remove any data from stations that have "bad" data. The standards that I set are that for a given year, a station should have at least 240 of the 365 days included for a year of data. Additionally, I applied a Hampel filter to identify outliers and if there are more than 14 consecutive outliers in the data i.e., days in a row with measurements that are abnormal, I considered the data from that station to not be suitable for inclusion in the final dataset.

NOTE: the daily values reported in the raw dataset are actually daily averages from periodic measurements made by the station throughout the day

After data from so called "bad" stations were removed, another function is applied which aggregates the daily observations into a monthly median The raw data is in the form of a single dataframe, which is fed into the important function raw_transform(). This function uses several functions to create a list of dataframes, one for each county, which represent the monthly median measurements for a certain pollutant in a given county in a given year.

QUALITY CHECK FUNCTION FOR STATION DATA: WANT TO ADDRESS OUTLIERS, MISSINGNESS

```{r}
#Given a dataset like CO2016 (list of 1000ish stations), check for 2/3 missing data and for strings of outliers (14 in a row)

station_quality_check = function(station_data){
  l = length(station_data)
  badindex = c()
  consecutive_outliers = list()
  
  for (i in 1:l){
    aqi = station_data[[i]]$aqi
    pollutant_level = station_data[[i]]$arithmetic_mean
    
    #check for outliers in AQI
    median.aqi = median(na.omit(aqi))
    mad.aqi = mad(na.omit(aqi))
    
    min.aqi = median.aqi-(3*mad.aqi)
    max.aqi = median.aqi+(3*mad.aqi)
    
    outliers.aqi = which(aqi < min.aqi | aqi > max.aqi)
    
    result.aqi = rle(diff(outliers.aqi))

    
    #check for outliers in pollutant measure
    median.pollutant = median(na.omit(pollutant_level))
    mad.pollutant = mad(na.omit(pollutant_level))
    
    min.pollutant = median.pollutant-(3*mad.pollutant)
    max.pollutant = median.pollutant+(3*mad.pollutant)
    
    outliers.pollutant = which(pollutant_level < min.pollutant | pollutant_level > max.pollutant)
    
    result.pollutants = rle(diff(outliers.pollutant))
    
    
    if (nrow(station_data[[i]]) < 240){
      badindex = c(badindex,i)
    }
    
    else if (any(result.aqi$lengths >= 14 & result.aqi$values == 1) == TRUE){
      badindex = c(badindex,i)
    }
    
    else if (any(result.pollutants$lengths >= 14 & result.pollutants$values == 1) == TRUE){
      badindex = c(badindex,i)
    }
    
    consecutive_outliers[[i]] = c("AQI",outliers.aqi,"POLLUTANTS",outliers.pollutant)
  }
  
  bad_list = list(badindex,consecutive_outliers)
  
  return(bad_list)
}


#Test on CO2016 and CO2017
# station_quality_check(CO2016) #returns 61 "bad stations" out of 1230
# 
# removeidx = station_quality_check(SO22017)[[1]] #returns 715 out of 1056 "bad stations" 
# 
# test = SO22017[- removeidx]
```

FUNCTION THAT AGGREGATES DAILY DATA INTO MONTHYLY MEDIANS

```{r}
monthly_agg = function(pollutantdata){
  #Aggregating all the station data at once
  date = ymd(pollutantdata$date_local)
  df2 <- pollutantdata                                   # Duplicate data
  df2$year_month <- floor_date(date,"month")  # Create year-month column
  df3 = df2 %>% select(county,site_number,arithmetic_mean,aqi,year_month) %>% as.data.frame()
  
  df3$arithmetic_mean = as.numeric(df3$arithmetic_mean)
  df3$aqi[which(df3$aqi == "NULL")] = NA
  df3$aqi = as.numeric(df3$aqi)
  
  df.agg = df3 %>% group_by(year_month) %>% dplyr::summarize(arithmetic_mean = median(na.omit(arithmetic_mean)),aqi = median(na.omit(aqi))) %>% as.data.frame()
  
  return(df.agg)
}

###NEW FUNCTION (6/20/25) TO GET MONTHLY TOTALS OF PM2.5
monthly_agg2 = function(pollutantdata){
  #Aggregating all the station data at once
  date = ymd(pollutantdata$date_local)
  df2 <- pollutantdata                                   # Duplicate data
  df2$year_month <- floor_date(date,"month")  # Create year-month column
  df3 = df2 %>% select(county,site_number,arithmetic_mean,aqi,year_month) %>% as.data.frame()
  
  df3$arithmetic_mean = as.numeric(df3$arithmetic_mean)
  df3$aqi[which(df3$aqi == "NULL")] = NA
  df3$aqi = as.numeric(df3$aqi)

  df.agg = df3 %>% group_by(year_month,site_number) %>% dplyr::summarize(monthly_total = sum(na.omit(arithmetic_mean)),aqi = median(na.omit(aqi))) %>% as.data.frame()
  
  df.agg2 = df.agg %>% group_by(year_month) %>% dplyr::summarize(avg_monthly_total = median(monthly_total), aqi = median(aqi))
  
  return(df.agg2)
}
```

IMPORTANT FUNCTION: TRANSFORMING RAW DATA TO FINAL FORM

```{r,eval = FALSE}
# Group 1: 14129 - Lead, 421012 - Carbon monoxide (CO), 42401 - Sulfur dioxide (SO2), 42602 - Nitrogen dioxide (NO2)
# Group 2: 44201 - Ozone (O3), 81102 - Total PM10, 88101 - PM2.5

raw_transform = function(rawdata,reference_list,standard){
  
  ###SEPARATE DF INTO A LIST OF DFs 
  
  matched_list = list()
  
  if(missing(standard)){
    for (i in 1:nrow(reference_list)){
    data = rawdata %>% filter(county_code == reference_list$county_code[i], site_number == reference_list$site_number[i])
  
    matched_list[[i]] = data
    }
  } else {
      for (i in 1:nrow(reference_list)){
      data = rawdata %>% filter(county_code == reference_list$county_code[i], site_number == reference_list$site_number[i],pollutant_standard == standard)

      matched_list[[i]] = data
    }
  }
  
  names(matched_list) = reference_list$measuring_county
  
  ###STATION QUALITY CHECK
  removeidx = station_quality_check(matched_list)[[1]]
  good_matched_list = matched_list[- removeidx]
  
  #Convert list back into one big dataframe
  temp = as.data.frame(do.call(rbind, good_matched_list)) #TOO MANY ROWS RIGHT?
  good_df = unique.data.frame(temp)
  
  
  ###MAKE A LIST OF COMBINED STATION DATA FOR EACH COUNTY
  mid_list = list()

  for (i in unique(reference_list$measuring_county)){
    
    df_new = data.frame(good_df[1,])
    subset = reference_list %>% filter(measuring_county == i) %>% select(county_code,site_number)
  
    for (j in 1:nrow(subset)){
      pull = good_df %>% filter(county_code == reference_list$county_code[j], site_number == reference_list$site_number[j])
      
      df_new = rbind(df_new,pull)
    }
    
    df_new = df_new[-1,]
    mid_list[[i]] = df_new
  }
  
  ###AGGREGATE DAILY DATA TO MONTHLY FOR EACH COUNTY
  
  final_list = lapply(mid_list,monthly_agg2)
  
  return(final_list)
}
```

Assembling daily averages for one pollutant, one year from mid_list: 

```{r,eval = FALSE}
Lead_daily19 = data.frame()

for (i in 1:58){
  data = mid_list[[i]]
  
  county_df = data19 %>% dplyr::group_by(date_local) %>% 
    dplyr::summarise(daily_avg = mean(arithmetic_mean,na.rm = TRUE),
                     daily_avg_aqi = mean(aqi,na.rm = TRUE))
  
  county = rep(counties[i],nrow(county_df))
  county_df = cbind(county, county_df)
  
  Lead_daily19 = rbind(Lead_daily19,county_df)
}
```


When assembling final datasets, note that certain pollutant standards are used bc they have values for AQI... the ones I used were:

Lead: Lead 3-Month 2009 ?? Has all NAs for AQI

CO: CO 8-hour 1971

SO2: SO2 1-hour 2010

NO2: NO2 1-hour 2010

O3: Ozone 8-hour 2015 ; sample duration should be 8 HR

PM10: PM10 24-hour 2006

PM2.5: PM25 24-hour 2012

APPLY RAW TRANSFORM FUNCTION TO ALL POLLUTANTS FOR ALL YEARS (BELOW SHOWS IT BEING DONE FOR 2019)

```{r,eval=FALSE}
#Load raw data
pollutants1_2019 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Raw/pollutants1_2019_8.18.RData")
pollutants2_2019 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Raw/pollutants2_2019_8.18.RData")

Lead2019 = pollutants1_2019 %>% filter(parameter_code == "14129")
CO2019 = pollutants1_2019 %>% filter(parameter_code == "42101")
SO22019 = pollutants1_2019 %>% filter(parameter_code == "42401")
NO22019 = pollutants1_2019 %>% filter(parameter_code == "42602")
O32019 = pollutants2_2019 %>% filter(parameter_code == "44201")
PM102019 = pollutants2_2019 %>% filter(parameter_code == "81102")
PM2.52019 = pollutants2_2019 %>% filter(parameter_code == "88101")

Lead2019_final = raw_transform(rawdata = Lead2019,reference_list = Lead_stations[[6]],standard = "Lead 3-Month 2009")

CO2019_final = raw_transform(rawdata = CO2019,reference_list = CO_stations[[6]],standard = "CO 8-hour 1971")

SO22019_final = raw_transform(rawdata = SO22019,reference_list = SO2_stations[[6]],standard = "SO2 1-hour 2010")

NO22019_final = raw_transform(rawdata = NO22019,reference_list = NO2_stations[[6]],standard = "NO2 1-hour 2010")

O32019_final = raw_transform(rawdata = O32019,reference_list = O3_stations[[6]],standard = "Ozone 8-hour 2015")

PM102019_final = raw_transform(rawdata = PM102019,reference_list = PM10_stations[[6]],standard = "PM10 24-hour 2006") 

PM2.52019_final = raw_transform(rawdata = PM2.52019,reference_list = PM2.5_stations[[6]],standard = "PM25 24-hour 2012")
```

NEED TOTAL PM2.5 PER MONTH FOR PAPER 2

```{r,eval = FALSE}
pollutants2_2014 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Raw/pollutants2_2014_8.17.RData")
PM2.52014 = pollutants2_2014 %>% filter(parameter_code == "88101")

PM2.52014_final = raw_transform(rawdata = PM2.52014,reference_list = PM2.5_stations[[1]],standard = "PM25 24-hour 2012")

p1 = as.data.frame(do.call(rbind, PM2.52019_final))
p1 = cbind(p1,rep(pollutants$parametercodes.code[7],nrow(p1))) #maybe change parameter codes to 1-7?
colnames(p1) = c("Year-Month","Value","AQI","Pollutant")

p2 = as.data.frame(do.call(rbind, PM2.52018_final))
p2 = cbind(p2,rep(pollutants$parametercodes.code[7],nrow(p2))) #maybe change parameter codes to 1-7?
colnames(p2) = c("Year-Month","Value","AQI","Pollutant")

p3 = as.data.frame(do.call(rbind, PM2.52017_final))
p3 = cbind(p3,rep(pollutants$parametercodes.code[7],nrow(p3))) #maybe change parameter codes to 1-7?
colnames(p3) = c("Year-Month","Value","AQI","Pollutant")

p4 = as.data.frame(do.call(rbind, PM2.52016_final))
p4 = cbind(p4,rep(pollutants$parametercodes.code[7],nrow(p4))) #maybe change parameter codes to 1-7?
colnames(p4) = c("Year-Month","Value","AQI","Pollutant")

p5 = as.data.frame(do.call(rbind, PM2.52015_final))
p5 = cbind(p5,rep(pollutants$parametercodes.code[7],nrow(p5))) #maybe change parameter codes to 1-7?
colnames(p5) = c("Year-Month","Value","AQI","Pollutant")

p6 = as.data.frame(do.call(rbind, PM2.52014_final))
p6 = cbind(p6,rep(pollutants$parametercodes.code[7],nrow(p6))) #maybe change parameter codes to 1-7?
colnames(p6) = c("Year-Month","Value","AQI","Pollutant")

total_pm25_1419 = rbind(p6,p5,p4,p3,p2,p1)
c = rep(counties,each=12)
total_pm25_1419$county = rep(c,6)

saveRDS(total_pm25_1419,"total_pm25_1419_df.rds")
```

COMBINING EACH POLLUTANTS DATASET INTO A SINGLE DATAFRAME FOR THE YEAR (BELOW SHOWS IT BEING DONE FOR 2019)

```{r,eval=FALSE}
###Combine final data into one dataframe for 2014 
test1 = as.data.frame(do.call(rbind, Lead2019_final))
test1 = cbind(test1,rep(pollutants$parametercodes.code[1],nrow(test1))) #maybe change parameter codes to 1-7?
colnames(test1) = c("Year-Month","Value","AQI","Pollutant")

test2 = as.data.frame(do.call(rbind, CO2019_final))
test2 = cbind(test2,rep(pollutants$parametercodes.code[2],nrow(test2))) #maybe change parameter codes to 1-7?
colnames(test2) = c("Year-Month","Value","AQI","Pollutant")

test3 = as.data.frame(do.call(rbind, SO22019_final))
test3 = cbind(test3,rep(pollutants$parametercodes.code[3],nrow(test3))) #maybe change parameter codes to 1-7?
colnames(test3) = c("Year-Month","Value","AQI","Pollutant")

test4 = as.data.frame(do.call(rbind, NO22019_final))
test4 = cbind(test4,rep(pollutants$parametercodes.code[4],nrow(test4))) #maybe change parameter codes to 1-7?
colnames(test4) = c("Year-Month","Value","AQI","Pollutant")

test5 = as.data.frame(do.call(rbind, O32019_final))
test5 = cbind(test5,rep(pollutants$parametercodes.code[5],nrow(test5))) #maybe change parameter codes to 1-7?
colnames(test5) = c("Year-Month","Value","AQI","Pollutant")

test6 = as.data.frame(do.call(rbind, PM102019_final))
test6 = cbind(test6,rep(pollutants$parametercodes.code[6],nrow(test6))) #maybe change parameter codes to 1-7?
colnames(test6) = c("Year-Month","Value","AQI","Pollutant")

test7 = as.data.frame(do.call(rbind, PM2.52019_final))
test7 = cbind(test7,rep(pollutants$parametercodes.code[7],nrow(test7))) #maybe change parameter codes to 1-7?
colnames(test7) = c("Year-Month","Value","AQI","Pollutant")

#Combine each pollutant dataset into one big dataset for the year
final_data19 = rbind(test1,test2,test3,test4,test5,test6,test7)

###SAVE FINAL DATASET LOCALLY
saveRDS(final_data19,file = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Aggregated/final_data19_9.1.RData")
```

HAVE TO CLEAN DATA BEFORE FINALIZING:

Having combined the monthly medians for every county for a single year for each pollutant into a single dataframe, one final cleaning step must be performed before putting each years' data together. The AQI value that is in each row corresponds to the AQI standardized measurement for that specific pollutant. Each pollutant has a different standardizing equation, but once they are all standardized as they are in the dataset and they can be compared against each other. The reported AQI measurement for a given day is just the maximum of the AQI values corresponding to each of the 7 pollutants. So the maximum AQI value (among 6 values bc Lead observations never have AQI values) was found for each month and that value was set as the actual AQI value for that month in all corresponding rows.

FIND MAX AQI (AMONG THE 7 POLLUTANTS) FOR EACH MONTH -\> SET AS ACTUAL AQI FOR THAT MONTH

(BELOW SHOWS IT BEING DONE FOR 2019)

```{r,eval=FALSE}
#Do for each year 
months = c("01","02","03","04","05","06","07","08","09","10","11","12")

###Do for each year 
for (i in 1:58){
  idx1 = which(stringr::str_starts(rownames(final_data19), counties[i]))
  subset1 = final_data19[idx1,]
  subset1$`Year-Month`= as.Date(subset1$`Year-Month`)
  
  for (j in months){ 
  #Filter by county and date
    date = paste0("2019-",j,"-01")
    date = as.Date(date)
    subset2 = subset1 %>% filter(`Year-Month` == as.Date(date)) 
    
    trueAQI = max(na.omit(subset2$AQI))
    
    idx2 = which(subset1$`Year-Month` == date)
    subset1$AQI[idx2] = trueAQI
  }
  
  final_data19[idx1,] = subset1
}

###SAVE FINAL DATASET LOCALLY
saveRDS(final_data19,file = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Aggregated/final_data19_9.1.RData")
```

COMBINING EACH YEARS DATASET INTO ONE BIG TIDY DATAFRAME FOR AIR QUALITY COVARIATES

FILL IN YOUR OWN FILE DIRECTORIES HERE! START WORKING W ACTUAL EPA DATA FROM HERE ON

```{r}
final_data14 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Aggregated/final_data14_9.1.RData")
final_data15 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Aggregated/final_data15_9.1.RData")
final_data16 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Aggregated/final_data16_9.1.RData")
final_data17 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Aggregated/final_data17_9.1.RData")
final_data18 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Aggregated/final_data18_9.1.RData")
final_data19 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Aggregated/final_data19_9.1.RData")

final_EPA_data = rbind(final_data14,final_data15,final_data16,final_data17,
                       final_data18,final_data19)
saveRDS(final_EPA_data,file = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Final/final_EPA_data_9.1.RData")


final_EPA_data = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 2 Code/EPA data/Final/final_EPA_data_9.1.RData")
head(final_EPA_data)
```
