---
title: "P2-EDA-v3"
author: "Jeffrey Wu"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    df-print: paged
---

```{r,eval = FALSE}
load("C:/Users/jeffr/Desktop/Paper 2/p2-eda-workspace-2.27.RData")
```


# Downloading relevant data

```{r,warning=FALSE,echo=FALSE}
set.seed(2)

library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(huge)
library(INLA)
library(HMMpa)
library(invgamma)
library(brinla)
library(reshape2)
library(patchwork)
library(jsonlite)
library(geosphere)
library(urbnmapr)
library(RAQSAPI)
library(con2aqi)
library(pscl)
library(reshape2)
library(corrplot)
library(superheat)
library(shapes)
library(scales)
library(splines)
library(viridis)
library(RColorBrewer)
library(ggh4x)
```

## SDI data

```{r,message=FALSE,cache=TRUE,collapse=TRUE}
###SoA data
soa.data = read_xlsx("SoA.data.1019.xlsx")

###county_flips are unique identifier for counties
soa.data$county_fips = as.character(soa.data$county_fips) ##change it to character

#IMPORTANT

# This shape file contains the coordinates for county boundaries
##counties is from urbanmap

CA.counties = urbnmapr::counties %>% filter(state_abbv == "CA")

###IF WE WANT TO BOIL DOWN TIME SERIES AND KEEP ALL DATA, SWITCH to CA_newdata below
soa_joint <- left_join(CA.counties, soa.data, by = "county_fips")

#Use with soa.data (full data)
CA_data = soa_joint %>% select(long, lat, county_name.y, Year, Score, Total_Pop,
                               EDUC_Lessthan9, EDUC_college, White_Collar,
                               Unemployment_Rate, Adj_HH_income, Income_Disparity,
                               Individuals_Below_Poverty, Median_Home_Value,
                               Median_Gross_Rent, Housing_No_Telephone,
                               Housing_Incomplete_Plumbing)

### Quick descriptions of SoA data variables
# Score: social deprivation index (SDI) score calculated from the following 11 subindices
# EDUC_Lessthan9: % of population older than 24 with less than 9 years of education
# EDUC_college: % of population older than 24 with at least four years of college education
# White_Collar: % of population older than 15 employed in a white collar occupation
# Unemployment_Rate: unemployment rate for population older than 15
# Adj_HH_income: median household income adjusted for local housing costs
# Income_Disparity: an income disparity ratio
# Individuals_Below_Poverty: % of population below the federal poverty line
# Median_Home_Value: median home value for owned, occupied units
# Median_Gross_Rent: median gross rent for rented units
# Housing_No_Telephone: % of households without a telephone
# Housing_Incomplete_Plumbing: % of households with incomplete plumbing

colnames(CA_data)[3] = "County"

CA_newdata = soa.data[1:58,]
CA_newdata = CA_newdata[,-c(4,7,8)]
```

## Census population data

Heatmap of population by county:

```{r}
#Initializing map and station locations
ca_map <- map_data("county", region = "california")

#Match population dataset with ca_map
#2010-2019 population data for CA 
USpops = read.csv("CA_census_pops1019.csv")
CApops = USpops %>% filter(STNAME == "California") %>% select(CTYNAME,POPESTIMATE2019)
CApops = CApops[-1,]

CApops$CTYNAME = unique(ca_map$subregion)
colnames(CApops) = c("subregion","pop")

merged_data <- merge(ca_map, CApops, by = "subregion", all.x = TRUE)

#Plot
gg_pop <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = pop), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  labs(title = "Heatmap of County Populations for 2019",fill = expression("Population")) +
  scale_fill_gradient(low = "yellow", high = "red")

print(gg_pop)
```


```{r}
# Create a data frame with county names and corresponding GEOIDs
california_geoids <- data.frame(
  County = c("Alameda", "Alpine", "Amador", "Butte", "Calaveras", "Colusa", 
             "Contra Costa", "Del Norte", "El Dorado", "Fresno", "Glenn", 
             "Humboldt", "Imperial", "Inyo", "Kern", "Kings", "Lake", 
             "Lassen", "Los Angeles", "Madera", "Marin", "Mariposa", 
             "Mendocino", "Merced", "Modoc", "Mono", "Monterey", "Napa", 
             "Nevada", "Orange", "Placer", "Plumas", "Riverside", 
             "Sacramento", "San Benito", "San Bernardino", "San Diego", 
             "San Francisco", "San Joaquin", "San Luis Obispo", "San Mateo", 
             "Santa Barbara", "Santa Clara", "Santa Cruz", "Shasta", 
             "Sierra", "Siskiyou", "Solano", "Sonoma", "Stanislaus", 
             "Sutter", "Tehama", "Trinity", "Tulare", "Tuolumne", 
             "Ventura", "Yolo", "Yuba"),
  GEOID = c("06001", "06003", "06005", "06007", "06009", "06011", 
            "06013", "06015", "06017", "06019", "06021", "06023", 
            "06025", "06027", "06029", "06031", "06033", "06035", 
            "06037", "06039", "06041", "06043", "06045", "06047", 
            "06049", "06051", "06053", "06055", "06057", "06059", 
            "06061", "06063", "06065", "06067", "06069", "06071", 
            "06073", "06075", "06077", "06079", "06081", "06083", 
            "06085", "06087", "06089", "06091", "06093", "06095", 
            "06097", "06099", "06101", "06103", "06105", "06107", 
            "06109", "06111", "06113", "06115")
)

county_map <- map_data("county", region = "california")
```

Heatmaps for deprivation score for each year:

```{r}
for (y in 2014:2019){
  SDI_data = soa_joint %>% filter(Year == y)
  wildfire_data = wildfires %>% filter(year == y) %>% select(lat,long)

  lat0=min(SDI_data$lat)
  lat1=max(SDI_data$lat)
  
  heatmap = ggplot() + geom_polygon(data = SDI_data, aes(x = long, y = lat, group = group, fill = Score), 
               color = "black") + geom_point(data = wildfire_data, aes(x = long, y = lat),
                                             color = "blue", size = 3) + 
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  labs(title = "",fill = sprintf("SDI Score %1.0f",y)) +
  scale_fill_gradient(low = "yellow", high = "red")
  
  print(heatmap)
}
```

# SKATER clustering

Learn SKATER clusters on SDI data. We ended up using queen adjacency matrix not graph estimated with HUGE

```{r}
###Setting up SPDF for CA counties 
CA_sf = st_read(getwd(),"CA_Counties_TIGER2016")
CA_spdf = as_Spatial(CA_sf)

# score_scaled = scale(CA_data$Score)

c_index = unique(CA_data$County)
y_index = unique(CA_data$Year)
SDI_df = matrix(nrow=10,ncol=58)
track1 = 1
track2 = 1

for (i in c_index){
  for (j in y_index){
    scores = CA_data %>% filter(County == i) %>% filter(Year == j) %>% select(Score) %>% unique()
    SDI_df[track1,track2] = scores$Score
    track1 = track1 + 1
  }
  track1 = 1
  track2 = track2 + 1
}

score_scaled = scale(SDI_df) #NEED TO SCALE THE DATA BEFORE FEEDING IT INTO SKATER

#covariates_scale = data.frame(apply(CA_data[,4:16],2,scale))
covariates_scale = data.frame(t(score_scaled))

CA_spdf@data = covariates_scale
```

```{r}
#Identify neighborhood list for counties 
CA_nb = poly2nb(CA_spdf)

#summary(CA_nb)

# plot(CA_spdf, main = "With queen")
# plot(CA_nb, coords = coordinates(CA_spdf), col="blue", add = TRUE)

#Calculate edge costs (dissimilarity matrix) based on Euclidean distance 
costs <- nbcosts(CA_nb, data = covariates_scale)

###Get adjacency matrix using nb2mat() (SEPARATE STEP FOR INLA)
adj = nb2mat(CA_nb,style = "B")

#Style means the coding scheme style used to create the weighting matrix 
# B: basic binary coding scheme
# W: row standardized coding scheme 
# C: globally standardized coding scheme  
# U: values of C / number of neighbors 
# S: variance stabilizing coding scheme 

#Transform edge costs to spatial weights 
ct_w <- nb2listw(CA_nb,costs,style="B")

#Create minimum spanning tree 
ct_mst <- mstree(ct_w)

plot(ct_mst,coordinates(CA_spdf),col="blue", cex.lab=0.5)
plot(CA_spdf, add=TRUE)

#Run SKATER algorithm to get 7 contiguous clusters (cluster idx is in order of CA_sf)
clus5 <- skater(edges = ct_mst[,1:2], data = covariates_scale, ncuts = 4)

#Determine an appropriate minimum population threshold based on???
pops_summary = summary(unique(CA_data$Total_Pop))
pops_summary

#Idea 1: Use median * (how many counties should be in a cluster at minimum)
min_pop = as.numeric(pops_summary[3] * 4)

#Idea 2: If we assume CA population is 39M, divide total pop by # clusters
min_pop2 = 39000000 / 5

#Add a min population constraint
clus5_min <- skater(edges = ct_mst[,1:2], 
                     data = covariates_scale, 
                     crit = 5000000, 
                     vec.crit = CA_data$Total_Pop,
                     ncuts = 4)

#Add a minimum number of areas in each cluster constraint 
clus5_minarea = skater(edges = ct_mst[,1:2], data = covariates_scale, ncuts = 4, 4)


CA_data_cluster = (CA_sf %>% mutate(clus = clus5_min$groups))

#Plot clustered CA
plot((CA_sf %>% mutate(clus = clus5$groups))['clus'], main = "5 cluster example")

plot((CA_sf %>% mutate(clus = clus5_min$groups))['clus'], main = "5 cluster example with population constraint")

plot((CA_sf %>% mutate(clus = clus5_minarea$groups))['clus'], main = "5 cluster example with minimum number of areas constraint")
```

For reference, here are the cluster labels for each county:

```{r}
clusterlabels = data.frame(CA_data_cluster$NAME,clus5_min$groups)
names(clusterlabels) = c("counties","Cluster")

o = order(clusterlabels$counties)
clusterlabels = clusterlabels[o,]
rownames(clusterlabels) = NULL

clusterlabels
CA_counties = clusterlabels$counties
num_clus = max(clus5_min$groups)
num_counties = length(unique(CA_counties))
```

# EPA data

```{r}
final_EPA_data = readRDS("final_EPA_data_9.1.RData")

head(final_EPA_data)

CA_counties = california_geoids$County

county_labels = rep(CA_counties,each=12)
county_labels = rep(county_labels,7)
county_labels = rep(county_labels,6)

final_EPA_data = cbind(county_labels,final_EPA_data)

final_EPA_data <- final_EPA_data %>%
  spread(key = Pollutant, value = Value)

colnames(final_EPA_data) = c("county_labels","date","AQI","lead","co",
                             "so2","no2","o3","pm10","pm25")

time = rep(1:72,58)
final_EPA_data$time = time
```

```{r}
#Attach total pm25 instead of median
total_pm25 = readRDS("total_pm25_1419_df.rds")

test = total_pm25 %>% arrange(by = county)
final_EPA_data$pm25 = test$Value
```

We have included the code for downloading and aggregating the EPA data in a different script (Download-EPA.qmd) because it takes a long time to run: 

## Get county and pollutant reference codes from EPA

Make sure each air pollutant time series is stationary:

```{r}
#Get county codes 
counties_url = "https://aqs.epa.gov/data/api/list/countiesByState?email=jeffreywu@ucsb.edu&key=copperheron86&state=06&"

countycodes = fromJSON(counties_url)
countycodes = countycodes[[2]]
california_counties = countycodes$code

#Get parameter codes
parameters_url = "https://aqs.epa.gov/data/api/list/parametersByClass?email=jeffreywu@ucsb.edu&key=copperheron86&pc=CRITERIA"

parametercodes = fromJSON(parameters_url)
parametercodes = parametercodes[[2]]

#Do we want to keep Lead PM10 or delete
parametercodes = parametercodes[-7,]

pollutants = data.frame(parametercodes$code)
labels = c("lead","co","so2","no2","o3","pm10","pm25")
pollutants = cbind(pollutants,labels)
```

```{r,eval=FALSE}
library(astsa)

for (i in CA_counties){
  for (j in pollutants$parametercodes.code){
    EPA_clus_ts = final_EPA_data %>% filter(Pollutant == j,county_labels == i) 

    title = sprintf("Pollutant %s - %s",j,i)
    tsplot(EPA_clus_ts$Value,main = title,xlab = "Time (months)",ylab = "Value")  
  }
}
```

It turns out that the time series for Lead were not stationary, so we need to detrend the data before using it

```{r,eval=FALSE}
for (i in CA_counties){
  for (j in pollutants$parametercodes.code){
    EPA_clus_ts = final_EPA_data %>% filter(Pollutant == j,county_labels == i) 
    
    lead_idx = which(final_EPA_data$Pollutant == "14129" & final_EPA_data$county_labels == i)
    
    if (j == "14129"){
      detrended1 = c(0,diff(EPA_clus_ts$Value,lag=1))
      final_EPA_agg_data[lead_idx,2] = detrended1
      
      title = sprintf("Pollutant %s - %s",j,i)
      tsplot(detrended1,main = title,xlab = "Time (months)",ylab = expression(nabla~Value))
    } else{
      title = sprintf("Pollutant %s - %s",j,i)
      tsplot(EPA_clus_ts$Value,main = title,xlab = "Time (months)",ylab = "Value")  
    }
  }
}
```

EPA Stations plotted on a map: 

```{r}
#Initializing map and station locations
pollutants1_2019 = readRDS("pollutants1_2019_8.18.RData")
pollutants2_2019 = readRDS("pollutants2_2019_8.18.RData")

stationlats = c(unique(pollutants1_2019$latitude),unique(pollutants2_2019$latitude))
stationlongs = c(unique(pollutants1_2019$longitude),unique(pollutants2_2019$longitude))

station_points = data.frame(stationlats,stationlongs)

#Initializing map and station locations
ca_map <- map_data("county", region = "california")

#Plot
gg_epa_stations <- ggplot() +
  geom_polygon(data = ca_map, aes(x = long, y = lat, group = group), 
               fill = "white",color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  geom_point(data = station_points, aes(x = stationlongs, y = stationlats), 
             color = "red", size = 1.5) + 
  theme_void() +
  labs(title = "Active EPA Stations")

print(gg_epa_stations)
```

# Mortality data

```{r}
###Cal-ViDa data
mortality = read.csv("respmortality1419.csv") #data from 2014-2019 bc we want to avoid COVID pandemic era
mortality = filter(mortality,Cause_of_Death %in% c("Chronic lower respiratory diseases","Influenza and pneumonia"))
mortality = mortality[,-c(1,4,9)]
Population = rep(100000,nrow(mortality))
mortality = cbind(mortality,Population)
```

## Clean and aggregate mortality data

### EM ALGORITHM TO IMPUTE CENSORED MORTALITY ROWS (\<10 DEATHS)

```{r}
population_age = read_xlsx("Population Categories.xlsx")
#head(population_age)

population_age = population_age[-(1:5),2]
population_age$...2 = as.numeric(population_age$...2)

#split under 5 category into < 1 and 1-4 years old 
less1 = floor(population_age$...2[1]*0.2)
onefour = floor(population_age$...2[1]*0.8)
death_byage = population_age$...2[-1]
death_byage = c(less1,onefour,death_byage)
death_byage = death_byage[1:19]

death_byage2 = death_byage[1:2]
idx = seq(from = 3, to = 17, by = 2)
for (i in idx){
  death_byage2 = c(death_byage2,(death_byage[i]+death_byage[i+1]))
}
death_byage2 = c(death_byage2,death_byage[19])


age_groups = unique(mortality$Age)
death_byage2 = data.frame(cbind(age_groups,death_byage2))
colnames(death_byage2) = c("Age_Group","Population_by_Age")
death_byage2$Population_by_Age = as.numeric(death_byage2$Population_by_Age)
#head(death_byage2)

#2010-2019 population data for CA 
USpops = read.csv("CA_census_pops1019.csv")
CApops = USpops %>% filter(STNAME == "California") %>% select(CTYNAME,POPESTIMATE2019)
counties = countycodes$value_represented #from EPA data file

weights = CApops[(2:59),2]
weights = weights/CApops[1,2]

groups = unique(mortality$Age)
step1 = 1
step2 = 1

for (i in counties){
  for (j in groups){
    idx = which(mortality$Age == j & mortality$County_of_Death == i)
    mortality$Population[idx] = ceiling(death_byage2$Population[step1]*weights[step2])
    step1 = step1+1
  }  
  step1 = 1
  step2 = step2+1
}

mortality$logpop = log(mortality$Population)

censored_idx = which(mortality$Total_Deaths == "<11")
censorTF = mortality$Total_Deaths == "<11"
mortality = cbind(mortality,censorTF)
#head(mortality)
```

```{r}
uncensored_mortality = mortality %>% filter(censorTF == FALSE) %>% select(Total_Deaths,Population)
uncensored_mortality$Total_Deaths = as.numeric(uncensored_mortality$Total_Deaths)

theta = mean(uncensored_mortality$Total_Deaths*100000/uncensored_mortality$Population)
```

FUNCTION FOR IMPUTING CENSORED VALUE BASED ON EXPECTATION GIVEN LAMBDA

```{r}
impute_small_values = function(lambda){
  x = 1:10
  p = dpois(x,lambda)
  
  value = sum(x*p)/sum(p)
  return(value)
}
```

INITIAL IMPUTATION

```{r}
mortality2 = mortality
mortality2$Total_Deaths[censored_idx] = 0.01
mortality2$Total_Deaths = as.numeric(mortality2$Total_Deaths)

for (i in censored_idx){
  lambda = theta*mortality2$Population[i] / 100000
  deaths = impute_small_values(lambda)
  
  mortality2$Total_Deaths[i] = floor(deaths)
}
```

INITIAL REGRESSION MODELS

```{r}
mortality2$Age = factor(mortality2$Age)
mortality2$Cause_of_Death = factor(mortality2$Cause_of_Death)
mortality2$Month = factor(mortality2$Month)

pois_reg = glm(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop), family = "poisson", data = mortality2)
# ZIP_reg = zeroinfl(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop) | 1, data = mortality2, dist = "poisson", link = "logit")

vec0 = coef(pois_reg)
# vec0 = coef(ZIP_reg)

summary(pois_reg)
# summary(ZIP_reg)
```

```{r}
mortality3 = mortality2
model = pois_reg

model_diff = 100
iter = 1
vec0 = coef(model)

while((model_diff > 0.01) & (iter < 10)){
  
  #impute data (should be between 1-10)
  fvs = fitted.values(model)
  
  for (i in censored_idx){
    deaths = impute_small_values(fvs[i])
    mortality3$Total_Deaths[i] = floor(deaths)
  }
  
  #fit model on "new" data
  model = glm(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop), family = "poisson", data = mortality3)
  vec1 = coef(model)
  
  model_diff = max(abs(vec1 - vec0))
  iter = iter+1
  vec0 = vec1
}

final_pois_reg = model
summary(final_pois_reg)

logLik(final_pois_reg)
```

Reformatting the data set (mainly aggregating)

```{r,collapse=TRUE}
#Aggregating by  cause of death
data = mortality3

data1 = data %>% filter(Cause_of_Death == "Chronic lower respiratory diseases")
data2 = data %>% filter(Cause_of_Death == "Influenza and pneumonia")

newdeaths = data1$Total_Deaths + data2$Total_Deaths
data1$Total_Deaths = newdeaths

respmortality = data1[,-5]


#Creates total deaths by adding deaths of all age groups together 
agg.respmortality = respmortality[1,]
agg.respmortality$Age = as.character(agg.respmortality$Age)
rows2 = seq(1,nrow(respmortality),11)

for (i in rows2){
  agg.respmortality[i,] = respmortality[i,]
  agg.respmortality$Total_Deaths[i] = sum(respmortality$Total_Deaths[i:(i+10)])
  agg.respmortality$Age[i] = "Everyone"
}

agg.respmortality = na.omit(agg.respmortality)
rownames(agg.respmortality) = NULL
total.respmortality = agg.respmortality[,-c(6:9)]
```

Reformat dataset into time series format (rows are counties, columns are months)

```{r,cache=FALSE}
months = unique(total.respmortality$Month_of_Death)
years = sort(unique(total.respmortality$Year_of_Death))
counties = unique(total.respmortality$County_of_Death)
x = 0

total.mortality.ts = matrix(1,nrow = 58, ncol = 72)

for (k in counties){
  county.ts = c()
  x = x+1
  
  for (i in years){
    for (j in months){
      deaths = total.respmortality %>% filter(County_of_Death == k) %>%  filter(Year_of_Death == i) %>% filter(Month_of_Death == j) %>% select(Total_Deaths) %>% as.numeric()
      county.ts = c(county.ts,deaths)
    }
  } 
  
  total.mortality.ts[x,] = county.ts
}

#Label time series data 
total.mortality.ts = as.data.frame(total.mortality.ts)

dates = c()
x=1
for (i in years){
  for (j in months){
    dates[x] = sprintf("%1.0f/%1.0f",j,i)
    x = x+1
  }
}

colnames(total.mortality.ts) = dates
ID = c(1:58)
total.mortality.ts = cbind(ID,counties,total.mortality.ts)
head(total.mortality.ts)
```

```{r}
# Convert total.mortality.ts to a long format data frame
library(tidyr)

# Assuming total.mortality.ts is already a data frame
total.mortality.ts_long <- total.mortality.ts %>%
  pivot_longer(cols = -c(ID, counties), names_to = "date", values_to = "deaths")

# Convert date to a proper date format
total.mortality.ts_long$date <- as.Date(paste0("01/", total.mortality.ts_long$date), format = "%d/%m/%Y")

# Create a time series plot for each county
ggplot(total.mortality.ts_long, aes(x = date, y = deaths, color = counties)) +
  geom_line() +
  facet_wrap(~ counties, scales = "free_y") +
  labs(title = "Time Series of Total Mortality by County",
       x = "Date",
       y = "Total Deaths") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
# Ensure the date column is in Date format
total.mortality.ts_long$date <- as.Date(total.mortality.ts_long$date, format = "%d/%m/%Y")

# Extract the year from the date
total.mortality.ts_long$Year <- format(total.mortality.ts_long$date, "%Y")
total.mortality.ts_long$Year = as.numeric(total.mortality.ts_long$Year)

# Merge total.mortality.ts_long with countypops to get the population for each county and year
countypops = CA_data %>% filter(Year > 2013) %>% select(Total_Pop,County,Year) %>% unique()
countypops$County = rep(california_geoids$County,each=6)

total.mortality.ts_long <- total.mortality.ts_long %>%
  left_join(countypops, by = c("counties" = "County", "Year" = "Year"))

# Calculate the death rate per 100,000 people
total.mortality.ts_long <- total.mortality.ts_long %>%
  mutate(death_rate = (deaths / Total_Pop) * 100000)

# Create a time series plot for each county
ggplot(total.mortality.ts_long, aes(x = date, y = death_rate, color = counties)) +
  geom_line() +
  facet_wrap(~ counties, scales = "free_y") +
  labs(title = "Time Series of Mortality Rate (per 100k) by County",
       x = "Date",
       y = "Mortality Rate (per 100k)") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
sb_test = total.mortality.ts_long %>% filter(counties == "Santa Barbara")

# Create a time series plot for each county
ggplot(sb_test, aes(x = date, y = death_rate, color = counties)) +
  geom_line() +
  facet_wrap(~ counties, scales = "free_y") +
  labs(title = "Time Series of Mortality Rate",
       x = "Date",
       y = "Respiratory Related Mortality Rate (per 100k)") +
  theme_minimal() +
  theme(legend.position = "none")

acf(sb_test$death_rate)
pacf(sb_test$death_rate)
```

Heatmap for deaths per 100,000 for Jan 2018: 

```{r}
deaths_df = total.mortality.ts_long %>% filter(date == "2018-01-01")
deaths_df$counties = tolower(deaths_df$counties)

#Plot onto a heatmap
# Merge county map data with the dataset containing deaths
county_map_deaths <- merge(county_map, deaths_df, by.x = "subregion", by.y = "counties")

ggplot() + geom_polygon(data = county_map_deaths, aes(x = long, y = lat, group = group, fill = death_rate), 
             color = "black") +
coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
theme_void() +
labs(title = "",fill = expression("Death rate (per 100,000 people) for Jan 2018")) +
scale_fill_gradient(low = "yellow", high = "red")
```

# Visualize correlations between different counties 

```{r}
library(corrplot)

#Ordered according to hclust
cor_df = total.mortality.ts_long |> pivot_wider(id_cols = date, names_from = counties, values_from=death_rate)
cor_df = cor_df %>% select(-date)

corrplot(cor(cor_df,method = "spearman"),order = 'hclust')


#Ordered according to SKATER clustering
clusterlabels2 = clusterlabels[order(clusterlabels$Cluster),]

cor_df2 = cor_df[,order(clusterlabels$Cluster)]
colnames(cor_df2) = clusterlabels2$counties

corrplot(cor(cor_df2,method = "spearman"))
```

## Cluster analysis of the time series correlations between counties

```{r}
county_corrs = cor(cor_df,method="spearman")

corr_pairs <- melt(county_corrs)
corr_pairs <- corr_pairs[upper.tri(county_corrs), ]  # Extract upper triangle without diagonal
colnames(corr_pairs) <- c("County1", "County2", "Correlation")

# Perform clustering to see if counties group like they do with SKATER

###Hierarchical clustering 
distance_matrix <- as.dist(1 - county_corrs)  

hc <- hclust(distance_matrix, method = "average")  # You can use "ward.D", "complete", etc.
clusters <- cutree(hc, k = 5)  

# Plot the dendrogram
plot(hc, main = "Hierarchical Clustering of Counties Based on Correlation", cex = 0.8)


###K-Means Clustering
set.seed(123)
km_clusters <- kmeans(county_corrs, centers = 5, nstart = 20)

#Check how well these two match -- not very well
km_clusters$cluster
clusterlabels
```


# Wildfire PM 2.5 data

Filter CA counties, attach county names to measurements

```{r,eval=FALSE}
# Load smokePM predictions on smoke days
preds = readRDS("smokePM2pt5_predictions_daily_county_20060101-20201231.rds")

# Load counties
ca_geoinfo = readRDS("ca_geodata.rds")

ca_geoid_counties = ca_geoinfo %>% select(GEOID,NAME)

# Load full set of dates
dates = seq.Date(ymd("20060101"), ymd("20201231"), by = "day")

# Get full combination of county-days
# Warning: this may require a large amount of memory
out = expand.grid(GEOID = ca_geoid_counties$GEOID, date = dates)

# Match smokePM predictions on smoke days to county-days
out = left_join(out, preds, by = c("GEOID", "date"))
out2 = left_join(ca_geoid_counties,out,by="GEOID")
out3 = out2 %>% filter(date > "2014-01-01" & date < "2020-01-01")

# Predict 0 for remaining county-days, which are non-smoke days
out = mutate(out3, smokePM_pred = replace_na(smokePM_pred, 0))

# test = out %>% filter(date > "2015-01-01",GEOID == "06111")
```

## Aggregate daily observations to monthly level by GEOID (takes a while)

```{r,eval=FALSE}
# Ensure the date column is in Date format
out$date <- ymd(out$date)

out_subset = out %>% filter(NAME == "Santa Barbara")

# Aggregate daily observations to monthly averages by GEOID
# monthly_avg = out %>%
#   group_by(GEOID, NAME, date = floor_date(date, "month")) %>%
#   summarize(monthly_avg_pm25 = mean(smokePM_pred, na.rm = TRUE))

# Aggregate daily observations to monthly totals by GEOID
# monthly_total = out %>%
#   group_by(GEOID, NAME, date = floor_date(date, "month")) %>%
#   summarize(monthly_total_pm25 = sum(smokePM_pred, na.rm = TRUE))

# Count number of "extreme" smoke days (>20) per month 
num_extreme_days10 = out %>% 
  group_by(GEOID, NAME, date = floor_date(date, "month")) %>%
  summarize(num_extreme_days = length(which(smokePM_pred > 10)))

num_extreme_days20 = out %>% 
  group_by(GEOID, NAME, date = floor_date(date, "month")) %>%
  summarize(num_extreme_days = length(which(smokePM_pred > 20)))

colnames(num_extreme_days10)[4] = "num_extreme_days10"
colnames(num_extreme_days20)[4] = "num_extreme_days20"


saveRDS(num_extreme_days10,"num_extreme_days10.rds")
saveRDS(num_extreme_days20,"num_extreme_days20.rds")

# View the aggregated dataframe
#View(monthly_avg)
```

```{r}
# monthly_avg = readRDS(file="wildfirepm25_monthly_avg.rds")

monthly_total = readRDS(file="wildfirepm25_monthly_total.rds")
```

### Notable wildfires from 2014-2019

```{r}
wildfires <- data.frame(
  wildfire_name = c("King Fire", "Happy Camp Complex Fire", "July Complex Fire", 
                    "Bald Fire", "Lodge Lightning Complex Fire","Butte Fire", 
                    "Valley Fire", "Rocky Fire", "Jerusalem Fire", "Rough Fire",
                    "Soberanes Fire", "Erskine Fire", "Sand Fire", "Blue Cut Fire", "Chimney Fire",
                    "Thomas Fire", "Tubbs Fire", "Nuns Fire", "Atlas Fire", "Detwiler Fire",
                    "Camp Fire", "Woolsey Fire", "Carr Fire", "Mendocino Complex Fire", "Ferguson Fire",
                    "Kincade Fire", "Tick Fire", "Walker Fire", "Maria Fire", "Saddle Ridge Fire"),
  year = c(2014, 2014, 2014, 2014, 2014,
           2015, 2015, 2015, 2015, 2015,
           2016, 2016, 2016, 2016, 2016,
           2017, 2017, 2017, 2017, 2017,
           2018, 2018, 2018, 2018, 2018,
           2019, 2019, 2019, 2019, 2019),
  lat = c(38.8000, 41.7900, 41.6000, 40.8000, 39.8000,
          38.3470, 38.7980, 38.8830, 38.8830, 36.9000,
          36.4600, 35.6100, 34.4000, 34.3000, 35.7500,
          34.4150, 38.5766, 38.4330, 38.4040, 37.6600,
          39.8136, 34.2350, 40.6540, 39.1773, 37.6540,
          38.7920, 34.4730, 40.0650, 34.3350, 34.3290),
  long = c(-120.6500, -123.3800, -123.1000, -121.5000, -123.5000,
           -120.6800, -122.5600, -122.4760, -122.5340, -118.9000,
           -121.9000, -118.4700, -118.5000, -117.5000, -120.9000,
           -119.0910, -122.5911, -122.6240, -122.2320, -120.2400,
           -121.4344, -118.7013, -122.6230, -123.2001, -119.8460,
           -122.7800, -118.3680, -120.5780, -119.0300, -118.4760)
)
```

# Weather data

## Download and aggregate

```{r,eval=FALSE}
weather_station_data = read.csv("cimis-monthly-weather.csv",header=TRUE)
weather_station_data$date = as.Date(paste("01", weather_station_data$Month.Year), format = "%d %b %Y")

library(cimir)

# Set your API key
set_key("ec67fb7d-c8cd-4ff7-b173-832a381b6968")

# Retrieve metadata for specific stations
station_ids <- c(2, 5, 6, 7, 12, 13, 15, 35, 39, 41, 43, 44, 47, 52, 64, 68, 70, 71, 75, 77, 78, 80, 83, 84, 87, 88, 90, 91, 99, 103, 104, 105, 106, 107, 113, 114, 116, 117, 124, 125, 126, 129, 131, 135, 136, 139, 140, 143, 144, 146, 147, 150, 151, 152, 153, 157, 158, 159, 160, 163, 165, 170) 

station_ids2 = c(171, 173, 174, 175, 178, 179, 181, 182, 183, 184, 187, 189, 191, 192, 193, 194, 195, 197, 199, 200, 202, 204, 205, 206, 207, 208, 209, 210, 211, 212, 213, 214, 215, 216, 217, 218, 220, 221, 222, 223, 224, 225, 226, 227, 228, 229, 231, 232, 233, 235, 236, 237, 239, 240, 241, 242, 243, 244, 245, 246, 247, 248, 249, 250, 251, 252, 253, 254, 256, 257, 258, 259, 260, 261, 262, 263, 264)

station_metadata <- cimis_station(station_ids)
station_metadata2 <- cimis_station(station_ids2)

station_metadata = rbind(station_metadata,station_metadata2)
station_lats = unique(station_metadata$HmsLatitude)
station_longs = unique(station_metadata$HmsLongitude)

station_coords = data.frame(station_id = as.numeric(unique(station_metadata$StationNbr)),
                            station_location = unique(station_metadata$Name),
                            lat = station_lats, long = station_longs)


weather_station_data = weather_station_data %>% left_join(station_coords,by=c("Stn.Id" = "station_id"))
```

## Plot CIMIS stations

```{r}
weather_station_data = readRDS("cimis_weather_station_data.rds")
station_coords = readRDS("cimis_station_coords.rds")
weather_station_data = weather_station_data[-8945,]

# Function to convert DMS to Decimal Degrees
dms_to_decimal <- function(dms) {
  # Extract degrees, minutes, and seconds using regular expressions
  matches <- regmatches(dms, gregexpr("(\\d+)[º°](\\d+)'(\\d+(\\.\\d+)?)", dms))[[1]]
  if (length(matches) == 0) {
    stop("No valid DMS format found in: ", dms)
  }
  
  # Split the matched components
  parts <- unlist(strsplit(matches, "[º°'\"NnSsEeWw]"))
  deg <- as.numeric(parts[1])
  min <- as.numeric(parts[2])
  sec <- as.numeric(parts[3])
  
  # Calculate decimal degrees
  decimal <- deg + min / 60 + sec / 3600
  
  # If direction is South or West, make it negative
  if (grepl("[SsWw]", dms)) {
    decimal <- -decimal
  }
  
  return(decimal)
}

weather_station_data$lat = sapply(weather_station_data$lat,dms_to_decimal)
weather_station_data$long = sapply(weather_station_data$long,dms_to_decimal)

station_coords$lat = sapply(station_coords$lat,dms_to_decimal)
station_coords$long = sapply(station_coords$long,dms_to_decimal)

# Plot the map with the station coordinates
gg_cimis_stations = ggplot() +
  # Plot the California map
  geom_polygon(data = ca_map, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  # Add the stations as points
  geom_point(data = weather_station_data, aes(x = long, y = lat), 
             color = "red", size = 1.5) +
  # Add labels for each station
  # geom_text(data = weather_station_data, aes(x = long, y = lat, label = Stn.Id), 
  #           nudge_y = 0.1, size = 3) +
  labs(title = "CIMIS Weather Stations", 
       x = "Longitude", y = "Latitude") +
  theme_void()

print(gg_cimis_stations)
```

```{r,eval=FALSE}
plot_list = list(gg_epa_stations,gg_cimis_stations)
combined = wrap_plots(plotlist = plot_list, ncol = 2)

# Save one combined figure
ggsave(
  filename = paste0("epa_cimis_stations_map_plot.png"),
  plot = combined,
  path = "/home/jovyan/work/important-plots",
  width = 8,
  height = 6,
  dpi = 600
)
```

## Identify subset of stations that has best spatial coverage wrt CA state (same process as EPA data)

Starting off with a list of all the stations measuring each parameter/pollutant, we want to identify a subset of stations that measure each pollutant for each county. This overall set of stations (subset for each county combined together) should have a good spatial coverage of the state.

This is difficult because there is not a station in every county measuring each pollutant. So in order to identify a good set of stations to query data from, I first looked up the two largest cities in each county, based on population.

Lat/long for 2 biggest cities (based on population) in each county

```{r}
citylats = c(37.8044,37.5485,38.7743,38.8071,38.3527,38.3488,39.7285,39.7596,
             38.1231,38.0678,39.2143,39.1546,37.9780,38.0049,41.7558,41.7548,
             38.9399,38.6688,36.7378,36.8252,39.7474,39.5243,40.8021,40.8665,
             32.7920,32.6789,37.3614,37.3855,35.3733,35.7688,36.3275,36.3008,
             38.9582,38.8080,40.4163,40.2840,34.0522,33.7701,36.9613,37.1230,
             37.9735,38.1074,37.4849,37.4320,39.4457,39.4096,37.3022,37.0583,
             41.4871,41.4099,37.6485,38.5149,36.6777,36.6149,38.2975,38.1749,
             39.3280,39.2191,33.8366,33.7455,38.7521,38.7907,39.9341,40.3063,
             33.9806,33.9425,38.5816,38.4088,36.8525,36.8125,34.1083,34.0922,
             32.7157,32.6401,37.7749,37.9780,37.9577,37.7396,35.2828,35.6369,
             37.6879,37.5630,34.9530,34.4208,37.3387,37.3688,36.9741,36.9102,
             40.5865,40.4482,39.6763,39.5595,41.7354,41.3099,38.1041,38.2492,
             38.4404,38.2324,37.6393,37.4946,39.1404,39.1165,40.1785,39.9277,
             40.7310,40.4156,36.3301,36.2077,38.0297,37.9829,34.1975,34.1706,
             38.5449,38.6785,39.1277,39.0954)

citylongs = c(122.2712,121.9886,119.8219,119.7960,120.9327,120.7741,121.8375,121.6219,
              120.8509,120.5385,122.0094,122.1494,122.0311,121.8058,124.2026,124.1580,
              119.9772,120.9872,119.7871,119.7029,122.1964,122.1936,124.1637,124.0828,
              115.5631,115.4989,118.3997,118.4105,119.0187,119.2471,119.6457,119.7829,
              122.6264,122.5583,120.6530,120.5394,118.2437,118.1937,120.0607,120.2602,
              122.5311,122.5697,119.9663,120.0985,123.8053,123.3556,120.4830,120.8499,
              120.5425,120.6791,118.9721,119.4768,121.6555,121.8221,122.2869,122.2608,
              120.1833,121.0611,117.9143,117.8677,121.2880,121.2358,120.8980,121.2319,
              117.3755,117.2297,121.4944,121.3716,121.4016,121.3658,117.2898,117.4350,
              117.1611,117.0842,122.4194,122.0311,121.2908,121.4260,120.6596,120.6545,
              122.4702,122.3255,120.4357,119.6982,121.8853,122.0363,122.0308,121.7569,
              122.3917,122.2978,120.2410,120.8277,122.6345,122.3106,122.2566,122.0405,
              122.7141,122.6367,120.9970,120.8460,121.6169,121.6380,122.2358,122.1792,
              122.9420,123.2100,119.2966,119.3473,119.9741,120.3822,119.1771,118.8376,
              121.7405,121.7733,121.5508,121.5522)
citylongs = -1*citylongs
```

Alameda: Oakland (429082) and Fremont

Alpine: Alpine Village (225) and Mesa Vista

Amador: Ione (8363) and Jackson

Butte: Chico (94776) and Paradise

Calaveras: Rancho Calaveras (5324) and Angels Camp

Colusa: Colusa (5911) and Williams

Contra Costa: Concord (129688) and Antioch

Del Norte: Crescent City (6805) and Bertsch-Oceanview

El Dorado: South Lake Tahoe (22036) and Cameron Park

Fresno: Fresno (530093) and Clovis

Glenn: Orland (7644) and Willows

Humboldt: Eureka (26998) and Arcata

Imperial: El Centro (44120) and Calexico

Inyo: Bishop (3746) and Dixon Lane-Meadow Creek

Kern: Bakersfield (383579) and Delano

Kings: Hanford (56910) and Lemoore

Lake: Clearlake (15384) and Hidden Valley Lake

Lassen: Susanville (15165) and Janesville

Los Angeles: Los Angeles (3990000) and Long Beach

Madera: Madera (65706) and Chowchilla

Marin: San Rafael (58704) and Novato

Mariposa: Mariposa (1526) and Catheys Valley

Mendocino: Fort Bragg (7359) and Willits

Merced: Merced (83316) and Los Banos

Modoc: Alturas (2509) and California Pines

Mono: Mammoth Lakes (8127) and Walker

Monterey: Salinas (156259) and Seaside

Napa: Napa (79263) and American Canyon

Nevada: Truckee (16561) and Grass Valley

Orange: Anaheim (352005) and Santa Ana

Placer: Roseville (139117) and Rocklin

Plumas: East Quincy (2489) and Chester

Riverside: Riverside (330063) and Moreno Valley

Sacramento: Sacramento (508529) and Elk Grove

San Benito: Hollister (39749) and Ridgemark

San Bernandino: San Bernandino (215941) and Fontana

San Diego: San Diego (1426000) and Chula Vista

San Francisco: San Francisco (810000) and Concord

San Joaquin: Stockton (311178) and Tracy

San Luis Obispo: San Luis Obispo (47446) and Paso Robles

San Mateo: Daly City (107008) and San Mateo

Santa Barbara: Santa Maria (107408) and Santa Barbara

Santa Clara: San Jose (1030000) and Sunnyvale

Santa Cruz: Santa Cruz (64725) and Watsonville

Shasta: Redding (91772) and Anderson

Sierra: Loyalton (700) and Downieville

Siskiyou: Yreka (7556) and Mount Shasta

Solano: Vallejo (121913) and Fairfield

Sonoma: Santa Rosa (177586) and Petaluma

Stanislaus: Modesto (215030) and Turlock

Sutter: Yuba City and South Yuba City

Tehama: Red Bluff (14283) and Corning

Trinity: Weaverville (3667) and Post Mountain

Tulare: Visalia (133800) and Tulare

Tuolumne: Phoenix Lake-Cedar Ridge (5108) and Sonora

Ventura: Oxnard (209877) and Thousand Oaks

Yolo: Davis (69289) and Woodland

Yuba: Linda (17773) and Olivehurst

Then, I created the function below to choose a group of stations that are within a certain distance (Haversine distance from the latitude and longitude) of each city that I identified in the previous step. If there are less than 5 stations associated to a given city, the distance threshold (which starts at 100km) is increased by 50km.

FUNCTION THAT SELECTS SET OF STATIONS CLOSEST TO A GIVEN LAT/LONG

```{r}
# Function to filter stations based on spatial coverage
subset_stations_by_spatial_coverage <- function(station_data, reference_lat, reference_lon, max_distance_km=100) {
  # Calculate distances between stations and reference location
  distances <- distHaversine(
    cbind(station_data$long, station_data$lat),
    c(reference_lon, reference_lat)
  )
  distances <- distances/1000
  
  # idx =  which(distances == min(distances))
  # #Identify station within min distance to centroid of county
  # station_data_subset <- station_data[idx, ]
  
  # Subset stations within the specified max_distance_km
  idx = which(distances <= max_distance_km)
  station_data_subset <- station_data[idx, ]
  station_data_subset <- cbind(station_data_subset,distances[idx])
  
  while (nrow(station_data_subset) < 5){
    max_distance_km = max_distance_km + 50
    station_data_subset = subset_stations_by_spatial_coverage(station_data, 
                              reference_lat, reference_lon, max_distance_km)
  }
  
  return(station_data_subset)
}

# # Construct the subset of stations based on spatial coverage criteria (test)
# reference_lat = citylats[3]
# reference_lon = citylongs[3]
# max_distance_km = 100
# 
# subset_stations <- subset_stations_by_spatial_coverage(station_data2014_pm2.5, reference_lat, reference_lon, max_distance_km)
# 
# # Print the subset of stations
# print(subset_stations)
```

```{r,eval=FALSE}
# Obtain centroid lat/longs for each county 
CA.counties2 = read.csv("counties.ca.data.csv")
ca.coordinates = data.frame(CA.counties2$county,CA.counties2$lat,CA.counties2$lng)
colnames(ca.coordinates) = c("county","lat","long")

ca.coordinates = ca.coordinates[order(ca.coordinates$county),]
row.names(ca.coordinates) = NULL
```

IMPORTANT FUNCTION: 

Given a dataset containing station locations/codes, the function below selects 5-20 stations that are closest to the lat/longs for the two biggest cities in each county and puts the station information (code, lat, long, etc) into a dataframe.

```{r}
best_stations = function(stationdata){

  subset_list = list()
  
  #Load lat/longs for 58x2 cities into dataframe
  CA.coords = data.frame(rep(countycodes$value_represented,each = 2),citylats,citylongs)
  colnames(CA.coords) = c("County","Lat","Long")
  
  #Find closest station for each county centroid using subset_stations_by_spatial_coverage function
  for (i in 1:nrow(CA.coords)){
  reference_lat = CA.coords$Lat[i]
  reference_lon = CA.coords$Long[i]
  max_distance_km = 50
  
  subset_stations <- subset_stations_by_spatial_coverage(stationdata, reference_lat, reference_lon, max_distance_km)
  subset_list[[i]] = subset_stations
  }
  
  #Combine pairs of city lists together 
  subset_list2 = list()
  sequence = seq(2,116,2)
  for(i in sequence){
    combine = rbind(subset_list[[i]],subset_list[[i-1]])
    subset_list2[[i-1]] = combine
  }
  subset_list2 =subset_list2[!sapply(subset_list2,is.null)]
  
  #Create a county label vector
  repnames = c()
  for(i in 1:58){
    repnames = c(repnames,nrow(subset_list2[[i]]))
  }  
  countylabels = rep(countycodes$value_represented,times = repnames)
  
  #Format the list into dataframe
  beststations = as.data.frame(do.call(rbind, subset_list2))
  beststations = cbind(countylabels,
                       beststations$`distances[idx]`,beststations)
  colnames(beststations)[1:2] = c("measuring_county","distance_apart")
  rownames(beststations) = NULL
  beststations = beststations[,-7]
  
  return(beststations)
}

beststations = best_stations(station_coords)
```

## Label which county each station is measuring (in order)

```{r}
weather_df = data.frame()

for(i in counties){
  county_data = beststations %>% filter(measuring_county == i) 
  idx = which(weather_station_data$Stn.Id %in% county_data$station_id)
  new_df = weather_station_data[idx,]
  new_df$measuring_county = i
  weather_df = rbind(weather_df,new_df)
}

unique(weather_df$measuring_county)
```

Some counties don't have a complete time series of 72... identify them and impute? 

```{r}
missing_list = list()

for (i in counties){
  county = weather_df %>% filter(measuring_county == i)
  missing_months = dates[!(dates %in% unique(county$date))]
  missing_list[[i]] = missing_months
}
```

Turns out it's just Del Norte...

```{r}
del_norte = weather_df %>% filter(measuring_county == "Del Norte")
dates[!(dates %in% unique(del_norte$date))]
```

## Aggregate data from different stations to get a value for each county 

```{r}
dates = unique(weather_df$date)
final_weather_df = data.frame()

for (i in counties){
  for (j in dates){
    county_month_data = weather_df %>% filter(measuring_county == i, date == j) %>% 
      select(Total.Precip..in.,Avg.Air.Temp..F.,Avg.Dew.Point..F.,Avg.Wind.Speed..mph.)
    
    county = i
    date = j
    total_precip = sum(county_month_data$Total.Precip..in.,na.rm = TRUE)
    avg_air_temp = mean(county_month_data$Avg.Air.Temp..F.,na.rm = TRUE)
    avg_dew_point = mean(county_month_data$Avg.Dew.Point..F.,na.rm = TRUE)
    avg_wind_speed = mean(county_month_data$Avg.Wind.Speed..mph.,na.rm = TRUE)
    
    row = cbind(county,date,total_precip,avg_air_temp,avg_dew_point,avg_wind_speed)
    final_weather_df = rbind(final_weather_df,row)
  }
}

final_weather_df$date = rep(dates,58)
final_weather_df$time = rep(c(1:72),58)
final_weather_df$total_precip = as.numeric(final_weather_df$total_precip)
final_weather_df$avg_air_temp = as.numeric(final_weather_df$avg_air_temp)
final_weather_df$avg_dew_point = as.numeric(final_weather_df$avg_dew_point)
final_weather_df$avg_wind_speed = as.numeric(final_weather_df$avg_wind_speed)
```

# Influenza like illness (ILI) prevalence data

```{r}
influenza_data <- read_csv("~/work/causal-wildfire-smoke-2-duplicate/flu-ili-byregion-fluseason.csv")

inla_df <- inla_df %>%
  mutate(
    ili_region = case_when(

      # -------------------
      # Bay Area
      # -------------------
      NAME %in% c(
        "Alameda", "Contra Costa", "Marin", "Napa", "San Francisco",
        "San Mateo", "Santa Clara", "Solano", "Sonoma"
      ) ~ "Bay Area",

      # -------------------
      # Central
      # (Central Valley + Central Coast)
      # -------------------
      NAME %in% c(
        # Central Valley
        "Butte", "Colusa", "Fresno", "Glenn", "Kern", "Kings", "Madera",
        "Merced", "Placer", "Sacramento", "San Joaquin", "Shasta",
        "Stanislaus", "Sutter", "Tehama", "Tulare", "Yolo", "Yuba",
        # Central Coast
        "Monterey", "San Benito", "San Luis Obispo", "Santa Barbara",
        "Santa Cruz", "Ventura"
      ) ~ "Central",

      # -------------------
      # Lower Southern
      # (LA + SD area)
      # -------------------
      NAME %in% c(
        "Los Angeles", "Orange", "San Diego", "Imperial"
      ) ~ "Lower Southern",

      # -------------------
      # Upper Southern
      # (Inland Empire + nearby)
      # -------------------
      NAME %in% c(
        "Riverside", "San Bernardino"
      ) ~ "Upper Southern",

      # -------------------
      # Northern
      # (All remaining northern counties)
      # -------------------
      TRUE ~ "Northern"
    )
  )


ili_monthly_region <- influenza_data %>%
  mutate(
    weekending = mdy(weekending),          # if this yields NAs, switch to ymd(weekending)
    date = floor_date(weekending, "month"),
    region = trimws(region)
  ) %>%
  filter(
    !is.na(date),
    region != "California",
    year(date) >= 2014, year(date) <= 2019
  ) %>%
  group_by(region, date) %>%
  summarize(
    percent_ili = mean(Percent_ILI, na.rm = TRUE),
    .groups = "drop"
  )

# Create monthly time series for California (2014–2019)
california_monthly_ili <- influenza_data %>%
  mutate(
    weekending = mdy(weekending),          # ensure Date format
    year = year(weekending),
    month = month(weekending)
  ) %>%
  filter(
    region == "California",
    year >= 2014,
    year <= 2019
  ) %>%
  group_by(year, month) %>%
  summarize(
    percent_ili = mean(Percent_ILI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    date = unique(inla_df$date)
  ) %>%
  arrange(date)
```

```{r,eval=FALSE}
percent_ili_ts = ts(california_monthly_ili$percent_ili,start=1,frequency=12)

inla_df = inla_df %>% left_join(california_monthly_ili %>% select(date,percent_ili),by = "date")

test = inla_df %>% filter(NAME == "Lake")

ccf(test$monthly_total_pm25,percent_ili_ts,lag.max = 36)
ccf(test$deaths,percent_ili_ts,lag.max = 36)
```

# Set up dataset for INLA 

```{r}
test = monthly_total %>% left_join(total.mortality.ts_long, by = c("NAME" = "counties","date"))
test2 = test %>% left_join(final_EPA_data, by = c("NAME" = "county_labels","date"))
inla_df = test2 %>% left_join(final_weather_df, by = c("NAME" = "county","date","time"))
inla_df <- inla_df %>% left_join(ili_monthly_region,by = c("ili_region" = "region", "date" = "date"))

inla_df$month = rep(c(1:12),348)
inla_df$month = as.factor(inla_df$month)
inla_df$id2 = c(1:nrow(inla_df))

# Create an intercept column for each county
for (i in 1:58) {
  county = california_geoids$County[i]
  inla_df[[paste0("Intercept_", i)]] <- ifelse(inla_df$NAME == county,
                                                    1, NA)
}

#saveRDS(inla_df,"inla_df_3.7.rds")
```

```{r}
#Attach cluster labels and SDI scores 
inla_df = inla_df %>% left_join(clusterlabels,by = c("NAME" = "counties"))

# Assign column names to your yearly df
colnames(SDI_df) <- CA_counties
rownames(SDI_df) <- 2010:2019

# Convert wide to long: one row per year-county pair
df_long <- as.data.frame(SDI_df) %>%
  mutate(year = as.integer(rownames(.))) %>%
  pivot_longer(-year, names_to = "county", values_to = "value")

# Add all 12 months to each row
df_monthly_expanded <- df_long %>%
  crossing(month = 1:12) %>%
  arrange(county, year, month) %>% 
  filter(year >= 2014)

inla_df$SDI = df_monthly_expanded$value
inla_df$Cluster = factor(inla_df$Cluster)

inla_df$nonsmoke_pm25 = inla_df$pm25 - inla_df$monthly_total_pm25
inla_df$nonsmoke_pm25[inla_df$nonsmoke_pm25 < 0] = 0
```

## Create county groups based on SDI

SKATER clusters by SDI:

```{r}
clusterlabels$Cluster = factor(clusterlabels$Cluster)
clusterlabels$subregion = tolower(clusterlabels$counties)

merged_data <- merge(ca_map, clusterlabels, by = "subregion", all.x = TRUE)

# Define distinct colors for each category
#distinct_colors = qualitative_hcl(7,palette = "dynamic")
distinct_colors <- c("red", "blue", "green", "purple", "yellow")
# distinct_colors <- c("red", "blue")

#Plot
gg_with_grouping <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = Cluster), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  scale_color_viridis(discrete=TRUE) +
  theme(
      title = element_text(size = 16),
      legend.text  = element_text(size = 20),  # legend labels
      legend.title = element_text(size = 20)   # legend title
    )

gg_with_grouping
```

Simplify clusters to quintiles of SDI score instead of SKATER

```{r}
# Calculate quintile breakpoints
SDI_groups = quantile(inla_df$SDI, c(0.2, 0.4, 0.6, 0.8))

# Calculate average SDI per county
avg_SDI = apply(SDI_df, 2, mean)
avg_SDI_df = data.frame(counties, avg_SDI)

# Assign group labels based on quintiles
grouplabels = rep(NA, nrow(avg_SDI_df))  # Initialize vector

grouplabels[avg_SDI <= SDI_groups[1]] = 1
grouplabels[avg_SDI > SDI_groups[1] & avg_SDI <= SDI_groups[2]] = 2
grouplabels[avg_SDI > SDI_groups[2] & avg_SDI <= SDI_groups[3]] = 3
grouplabels[avg_SDI > SDI_groups[3] & avg_SDI <= SDI_groups[4]] = 4
grouplabels[avg_SDI > SDI_groups[4]] = 5

# Add to dataframe
avg_SDI_df$Group = grouplabels
group_labels = avg_SDI_df[,c(1,3)]

inla_df = inla_df %>% left_join(group_labels,by = c("NAME" = "counties"))
inla_df$Group = factor(inla_df$Group)
```

```{r}
#Attach group labels to each county
group_labels$Group = factor(group_labels$Group)
group_labels$subregion = tolower(group_labels$counties)

merged_data <- merge(ca_map, group_labels, by = "subregion", all.x = TRUE)

# Define distinct colors for each category
#distinct_colors = qualitative_hcl(7,palette = "dynamic")
distinct_colors <- c("red", "blue", "green", "purple", "yellow")
# distinct_colors <- c("red", "blue")

#Plot
gg_with_grouping <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = Group), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  scale_color_viridis(discrete=TRUE) +
  theme(
      title = element_text(size = 16),
      legend.text  = element_text(size = 20),  # legend labels
      legend.title = element_text(size = 20)   # legend title
    )

gg_with_grouping
```

# Exploratory Data Analysis 

## Time series plots of treatment, response, and confounders

Preview some time series plots for some random counties

```{r}
plot_counties = c('Fresno','Modoc','Humboldt','San Diego','Alameda')
idx = which(california_geoids$County %in% plot_counties)

# idx = sample(1:nrow(california_geoids),10)
plot_counties = california_geoids[idx,]

county_colors <- c(
  "Fresno"    = "#F57B63",  # red / coral
  "Modoc"     = "#A6A431",  # yellow-green
  "Humboldt"  = "#00996F",  # green-teal
  "San Diego" = "#2BA5B1",  # blue-teal
  "Alameda"   = "#C67ACD"   # purple
)
```

```{r}
for(i in plot_counties$GEOID){
  data = monthly_total %>% filter(GEOID == i)

  # Create the time series plot
  treatment.tsplot = ggplot(data, aes(x = date, y = monthly_total_pm25, 
  color = NAME)) +
    geom_line() +
    labs(title = "Time Series of smokePM_pred",
        x = "Date",
        y = "smokePM_pred",
        color = "County") +
    theme_minimal()

  print(treatment.tsplot)
}

# #Santa Barbara
# data = monthly_total %>% filter(GEOID == "06083")
# 
# # Create the time series plot
# treatment.tsplot = ggplot(data, aes(x = date, y = monthly_total_pm25, 
# color = NAME)) +
#   geom_line() +
#   labs(title = "Time Series of smokePM_pred",
#       x = "Date",
#       y = "smokePM_pred",
#       color = "County") +
#   theme_minimal()
# 
# print(treatment.tsplot)
```

Some spikes here and there but usually 0, so probably not normally distributed. Not stationary either 

```{r}
combined_data = monthly_total %>% filter(NAME %in% plot_counties$County) 

combined_data$NAME <- factor(combined_data$NAME,levels = plot_counties$County)

exposure_label = expression(paste("Monthly Wildfire Specific PM2.5 ("*mu*"g/m"^3*")"))

# Create the time series plot
p = ggplot(combined_data, aes(x = date, y = monthly_total_pm25, color = NAME)) +
  geom_line() + facet_wrap2(~NAME,scales = "free_y",axes="all") +
  labs(title = "Time Series of Wildfire Specific PM2.5",
       x = "Date",
       y = exposure_label,
       color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "exposure-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

The spikes tend to match up between different counties 

Weather variables: 

```{r}
#Precipitation
for(i in plot_counties$County){
  data = final_weather_df %>% filter(county == i)

  # Create the time series plot
  precip.tsplot = ggplot(data, aes(x = date, y = total_precip, 
  color = county)) +
    geom_line() +
    labs(title = "Time Series of Total Precipitation",
        x = "Date",
        y = "Total Precipitation (in)",
        color = "County") +
    theme_minimal()

  print(precip.tsplot)
}
```

Different ranges of values for different counties, all follow similar seasonal patterns. Relatively stationary 

```{r}
combined_data = final_weather_df %>% filter(county %in% plot_counties$County) 
combined_data$county <- factor(combined_data$county,levels = plot_counties$County)

# Create the time series plot
p = ggplot(combined_data, aes(x = date, y = total_precip, color = county)) +
  geom_line() + facet_wrap2(~county,scales = "free_y",axes="all") + 
  labs(title = "Time Series of Total Precipitation",
       x = "Date",
       y = "Monthly Total Precipitation (in)",
       color = "County") + 
  theme_minimal()

p

# ggsave(
#   filename = "precip-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

```{r}
#Temperature
for(i in plot_counties$County){
  data = final_weather_df %>% filter(county == i)

  # Create the time series plot
  temp.tsplot = ggplot(data, aes(x = date, y = avg_air_temp, 
  color = county)) +
    geom_line() +
    labs(title = "Time Series of Avg Temperature",
        x = "Date",
        y = "Avg Air Temperature (F)",
        color = "County") +
    theme_minimal()

  print(temp.tsplot)
}
```

Very stationary and seasonal, no surprise 

```{r}
combined_data = final_weather_df %>% filter(county %in% plot_counties$County) 
combined_data$county <- factor(combined_data$county,levels = plot_counties$County)

# Create the time series plot
p = ggplot(combined_data, aes(x = date, y = avg_air_temp, color = county)) +
  geom_line() + facet_wrap2(~county,scales = "free_y",axes="all") + 
  labs(title = "Time Series of Avg Temperature",
       x = "Date",
       y = "Monthly Average Air Temperature (F)",
       color = "County") +
  theme_minimal()

p

# ggsave(
#   filename = "temp-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

```{r}
#Humidity
for(i in plot_counties$County){
  data = final_weather_df %>% filter(county == i)

  # Create the time series plot
  humidity.tsplot = ggplot(data, aes(x = date, y = avg_dew_point, 
  color = county)) +
    geom_line() +
    labs(title = "Time Series of Avg Humidity",
        x = "Date",
        y = "Avg Dew Point (F)",
        color = "County") +
    theme_minimal()

  print(humidity.tsplot)
}
```

Very similar scales across counties. Same seasonal pattern as well. Higher dew point = more humid 

```{r}
combined_data = final_weather_df %>% filter(county %in% plot_counties$County) 
combined_data$county <- factor(combined_data$county,levels = plot_counties$County)

# Create the time series plot
p = ggplot(combined_data, aes(x = date, y = avg_dew_point, color = county)) +
  geom_line() + facet_wrap2(~county,scales = "free_y",axes="all") + 
  labs(title = "Time Series of Avg Humidity",
       x = "Date",
       y = "Monthly Average Dew Point (F)",
       color = "County") +
  theme_minimal()

p

# ggsave(
#   filename = "humidity-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

Big dip in Glenn in 2015, very unusual observation? Del Norte seems to be a little colder than the other counties

```{r}
#Wind
for(i in plot_counties$County){
  data = final_weather_df %>% filter(county == i)

  # Create the time series plot
  wind.tsplot = ggplot(data, aes(x = date, y = avg_wind_speed, 
  color = county)) +
    geom_line() +
    labs(title = "Time Series of Avg Wind",
        x = "Date",
        y = "Avg Wind Speed (mph)",
        color = "County") +
    theme_minimal()

  print(wind.tsplot)
}
```

All average wind speeds pretty much less than 10. Dip in fall and winter and then spike winter into spring 

```{r}
combined_data = final_weather_df %>% filter(county %in% plot_counties$County) 
combined_data$county <- factor(combined_data$county,levels = plot_counties$County)

# Create the time series plot
p = ggplot(combined_data, aes(x = date, y = avg_wind_speed, color = county)) +
  geom_line() + facet_wrap2(~county,scales = "free_y",axes="all") +
  labs(title = "Time Series of Avg Wind Speed",
       x = "Date",
       y = "Monthly Average Wind Speed (mph)",
       color = "County") +
  theme_minimal()

p

# ggsave(
#   filename = "wind-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

Similar patterns between counties but more messy than the other weather variables 

ILI prevalence: 

```{r}
ggplot(california_monthly_ili,
       aes(x = date, y = percent_ili)) +
  geom_line() +
  labs(
    x = "Date (2014-2019)",
    y = "Average Percent ILI",
    title = "California Monthly ILI Prevalence"
  ) +
  theme_bw()
```

```{r}
ggplot(inla_df,
       aes(x = date, y = percent_ili,color = ili_region)) +
  geom_line() + facet_wrap(~ili_region) + 
  labs(
    x = "Date (2014-2019)",
    y = "Average Percent ILI",
    title = "California Monthly ILI Prevalence"
  ) +
  theme_bw()
```

Shows some years have higher ILI prevalence than others - might confound our outcome model

```{r}
plot_data <- inla_df %>%
  filter(NAME %in% plot_counties$County) %>%
  mutate(NAME = factor(NAME, levels = plot_counties$County))

p = ggplot(plot_data, aes(x = date, y = percent_ili, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of ILI Prevalence",
    x = "Date",
    y = "Monthly ILI prevalence (%)",
    color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "iliprev-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```
Lead: 

```{r}
lead_label = expression(paste("Monthly Median Lead Levels ("*mu*"g/m"^3*")"))

p = ggplot(plot_data, aes(x = date, y = lead, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of Lead",
    x = "Date",
    y = lead_label,
    color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "lead-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

CO: 

```{r}
p = ggplot(plot_data, aes(x = date, y = co, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of CO",
    x = "Date",
    y = "Monthly Median CO Levels (ppm)",
    color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "co-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```
SO2: 

```{r}
p = ggplot(plot_data, aes(x = date, y = so2, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of SO2",
    x = "Date",
    y = "Monthly Median SO2 Levels (ppb)",
    color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "so2-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```
NO2: 

```{r}
p = ggplot(plot_data, aes(x = date, y = no2, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of NO2",
    x = "Date",
    y = "Monthly Median NO2 Levels (ppb)",
    color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "no2-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

O3: 

```{r}
p = ggplot(plot_data, aes(x = date, y = o3, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of O3",
    x = "Date",
    y = "Monthly Median O3 Levels (ppm)",
    color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "o3-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

PM10: 

```{r}
pm10_label = expression(paste("Monthly Median Ambient PM10 ("*mu*"g/m"^3*")"))

p = ggplot(plot_data, aes(x = date, y = pm10, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of PM10",
    x = "Date",
    y = pm10_label,
    color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "pm10-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

PM2.5: 

```{r}
pm25_label = expression(paste("Monthly Total Ambient PM2.5 ("*mu*"g/m"^3*")"))

p = ggplot(plot_data, aes(x = date, y = pm25, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of PM2.5",
    x = "Date",
    y = pm25_label,
    color = "County") +
  theme_minimal() 

p

ggsave(
  filename = "pm25-county-ts.png",
  plot = p,
  path = "~/work/important-plots",
  width = 8,
  height = 10,
  dpi = 600
)
```

Deaths: 

```{r}
p = ggplot(plot_data, aes(x = date, y = deaths, color = NAME, group = NAME)) +
  geom_line() +
  facet_wrap2(~NAME,scales = "free_y",axes="all") +
  #scale_color_manual(values = county_colors, drop = FALSE) +
  labs(
    title = "Time Series of Respiratory Related Deaths",
    x = "Date",
    y = "Monthly Respiratory Related Deaths",
    color = "County") +
  theme_minimal() 

p

# ggsave(
#   filename = "deaths-county-ts.png",
#   plot = p,
#   path = "~/work/important-plots",
#   width = 8,
#   height = 10,
#   dpi = 600
# )
```

## ACF/PACF plots for important variables: 

ACF measures the correlation between a time series and its lagged values (direct and indirect effects)

PACF measures the direct correlation between a time series and its lagged values after removing the effects of intermediate lags SO PACF is prob more useful in our case

### Check for stationarity

```{r}
county_df = inla_df %>% filter(NAME == "Santa Barbara")

#Mortality
plot.ts(county_df$death_rate)
transformed_outcome_ts = log(county_df$death_rate+0.01)
plot.ts(transformed_outcome_ts)

#Exposure (needs transformation)
plot.ts(county_df$monthly_total_pm25)

transformed_treatment_ts = log(county_df$monthly_total_pm25+0.01)
plot.ts(transformed_treatment_ts)

#Air pollutants
plot.ts(county_df$co)
plot.ts(county_df$no2)
plot.ts(county_df$so2)
plot.ts(county_df$o3)

#ILI prevalence (needs transformation)
plot.ts(county_df$percent_ili)
plot.ts(sqrt(county_df$percent_ili))

#Weather
plot.ts(county_df$total_precip)
plot.ts(county_df$avg_air_temp)
plot.ts(county_df$avg_dew_point)
plot.ts(county_df$avg_wind_speed)
```

Need to analyze one county at a time -> use SB as a case study (will try San Diego, Shasta, Alameda too)

```{r}
acf(transformed_outcome_ts,main = "ACFs for Respiratory Related Mortality")
pacf(transformed_outcome_ts,main = "PACFs for Respiratory Related Mortality")
```

Death: 

- ACF significant lags are 1,4,5,6,7,8,11,12,13,17,18
- PACF significant lags is 6,18,12

SO ARROWS BETWEEN Y's IS VALID 

```{r}
# acf(county_df$monthly_total_pm25)
# pacf(county_df$monthly_total_pm25)

acf(transformed_treatment_ts,main = "ACFs for Wildfire Specific PM2.5")
pacf(transformed_treatment_ts,main = "PACFs for Wildfire Specific PM2.5")
```

Wildfire PM2.5

- ACF: 1,6,7,11,12,13 
- PACF: 1,6,11

SO ARROWS BETWEEN A's IS VALID 

```{r}
acf(county_df$AQI,main = "ACFs for AQI")
pacf(county_df$AQI,main = "PACFs for AQI")
```

AQI

- ACF significant lags are 6,12
- PACF significant lags is 6 (12 almost significant)

```{r}
# acf(county_df$percent_ili,main = "ACFs for ILI Prevalence")
# pacf(county_df$percent_ili,main = "PACFs for ILI Prevalence")

acf(sqrt(county_df$percent_ili),main = "ACFs for ILI Prevalence")
pacf(sqrt(county_df$percent_ili),main = "PACFs for ILI Prevalence")
```

ILI prevalence

- ACF: 1,2,5,6,7,8,11,12,13,14,17,18
- PACF significant lags is 1,2,3,4

```{r}
acf(county_df$total_precip,main = "ACFs for Precipitation")
pacf(county_df$total_precip,main = "PACFs for Precipitation")
```

Precipitation

- ACF significant lags are 1,5,6,7 (10,12,18 are almost significant)
- PACF significant lags is 1

```{r}
acf(county_df$avg_air_temp,main = "ACFs for Temperature")
pacf(county_df$avg_air_temp,main = "PACFs for Temperature")
```

Temperature

- ACF significant lags are 1,2,4,5,6,7,8,10,11,12,13,14,16,17,18
- PACF significant lags is 1,2,3,4,6

```{r}
acf(county_df$avg_dew_point,main = "ACFs for Humidity")
pacf(county_df$avg_dew_point,main = "PACFs for Humidity")
```

Humidity

- ACF significant lags are 1,2,4,5,6,7,8,10,11,12,13,16,17,18
- PACF significant lags is 1,3,4

```{r}
acf(county_df$avg_wind_speed,main = "ACFs for Wind")
pacf(county_df$avg_wind_speed,main = "PACFs for Wind")
```

Wind

- ACF significant lags are 1,2,5,6,7,11,12,13
- PACF: 1,3,10

SEEMS LIKE CONFOUNDER VALUES DEPENDENT ON PREVIOUS VALUES


## Correlation plots (Spearman's, Cross Correlation, Granger causality):

Granger causality: is one time series useful in forecasting another time series (temporally related)

```{r}
#mortality rate + wildfire pm25
# ccf(county_df$death_rate,county_df$monthly_total_pm25,lag.max = 36)
# 
# cor(county_df$death_rate,county_df$monthly_total_pm25,method = "spearman")
# 
# lmtest::grangertest(county_df$death_rate~county_df$monthly_total_pm25,order = 12)

###with transformed treatment ts
ccf(transformed_outcome_ts,transformed_treatment_ts,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & Exposure")

cor(transformed_outcome_ts,transformed_treatment_ts,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~transformed_treatment_ts,order = 12)
```

Mortality rate + treatment

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.407
- Granger causality hypothesis test: not significant

```{r}
#wildfire pm25 + AQI
# ccf(county_df$monthly_total_pm25,county_df$AQI,lag.max = 36)
# 
# cor(county_df$monthly_total_pm25,county_df$AQI,method = "spearman")
# 
# lmtest::grangertest(county_df$AQI~county_df$monthly_total_pm25,order = 12)

###with transformed treatment ts
ccf(transformed_treatment_ts,county_df$AQI,lag.max = 36,main = "CCFs for Exposure & AQI")

cor(transformed_treatment_ts,county_df$AQI,method = "spearman")

lmtest::grangertest(county_df$AQI~transformed_treatment_ts,order = 12)
```

Exposure + AQI
- CCF: significant lags at -28,-17,-16,-10,-4,7,8,20,32

Interpretation: peak at -4 means that past values of treatment are correlated with AQI levels now, peak at 7 means that future values of treatment are correlated with past values of AQI

- Spearman's: 0.142
- Granger causality hypothesis test: Not significant

```{r}
#wildfire pm25 + CO
# ccf(county_df$monthly_total_pm25,county_df$co,lag.max = 36)
# 
# cor(county_df$monthly_total_pm25,county_df$co,method = "spearman")
# 
# lmtest::grangertest(county_df$co~county_df$monthly_total_pm25,order = 12)

###with transformed treatment ts
ccf(transformed_treatment_ts,county_df$co,lag.max = 36,main = "CCFs for Exposure & CO")

cor(transformed_treatment_ts,county_df$co,method = "spearman")

lmtest::grangertest(county_df$co~transformed_treatment_ts,order = 12)


#wildfire pm25 + NO2
# ccf(county_df$monthly_total_pm25,county_df$no2,lag.max = 36)
# 
# cor(county_df$monthly_total_pm25,county_df$no2,method = "spearman")
# 
# lmtest::grangertest(county_df$no2~county_df$monthly_total_pm25,order = 12)

###with transformed treatment ts
ccf(transformed_treatment_ts,county_df$no2,lag.max = 36,main = "CCFs for Exposure & NO2")

cor(transformed_treatment_ts,county_df$no2,method = "spearman")

lmtest::grangertest(county_df$no2~transformed_treatment_ts,order = 12)


#wildfire pm25 + SO2
# ccf(county_df$monthly_total_pm25,county_df$so2,lag.max = 36)
# 
# cor(county_df$monthly_total_pm25,county_df$so2,method = "spearman")
# 
# lmtest::grangertest(county_df$so2~county_df$monthly_total_pm25,order = 12)

###with transformed treatment ts
ccf(transformed_treatment_ts,county_df$so2,lag.max = 36,main = "CCFs for Exposure & SO2")

cor(transformed_treatment_ts,county_df$so2,method = "spearman")

lmtest::grangertest(county_df$so2~transformed_treatment_ts,order = 12)


#wildfire pm25 + O3
# ccf(county_df$monthly_total_pm25,county_df$o3,lag.max = 36)
# 
# cor(county_df$monthly_total_pm25,county_df$o3,method = "spearman")
# 
# lmtest::grangertest(county_df$o3~county_df$monthly_total_pm25,order = 12)

###with transformed treatment ts
ccf(transformed_treatment_ts,county_df$o3,lag.max = 36,main = "CCFs for Exposure & O3")

cor(transformed_treatment_ts,county_df$o3,method = "spearman")

lmtest::grangertest(county_df$o3~transformed_treatment_ts,order = 12)
```

Exposure + air pollutants: 

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: all weak correlation levels
- Granger causality hypothesis test: none significant 

```{r}
#wildfire pm25 + ILI prevalence

###with transformed treatment ts
ccf(transformed_treatment_ts,sqrt(county_df$percent_ili),lag.max = 36,main = "CCFs for Exposure & ILI Prevalence")

cor(transformed_treatment_ts,sqrt(county_df$percent_ili),method = "spearman")

lmtest::grangertest(sqrt(county_df$percent_ili)~transformed_treatment_ts,order = 12)
```

Exposure + ILI prevalence: 

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.354
- Granger causality hypothesis test: not significant 

```{r}
#mortality rate + AQI
ccf(transformed_outcome_ts,county_df$AQI,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & AQI")

cor(transformed_outcome_ts,county_df$AQI,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df$AQI,order = 12)
```

Mortality rate + AQI

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: 0.231
- Granger causality hypothesis test: Not significant

```{r}
#wildfire pm25 + precipitation
ccf(transformed_treatment_ts,county_df$total_precip,lag.max = 36,main = "CCFs for Exposure & Precipitation")

cor(transformed_treatment_ts,county_df$total_precip,method = "spearman")

lmtest::grangertest(transformed_treatment_ts~county_df$total_precip,order = 12)
```

Exposure + Precipitation

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.387
- Granger causality hypothesis test: not significant

```{r}
#wildfire pm25 + temp
ccf(transformed_treatment_ts,county_df$avg_air_temp,lag.max = 36,main = "CCFs for Exposure & Temperature")

cor(transformed_treatment_ts,county_df$avg_air_temp,method = "spearman")

lmtest::grangertest(transformed_treatment_ts~county_df$avg_air_temp,order = 12)
```

Exposure + Temperature

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: 0.413
- Granger causality hypothesis test: Not significant

```{r}
#wildfire pm25 + humidity
ccf(transformed_treatment_ts,county_df$avg_dew_point,lag.max = 36,main = "CCFs for Exposure & Humidity")

cor(transformed_treatment_ts,county_df$avg_dew_point,method = "spearman")

lmtest::grangertest(transformed_treatment_ts~county_df$avg_dew_point,order = 12)
```

Exposure + Humidity

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns 
- Spearman's: 0.243
- Granger causality hypothesis test: Not significant

```{r}
#wildfire pm25 + wind
ccf(transformed_treatment_ts,county_df$avg_wind_speed,lag.max = 36,main = "CCFs for Exposure & Wind")

cor(transformed_treatment_ts,county_df$avg_wind_speed,method = "spearman")

lmtest::grangertest(transformed_treatment_ts~county_df$avg_wind_speed,order = 12)
```

Exposure + Wind 

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.292
- Granger causality hypothesis test: Not significant 

```{r}
#temperature + wind
ccf(county_df$avg_air_temp,county_df$avg_wind_speed,lag.max = 36,main = "CCFs for Temperature & Wind")

cor(county_df$avg_air_temp,county_df$avg_wind_speed,method = "spearman")

lmtest::grangertest(county_df$avg_air_temp~county_df$avg_wind_speed,order = 12)
```

Temperature + Wind

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.076
- Granger causality hypothesis test: Not significant

```{r}
#humidity + wind
ccf(county_df$avg_dew_point,county_df$avg_wind_speed,lag.max = 36,main = "CCFs for Humidity & Wind")

cor(county_df$avg_dew_point,county_df$avg_wind_speed,method = "spearman")

lmtest::grangertest(county_df$avg_dew_point~county_df$avg_wind_speed,order = 12)
```

Humidity + Wind

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.065
- Granger causality hypothesis test: not significant

```{r}
#precipitation + humidity
ccf(county_df$total_precip,county_df$avg_dew_point,lag.max = 36,main = "CCFs for Precipitation & Humidity")

cor(county_df$total_precip,county_df$avg_dew_point,method = "spearman")

lmtest::grangertest(county_df$total_precip~county_df$avg_dew_point,order = 12)
```

Precipitation + Humidity: 

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.493
- Granger causality hypothesis test: Not significant

```{r}
#mortality rate + temperature
ccf(transformed_outcome_ts,county_df$avg_air_temp,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & Temperature")

cor(transformed_outcome_ts,county_df$avg_air_temp,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df$avg_air_temp,order = 12)
```

Mortality rate + Temperature

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.751
- Granger causality hypothesis test: Not significant

```{r}
#mortality rate + humidity
ccf(transformed_outcome_ts,county_df$avg_dew_point,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & Humidity")

cor(transformed_outcome_ts,county_df$avg_dew_point,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df$avg_dew_point,order = 12)
```

Mortality rate + Humidity

- CCF: clusters of significant negative and positive lags alternating, exhibiting seasonal patterns
- Spearman's: -0.663
- Granger causality hypothesis test: not significant


Mortality rate + other weather confounders had similar patterns 


```{r,eval=FALSE}
ccfvalues = ccf(transformed_outcome_ts,county_df$AQI,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & AQI")
ccfvalues
```

Not very strong correlations overall, moderate at best. BUT they are significant. I also ran the analysis above for San Bernandino, Alameda, and Alpine. Seems like we can say that past values of confounders affects current value of treatment and then that value also affects future values of confounders e.g. precipitation and wind. I think it is safe to say that there should be arrows between most confounders and then also between confounders and response.


## Grouped time series plot for important variables:

```{r}
colors <- c("pm25" = "blue", "monthly_total_pm25" = "red")

ggplot(county_df, aes(x= date)) + 
  geom_line(aes(y = monthly_total_pm25, color = "monthly_total_pm25")) +
  geom_line(aes(y = pm25, color = "pm25")) + 
  labs(x = "Year",
       y = "Value",
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```

```{r}
colors <- c("monthly_total_pm25" = "blue", "death_rate" = "red", "AQI" = "green")

ggplot(county_df, aes(x= date)) + 
  geom_line(aes(y = monthly_total_pm25, color = "monthly_total_pm25")) +
  geom_line(aes(y = death_rate, color = "death_rate")) + 
  geom_line(aes(y = AQI, color = "AQI")) + 
  labs(x = "Year",
       y = "Value",
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```


Mortality rate + wildfire pm25

```{r}
colors <- c("monthly_total_pm25" = "blue", "death_rate" = "red")

ggplot(county_df, aes(x= date)) + 
  geom_line(aes(y = monthly_total_pm25, color = "monthly_total_pm25")) +
  geom_line(aes(y = death_rate, color = "death_rate")) + 
  labs(x = "Year",
       y = "Value",
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```


wildfire pm25 + aqi

```{r}
colors <- c("monthly_total_pm25" = "blue", "AQI" = "green")

ggplot(county_df, aes(x= date)) + 
  geom_line(aes(y = monthly_total_pm25, color = "monthly_total_pm25")) +
  geom_line(aes(y = AQI, color = "AQI")) + 
  labs(x = "Year",
       y = "Value",
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```
Exposure + air pollutants

```{r}
colors <- c("monthly_total_pm25" = "red", "co" = "blue", "no2" = "green", "so2" = "pink", "o3" = "purple")

y_label = expression(paste("Ambient Air Concentration ("*mu*"g/m"^3*")"))

ggplot(county_df, aes(x= date)) + 
  geom_line(aes(y = monthly_total_pm25, color = "monthly_total_pm25")) +
  # geom_line(aes(y = co, color = "co")) + 
  geom_line(aes(y = no2, color = "no2")) + 
  # geom_line(aes(y = so2, color = "so2")) + 
  # geom_line(aes(y = o3, color = "o3")) + 
  labs(x = "Year",
       y = y_label,
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```

```{r}
ggplot(county_df, aes(x= date)) + 
  # geom_line(aes(y = monthly_total_pm25, color = "monthly_total_pm25")) +
  geom_line(aes(y = co, color = "co")) + 
  # geom_line(aes(y = no2, color = "no2")) + 
  geom_line(aes(y = so2, color = "so2")) + 
  geom_line(aes(y = o3, color = "o3")) + 
  labs(x = "Year",
       y = y_label,
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```

Exposure + ILI prevalence

```{r}
colors <- c("monthly_total_pm25" = "red", "percent_ili" = "blue")

ggplot(county_df, aes(x= date)) + 
  geom_line(aes(y = monthly_total_pm25, color = "monthly_total_pm25")) +
  geom_line(aes(y = percent_ili, color = "percent_ili")) + 
  labs(x = "Year",
       y = "Value",
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```

Mortality rate + ILI prevalence

```{r}
colors <- c("death_rate" = "red", "percent_ili" = "blue")

ggplot(county_df, aes(x= date)) + 
  geom_line(aes(y = death_rate, color = "death_rate")) +
  geom_line(aes(y = percent_ili, color = "percent_ili")) + 
  labs(x = "Year",
       y = "Value",
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```

Mortality rate + aqi

```{r}
colors <- c("death_rate" = "red", "AQI" = "green")

ggplot(county_df, aes(x= date)) + 
  geom_line(aes(y = death_rate, color = "death_rate")) + 
  geom_line(aes(y = AQI, color = "AQI")) + 
  labs(x = "Year",
       y = "Value",
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```


Wildfire pm25 + weather confounders

```{r}
colors <- c("monthly_total_pm25" = "black", "total_precip" = "blue", "avg_air_temp" = "red",
            "avg_dew_point" = "darkgreen", "avg_wind_speed" = "purple")

ggplot(county_df, aes(x = date)) + 
  geom_line(aes(y = monthly_total_pm25, color = "monthly_total_pm25")) +
  geom_line(aes(y = total_precip, color = "total_precip")) + 
  geom_line(aes(y = avg_air_temp, color = "avg_air_temp")) + 
  geom_line(aes(y = avg_dew_point, color = "avg_dew_point")) + 
  geom_line(aes(y = avg_wind_speed, color = "avg_wind_speed")) + 
  labs(x = "Year",
       y = "Value",
       color = "Legend",title = "Santa Barbara") +
  scale_color_manual(values = colors) + theme_minimal()
```

# Analysis of residuals (IMPORTANT)

## Use random component from time series decomposition and compare against residual

### Create new columns for residuals of important variables

```{r}
test = county_df %>% select(death_rate,monthly_total_pm25,
                    percent_ili,total_precip,avg_air_temp,
                    avg_dew_point,avg_wind_speed,AQI,
                    lead,co,so2,no2,o3,
                    pm10,nonsmoke_pm25) %>% st_drop_geometry()

test_ts = ts(test,start=1,frequency = 12)

ts_decomposed = lapply(test_ts,decompose)

county_df_resid = data.frame(county_df$date)

for(i in 1:15){
  # plot(ts_decomposed[[i]])
  resid_col = ts_decomposed[[i]]$random
  county_df_resid = cbind(county_df_resid,resid_col)
}

###Could add linear trend in but prob not good to add otherwise correlations will pick it up
# for(i in 1:13){
#   # plot(ts_decomposed[[i]])
#   resid_col = ts_decomposed[[i]]$random + ts_decomposed[[i]]$trend
#   county_df_resid = cbind(county_df_resid,resid_col)
# }

colnames(county_df_resid) = c("date","death_rate","monthly_total_pm25",
                          "percent_ili","total_precip","avg_air_temp",
                          "avg_dew_point","avg_wind_speed",
                          "AQI","lead","co","so2","no2","o3",
                          "pm10","nonsmoke_pm25")

county_df_resid = na.omit(county_df_resid)
```

### Check for stationarity

```{r}
#Mortality (needs transformation)
plot.ts(county_df_resid$death_rate)

transformed_outcome_ts = sqrt(county_df_resid$death_rate+3)
plot.ts(transformed_outcome_ts)

#Exposure (needs transformation)
plot.ts(county_df_resid$monthly_total_pm25)

minval = min(county_df_resid$monthly_total_pm25)

transformed_treatment_ts = sqrt(county_df_resid$monthly_total_pm25-minval)
plot.ts(transformed_treatment_ts)

#Air pollutants
plot.ts(county_df_resid$co)
plot.ts(county_df_resid$no2)
plot.ts(county_df_resid$so2)
plot.ts(county_df_resid$o3)

#ILI prevalence (needs transformation)
plot.ts(county_df_resid$percent_ili)

transformed_ili_prev_ts = sqrt(county_df_resid$percent_ili + 2)
plot.ts(transformed_ili_prev_ts)

#Weather
plot.ts(county_df_resid$total_precip)
plot.ts(county_df_resid$avg_air_temp)
plot.ts(county_df_resid$avg_dew_point)
plot.ts(county_df_resid$avg_wind_speed)
```


### ACF/PACFs

```{r}
# acf(county_df_resid$death_rate)
# pacf(county_df_resid$death_rate)

acf(transformed_outcome_ts)
pacf(transformed_outcome_ts)
```

Resid_death: PACF has significant lags at 3 and 4 (-3ish) why negative? No significant ACF lags

```{r}
# acf(county_df_resid$monthly_total_pm25)
# pacf(county_df_resid$monthly_total_pm25)

acf(transformed_treatment_ts)
pacf(transformed_treatment_ts)
```

Resid_treatment: ACF has significant lag at 12 (-0.25), PACF has significant lag at 12 (-0.3)

```{r}
acf(county_df_resid$AQI)
pacf(county_df_resid$AQI)
```

Resid_aqi: PACF has significant lags at 12 (-0.27ish) why negative? No significant ACF lags

```{r}
# acf(county_df_resid$percent_ili)
# pacf(county_df_resid$percent_ili)

acf(transformed_ili_prev_ts)
pacf(transformed_ili_prev_ts)
```

Resid_ili_prevalence: ACF has significant lag at 1,5 (0.3,-0.25), PACF has significant lag at 1,11 (0.33,-0.3)

```{r}
acf(county_df_resid$total_precip)
pacf(county_df_resid$total_precip)
```

Resid_precip: ACF has significant lag at 12 (-0.25), PACF has significant lag at 5 and 12 (-0.27ish)

```{r}
acf(county_df_resid$avg_air_temp)
pacf(county_df_resid$avg_air_temp)
```

Resid_temp: ACF has significant lag at 3 (-0.4), PACF has significant lag at 2,3, and 5 (all -0.3ish)

```{r}
acf(county_df_resid$avg_dew_point)
pacf(county_df_resid$avg_dew_point)
```

Resid_dew: ACF has significant lag at 12 (-0.25), PACF has significant lag at 12 (-0.27)

```{r}
acf(county_df_resid$avg_wind_speed)
pacf(county_df_resid$avg_wind_speed)
```

Resid_wind: ACF has significant lags at 4 and 5 (-0.4) and 18 (0.35), PACF has significant lags as 12 and 14 (-0.3)

```{r}
acf(county_df_resid$co)
pacf(county_df_resid$co)
```

Resid_co: no significant lags 

```{r}
acf(county_df_resid$no2)
pacf(county_df_resid$no2)
```

Resid_no2: no significant lags (maybe 14 for ACF and 1 for PACF)

```{r}
acf(county_df_resid$so2)
pacf(county_df_resid$so2)
```

Resid_so2: significant lags for ACF at 3,4,5 (-0.3) and 9 (0.33) and for PACF 3,5,17

```{r}
acf(county_df_resid$o3)
pacf(county_df_resid$o3)
```

Resid_o3: significant lags for ACF at 4 (-0.4) and for PACF 2,4,14 (-0.29,-0.4,-0.4)

Negative autocorrelations: current values are negative correlated with past values - year to year variation? 


### Correlation plots (Cross Correlation, Spearman's, Granger causality):

Significant negative lags in CCF: lags of first TS are correlated with present value of second TS 

Significant positive lags in CCF: lags of second TS are correlated with present value of first TS

Granger causality: is one time series useful in forecasting another time series (temporally related)

```{r}
save_ccf <- function(x, y, file, lag.max = 36, width = 6, height = 3, res = 600, title){
  png(file, width = width, height = height, units = "in", res = res)
  ccf(x, y, lag.max = 36,main = title)
  dev.off()
}
```


```{r}
#mortality rate + wildfire pm25
ccf(transformed_outcome_ts,transformed_treatment_ts,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & Exposure")

save_ccf(
  x = transformed_outcome_ts,
  y = transformed_treatment_ts,
  file = "/home/jovyan/work/important-plots/ccf_outcome_treatment.png",
  title = "CCFs for Respiratory Related Mortality Rate & Exposure"
)

cor(transformed_outcome_ts,transformed_treatment_ts,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~transformed_treatment_ts,order = 12)
```

Mortality rate + treatment

- CCF: Only significant peak is at +1 (0.6) which means that mortality rates one month in the future is positively correlated with wildfire pm2.5 levels now
- Spearman's: 0.164
- Granger causality hypothesis test: Almost significant but no

```{r}
#wildfire pm25 + AQI
ccf(transformed_treatment_ts,county_df_resid$AQI,lag.max = 36,main = "CCFs for Exposure & AQI")

cor(transformed_treatment_ts,county_df_resid$AQI,method = "spearman")

lmtest::grangertest(county_df_resid$AQI~transformed_treatment_ts,order = 12)
```

Exposure + AQI

- CCF: Significant peaks at 0,1,35. Positively correlated at 0,35. Negatively correlated at 1 - How do we interpret negative correlation at lag +1? 
- Spearman's: 0.397
- Granger causality hypothesis test: significant! 

```{r}
#wildfire pm25 + CO
# ccf(county_df_resid$monthly_total_pm25,county_df_resid$co,lag.max = 36)
# 
# cor(county_df_resid$monthly_total_pm25,county_df_resid$co,method = "spearman")
# 
# lmtest::grangertest(county_df_resid$co~county_df_resid$monthly_total_pm25,order = 12)

#with transformed treatment
ccf(transformed_treatment_ts,county_df_resid$co,lag.max = 36,main = "CCFs for Exposure & CO")

save_ccf(
  x = transformed_treatment_ts,
  y = county_df_resid$co,
  file = "/home/jovyan/work/important-plots/ccf_treatment_co.png",
  title = "CCFs for Exposure & CO"
)

cor(transformed_treatment_ts,county_df_resid$co,method = "spearman")

lmtest::grangertest(county_df_resid$co~transformed_treatment_ts,order = 12)
```

Exposure + CO 

- CCF: Significant peaks at -11,0,1,35. Positively correlated at -11, 0, and 35. Negatively correlated at 1 - how do we interpret negative correlation at lag 1? 
- Spearman's: 0.427
- Granger causality hypothesis test: significant! 

```{r}
#wildfire pm25 + NO2
# ccf(county_df_resid$monthly_total_pm25,county_df_resid$no2,lag.max = 36)
# 
# cor(county_df_resid$monthly_total_pm25,county_df_resid$no2,method = "spearman")
# 
# lmtest::grangertest(county_df_resid$no2~county_df_resid$monthly_total_pm25,order = 12)

#with transformed treatment
ccf(transformed_treatment_ts,county_df_resid$no2,lag.max = 36,main = "CCFs for Exposure & NO2")

cor(transformed_treatment_ts,county_df_resid$no2,method = "spearman")

lmtest::grangertest(county_df_resid$no2~transformed_treatment_ts,order = 12)
```

Exposure + NO2

- CCF: Significant peaks at -12,-11,0,1,35. Positively correlated at -11, 0, and 35. Negatively correlated at -12 and +1.  
- Spearman's: 0.441
- Granger causality hypothesis test: not significant 

```{r}
#wildfire pm25 + SO2
# ccf(county_df_resid$monthly_total_pm25,county_df_resid$so2,lag.max = 36)
# 
# cor(county_df_resid$monthly_total_pm25,county_df_resid$so2,method = "spearman")
# 
# lmtest::grangertest(county_df_resid$so2~county_df_resid$monthly_total_pm25,order = 12)

#with transformed treatment
ccf(transformed_treatment_ts,county_df_resid$so2,lag.max = 36,main = "CCFs for Exposure & SO2")

cor(transformed_treatment_ts,county_df_resid$so2,method = "spearman")

lmtest::grangertest(county_df_resid$so2~transformed_treatment_ts,order = 12)
```

Exposure + SO2

- CCF: Significant peaks at -12,-7,0,21,35. Positively correlated at -7, 0, and 35. Negatively correlated at -12, and 21.  
- Spearman's: 0.304
- Granger causality hypothesis test: significant! 

```{r}
#wildfire pm25 + O3
# ccf(county_df_resid$monthly_total_pm25,county_df_resid$o3,lag.max = 36)
# 
# cor(county_df_resid$monthly_total_pm25,county_df_resid$o3,method = "spearman")
# 
# lmtest::grangertest(county_df_resid$o3~county_df_resid$monthly_total_pm25,order = 12)

#with transformed treatment
ccf(transformed_treatment_ts,county_df_resid$o3,lag.max = 36,main = "CCFs for Exposure & O3")

cor(transformed_treatment_ts,county_df_resid$o3,method = "spearman")

lmtest::grangertest(county_df_resid$o3~transformed_treatment_ts,order = 12)
```

Exposure + O3

- CCF: Significant peaks at 6,10,11,35. Positively correlated at 10 and 11. Negatively correlated at 6,35.  
- Spearman's: -0.056
- Granger causality hypothesis test: significant! 

```{r}
#mortality rate + AQI
ccf(transformed_outcome_ts,county_df_resid$AQI,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & AQI")

cor(transformed_outcome_ts,county_df_resid$AQI,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df_resid$AQI,order = 12)
```

Mortality rate + AQI

- CCF: Significant peaks at +12 (-0.3) and 36 (0.3)? Death rate one year in the future is negatively correlated with current AQI? 
- Spearman's: 0.069
- Granger causality hypothesis test: significant! 

```{r}
#wildfire pm25 + ILI prevalence

###with transformed treatment ts
# ccf(transformed_treatment_ts,county_df_resid$percent_ili,lag.max = 36, main = "CCFs for Exposure & ILI Prevalence")
# 
# cor(transformed_treatment_ts,county_df_resid$percent_ili,method = "spearman")
# 
# lmtest::grangertest(county_df_resid$percent_ili~transformed_treatment_ts,order = 12)


ccf(transformed_treatment_ts,transformed_ili_prev_ts,lag.max = 36, main = "CCFs for Exposure & ILI Prevalence")

save_ccf(
  x = transformed_treatment_ts,
  y = transformed_ili_prev_ts,
  file = "/home/jovyan/work/important-plots/ccf_treatment_ili.png",
  title = "CCFs for Exposure & ILI Prevalence"
)

cor(transformed_treatment_ts,transformed_ili_prev_ts,method = "spearman")

lmtest::grangertest(transformed_ili_prev_ts~transformed_treatment_ts,order = 12)
```

Exposure + ILI prevalence: 

- CCF: Significant peak at -1 (0.45)
- Spearman's: 0.062
- Granger causality hypothesis test: not significant 

```{r}
#wildfire pm25 + precipitation
ccf(transformed_treatment_ts,county_df_resid$total_precip,lag.max = 36,main = "CCFs for Exposure & Precipitation")

save_ccf(
  x = transformed_treatment_ts,
  y = county_df_resid$total_precip,
  file = "/home/jovyan/work/important-plots/ccf_treatment_precip.png",
  title = "CCFs for Exposure & Precipitation"
)

cor(transformed_treatment_ts,county_df_resid$total_precip,method = "spearman")

lmtest::grangertest(transformed_treatment_ts~county_df_resid$total_precip,order = 12)
```

Exposure + Precipitation

- CCF: Significant peaks at -3,9,11,21,36. Positively correlated at -3,11,36 and negatively correlated at +9,35
- Spearman's: -0.111
- Granger causality hypothesis test: not significant

```{r}
#wildfire pm25 + temp
ccf(transformed_treatment_ts,county_df_resid$avg_air_temp,lag.max = 36,main = "CCFs for Exposure & Temperature")

cor(transformed_treatment_ts,county_df_resid$avg_air_temp,method = "spearman")

lmtest::grangertest(transformed_treatment_ts~county_df_resid$avg_air_temp,order = 12)
```

Exposure + Temperature

- CCF: no significant lags
- Spearman's: 0.141
- Granger causality hypothesis test: not significant

```{r}
#wildfire pm25 + humidity
ccf(transformed_treatment_ts,county_df_resid$avg_dew_point,lag.max = 36,main = "CCFs for Exposure & Humidity")

cor(transformed_treatment_ts,county_df_resid$avg_dew_point,method = "spearman")

lmtest::grangertest(transformed_treatment_ts~county_df_resid$avg_dew_point,order = 12)
```

Exposure + Humidity

- CCF: Significant peaks at 0,1. Notable negative correlation at lag 0 (-0.45) which makes sense, because higher dew point means more humid. But for lag 1, past values of humidity are positively correlated with current values of treatment 
- Spearman's: -0.314
- Granger causality hypothesis test: not significant

```{r}
#wildfire pm25 + wind
ccf(transformed_treatment_ts,county_df_resid$avg_wind_speed,lag.max = 36,main = "CCFs for Exposure & Wind")

cor(transformed_treatment_ts,county_df_resid$avg_wind_speed,method = "spearman")

lmtest::grangertest(transformed_treatment_ts~county_df_resid$avg_wind_speed,order = 12)
```

Exposure + Wind 

- CCF: Significant peaks at -4 and +1. Notable negative correlations at lag +1 (-0.4). Positive correlation at -4 (0.28). Negative correlations make sense bc higher wind speeds would probably lead to smaller treatment value  
- Spearman's: -0.397
- Granger causality hypothesis test: not significant

```{r}
#temperature + wind
ccf(county_df_resid$avg_air_temp,county_df_resid$avg_wind_speed,lag.max = 36,main = "CCFs for Temperature & Wind")

cor(county_df_resid$avg_air_temp,county_df_resid$avg_wind_speed,method = "spearman")

lmtest::grangertest(county_df_resid$avg_air_temp~county_df_resid$avg_wind_speed,order = 12)
```

Temperature + Wind

- CCF: Significant peaks at 0,+1,3,4. Lags 0 and 1 have negative correlations (-0.45 and -0.3) and lags 3 and 4 have positive correlations (0.4 and 0.2)
- Spearman's: -0.432
- Granger causality hypothesis test: not significant

```{r}
#humidity + wind
ccf(county_df_resid$avg_dew_point,county_df_resid$avg_wind_speed,lag.max = 36,main = "CCFs for Humidity & Wind")

cor(county_df_resid$avg_dew_point,county_df_resid$avg_wind_speed,method = "spearman")

lmtest::grangertest(county_df_resid$avg_dew_point~county_df_resid$avg_wind_speed,order = 12)
```

Humidity + Wind

- CCF: Significant peaks at 0,+2. Lags 0 has a negative correlation of -0.27 and lag 2 has a positive correlation of 0.27
- Spearman's: -0.202
- Granger causality hypothesis test: not significant

```{r}
#precipitation + humidity
ccf(county_df_resid$total_precip,county_df_resid$avg_dew_point,lag.max = 36,main = "CCFs for Precipitation & Humidity")

cor(county_df_resid$total_precip,county_df_resid$avg_dew_point,method = "spearman")

lmtest::grangertest(county_df_resid$total_precip~county_df_resid$avg_dew_point,order = 12)
```

Precipitation + Humidity: 

- CCF: Many significant peaks (-35,-2,+3,16,28), lags -35,-2, and +16 are positive correlations, lags 3 and +28 have negative correlations
- Spearman's: 0.108
- Granger causality hypothesis test: not significant

```{r}
#mortality rate + temperature
ccf(transformed_outcome_ts,county_df_resid$avg_air_temp,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & Temperature")

cor(transformed_outcome_ts,county_df_resid$avg_air_temp,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df_resid$avg_air_temp,order = 12)
```

Mortality rate + Temperature

- CCF: Significant peaks at -2,1,25. Lags -2,25 have negative correlations of about -0.25 and lag 1 has a positive correlation of 0.3
- Spearman's: 0.210
- Granger causality hypothesis test: not significant

```{r}
#mortality rate + humidity
ccf(transformed_outcome_ts,county_df_resid$avg_dew_point,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & Humidity")

cor(transformed_outcome_ts,county_df_resid$avg_dew_point,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df_resid$avg_dew_point,order = 12)
```

Mortality rate + Humidity

- CCF: Significant peaks at 2,3, and 25. Lags 3,25 have negative correlations of -0.3 and lag 2 has a positive correlation of 0.33
- Spearman's: 0.121
- Granger causality hypothesis test: significant!

```{r}
#mortality rate + precipitation
ccf(transformed_outcome_ts,county_df_resid$total_precip,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & Precipitation")

save_ccf(
  x = transformed_outcome_ts,
  y = county_df_resid$total_precip,
  file = "/home/jovyan/work/important-plots/ccf_outcome_precip.png",
  title = "CCFs for Respiratory Related Mortality Rate & Precipitation"
)

cor(transformed_outcome_ts,county_df_resid$total_precip,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df_resid$total_precip,order = 12)
```

Mortality rate + Precipitation

- CCF: Significant peaks at 11 (0.35)
- Spearman's: -0.112
- Granger causality hypothesis test: not significant

```{r}
#mortality rate + wind
ccf(transformed_outcome_ts,county_df_resid$avg_wind_speed,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & Wind")

cor(transformed_outcome_ts,county_df_resid$avg_wind_speed,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df_resid$avg_wind_speed,order = 12)
```

Mortality rate + Wind

- CCF: Significant lags -13 and +1,2 have correlations of 0.4, -0.35, and -0.36 respectively
- Spearman's: -0.172
- Granger causality hypothesis test: significant!

```{r}
#mortality rate + co
ccf(transformed_outcome_ts,county_df_resid$co,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & CO")

save_ccf(
  x = transformed_outcome_ts,
  y = county_df_resid$co,
  file = "/home/jovyan/work/important-plots/ccf_outcome_co.png",
  title = "CCFs for Respiratory Related Mortality Rate & CO"
)

cor(transformed_outcome_ts,county_df_resid$co,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~county_df_resid$co,order = 12)
```

Mortality rate + CO

- CCF: Almost significant lags -10,3 have correlations of 0.3
- Spearman's: basically 0
- Granger causality hypothesis test: significant!

```{r}
#mortality rate + ILI prevalence
ccf(transformed_outcome_ts,transformed_ili_prev_ts,lag.max = 36,main = "CCFs for Respiratory Related Mortality Rate & ILI Prevalence")

save_ccf(
  x = transformed_outcome_ts,
  y = transformed_ili_prev_ts,
  file = "/home/jovyan/work/important-plots/ccf_outcome_ili.png",
  title = "CCFs for Respiratory Related Mortality Rate & ILI Prevalence"
)

cor(transformed_outcome_ts,transformed_ili_prev_ts,method = "spearman")

lmtest::grangertest(transformed_outcome_ts~transformed_ili_prev_ts,order = 12)
```

Mortality rate + ILI prevalence

- CCF: Significant lags -4,-1,0 have correlations of -0.3, -0.31, and 0.4 respectively
- Spearman's: 0.157
- Granger causality hypothesis test: Not significant


ARE THE RESULTS SIMILAR BETWEEN COUNTIES? OR ARE THERE CLUSTERS 

- seems like results are more or less similar... there are significant autocorrelations most of the time
- seems like patterns are more geographic based than SKATER (socioeconomic) based 

### Function to calculate CCF for each county and create a heatmap

```{r}
spatial_ccf_heatmap <- function(var1 = 'death_rate', var2 = 'monthly_total_pm25', lag = 0) {
  county_ccfs <- data.frame(county = california_geoids$County, ccf = NA)
  
  for (i in 1:58) {
    df = inla_df %>% filter(NAME == california_geoids$County[i]) %>% select(var1, var2) %>%
      st_drop_geometry() %>% na.omit()
    
    ccf_vals = ccf(df[[var1]], df[[var2]], plot = FALSE, lag.max = 36)
    county_ccfs$ccf[i] = ccf_vals$acf[lag + ccf_vals$lag == 0]
  }
  
  # Merge county_ccfs with county_map for plotting
  county_ccfs$county = tolower(county_ccfs$county)
  county_map_ccfs = merge(county_map, county_ccfs, by.x = "subregion", by.y = "county")
  
  # Create the heatmap
  ggplot() + 
    geom_polygon(data = county_map_ccfs, aes(x = long, y = lat, group = group, fill = ccf), 
                 color = "black") +
    coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
    theme_void() +
    labs(title = paste("CCF between", var1, "and", var2, "at lag", lag),
         fill = "CCF") +
    scale_fill_gradient(low = "yellow", high = "red")
}

# Example usage
spatial_ccf_heatmap(var1 = 'death_rate', var2 = 'total_precip', lag = 12)
```


