---
title: "GE-IPW-Simulation"
author: "Jeffrey Wu"
format: html
---

# Generate synthetic data

```{r}
library(INLA)
library(dplyr)
library(truncnorm)
```


```{r}
# Set random seed for reproducibility
set.seed(42)

# Parameters
n_individuals <- 50           # Number of individuals
n_timepoints <- 14              # Number of time points
baseline_covariate_effects <- seq(0.2,1,length.out = 5) # Effect of baseline covariate on treatment and outcome

#Assumes same treatment effect at each time point 
true_treatment_linear <- 0.3     # Linear component of treatment effect on outcome
true_treatment_quadratic <- -0.05 # Quadratic component for nonlinear effect

# Define 5 unique baseline covariate values
unique_baseline_values <- seq(1, 5, length.out = 5)

# Initialize empty list to store data
data_records <- list()

# Generate data
for (i in 1:n_individuals) {
  baseline_covariate <- sample(unique_baseline_values, 1) # Draw from 5 unique values
  baseline_covariate_effect <- baseline_covariate_effects[baseline_covariate]
  
  # Initial values for three time-varying confounders
  confounder_1 <- rnorm(1)
  confounder_2 <- rnorm(1)
  confounder_3 <- rnorm(1)
  outcome_prev <- 0
  treatment_prev <- 0 # Initialize previous treatment value
  
  for (t in 1:n_timepoints) {
    
    # Treatment as a function of baseline covariate and time-varying confounders
    treatment <- 0.5 * treatment_prev + baseline_covariate_effect +
                 0.2 * confounder_1 + 0.3 * confounder_2 + 0.4 * confounder_3 + rnorm(1)
    treatment <- ifelse(treatment < 0, 0, treatment)

    
    # Calculate the mean outcome for exponential distribution based on treatment and covariates
    outcome <- true_treatment_linear * treatment + 
                      true_treatment_quadratic * treatment^2 +
                      0.6 * treatment_prev + baseline_covariate_effect -
                      0.2 * confounder_1 + 0.3 * confounder_2 + 0.4 * confounder_3 + 
                      0.3 * outcome_prev
    
    # Generate outcome from truncated normal distribution
    # outcome <- rtruncnorm(1, a=0, b=Inf, mean=outcome_mean, sd=1)
    
    # Append record
    data_records[[length(data_records) + 1]] <- data.frame(
      id = i,
      time = t,
      baseline_covariate = baseline_covariate,
      # baseline_covariate_effect = baseline_covariate_effect,
      confounder_1 = confounder_1,
      confounder_2 = confounder_2,
      confounder_3 = confounder_3,
      treatment = treatment,
      treatment_prev = treatment_prev,
      outcome = outcome,
      outcome_prev = outcome_prev
    )
    
    # Update confounders and outcome_prev for next timepoint
    confounder_1 <- 0.6 * treatment + 0.4 * outcome + rnorm(1)
    confounder_2 <- 0.5 * treatment + 0.3 * outcome + rnorm(1)
    confounder_3 <- 0.4 * treatment + 0.2 * outcome + rnorm(1)
    outcome_prev <- outcome
    treatment_prev <- treatment # Update previous treatment for the next iteration
  }
}

# Combine list into a single data frame
simulated_data <- do.call(rbind, data_records)
simulated_data$baseline_covariate = factor(simulated_data$baseline_covariate)

# Display the first few rows of the simulated dataset
head(simulated_data)
```

# Plot outcome, treatment, and confounders

```{r,eval=FALSE}
subj1_outcome_ts = synthetic_df %>% filter(id == 1) %>% select(outcome) %>% ts()

plot.ts(diff(subj1_outcome_ts))
acf(diff(subj1_outcome_ts))
pacf(diff(subj1_outcome_ts))

subj1_treatment_ts = synthetic_df %>% filter(id == 1) %>% select(treatment) %>% ts()

plot.ts(diff(subj1_treatment_ts))
acf(diff(subj1_treatment_ts))
pacf(diff(subj1_treatment_ts))

subj1_confounder1_ts = synthetic_df %>% filter(id == 1) %>% select(confounder_1) %>% ts()

plot.ts(diff(subj1_confounder1_ts))
acf(diff(subj1_confounder1_ts))
pacf(diff(subj1_confounder1_ts))
```

Make outcome, treatment, and confounder variables stationary: 

```{r,eval=FALSE}
diff_ts_list = list()

for (i in 1:max(simulated_data$id)){
  subj_ts = simulated_data %>% filter(id == i) %>% ts()
  subj_baseline_covariate = unique(subj_ts[,3])
  diff_ts = diff(subj_ts[,4:10])
  
  subj_df_new = cbind(rep(i,nrow(diff_ts)),c(1:nrow(diff_ts)),
                      rep(subj_baseline_covariate,nrow(diff_ts)),diff_ts)
  diff_ts_list[[i]] = data.frame(subj_df_new)
}

diff_simulated_data = do.call(rbind,diff_ts_list)
colnames(diff_simulated_data) = colnames(simulated_data)
```


Attach lagged version of each confounder: 

```{r}
#use diff_simulated_data if wanna test
synthetic_df <- simulated_data %>%
  group_by(id) %>%
  mutate(
    confounder_1_lag2 = lag(confounder_1, 2),
    confounder_1_lag4 = lag(confounder_1, 4),
    confounder_1_lag6 = lag(confounder_1, 6),
    confounder_2_lag2 = lag(confounder_2, 2),
    confounder_2_lag4 = lag(confounder_2, 4),
    confounder_2_lag6 = lag(confounder_2, 6),
    confounder_3_lag2 = lag(confounder_3, 2),
    confounder_3_lag4 = lag(confounder_3, 4),
    confounder_3_lag6 = lag(confounder_3, 6)
  ) %>%
  ungroup()

head(synthetic_df)
```

# Naive approach

Just fit a regression model and interpret the treatment coefficient 

```{r}
naive_model = lm(outcome ~ baseline_covariate + confounder_1 + confounder_2 + 
                   confounder_3 + treatment + I(treatment^2) + outcome_prev, data = synthetic_df)

summary(naive_model)
```

# MSMs with IPW for initial estimate

Step 1: Model propensity score for exposure (expected value for continuous exposure) -> create stablized IPWs 

```{r}
ps_df = matrix(ncol=2)

for (i in 1:n_individuals){
  subject_data = synthetic_df %>% filter(id == i)
  ps_subject = cbind(subject_data$treatment,NA)
    
  ps_lm_num = lm(subject_data$treatment ~ subject_data$treatment_prev, 
                 data = subject_data)
  ps_lm_denom = lm(subject_data$treatment ~ subject_data$treatment_prev + 
                     confounder_1 + confounder_2 + confounder_3, data = subject_data)
  
  ###Situation where we cannot observe confounders 2+3
  # ps_lm_denom = lm(subject_data$treatment ~ subject_data$treatment_prev + 
  #                    confounder_1, data = subject_data)
  
  preds_num = predict(ps_lm_num,newdata = subject_data)
  preds_denom = predict(ps_lm_denom,newdata = subject_data)
  
  ps_subject[,2] = preds_num/preds_denom
  ps_df = rbind(ps_df,ps_subject)
}

ps_df = data.frame(ps_df[-1,])
colnames(ps_df) = c("exposure","stabilized_IPW")
```

Step 2: Model outcome with covariates and propensity score

```{r}
stabilized_IPW = ps_df$stabilized_IPW
synthetic_df = cbind(synthetic_df,stabilized_IPW)

#Robins and Brumback say to not include time varying covariates but Vansteelandt does? And including them leads to correct estimate
msm_model = lm(outcome ~ baseline_covariate + stabilized_IPW + treatment +
                    I(treatment^2) + treatment_prev + confounder_1 + 
                    confounder_2 + confounder_3 + outcome_prev, data = synthetic_df)
summary(msm_model)

###Situation where we cannot observe confounders 2+3
# msm_model2 = lm(outcome ~ baseline_covariate + stabilized_IPW + treatment +
#                     I(treatment^2) + confounder_1 + treatment_prev + outcome_prev, data = synthetic_df)
# summary(msm_model2)

causal_params = coef(msm_model)[7:8]
```

# G estimation 

Step 3: Create H terms from Y and causal parameters for each individual and time point 

```{r}
### If I choose biased initial guess, g-estimation corrects it (initial guess can't be too off)
#causal_params = c(0.1,0.1) 
Hst = rep(NA,3)

for(s in 2:n_timepoints){
  Ys = synthetic_df %>% filter(time == s) %>% select(outcome)
  post_t_effect = 0
  
for(t in 1:(s-1)){
    post_t_effect = 0
    for (i in t:(s-1)){
      treatment = synthetic_df %>% filter(time == i) %>% select(treatment) 
      add = (causal_params[1] * treatment$treatment) + (causal_params[2] * treatment$treatment^2)
      post_t_effect = post_t_effect + add
    }
    
    newH = cbind((Ys-post_t_effect),s,t)
    Hst = rbind(Hst,newH)
  }
}

Hst = Hst[-1,]

rownames(Hst) = NULL
colnames(Hst) = c("residual_effect","s","t")
```


Step 4: Fit a regression model for Hs on treatment, covariate history and PS

```{r}
H_df = rep(NA,11)

# Attach treatment @ t, treatment @ t-1, covariates @ t, PS @ t
for (i in seq(1,nrow(Hst),by=50)){
  time_val = Hst$t[i]
  time_df = synthetic_df %>% filter(time == time_val)
  
  attach = cbind(Hst[i:(i+49),],time_df$baseline_covariate,time_df$confounder_1,
                 time_df$confounder_2,time_df$confounder_3,time_df$treatment,
                 time_df$treatment_prev,time_df$outcome_prev,time_df$stabilized_IPW)
  H_df = rbind(H_df,attach)
}

H_df = H_df[-1,]

rownames(H_df) = NULL
colnames(H_df) = c("Hst","s","t",colnames(synthetic_df)[c(3:8,10,20)])
```

```{r}
# H_df1 = H_df %>% filter(baseline_covariate == 5)

#Vansteelandt paper says include treatment_prev but that leads to biased estimate? 
gestimation_model = lm(Hst ~ baseline_covariate + confounder_1 + confounder_2 + 
                         confounder_3 + treatment +
                         I(treatment^2) + treatment_prev + 
                         outcome_prev + stabilized_IPW, data = H_df)

summary(gestimation_model)
```


