---
title: "INLA-Hyper-Learning"
author: "Jeffrey Wu"
date: "2024-08-23"
output: html_document
---

Load all relevant packages:

```{r,warning=FALSE,echo=FALSE}
library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(huge)
library(INLA)
library(HMMpa)
library(invgamma)
library(brinla)
library(reshape2)
library(patchwork)
library(jsonlite)
library(geosphere)
library(urbnmapr)
library(RAQSAPI)
library(con2aqi)
library(pscl)
library(randtoolbox)
library(lhs)
```

# Load workspace from Paper2-ToyEx

```{r}
load("~/work/toyex-8.23.RData")
```

# Functions needed for INLA models

```{r}
EPA_kernel2 = function(EPA_data = inla_dataset,unique_dates,rho_rbf,rho_periodic,sigma2){

  K_EPA = matrix(0,nrow=length(unique_dates),ncol=length(unique_dates))
  i = 1
  j = 1
  
  for(t1 in unique_dates){
    for (t2 in unique_dates){
      A = EPA_data %>% filter(date == t1)
      B = EPA_data %>% filter(date == t2)
      
      ABtest = c(A$wildfire_pm25-B$wildfire_pm25)^2 
      
      # K_EPA[i,j] = exp(-sum(ABtest) / (2*rho_rbf)) * sigma2
      
      K_EPA[i,j] = exp(- (mean(ABtest))
                     / (2*rho_rbf)) * exp(- (2*sin(mean(abs(ABtest))*pi/12)^2)
                     / (rho_periodic)) * sigma2
      
      j = j+1
    }
    
    j = 1
    i = i+1
  }
  
  return(K_EPA)
}
```


## Ensuring precision matrix is not computationally singular, so we jitter eigenvalues, using reciprocal condition number as constraint

```{r,eval=FALSE}
desingularize = function(covmatrix,threshold = 1e-2,increment = 0.01){
  
  tracker = 0
  
  while (rcond(covmatrix) <= threshold){
    #Perform spectral decomposition
    ev = eigen(covmatrix)
    L = ev$values
    V = ev$vectors
    
    # #Add a little noise to eigenvalues to bring away from 0
    L = L + increment
    
    # #Add 0.01 to eigenvalues in bottom five percentile to bring away from 0
    # cutoff = quantile(abs(L),0.05)
    # L[which(abs(L) < cutoff)] = L[which(abs(L) < cutoff)] + 0.01
    
    #Calculate new precision matrix 
    covmatrix = V %*% diag(L) %*% t(V)
    
    tracker = tracker + 1
  }
  
  results_list = list(covmatrix,tracker)
  #sprintf("%s iterations of desingularizer applied",tracker)
  return(results_list)
}

# test = desingularize(K_time)
```

## Calculating simple time kernel K

```{r,eval=FALSE}
time_kernel = function(time_span,rho_rbf,rho_periodic,sigma2){
  K_time = matrix(NA,nrow = time_span, ncol = time_span)
  
  for (i in 1:time_span){
    for (j in 1:time_span){
      # K_time[i,j] = exp(- (abs(i-j)^2) / (2*rho)) * sigma2
      
      K_time[i,j] = exp(- (abs(i-j)^2) / (2*rho_rbf)) * exp(- (2*sin(sum(abs(i-j))*3.14/12)^2)
                     / (rho_periodic)) * sigma2
    }
  }
  
  return(K_time)
}
```


# IN SAMPLE FITTING ANALYSIS

## Ref model 2

```{r}
ref_model2 = function(dataset,a_prec_prior = 1,b_prec_prior = 5e-04,a_phi_prior = 1,b_phi_prior = 5e-04,link=1){
  ###Fit INLA model 
  bym_prior <- list(
  prec.unstruct = list(
    prior = "loggamma",
    param = c(a_prec_prior,b_prec_prior)),
  prec.spatial = list(
    prior = "loggamma",
    param = c(a_phi_prior,b_phi_prior))
  )
  
  ref_formula2 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(county_id, model = "bym", graph = huge.est, hyper = bym_prior) 
  
  model = inla(formula = ref_formula2,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  
  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  ref_model2_results = list(
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model
  )
  
  return(ref_model2_results)
}
```

# Defining grid of hyperparameter value pairs

```{r}
a_phi_prior = seq(1,5,length=5)
b_phi_prior = seq(0.001,1,length=5)

hyper_grid = matrix(ncol=2)

for (i in a_phi_prior){
  for(j in b_phi_prior){
    hyper_grid = rbind(hyper_grid,c(i,j))
  }
} 

hyper_grid = hyper_grid[-1,]
colnames(hyper_grid) = c("alpha","beta")
```

```{r,warning=FALSE}
#Need to try apply on this function(), store each list of results as item of a bigger list
# Define a function to apply kgr_model3 to each row
apply_function <- function(row) {
  a_phi_prior <- row[1]
  b_phi_prior <- row[2]
  
  # Call your ref_model1 function with the specific hyperparameters
  result <- ref_model2(dataset = inla_train_dataset,a_prec_prior = 1,b_prec_prior = 0.001,
                       a_phi_prior = a_phi_prior,b_phi_prior = b_phi_prior)
  
  # Return the result
  return(result)
}

# Apply the function to each row of the matrix
ref_model2_results_list <- apply(hyper_grid, 1, apply_function)

# #Extracting DIC values 
# ref_model2_results_DIC = c()
# 
# for (i in 1:length(ref_model2_results_list)){
#   ref_model2_results_DIC[i] = pred_data = ref_model2_results_list[[i]]$model_DIC
# }
# 
# hist(ref_model2_results_DIC)
# top5 = head(sort(ref_model2_results_DIC))
# top5
# which(ref_model2_results_DIC <= top5[5])
# 
# ref_model2_results_DIC = cbind(hyper_grid,ref_model2_results_DIC)

#Extracting WAIC values 
ref_model2_results_WAIC = c()

for (i in 1:length(ref_model2_results_list)){
  ref_model2_results_WAIC[i] = pred_data = ref_model2_results_list[[i]]$model_WAIC
}

hist(ref_model2_results_WAIC)
top5 = head(sort(ref_model2_results_WAIC))
top5
top5_idx = which(ref_model2_results_WAIC <= top5[5])

ref_model2_results_WAIC = cbind(hyper_grid,ref_model2_results_WAIC)
colnames(ref_model2_results_WAIC) = c(colnames(hyper_grid),"WAIC")
ref_model2_results_WAIC[top5_idx,]

# #Plot WAIC values for each hyperparameter grid 
# for (i in 1:2){
#   plot(x = ref_model2_results_WAIC[,i],y = ref_model2_results_WAIC[,3],xlab = colnames(ref_model2_results_WAIC)[i],ylab = "WAIC")
# }
```

Best hyperparameter values (so far): (0.1,0.1) for prec and (4,0.5005) for phi , WAIC = 4648.238 (very flat no difference)

# Fitting kernel graph regression models

## Proposed model 1

$vec(Y) | \Lambda, \textbf{X} \sim Pois(vec(\Lambda))$ and $\Lambda_{i,t} | \textbf{X} = exp(\beta_0 + F_{i,t})$

where the graph signal $\textbf{F} | \textbf{X} \sim GP(0,Ktime \otimes H^2)$ with $Cov[F_{n_1,t_1},F_{n_2,t_2}] = k(t_1,t_2)(H^2)_{n_1,n_2}$.

```{r,warning=FALSE}
kgr_model1 = function(dataset, rho_time_rbf = 1, rho_time_periodic = 1, sigma2_time = 1, link=1){
  
  #Calculating gram matrix K_time
  K_time = time_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_time_rbf, 
                        rho_periodic = rho_time_periodic, sigma2 = sigma2_time)
  
  #Heatmap of resulting K 
  K_time_heatmap = matrix_heatmap(K_time,title = "K_time heatmap")
  
  #Calculate trace norm of gram matrix
  K_time_weight = norm((1/length(unique(dataset$date)))*K_time,type = "F")
  
  #Calculate proposed kernel
  covGP2 = kronecker(K_time/length(unique(dataset$date)),(H^2)/num_clus)
  
  #Need to ensure precision matrix is not computationally singular i.e det > 0
  covGP_jittered = desingularize(covGP2,threshold = 1e-2,increment = 0.01)
  covGP2 = covGP_jittered[[1]]
  
  inv_covGP2 = solve(covGP2)
  
  #Heatmap of resulting inv_covGP2 
  inv_covGP2_heatmap = matrix_heatmap(inv_covGP2,title = "")
  
  ###Fit INLA model 
  kgr_formula1 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP2)
  
  model = inla(formula = kgr_formula1,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  # model_summary <- model$summary.fixed
  # bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  # preds_model <- model$summary.fitted.values
  # preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  # colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  # marginal_fvs <- model$marginals.fitted.values
  # 
  # #Exponentiating parameter to get better interpretation of estimates 
  # multeff <- exp(model$summary.fixed$mean)
  # names(multeff) <- model$names.fixed
  # 
  # #Plot of each parameters' posterior density 
  # mf <- melt(model$marginals.fixed)
  # cf <- spread(mf,Var2,value)
  # names(cf)[2] <- 'parameter'
  # param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
  #            scales="free") + geom_vline(xintercept=0) + ylab("density")
  # 
  # #Plot of precision of random effect (main hyperparameter of interest)
  # sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  # hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
  #            xlab("linear predictor")
  
  #Store the results in the list
  kgr_model1_results = list(
    # K_time_heatmap = K_time_heatmap,
    # K_time_weight = K_time_weight/(K_time_weight + gfilter_weight),
    # gfilter_weight = gfilter_weight/(K_time_weight + gfilter_weight),
    # covmatrix = covGP2,
    # prec = inv_covGP2,
    # num_jitters = covGP_jittered[[2]],
    # prec_heatmap = inv_covGP2_heatmap,
    # model_summary = model_summary,
    # bri_hyperpar_summary = bri_hyperpar_summary,
    # exp_effects = multeff,
    # param_plot = param_plot,
    # hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC
    # fitted_values = preds_model,
    # marg_fitted_values = marginal_fvs
  )
  
  return(kgr_model1_results)
}
```

```{r}
hyper_grid2 = randomLHS(100,3)
colnames(hyper_grid2) = c("rho_time_rbf","rho_time_periodic","sigma2_time")
hyper_grid2[,1:2] = hyper_grid2[,1:2]*600 
hyper_grid2[,3] = hyper_grid2[,3]*5
```

```{r,warning=FALSE}
apply_function <- function(row, index) {
  rho_time_rbf <- row[1]
  rho_time_periodic <- row[2]
  sigma2_time <- row[3]

  result <- tryCatch(
    {
      kgr_model1(dataset = inla_train_dataset, rho_time_rbf = rho_time_rbf,
                       rho_time_periodic = rho_time_periodic, sigma2_time = sigma2_time)
    },
    error = function(e) {
      # Handle the error by printing a message and returning NULL
      cat("Error in model", index, ":", conditionMessage(e), "\n")
      return(NULL)
    }
  )

  # If result is not NULL, process the result
  if (!is.null(result)) {
    # Print and save WAIC every 5th model
    if (index %% 10 == 0) {
      cat("Model", index, "WAIC:", result$model_WAIC, "\n")
      kgr_model1_running_results <<- c(kgr_model1_running_results, result$model_WAIC)
      saveRDS(kgr_model1_running_results, paste0("kgr_model1_running_results_WAIC.rds"))
    }
  }

  # Return the result (which could be NULL if there was an error)
  return(result)
}

kgr_model1_running_results <- vector()  # Initialize an empty vector to store WAIC results

kgr_model1_results_list <- lapply(seq_len(nrow(hyper_grid2)), function(i) {
  apply_function(hyper_grid2[i, ], i)
})

# Remove NULL results from the list if needed
kgr_model1_results_list <- Filter(Negate(is.null), kgr_model1_results_list)

#Extracting DIC values 
kgr_model1_results_DIC = c()

for (i in 1:length(kgr_model1_results_list)){
  kgr_model1_results_DIC[i] = pred_data = kgr_model1_results_list[[i]]$model_DIC
}

hist(kgr_model1_results_DIC)
top5 = head(sort(kgr_model1_results_DIC))
top5
which(kgr_model1_results_DIC <= top5[5])

kgr_model1_results_DIC = cbind(hyper_grid2,kgr_model1_results_DIC)

#Extracting WAIC values 
kgr_model1_results_WAIC = c()

for (i in 1:length(kgr_model1_results_list)){
  kgr_model1_results_WAIC[i] = pred_data = kgr_model1_results_list[[i]]$model_WAIC
}


saveRDS(kgr_model1_results_WAIC, paste0("kgr_model1_results_WAIC.rds"))


hist(kgr_model1_results_WAIC)
top5 = head(sort(kgr_model1_results_WAIC))
top5
top5_idx = which(kgr_model1_results_WAIC <= top5[5])

kgr_model1_results_WAIC = cbind(hyper_grid2,kgr_model1_results_WAIC)
colnames(kgr_model1_results_WAIC) = c(colnames(hyper_grid2),"WAIC")
kgr_model1_results_WAIC[top5_idx,]
```

Best hyperparameter values (so far): (rho_time_rbf, rho_time_periodic, sigma2_time) = (536.519,5.707,2.773) , WAIC = 16406.52 

Finally, we could also increase the complexity of our proposed model by including our time kernel in the covariance structure of the underlying GP. Notice that K does not explicitly have a temporal dependence structure; instead, it represents EPA covariate similarity compared across months. We can explicitly include the time kernel above by either element wise adding or multiplying K_EPA and K_time together before taking the kronecker product with $H^2$

## Proposed model 3

$vec(Y) | \Lambda, \textbf{X} \sim Pois(vec(\Lambda))$ and $\Lambda_{it} | \textbf{X} = exp(\beta_0 + F_{it})$

where the graph signal $\textbf{F} | \textbf{X} \sim GP(0,(Ktime \odot K_{EPA}) \otimes H^2)$ with $Cov[F_{n_1,t_1},F_{n_2,t_2}] = k(t_1,t_2)k(x_{t_1},x_{t_2})(H^2)_{n_1,n_2}$.

```{r,warning=FALSE}
kgr_model3 = function(dataset,rho_EPA_rbf = 1,rho_EPA_periodic = 1,rho_time_rbf = 1,rho_time_periodic = 1,sigma2 = 1,link=1){
  
  ###Calculating gram matrix K_EPA
  K_EPA = EPA_kernel2(unique_dates = unique(dataset$date),rho_rbf = rho_EPA_rbf,
                     rho_periodic = rho_EPA_periodic,sigma2 = 1)
  
  #Heatmap of resulting K 
  K_EPA_heatmap = matrix_heatmap(K_EPA,title = "K_EPA heatmap")
  
  ###Calculating gram matrix K_time
 K_time = time_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_time_rbf, 
                      rho_periodic = rho_time_periodic, sigma2 = sigma2)
  
  #Heatmap of resulting K 
  K_time_heatmap = matrix_heatmap(K_time,title = "K_time heatmap")
  
  K_EPA_norm = norm(K_EPA,type = "F")
  K_time_norm = norm(K_time,type = "F")
  
  #Calculate trace norm of gram matrix
  gram = (K_EPA*K_time)/sigma2
  K_weight = norm((1/length(unique(dataset$date)))*gram,type = "F")
  
  
  ###Load graph regression kernel 
  # covGP3 = kronecker(gram,H^2)
  covGP3 = kronecker(gram/length(unique(dataset$date)),(H^2)/num_clus)
  
  #Need to ensure precision matrix is not computationally singular i.e det > 0
  covGP_jittered = desingularize(covGP3,threshold = 1e-2,increment = 0.1)
  covGP3 = covGP_jittered[[1]]
  
  inv_covGP3 = solve(covGP3)
  
  #Heatmap of resulting K 
  inv_covGP3_heatmap = matrix_heatmap(inv_covGP3,title = "")
  
  ###Fit INLA model 
  # kgr_formula3 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
  #   b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
  #   Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP3)
  
  kgr_formula3 = deaths ~ -1 + month + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP3)
  
  model = inla(formula = kgr_formula3,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "gaussian"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  # model_summary <- model$summary.fixed
  # bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  # preds_model <- model$summary.fitted.values
  # preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  # colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  # marginal_fvs <- model$marginals.fitted.values
  # 
  # #Exponentiating parameter to get better interpretation of estimates 
  # multeff <- exp(model$summary.fixed$mean)
  # names(multeff) <- model$names.fixed
  # 
  # #Plot of each parameters' posterior density 
  # mf <- melt(model$marginals.fixed)
  # cf <- spread(mf,Var2,value)
  # names(cf)[2] <- 'parameter'
  # param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
  #            scales="free") + geom_vline(xintercept=0) + ylab("density")
  # 
  # #Plot of precision of random effect (main hyperparameter of interest)
  # sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  # hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
  #            xlab("linear predictor")
  
  #Store the results in the list
  kgr_model3_results = list(
    # K_EPA_heatmap = K_EPA_heatmap,
    # K_time_heatmap = K_time_heatmap,
    # K_EPA_weight = K_EPA_norm / (K_EPA_norm + K_time_norm),
    # K_time_weight = K_time_norm / (K_EPA_norm + K_time_norm),
    # K_weight = K_weight/(K_weight + gfilter_weight),
    # gfilter_weight = gfilter_weight/(K_weight + gfilter_weight),
    # covmatrix = covGP3,
    # prec = inv_covGP3,
    # num_jitters = covGP_jittered[[2]],
    # prec_heatmap = inv_covGP3_heatmap,
    # model_summary = model_summary,
    # bri_hyperpar_summary = bri_hyperpar_summary,
    # exp_effects = multeff,
    # param_plot = param_plot,
    # hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC
    # fitted_values = preds_model,
    # marg_fitted_values = marginal_fvs
  )
  
  return(kgr_model3_results)
}
```

# Defining grid of hyperparameter value pairs

```{r}
hyper_grid2 = randomLHS(100,5)
colnames(hyper_grid2) = c("rho_EPA_rbf","rho_EPA_periodic","rho_time_rbf",
                         "rho_time_periodic","sigma2")

hyper_grid2[,1:4] = hyper_grid2[,1:4]*600000
hyper_grid2[,5] = hyper_grid2[,5]*5 
```

```{r,warning=FALSE}
apply_function <- function(row, index) {
  rho_EPA_rbf <- row[1]
  rho_EPA_periodic <- row[2]
  rho_time_rbf <- row[3]
  rho_time_periodic <- row[4]
  sigma2 <- row[5]

  result <- tryCatch(
    {
      kgr_model3(dataset = inla_train_dataset, rho_EPA_rbf = rho_EPA_rbf,
                       rho_EPA_periodic = rho_EPA_periodic, rho_time_rbf = rho_time_rbf,
                       rho_time_periodic = rho_time_periodic, sigma2 = sigma2)
    },
    error = function(e) {
      # Handle the error by printing a message and returning NULL
      cat("Error in model", index, ":", conditionMessage(e), "\n")
      return(NULL)
    }
  )

  # If result is not NULL, process the result
  if (!is.null(result)) {
    # Print and save WAIC every 5th model
    if (index %% 20 == 0) {
      cat("Model", index, "WAIC:", result$model_WAIC, "\n")
      kgr_model3_running_results <<- c(kgr_model3_running_results, result$model_WAIC)
      saveRDS(kgr_model3_running_results, paste0("kgr_model3_running_results_WAIC.rds"))
    }
  }

  # Return the result (which could be NULL if there was an error)
  return(result)
}

kgr_model3_running_results <- vector()  # Initialize an empty vector to store WAIC results

kgr_model3_results_list <- lapply(seq_len(nrow(hyper_grid2)), function(i) {
  apply_function(hyper_grid2[i, ], i)
})

# Remove NULL results from the list if needed
kgr_model3_results_list <- Filter(Negate(is.null), kgr_model3_results_list)

#Extracting DIC values 
kgr_model3_results_DIC = c()

for (i in 1:length(kgr_model3_results_list)){
  kgr_model3_results_DIC[i] = pred_data = kgr_model3_results_list[[i]]$model_DIC
}

hist(kgr_model3_results_DIC)
top5 = head(sort(kgr_model3_results_DIC))
top5
which(kgr_model3_results_DIC <= top5[5])

kgr_model3_results_DIC = cbind(hyper_grid2,kgr_model3_results_DIC)

#Extracting WAIC values 
kgr_model3_results_WAIC = c()

for (i in 1:length(kgr_model3_results_list)){
  kgr_model3_results_WAIC[i] = pred_data = kgr_model3_results_list[[i]]$model_WAIC
}

saveRDS(kgr_model3_results_WAIC, paste0("kgr_model3_results_WAIC.rds"))


hist(kgr_model3_results_WAIC)
top5 = head(sort(kgr_model3_results_WAIC))
top5
top5_idx = which(kgr_model3_results_WAIC <= top5[5])

kgr_model3_results_WAIC = cbind(hyper_grid2,kgr_model3_results_WAIC)
colnames(kgr_model3_results_WAIC) = c(colnames(hyper_grid2),"WAIC")
kgr_model3_results_WAIC[top5_idx,]

# #Plot WAIC values for each hyperparameter grid 
# for (i in 1:5){
#   plot(x = kgr_model3_results_WAIC[,i],y = kgr_model3_results_WAIC[,4],xlab = colnames(kgr_model3_results_WAIC)[i],ylab = "WAIC")
# }
```

Best hyperparameter values (so far): (rho_EPA_rbf, rho_EPA_periodic, rho_time_rbf, rho_time_periodic, sigma2) 
= (37801.48, 571512.56. 39.5276, 380794.8, 1.124888) , WAIC= 16419.36

## Proposed model 5

```{r,warning=FALSE}
kgr_model5 = function(dataset, rho_AR_rbf = 1, rho_AR_periodic = 1, rho_DL_rbf = 1, rho_DL_periodic = 1, rho_int_rbf = 1, rho_int_periodic = 1, sigma2_AR = 1, sigma2_DL = 1, sigma2_int = 1, link = 1){

  ###Calculating gram matrix K_AR
  K_AR_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab S_random data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    S_random_county = county_data$S_random
    
    #Create a list to contain covariance matrix for each col (should I get rid of the for loop and just do wildfire_pm2.5)
    K_AR_periodic_list = list()
    
    rho_AR = 1
    sigma2_AR = 1
    
    time_span = nrow(S_random_county)
    
    #Calculate a AR 1 covariance matrix for each col and store in list
    for (i in 1:2){
      ts =  S_random_county[,i]
      K_covariate_periodic = matrix(nrow=time_span,ncol=time_span)
      
      for(j in 1:time_span){
        for (k in 1:time_span){
          if (abs(j-k) <= 1){
            K_covariate_periodic[j,k] = exp(- ((ts[j] - ts[k])^2) / (2*rho_AR)) * exp(- (2*sin((abs(ts[j] - ts[k]))*pi/12)^2)
                        / (rho_AR)) * sigma2_AR
          }
          else{
            K_covariate_periodic[j,k] = 0
            }
        }
      }
      K_AR_periodic_list[[i]] = K_covariate_periodic
    }

    names(K_AR_periodic_list) = colnames(S_random_county)
    
    #Add each pollutant's covariance matrix to get AR 1 matrix for each county
    K_AR_periodic = (1/2)*(K_AR_periodic_list[[1]] + K_AR_periodic_list[[2]])
    K_AR_periodic_weights = c(norm(K_AR_periodic_list[[1]],type = "F"),
    norm(K_AR_periodic_list[[2]],type = "F"))

    # K_AR_periodic = matrix(0,nrow=time_span,ncol=time_span)
    
    # K_AR_periodic_weights = rep(1,length(K_AR_periodic_list))
    
    # for(i in 1:length(K_AR_periodic_list)){
    #   K_AR_periodic = K_AR_periodic + ((1/2)*K_AR_periodic_list[[i]])
      
    #   K_AR_periodic_weights[i] = norm(K_AR_periodic_list[[i]],type = "F")
    # }
    
    K_AR_periodic_weights = K_AR_periodic_weights / sum(K_AR_periodic_weights)
    K_AR_periodic_county[[c]] = K_AR_periodic
  }

  K_AR_periodic = matrix(0,nrow=time_span,ncol=time_span)

  K_AR_periodic_weights = rep(1,num_counties)

  for(i in 1:num_counties){
    K_AR_periodic = K_AR_periodic + ((1/num_counties)*K_AR_periodic_county[[i]])
    
    K_AR_periodic_weights[i] = norm(K_AR_periodic_county[[i]],type = "F")
  }

  ###Calculating gram matrix K_DL
  K_DL_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab S_DL data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    S_DL_county = county_data$S_DL #has original and backshifted cols if needed
    S_DL_county = S_DL_county[,1] #only take original ts for now
    
    #Create a list to store covariance matrix for each DL 
    K_DL_periodic_list = list()
    
    dl_lags = c(3,6,12)
    tracker = 1
    
    for (i in dl_lags){
      
      K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
      
      rho_DL = 1
      sigma2_DL = 1
      
      #Calculate DL covariance matrix for specified lag   
      for(j in 1:length(S_DL_county)){
        for (k in 1:length(S_DL_county)){
          
          if ((abs(j-k) == 0) || (abs(j-k) == i)){
            K_DL_periodic[j,k] = exp(- (sum(S_DL_county[j] - S_DL_county[k])^2)
                                / (2*rho_DL)) * exp(- (2*sin(sum(abs(S_DL_county[j] - S_DL_county[k]))*pi/12)^2)
                                / (rho_DL)) * sigma2_DL
          } 
          else{
            K_DL_periodic[j,k] = 0
            }
        }
      }
      K_DL_periodic_list[[tracker]] = K_DL_periodic
      tracker = tracker+1
    }
    
    #Combine the 3 DL covariance matrices together
    K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
    
    K_DL_periodic_weights = rep(1,length(K_DL_periodic_list))
    
    for(i in 1:length(K_DL_periodic_list)){
      K_DL_periodic = K_DL_periodic + ((1/3)*K_DL_periodic_list[[i]])
      K_DL_periodic_weights[i] = norm(K_DL_periodic_list[[i]],type = "F")
    }
    
    K_DL_periodic_weights = K_DL_periodic_weights / sum(K_DL_periodic_weights)
   
    #Store DL(3,6,12) covariance matrix for each county 
    K_DL_periodic_county[[c]] = K_DL_periodic
  }

  K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
  K_DL_periodic_weights = rep(1,num_counties)

  for(i in 1:num_counties){
    K_DL_periodic = K_DL_periodic + ((1/num_counties)*K_DL_periodic_county[[i]])
    K_DL_periodic_weights[i] = norm(K_DL_periodic_county[[i]],type = "F")
  }
  
  ###Calculating gram matrix K_Interaction
  K_Interaction_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab interaction pair data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    W_county = county_data$W
    time_sq = length(W_county)

    K_interaction_periodic_list = list()
    
    #Create sequence of indices corresponding to comparisons for real time and one lag interaction effects
    idx = seq(1,time_sq,by=70)
    
    #Calculate a kernel for each interaction pair  
    K_int0_periodic = matrix(nrow = time_span,ncol = time_span)
    
    rho_int = 1
    sigma2_int = 1
    
    for (i in 1:time_span){
      for (j in 1:time_span){       
        #Locally periodic kernels 
        K_int0_periodic[i,j] = exp(- ((W_county[idx[i]] - W_county[idx[j]])^2) / (2*rho_int)) * 
        exp(- (2*sin((abs(W_county[idx[i]] - W_county[idx[j]]))*pi/12)^2) / (rho_int)) * sigma2_int
      }
    }
    
    # matrix_heatmap(K_int0_periodic,title = "Periodic Interaction Covariance Structure")

    K_Interaction_periodic_county[[c]] = K_int0_periodic
  }

  K_interaction_periodic = matrix(0,nrow=time_span,ncol=time_span)
  K_interaction_periodic_weights = rep(1,num_counties)

  for(i in 1:num_counties){
    K_interaction_periodic = K_interaction_periodic + ((1/num_counties)*K_Interaction_periodic_county[[i]])

    K_interaction_periodic_weights[i] = norm(K_Interaction_periodic_county[[i]],type = "F")
  }
  
  gram = (1/3)*(K_AR_periodic+K_DL_periodic+K_interaction_periodic)
  K_weight = norm((1/time_span)*gram,type = "F")
  
  ###Load graph regression kernel 
  # covGP5 = kronecker(gram,H^2)
  covGP5 = kronecker(gram/time_span,(H^2)/num_clus)
  
  #Need to ensure precision matrix is not computationally singular i.e det > 0
  covGP_jittered = desingularize(covGP5,threshold = 1e-2,increment = 0.1)
  covGP5 = covGP_jittered[[1]]
  
  inv_covGP5 = solve(covGP5)
  
  #Heatmap of resulting K 
  # inv_covGP5_heatmap = matrix_heatmap(inv_covGP5,title = "")
  
  ###Fit INLA model 
  # kgr_formula5 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
  #   b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
  #   Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP3)

  kgr_formula5.2 = deaths ~ -1 + month + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP5)
  
  model = inla(formula = kgr_formula5.2,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  # model_summary <- model$summary.fixed
  # bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  # preds_model <- model$summary.fitted.values
  # preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  # colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  # marginal_fvs <- model$marginals.fitted.values
  # 
  # #Exponentiating parameter to get better interpretation of estimates 
  # multeff <- exp(model$summary.fixed$mean)
  # names(multeff) <- model$names.fixed
  # 
  # #Plot of each parameters' posterior density 
  # mf <- melt(model$marginals.fixed)
  # cf <- spread(mf,Var2,value)
  # names(cf)[2] <- 'parameter'
  # param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
  #            scales="free") + geom_vline(xintercept=0) + ylab("density")
  # 
  # #Plot of precision of random effect (main hyperparameter of interest)
  # sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  # hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
  #            xlab("linear predictor")
  
  #Store the results in the list
  kgr_model5_results = list(
    # K_AR_heatmap = K_AR_heatmap, 
    # K_DL_heatmap = K_DL_heatmap,
    # K_Interaction_heatmap = K_Interaction_heatmap,
    # K_AR_weight = K_AR_norm / (K_AR_norm + K_DL_norm + K_Int_norm),
    # K_DL_weight = K_DL_norm / (K_AR_norm + K_DL_norm + K_Int_norm),
    # K_Int_weight = K_Int_norm / (K_AR_norm + K_DL_norm + K_Int_norm),
    # K_weight = K_weight/(K_weight + gfilter_weight),
    # gfilter_weight = gfilter_weight/(K_weight + gfilter_weight),
    # covmatrix = covGP5,
    # prec = inv_covGP5,
    # num_jitters = covGP_jittered[[2]],
    # prec_heatmap = inv_covGP5_heatmap,
    # model_summary = model_summary,
    # bri_hyperpar_summary = bri_hyperpar_summary,
    # exp_effects = multeff,
    # param_plot = param_plot,
    # hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC
    # fitted_values = preds_model,
    # marg_fitted_values = marginal_fvs
  )
  
  return(kgr_model5_results)
}
```

# Defining grid of hyperparameter value pairs

```{r}
hyper_grid2 = randomLHS(100,9)
colnames(hyper_grid2) = c("rho_AR_rbf","rho_AR_periodic","rho_DL_rbf","rho_DL_periodic",
                         "rho_int_rbf","rho_int_periodic","sigma2_AR","sigma2_DL","sigma2_int")
hyper_grid2[,1:6] = hyper_grid2[,1:6]*0.0009 
hyper_grid2[,7:9] = hyper_grid2[,7:9]*5
```


```{r,warning=FALSE}
apply_function <- function(row,index) {
  rho_AR_rbf <- row[1]
  rho_AR_periodic <- row[2]
  rho_DL_rbf <- row[3]
  rho_DL_periodic <- row[4]
  rho_int_rbf <- row[5]
  rho_int_periodic <- row[6]
  sigma2_AR <- row[7]
  sigma2_DL <- row[8]
  sigma2_int <- row[9]
  
  # Call kgr_model5 function with the specific hyperparameters
  result <- kgr_model5(dataset = inla_train_dataset, rho_AR_rbf = rho_AR_rbf, 
                       rho_AR_periodic = rho_AR_periodic, rho_DL_rbf = rho_DL_rbf, 
                       rho_DL_periodic = rho_DL_periodic, rho_int_rbf = rho_int_rbf,
                       rho_int_periodic = rho_int_periodic, sigma2_AR = sigma2_AR, 
                       sigma2_DL = sigma2_DL, sigma2_int = sigma2_int, link=1)
  
# Print and save WAIC every 100th model
  if (index %% 5 == 0) {
    cat("Model", index, "WAIC:", result$model_WAIC, "\n")
    kgr_model5_running_results = c(kgr_model5_running_results , result$model_WAIC)
    saveRDS(kgr_model5_running_results, paste0("kgr_model5_running_results_WAIC.rds"))
  }
  
  # Return the result
  return(result)
}

kgr_model5_running_results <- vector()  # Initialize an empty vector to store WAIC results

kgr_model5_results_list <- lapply(seq_len(nrow(hyper_grid2)), function(i) {
  apply_function(hyper_grid2[i, ], i)
})

# #Posterior predictive plots
# for (i in 1:length(kgr_model5_results_list)){
#   pp_insample_plot(pred_data = kgr_model5_results_list[[i]]$fitted_values)
# }


# #Extracting DIC values 
# kgr_model5_results_DIC = c()
# 
# for (i in 1:length(kgr_model5_results_list)){
#   kgr_model5_results_DIC[i] = pred_data = kgr_model5_results_list[[i]]$model_DIC
# }
# 
# hist(kgr_model5_results_DIC)
# top5 = head(sort(kgr_model5_results_DIC))
# top5
# which(kgr_model5_results_DIC <= top5[5])
# 
# kgr_model5_results_DIC = cbind(hyper_grid,kgr_model5_results_DIC)

#Extracting WAIC values 
kgr_model5_results_WAIC = c()

for (i in 1:length(kgr_model5_results_list)){
  kgr_model5_results_WAIC[i] = pred_data = kgr_model5_results_list[[i]]$model_WAIC
}

saveRDS(kgr_model5_results_WAIC, paste0("kgr_model5_results_WAIC.rds"))

hist(kgr_model5_results_WAIC)
top5 = head(sort(kgr_model5_results_WAIC))
top5
top5_idx = which(kgr_model5_results_WAIC <= top5[5])

kgr_model5_results_WAIC = cbind(hyper_grid2,kgr_model5_results_WAIC)
colnames(kgr_model5_results_WAIC) = c(colnames(hyper_grid2),"WAIC")
kgr_model5_results_WAIC[top5_idx,]

#Plot WAIC values for each hyperparameter grid 
# for (i in 1:9){
#   plot(x = kgr_model5_results_WAIC[,i],y = kgr_model5_results_WAIC[,5],xlab = colnames(kgr_model5_results_WAIC)[i],ylab = "WAIC")
# }
```

Best hyperparameter values (so far): (rho_EPA_rbf, rho_EPA_periodic, rho_time_rbf, rho_time_periodic, sigma2) 
= (1.019648e-04,3.904021e-04,6.926401e-04,1.378534e-05,1.494036e-04,3.228066e-04,
5.567759e-01,5.955738e-01,2.712031) , WAIC= 16430.73 (all hyperparameter values lead to same WAIC?) 

