---
title: "LengthScale_Bounds"
author: "Jeffrey Wu"
date: "2024-08-23"
output: html_document
---

Finding reasonable upper bound for each model's length scale parameters - for a given lag, what values of rho lead to a decay in autocorrelation < 0.05

# Proposed model 1 (time kernel)

```{r}
kernel1 = function(t1=1,t2=60,rho){
  correlation = exp(- (abs(t1-t2)^2) / (2*rho)) * exp(- (2*sin(sum(abs(t1-t2))*pi/12)^2)
                     / (rho))
  return(correlation)
}

#Less than 0.05 correlation at 3 yr lag
corr3 = kernel1(1,36,1:1000)
plot(corr3)

length(which(corr3 < 0.05))

kernel1(1,36,204)

#Less than 0.05 correlation at 5 yr lag
corr5 = kernel1(1,60,1:1000)
plot(corr5)

length(which(corr5 < 0.05))

kernel1(1,60,581)
```

If memoryless after 3 years, rho upper bound is 204

If memoryless after 5 years, rho upper bound is 581

# Proposed model 3 (EPA kernel * time kernel)

```{r}
kernel3 = function(EPA_data = final_EPA_data,t1,t2,rho){
  A = EPA_data %>% filter(time == t1)
  B = EPA_data %>% filter(time == t2)
  
  ABtest = c((A$AQI-B$AQI)^2,(A$lead-B$lead)^2,(A$co-B$co)^2,(A$so2-B$so2)^2,(A$no2-B$no2)^2,
             (A$o3-B$o3)^2,(A$pm10-B$pm10)^2,(A$pm25-B$pm25)^2) #58 counties * 8 measurements
  
  correlation1 = exp(- (mean(ABtest)) / (2*rho)) * exp(- (2*sin(mean(abs(ABtest))*pi/12)^2) / (rho)) 
  
  correlation2 = exp(- (abs(t1-t2)^2) / (2*rho)) * exp(- (2*sin(sum(abs(t1-t2))*pi/12)^2) / (rho))
  
  return(correlation1*correlation2)
}

#Less than 0.05 correlation at 3 yr lag
corr3 = kernel3(t1=1,t2=36,rho=1:500)
plot(corr3)

length(which(corr3 < 0.05))

kernel3(t1=1,t2=36,rho=218)

#Less than 0.05 correlation at 5 yr lag
corr5 = kernel3(t1=1,t2=60,rho=1:1000)
plot(corr5)

length(which(corr5 < 0.05))

kernel3(t1=1,t2=60,rho=598)
```

If memoryless after 3 years, rho upper bound is 218

If memoryless after 5 years, rho upper bound is 598

```{r}
kernel3.2 = function(EPA_data = inla_dataset,t1,t2,rho){
  A = EPA_data %>% filter(date == t1)
  B = EPA_data %>% filter(date == t2)
  
  # ABtest = c((A$AQI-B$AQI)^2,(A$lead-B$lead)^2,(A$co-B$co)^2,(A$so2-B$so2)^2,(A$no2-B$no2)^2,
  #            (A$o3-B$o3)^2,(A$pm10-B$pm10)^2,(A$pm25-B$pm25)^2) #58 counties * 8 measurements 
  
  ABtest = c(A$wildfire_pm25-B$wildfire_pm25)^2 #58 counties * 8 measurements 
  
  correlation1 = exp(- (mean(ABtest)) / (2*rho)) * exp(- (2*sin(mean(abs(ABtest))*pi/12)^2) / (rho)) 
  
  correlation2 = exp(- (abs(as.numeric(difftime(t1,t2)))^2) / (2*rho)) * 
    exp(- (2*sin(sum(abs(as.numeric(difftime(t1,t2))))*pi/12)^2) / (rho))
  
  return(correlation1*correlation2)
}

#Less than 0.05 correlation at 3 yr lag
corr3 = kernel3.2(t1="2014-04-01",t2="2017-04-01",rho=seq(100000,300000,length.out=1000))
plot(corr3)

length(which(corr3 < 0.05))

kernel3.2(t1="2014-04-01",t2="2017-04-01",rho=200300)

#Less than 0.05 correlation at 5 yr lag
corr5 = kernel3.2(t1="2014-04-01",t2="2019-04-01",rho=seq(500000,1000000,length.out=1000))
plot(corr5)

length(which(corr5 < 0.05))

kernel3.2(t1="2014-04-01",t2="2019-04-01",rho=556056)
```

If memoryless after 3 years, rho upper bound is 200300

If memoryless after 5 years, rho upper bound is 556056


# Proposed model 5 (AR kernel + DL kernel + Interaction kernel)

```{r}
kernel5 = function(data = decomposed_wildfire_county_data,t1,t2,rho){
  ###Calculating gram matrix K_AR
  K_AR_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab S_random data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    S_random_county = county_data$S_random
    
    #Create a list to contain covariance value for each col (should I get rid of the for loop and just do wildfire_pm2.5)
    K_covariate_periodic = c()
    rho = rho
    time_span = nrow(S_random_county)
    
    #Calculate a AR 1 covariance value for each col and store in list
    for (i in 1:2){
      ts =  S_random_county[,i]
      if (abs(t1-t2) <= 1){
        K_covariate_periodic[i] = exp(- ((ts[t1] - ts[t2])^2) / (2*rho)) * exp(- (2*sin((abs(ts[t1] - ts[t2]))*pi/12)^2)
                    / (rho))
      }
      else{
        K_covariate_periodic[i] = 0
        }
    }

    #Add each pollutant's covariance value to get AR 1 matrix for each county
    K_AR_periodic = (1/2)*(K_covariate_periodic[1] + K_covariate_periodic[2])

    # K_AR_periodic = matrix(0,nrow=time_span,ncol=time_span)
    
    # K_AR_periodic_weights = rep(1,length(K_AR_periodic_list))
    
    # for(i in 1:length(K_AR_periodic_list)){
    #   K_AR_periodic = K_AR_periodic + ((1/2)*K_AR_periodic_list[[i]])
      
    #   K_AR_periodic_weights[i] = norm(K_AR_periodic_list[[i]],type = "F")
    # }
    
    K_AR_periodic_county[[c]] = K_AR_periodic
  }

  K_AR_periodic = 0

  for(i in 1:num_counties){
    K_AR_periodic = K_AR_periodic + ((1/num_counties)*K_AR_periodic_county[[i]])
  }

  ###Calculating gram matrix K_DL
  K_DL_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab S_DL data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    S_DL_county = county_data$S_DL #has original and backshifted cols if needed
    S_DL_county = S_DL_county[,1] #only take original ts for now
    
    #Create a list to store covariance value for each DL 
    K_DL_periodic_vec = c()
    
    dl_lags = c(3,6,12)
    tracker = 1
    rho = rho
    
    for (i in dl_lags){
      
      #Calculate DL covariance value for specified lag   
      if ((abs(t1-t2) == 0) || (abs(t1-t2) == i)){
        K_DL_periodic_vec[tracker] = exp(- (sum(S_DL_county[t1] - S_DL_county[t2])^2)
                            / (2*rho)) * exp(- (2*sin(sum(abs(S_DL_county[t1] - S_DL_county[t2]))*pi/12)^2)
                            / (rho))
      } 
      else{
        K_DL_periodic_vec[tracker] = 0
        }
      tracker = tracker+1
    }
    
    #Combine the 3 DL covariance values together
    K_DL_periodic = 0
    
    for(i in 1:length(K_DL_periodic_vec)){
      K_DL_periodic = K_DL_periodic + ((1/3)*K_DL_periodic_vec[i])
    }
    
    #Store DL(3,6,12) covariance matrix for each county 
    K_DL_periodic_county[[c]] = K_DL_periodic
  }

  K_DL_periodic = 0

  for(i in 1:num_counties){
    K_DL_periodic = K_DL_periodic + ((1/num_counties)*K_DL_periodic_county[[i]])
  }
  
  ###Calculating gram matrix K_Interaction
  K_Interaction_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab interaction pair data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    W_county = county_data$W
    time_sq = length(W_county)
    
    #Create sequence of indices corresponding to comparisons for real time and one lag interaction effects
    idx = seq(1,time_sq,by=70)
    
    #Calculate a kernel for each interaction pair  
    rho = rho
    
    #Locally periodic kernels 
    K_int0_periodic = exp(- ((W_county[idx[t1]] - W_county[idx[t2]])^2) / (2*rho)) * 
    exp(- (2*sin((abs(W_county[idx[t1]] - W_county[idx[t2]]))*pi/12)^2) / (rho))

    # matrix_heatmap(K_int0_periodic,title = "Periodic Interaction Covariance Structure")

    K_Interaction_periodic_county[[c]] = K_int0_periodic
  }

  K_interaction_periodic = 0

  for(i in 1:num_counties){
    K_interaction_periodic = K_interaction_periodic + ((1/num_counties)*K_Interaction_periodic_county[[i]])
  }
  
  correlation = (1/3)*(K_AR_periodic+K_DL_periodic+K_interaction_periodic)

return(correlation)
}
```

```{r}
#Less than 0.05 correlation at 3 yr lag
corr3 = c()
grid = seq(0.00001,0.0005,length.out=100)

for (i in grid){
  corr3 = c(corr3,kernel5(data = decomposed_cluster_data, t1 = 1,t2=36,rho=i))
}

plot(corr3)

grid[length(which(corr3 < 0.05))]

kernel5(t1=1,t2=36,rho=8.424242e-05)

#Less than 0.05 correlation at 5 yr lag
corr5 = c()
grid = seq(0.0001,0.001,length.out=100)

for (i in grid){
  corr5 = c(corr5,kernel5(data = decomposed_cluster_data, t1 = 1,t2=60,rho=i))
}

plot(corr5)

grid[length(which(corr5 < 0.05))]

kernel5(t1=1,t2=60,rho=0.0008545455)
```




