---
title: "Wildfire KGR Causal Exercise"
author: "Jeffrey Wu"
date: "2024-09-18"
format:
  html:
   code-fold: true
   code-summary: "Show the code"
   embed-resources: true
   toc: true
   output-ext: "html"
   grid:
    sidebar-width: 0px
    body-width: 1000px
self-contained: true
---

# Downloading relevant data

```{r,warning=FALSE,echo=FALSE}
library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(huge)
library(INLA)
library(HMMpa)
library(invgamma)
library(brinla)
library(reshape2)
library(patchwork)
library(jsonlite)
library(geosphere)
library(urbnmapr)
library(RAQSAPI)
library(con2aqi)
library(pscl)
library(reshape2)
library(corrplot)
library(superheat)
library(shapes)
library(scales)
library(kableExtra)
library(splines)
```

## SDI data

```{r,message=FALSE,cache=TRUE,collapse=TRUE}
###SoA data
soa.data = read_xlsx("SoA.data.1019.xlsx")

###county_flips are unique identifier for counties
soa.data$county_fips = as.character(soa.data$county_fips) ##change it to character

#IMPORTANT

# This shape file contains the coordinates for county boundaries
##counties is from urbanmap

CA.counties = urbnmapr::counties %>% filter(state_abbv == "CA")

###IF WE WANT TO BOIL DOWN TIME SERIES AND KEEP ALL DATA, SWITCH to CA_newdata below
soa_joint <- left_join(CA.counties, soa.data, by = "county_fips")

#Use with soa.data (full data)
CA_data = soa_joint %>% select(long, lat, county_name.y, Year, Score, Total_Pop,
                               EDUC_Lessthan9, EDUC_college, White_Collar,
                               Unemployment_Rate, Adj_HH_income, Income_Disparity,
                               Individuals_Below_Poverty, Median_Home_Value,
                               Median_Gross_Rent, Housing_No_Telephone,
                               Housing_Incomplete_Plumbing)

### Quick descriptions of SoA data variables
# Score: social deprivation index (SDI) score calculated from the following 11 subindices
# EDUC_Lessthan9: % of population older than 24 with less than 9 years of education
# EDUC_college: % of population older than 24 with at least four years of college education
# White_Collar: % of population older than 15 employed in a white collar occupation
# Unemployment_Rate: unemployment rate for population older than 15
# Adj_HH_income: median household income adjusted for local housing costs
# Income_Disparity: an income disparity ratio
# Individuals_Below_Poverty: % of population below the federal poverty line
# Median_Home_Value: median home value for owned, occupied units
# Median_Gross_Rent: median gross rent for rented units
# Housing_No_Telephone: % of households without a telephone
# Housing_Incomplete_Plumbing: % of households with incomplete plumbing

colnames(CA_data)[3] = "County"

CA_newdata = soa.data[1:58,]
CA_newdata = CA_newdata[,-c(4,7,8)]
```

## Census population data

Heatmap of population by county:

```{r}
#Initializing map and station locations
ca_map <- map_data("county", region = "california")

#Match population dataset with ca_map
#2010-2019 population data for CA 
USpops = read.csv("CA_census_pops1019.csv")
CApops = USpops %>% filter(STNAME == "California") %>% select(CTYNAME,POPESTIMATE2019)
CApops = CApops[-1,]

CApops$CTYNAME = unique(ca_map$subregion)
colnames(CApops) = c("subregion","pop")

merged_data <- merge(ca_map, CApops, by = "subregion", all.x = TRUE)

#Plot
gg_pop <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = pop), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  labs(title = "Heatmap of County Populations for 2019",fill = expression("Population")) +
  scale_fill_gradient(low = "yellow", high = "red")

print(gg_pop)
```

# SKATER clustering

Learn SKATER clusters on SDI data (but don't aggregate):

```{r}
###Setting up SPDF for CA counties 
CA_sf = st_read(getwd(),"CA_Counties_TIGER2016")
CA_spdf = as_Spatial(CA_sf)

# score_scaled = scale(CA_data$Score)

c_index = unique(CA_data$County)
y_index = unique(CA_data$Year)
SDI_df = matrix(nrow=10,ncol=58)
track1 = 1
track2 = 1

for (i in c_index){
  for (j in y_index){
    scores = CA_data %>% filter(County == i) %>% filter(Year == j) %>% select(Score) %>% unique()
    SDI_df[track1,track2] = scores$Score
    track1 = track1 + 1
  }
  track1 = 1
  track2 = track2 + 1
}

score_scaled = scale(SDI_df) #NEED TO SCALE THE DATA BEFORE FEEDING IT INTO SKATER

#covariates_scale = data.frame(apply(CA_data[,4:16],2,scale))
covariates_scale = data.frame(t(score_scaled))

CA_spdf@data = covariates_scale
```

```{r}
#Identify neighborhood list for counties 
CA_nb = poly2nb(CA_spdf)

#summary(CA_nb)

# plot(CA_spdf, main = "With queen")
# plot(CA_nb, coords = coordinates(CA_spdf), col="blue", add = TRUE)

#Calculate edge costs (dissimilarity matrix) based on Euclidean distance 
costs <- nbcosts(CA_nb, data = covariates_scale)

###Get adjacency matrix using nb2mat() (SEPARATE STEP FOR INLA)
adj = nb2mat(CA_nb,style = "B")

#Style means the coding scheme style used to create the weighting matrix 
# B: basic binary coding scheme
# W: row standardized coding scheme 
# C: globally standardized coding scheme  
# U: values of C / number of neighbors 
# S: variance stabilizing coding scheme 

#Transform edge costs to spatial weights 
ct_w <- nb2listw(CA_nb,costs,style="B")

#Create minimum spanning tree 
ct_mst <- mstree(ct_w)

plot(ct_mst,coordinates(CA_spdf),col="blue", cex.lab=0.5)
plot(CA_spdf, add=TRUE)

#Run SKATER algorithm to get 7 contiguous clusters (cluster idx is in order of CA_sf)
clus5 <- skater(edges = ct_mst[,1:2], data = covariates_scale, ncuts = 4)

#Determine an appropriate minimum population threshold based on???
pops_summary = summary(unique(CA_data$Total_Pop))
pops_summary

#Idea 1: Use median * (how many counties should be in a cluster at minimum)
min_pop = as.numeric(pops_summary[3] * 4)

#Idea 2: If we assume CA population is 39M, divide total pop by # clusters
min_pop2 = 39000000 / 5

#Add a min population constraint
clus5_min <- skater(edges = ct_mst[,1:2], 
                     data = covariates_scale, 
                     crit = 5000000, 
                     vec.crit = CA_data$Total_Pop,
                     ncuts = 64)

#Add a minimum number of areas in each cluster constraint 
clus5_minarea = skater(edges = ct_mst[,1:2], data = covariates_scale, ncuts = 4, 4)


CA_data_cluster = (CA_sf %>% mutate(clus = clus5_min$groups))

#Plot clustered CA
plot((CA_sf %>% mutate(clus = clus5$groups))['clus'], main = "5 cluster example")

plot((CA_sf %>% mutate(clus = clus5_min$groups))['clus'], main = "5 cluster example with population constraint")

plot((CA_sf %>% mutate(clus = clus5_minarea$groups))['clus'], main = "5 cluster example with minimum number of areas constraint")
```

For reference, here are the cluster labels for each county:

```{r}
clusterlabels = data.frame(CA_data_cluster$NAME,clus5_minarea$groups)
names(clusterlabels) = c("counties","Cluster")

o = order(clusterlabels$counties)
clusterlabels = clusterlabels[o,]
rownames(clusterlabels) = NULL

clusterlabels
CA_counties = clusterlabels$counties
num_clus = max(clus5_minarea$groups)
num_counties = length(unique(CA_counties))
```

# Learn a graph on SDI data with HUGE

```{r}
huge_input = as.matrix(score_scaled)
#huge_input = as.matrix(SDI_df)

#Graph learning w HUGE
out.glasso = huge(huge_input,lambda = seq(0.95,0.05,by=-0.05),method="glasso")

glasso.stars = huge.select(out.glasso,criterion = "stars",stars.thresh = 0.1)
glasso.ric = huge.select(out.glasso,criterion = "ric")
glasso.ebic = huge.select(out.glasso,criterion = "ebic")

plot(glasso.stars)
plot(glasso.ric)
plot(glasso.ebic)

huge.est = glasso.ebic$refit
huge.est

degree_connectivity = data.frame(colSums(huge.est))
colnames(degree_connectivity) = "node_connections"
degree_connectivity = cbind(c(1:num_counties),degree_connectivity)
```

## Transforming estimated adjacency matrix to graph filter H

The code below takes the adjacency matrix estimated in the previous step and transforms it into a graph filter H. The steps are explained in Antonian et al (Gareth's paper). The cutoff transformation is used for the eigenvalues.

```{r}
A = as.matrix(huge.est)
p = nrow(A)

#obtain graph Laplacian L
D = diag(p)
for (i in 1:p){
  d = sum(A[,i])
  D[i,i] = d
}

L = D - A

#eigendecomposition of L
Ldecomp = eigen(L)
U = as.matrix(Ldecomp$vectors)
Lambdas = Ldecomp$values

#test
#U %*% (diag(p)*Lambdas) %*% t(U)

#Function implementing cutoff tranform for eigenvalues 
cutoff.transform = function(lambdas,q){
  transformed = c()
  cutoff = quantile(lambdas,q)
  for (i in lambdas){
    if(i <= cutoff){
      transformed = c(transformed,1)
    }
    else{
      transformed = c(transformed,0)
    }
  }
  
  return(transformed)
}

#quantile(Lambdas,2/3)
transformed.L = cutoff.transform(Lambdas,2/3)
eta.L = diag(p)*transformed.L

#obtain graph filter
H = U %*% eta.L %*% t(U)

gfilter_weight = norm((1/58)*H^2,type = "F")
# gfilter_weight = norm(H^2,type = "F")
```

Create a function to generate heatmap plots of matrices:

```{r}
matrix_heatmap= function(matrix,title = "",gradient_zones = c(0,0.8,0.999)){
  r = nrow(matrix)
  df = as_tibble(cbind(expand.grid(rev(seq_len(r)),seq_len(r)),c(matrix))) %>% setNames(c("row","col","value"))
  df$value[df$value == 1] = 0.999
  
  plot = ggplot(df,mapping = aes(x=row,y=col,fill=value)) + geom_tile() + 
    scale_fill_gradientn(colors = c("yellow","orange","red"),
                       values = rescale(gradient_zones),
                       limits = c(0, 0.99),
                       oob = squish) + ggtitle(title) + theme_void()
  return(plot)
}
```

Heatmap plot of graph filter:

```{r}
#Heatmap of resulting H 
#corrplot(H, order = 'original', cl.pos = 'b', tl.pos = 'n',method = "color", col = COL1('YlOrRd',10),
         #title = "Graph filter")

matrix_heatmap(H,title = "Graph filter")
```

# EPA data

```{r}
final_EPA_data = readRDS("final_EPA_data_9.1.RData")
head(final_EPA_data)

county_labels = rep(CA_counties,each=12)
county_labels = rep(county_labels,7)
county_labels = rep(county_labels,6)

final_EPA_data = cbind(county_labels,final_EPA_data)

final_EPA_data <- final_EPA_data %>%
  spread(key = Pollutant, value = Value)

colnames(final_EPA_data) = c("county_labels","date","AQI","lead","co",
                             "so2","no2","o3","pm10","pm25")

time = rep(1:72,58)
final_EPA_data$time = time
```

## Get county and pollutant reference codes from EPA

Make sure each air pollutant time series is stationary:

```{r}
#Get county codes 
counties_url = "https://aqs.epa.gov/data/api/list/countiesByState?email=jeffreywu@ucsb.edu&key=copperheron86&state=06&"

countycodes = fromJSON(counties_url)
countycodes = countycodes[[2]]
california_counties = countycodes$code

#Get parameter codes
parameters_url = "https://aqs.epa.gov/data/api/list/parametersByClass?email=jeffreywu@ucsb.edu&key=copperheron86&pc=CRITERIA"

parametercodes = fromJSON(parameters_url)
parametercodes = parametercodes[[2]]

#Do we want to keep Lead PM10 or delete
parametercodes = parametercodes[-7,]

pollutants = data.frame(parametercodes$code)
labels = c("lead","co","so2","no2","o3","pm10","pm25")
pollutants = cbind(pollutants,labels)
```

```{r,eval=FALSE}
library(astsa)

for (i in CA_counties){
  for (j in pollutants$parametercodes.code){
    EPA_clus_ts = final_EPA_data %>% filter(Pollutant == j,county_labels == i) 

    title = sprintf("Pollutant %s - %s",j,i)
    tsplot(EPA_clus_ts$Value,main = title,xlab = "Time (months)",ylab = "Value")  
  }
}
```

It turns out that the time series for Lead were not stationary, so we need to detrend the data before using it

```{r,eval=FALSE}
for (i in CA_counties){
  for (j in pollutants$parametercodes.code){
    EPA_clus_ts = final_EPA_data %>% filter(Pollutant == j,county_labels == i) 
    
    lead_idx = which(final_EPA_data$Pollutant == "14129" & final_EPA_data$county_labels == i)
    
    if (j == "14129"){
      detrended1 = c(0,diff(EPA_clus_ts$Value,lag=1))
      final_EPA_agg_data[lead_idx,2] = detrended1
      
      title = sprintf("Pollutant %s - %s",j,i)
      tsplot(detrended1,main = title,xlab = "Time (months)",ylab = expression(nabla~Value))
    } else{
      title = sprintf("Pollutant %s - %s",j,i)
      tsplot(EPA_clus_ts$Value,main = title,xlab = "Time (months)",ylab = "Value")  
    }
  }
}
```

# Mortality data

```{r}
###Cal-ViDa data
mortality = read.csv("respmortality1419.csv") #data from 2014-2019 bc we want to avoid COVID pandemic era
mortality = filter(mortality,Cause_of_Death %in% c("Chronic lower respiratory diseases","Influenza and pneumonia"))
mortality = mortality[,-c(1,4,9)]
Population = rep(100000,nrow(mortality))
mortality = cbind(mortality,Population)
```

```{r}
population_age = read_xlsx("Population Categories.xlsx")
#head(population_age)

population_age = population_age[-(1:5),2]
population_age$...2 = as.numeric(population_age$...2)

#split under 5 category into < 1 and 1-4 years old 
less1 = floor(population_age$...2[1]*0.2)
onefour = floor(population_age$...2[1]*0.8)
death_byage = population_age$...2[-1]
death_byage = c(less1,onefour,death_byage)
death_byage = death_byage[1:19]

death_byage2 = death_byage[1:2]
idx = seq(from = 3, to = 17, by = 2)
for (i in idx){
  death_byage2 = c(death_byage2,(death_byage[i]+death_byage[i+1]))
}
death_byage2 = c(death_byage2,death_byage[19])


age_groups = unique(mortality$Age)
death_byage2 = data.frame(cbind(age_groups,death_byage2))
colnames(death_byage2) = c("Age_Group","Population_by_Age")
death_byage2$Population_by_Age = as.numeric(death_byage2$Population_by_Age)
#head(death_byage2)

#2010-2019 population data for CA 
USpops = read.csv("CA_census_pops1019.csv")
CApops = USpops %>% filter(STNAME == "California") %>% select(CTYNAME,POPESTIMATE2019)
counties = countycodes$value_represented #from EPA data file

weights = CApops[(2:59),2]
weights = weights/CApops[1,2]

groups = unique(mortality$Age)
step1 = 1
step2 = 1

for (i in counties){
  for (j in groups){
    idx = which(mortality$Age == j & mortality$County_of_Death == i)
    mortality$Population[idx] = ceiling(death_byage2$Population[step1]*weights[step2])
    step1 = step1+1
  }  
  step1 = 1
  step2 = step2+1
}

mortality$logpop = log(mortality$Population)

censored_idx = which(mortality$Total_Deaths == "<11")
censorTF = mortality$Total_Deaths == "<11"
mortality = cbind(mortality,censorTF)
#head(mortality)
```

## Clean and aggregate mortality data

EM ALGORITHM TO IMPUTE CENSORED MORTALITY ROWS (\<10 DEATHS)

```{r}
uncensored_mortality = mortality %>% filter(censorTF == FALSE) %>% select(Total_Deaths,Population)
uncensored_mortality$Total_Deaths = as.numeric(uncensored_mortality$Total_Deaths)

theta = mean(uncensored_mortality$Total_Deaths*100000/uncensored_mortality$Population)
```

FUNCTION FOR IMPUTING CENSORED VALUE BASED ON EXPECTATION GIVEN LAMBDA

```{r}
impute_small_values = function(lambda){
  x = 1:10
  p = dpois(x,lambda)
  
  value = sum(x*p)/sum(p)
  return(value)
}
```

INITIAL IMPUTATION

```{r}
mortality2 = mortality
mortality2$Total_Deaths[censored_idx] = 0.01
mortality2$Total_Deaths = as.numeric(mortality2$Total_Deaths)

for (i in censored_idx){
  lambda = theta*mortality2$Population[i] / 100000
  deaths = impute_small_values(lambda)
  
  mortality2$Total_Deaths[i] = floor(deaths)
}
```

INITIAL REGRESSION MODELS

```{r}
mortality2$Age = factor(mortality2$Age)
mortality2$Cause_of_Death = factor(mortality2$Cause_of_Death)
mortality2$Month = factor(mortality2$Month)

pois_reg = glm(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop), family = "poisson", data = mortality2)
# ZIP_reg = zeroinfl(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop) | 1, data = mortality2, dist = "poisson", link = "logit")

vec0 = coef(pois_reg)
# vec0 = coef(ZIP_reg)

summary(pois_reg)
# summary(ZIP_reg)
```

```{r}
mortality3 = mortality2
model = pois_reg

model_diff = 100
iter = 1
vec0 = coef(model)

while((model_diff > 0.01) & (iter < 10)){
  
  #impute data (should be between 1-10)
  fvs = fitted.values(model)
  
  for (i in censored_idx){
    deaths = impute_small_values(fvs[i])
    mortality3$Total_Deaths[i] = floor(deaths)
  }
  
  #fit model on "new" data
  model = glm(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop), family = "poisson", data = mortality3)
  vec1 = coef(model)
  
  model_diff = max(abs(vec1 - vec0))
  iter = iter+1
  vec0 = vec1
}

final_pois_reg = model
summary(final_pois_reg)

logLik(final_pois_reg)
```

Reformatting the data set (mainly aggregating)

```{r,collapse=TRUE}
#Aggregating by  cause of death
data = mortality3

data1 = data %>% filter(Cause_of_Death == "Chronic lower respiratory diseases")
data2 = data %>% filter(Cause_of_Death == "Influenza and pneumonia")

newdeaths = data1$Total_Deaths + data2$Total_Deaths
data1$Total_Deaths = newdeaths

respmortality = data1[,-5]


#Creates total deaths by adding deaths of all age groups together 
agg.respmortality = respmortality[1,]
agg.respmortality$Age = as.character(agg.respmortality$Age)
rows2 = seq(1,nrow(respmortality),11)

for (i in rows2){
  agg.respmortality[i,] = respmortality[i,]
  agg.respmortality$Total_Deaths[i] = sum(respmortality$Total_Deaths[i:(i+10)])
  agg.respmortality$Age[i] = "Everyone"
}

agg.respmortality = na.omit(agg.respmortality)
rownames(agg.respmortality) = NULL
total.respmortality = agg.respmortality[,-c(6:9)]
```

Reformat dataset into time series format (rows are counties, columns are months)

```{r,cache=FALSE}
months = unique(total.respmortality$Month_of_Death)
years = sort(unique(total.respmortality$Year_of_Death))
counties = unique(total.respmortality$County_of_Death)
x = 0

total.mortality.ts = matrix(1,nrow = 58, ncol = 72)

for (k in CA_counties){
  county.ts = c()
  x = x+1
  
  for (i in years){
    for (j in months){
      deaths = total.respmortality %>% filter(County_of_Death == k) %>%  filter(Year_of_Death == i) %>% filter(Month_of_Death == j) %>% select(Total_Deaths) %>% as.numeric()
      county.ts = c(county.ts,deaths)
    }
  } 
  
  total.mortality.ts[x,] = county.ts
}

#Label time series data 
total.mortality.ts = as.data.frame(total.mortality.ts)

dates = c()
x=1
for (i in years){
  for (j in months){
    dates[x] = sprintf("%1.0f/%1.0f",j,i)
    x = x+1
  }
}

colnames(total.mortality.ts) = dates
ID = c(1:58)
total.mortality.ts = cbind(ID,counties,total.mortality.ts)

total.mortality.ts = left_join(clusterlabels,total.mortality.ts,by = "counties")
head(total.mortality.ts)
```

# Wildfire PM 2.5 data

Filter CA counties, attach county names to measurements

```{r}
# Load smokePM predictions on smoke days
preds = readRDS("smokePM2pt5_predictions_daily_county_20060101-20201231.rds")

# Load counties
ca_geoinfo = readRDS("ca_geodata.rds")

ca_geoid_counties = ca_geoinfo %>% select(GEOID,NAME)

# Load full set of dates
dates = seq.Date(ymd("20060101"), ymd("20201231"), by = "day")

# Get full combination of county-days
# Warning: this may require a large amount of memory
out = expand.grid(GEOID = ca_geoid_counties$GEOID, date = dates)

# Match smokePM predictions on smoke days to county-days
out = left_join(out, preds, by = c("GEOID", "date"))
out2 = left_join(ca_geoid_counties,out,by="GEOID")
out3 = out2 %>% filter(date > "2014-01-01" & date < "2020-01-01")

# Predict 0 for remaining county-days, which are non-smoke days
out = mutate(out3, smokePM_pred = replace_na(smokePM_pred, 0))

test = out %>% filter(date > "2015-01-01",GEOID == "06111")
```

## Aggregate daily observations to monthly averages by GEOID

```{r,eval=FALSE}
# Ensure the date column is in Date format
out$date <- ymd(out$date)

out_subset = out %>% filter(NAME == "Santa Barbara")

# Aggregate daily observations to monthly averages by GEOID
monthly_avg <- out %>%
  group_by(GEOID, NAME, date = floor_date(date, "month")) %>%
  summarize(monthly_avg_pm25 = mean(smokePM_pred, na.rm = TRUE))

# View the aggregated dataframe
#View(monthly_avg)
```

Preview some time series plots for SB and SD county

```{r,eval=FALSE}
# Ensure the date column is in Date format
out$date <- ymd(out$date)

# Filter data for Santa Barbara (GEOID: 06083) and San Diego (GEOID: 06073) counties
santa_barbara_data <- monthly_avg %>% filter(GEOID == "06083")
san_diego_data <- monthly_avg %>% filter(GEOID == "06073")

# Combine the data for plotting
combined_data <- bind_rows(
  santa_barbara_data %>% mutate(NAME = "Santa Barbara"),
  san_diego_data %>% mutate(NAME = "San Diego")
)

# Create the time series plot
ggplot(combined_data, aes(x = date, y = monthly_avg_pm25, color = NAME)) +
  geom_line() +
  labs(title = "Time Series of smokePM_pred",
       x = "Date",
       y = "smokePM_pred",
       color = "NAME") +
  theme_minimal()
```

# Assemble final dataset to feed into INLA (EPA, mortality, smoke datasets)

```{r,eval=FALSE}
# Combine Month_of_Death and Year_of_Death into a single date column
total.respmortality2 <- total.respmortality %>%
  mutate(date = make_date(Year_of_Death, Month_of_Death, 1))


# Ensure date columns are in Date format
final_EPA_data$date <- ymd(final_EPA_data$date)
total.respmortality2$date <- ymd(total.respmortality2$date)
monthly_avg$date <- ymd(monthly_avg$date)

# epa_test = final_EPA_data %>% filter(county_labels == "Santa Barbara")
# mortality_test = total.respmortality2 %>% filter(County_of_Death == "Santa Barbara")
#monthly_avg2 = monthly_avg %>% filter(month < "2020-01-01")


# Join the datasets
combined_data <- final_EPA_data %>%
  inner_join(total.respmortality2, by = c("date", "county_labels" = "County_of_Death"))
combined_data2 = cbind(combined_data,monthly_avg)

inla_dataset = combined_data2[,-c(11,13,16,17)]
colnames(inla_dataset)[c(1,11,12)] = c("county","month","deaths")

# View the combined data
head(inla_dataset)

saveRDS(inla_dataset,file="inla_dataset_8.20")
```

# Fit INLA models (BYM and KGR):

Create training and test dataset:

```{r}
inla_dataset = readRDS("inla_dataset_8.20")

#Examine effect of wildfire smoke levels 3 and 6 months before 
#So create B3 and B6 columns for lag air pollutant covariates 

b3_inla_dataset = inla_dataset %>% filter(date >= "2014-04-01")
b6_inla_dataset = inla_dataset %>% filter(date >= "2014-07-01")

b3_rows = which(inla_dataset$date < "2019-10-01")
b6_rows = which(inla_dataset$date < "2019-07-01")

b3_covariates = inla_dataset[b3_rows,c(3:10,14)]
b6_covariates = inla_dataset[b6_rows,c(3:10,14)]

b3_inla_dataset = cbind(b3_inla_dataset,b3_covariates)
b6_inla_dataset = cbind(b6_inla_dataset,b6_covariates) 

colnames(b3_inla_dataset)[c(14,16:24)] = c("wildfire_pm25","b3_AQI","b3_lead","b3_co","b3_so2","b3_no2","b3_o3","b3_pm10","b3_pm25","b3_wildfire_pm25")
colnames(b6_inla_dataset)[c(14,16:24)] = c("wildfire_pm25","b6_AQI","b6_lead","b6_co","b6_so2","b6_no2","b6_o3","b6_pm10","b6_pm25","b6_wildfire_pm25")
```

Add county IDs and cluster level fixed effects

```{r}
county_ids <- b3_inla_dataset %>%
  distinct(county) 
county_ids$county_id = c(1:58)
county_ids$cluster = clusterlabels$Cluster

# Merge the county_id back into the original dataset
b3_inla_dataset <- b3_inla_dataset %>%
  left_join(county_ids, by = "county")

# b3_inla_dataset = b3_inla_dataset %>%
#   left_join(clusterlabels, by = c("county" = "counties"))

inla_dataset = b3_inla_dataset


#Add multiple intercept columns, one for each cluster 
b3_inla_dataset <- b3_inla_dataset %>%
  mutate(
    Intercept1 = ifelse(cluster == 1, 1, NA),
    Intercept2 = ifelse(cluster == 2, 1, NA),
    Intercept3 = ifelse(cluster == 3, 1, NA),
    Intercept4 = ifelse(cluster == 4, 1, NA),
    Intercept5 = ifelse(cluster == 5, 1, NA)
  ) 
```

Say we are examining the Thomas fire in SB (12/4/17-1/12/-18) -\> split data into

b3: Apr 2014-Apr 2018 training and May 2018-2019 test b6: July 2014-July 2018 training and Aug 2018-2019 test

```{r}
inla_train_dataset = b3_inla_dataset %>% filter(date < "2018-05-01")
inla_test_dataset = b3_inla_dataset %>% filter(date >= "2018-05-01")

inla_dataset_toimpute = b3_inla_dataset

#omit_idx = which(inla_dataset_toimpute$date >= "2018-05-01" & inla_dataset_toimpute$date <= "2018-12-01")

omit_idx = which(inla_dataset_toimpute$date >= "2018-05-01")
inla_dataset_toimpute$deaths[omit_idx] = NA

idx1 = nrow(inla_train_dataset)
idx2 = nrow(inla_test_dataset)
idx3 = nrow(inla_dataset_toimpute)

inla_train_dataset$idx = c(1:idx1)
inla_test_dataset$idx = c(1:idx2)
inla_dataset_toimpute$idx = c(1:idx3)
```

# Fit models in sample: WITH WILDFIRE PM2.5 AS PREDICTOR (start w 1 month before, then 3, 6, 12)

```{r,eval=FALSE}
load("~/work/P2-ToyEx-premodeling.RData")
```

Functions for generating posterior predictive plots:

```{r,warning=FALSE}
#Write a function to make plot of posterior predictive estimates with credible interval bands OVERLAID on response (default is random, FALSE for all counties)
pp_insample_plot = function(random_or_all = TRUE, ref_data = inla_train_dataset, pred_data){
  
  if(random_or_all){
    idx = ceiling(runif(10,0,58))
    
    for (i in idx){
      df = ref_data %>% filter(county_id == i) %>% select(deaths)
      preds = pred_data %>% filter(county_id == i) 
      df = cbind(df,preds)
      
      # title = sprintf("Posterior Predictive Fits for Cluster %s",i)
      title = sprintf("%s County (Cluster %s)",
                      county_ids$county[i], county_ids$cluster[i])
      
      post_pred_plot = df %>% ggplot(aes(x=date,y=deaths)) + 
        geom_point() + geom_line(aes(y=mean),color = "red") + 
        geom_ribbon(aes(ymin = `0.025quant`,
                        ymax = `0.975quant`),alpha = 0.3) +
        ggtitle(title)
      
      print(post_pred_plot)
    }
  } else{
    for (i in 1:num_plots){
      df = ref_data %>% filter(county_id == i) %>% select(deaths)
      preds = pred_data %>% filter(county_id == i) 
      df = cbind(df,preds)
      
      title = sprintf("%s County (Cluster %s)",
                      county_ids$county[i], county_ids$cluster[i])
      
      post_pred_plot = df %>% ggplot(aes(x=date,y=deaths)) + 
        geom_point() + geom_line(aes(y=mean),color = "red") + 
        geom_ribbon(aes(ymin = `0.025quant`,
                        ymax = `0.975quant`),alpha = 0.3) + 
        ggtitle(title)

      
      print(post_pred_plot)
    }  
  }
}
```

```{r}
#Maybe modify to be useful for inla_dataset_toimpute as well
pp_outsample_plot = function(random_or_all = TRUE, ref_data = inla_test_dataset, pred_data){
  if(random_or_all){
    idx = ceiling(runif(10,0,58))
    
    for (i in idx){
      df = ref_data %>% filter(county_id == i) %>% select(deaths)
      preds = pred_data %>% filter(county_id == i) 
      df = cbind(df,preds)
      
      title = sprintf("%s County (Cluster %s)",
                      county_ids$county[i], county_ids$cluster[i])
      
      post_pred_plot = df %>% ggplot(aes(x=date,y=deaths)) + 
        geom_point() + geom_line(aes(y=mean),color = "red") + 
        geom_ribbon(aes(ymin = `0.025quant`,
                        ymax = `0.975quant`),alpha = 0.3) + 
        ggtitle(title)

      
      print(post_pred_plot)
    }
  } else{
    for (i in 1:num_plots){
      df = ref_data %>% filter(county_id == i) %>% select(deaths)
      preds = pred_data %>% filter(county_id == i) 
      df = cbind(df,preds)
      
      title = sprintf("%s County (Cluster %s)",
                      county_ids$county[i], county_ids$cluster[i])
      
      post_pred_plot = df %>% ggplot(aes(x=date,y=deaths)) + 
        geom_point() + geom_line(aes(y=mean),color = "red") + 
        geom_ribbon(aes(ymin = `0.025quant`,
                        ymax = `0.975quant`),alpha = 0.3) + 
        ggtitle(title)
      
      print(post_pred_plot)
    }  
  }
}
```

## BYM model

Y: response Z: treatment (b3_wildfire_pm2.5) X: other EPA pollutant covariates S: SDI data

$\textbf{Y} | \boldsymbol{\Lambda}, \boldsymbol{\beta} ,\textbf{F}, \textbf{Z}, \textbf{X_{EPA]}, \textbf{S}} \sim Poisson(\boldsymbol{\Lambda})$

$\Lambda_{i,t} | \textbf{F}, \textbf{Z}, \textbf{X}_{EPA} = \exp(\beta_{c1} I \{ c=1 \} + \beta_{c2} I \{ c=2 \} + ... + \beta_{c5} I \{ c=5 \} + \beta_1 I \{t \, mod \, 12 = 1 \} + ... + \beta_{11} I \{t \, mod \, 12 = 11 \} + \beta_{12} Z + \boldsymbol{\beta}^T \textbf{X_{EPA}} + \phi_i + \textbf{F}_i)$

$\textbf{F} | \textbf{S}, \tau \sim \mathcal{GP}(\textbf{0},\tau \boldsymbol{\Sigma})$

$p(\boldsymbol{\phi}) \propto \exp(-\frac{1}{2} \sum_{c_1 \sim c_2} (\phi_{c_1} - \phi_{c_2})^2)$

```{r}
#Write a function to fit our BYM model in INLA 
ref_model2 = function(dataset,a_prec_prior = 1,b_prec_prior = 5e-04,a_phi_prior = 1,b_phi_prior = 5e-04,link=1){
  ###Fit INLA model 
  bym_prior <- list(
  prec.unstruct = list(
    prior = "loggamma",
    param = c(a_prec_prior,b_prec_prior)),
  prec.spatial = list(
    prior = "loggamma",
    param = c(a_phi_prior,b_phi_prior))
  )
  
  ref_formula2 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(county_id, model = "bym", graph = huge.est, hyper = bym_prior) 
  
  model = inla(formula = ref_formula2,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  
  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  ref_model2_results = list(
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model
  )
  
  return(ref_model2_results)
}

#Fit ref_model2
ref_model2_fit = ref_model2(dataset = inla_train_dataset,a_prec_prior = 1,b_prec_prior = 5e-4,a_phi_prior = 1,b_phi_prior = 5e-4)

#Extract DIC and WAIC
ref_model2_DIC = ref_model2_fit$model_DIC
ref_model2_WAIC = ref_model2_fit$model_WAIC

#Get summaries of parameter estimates
ref_model2_fit$model_summary
ref_model2_fit$bri_hyperpar_summary
ref_model2_fit$exp_effects

#Show plots
ref_model2_fit$param_plot
ref_model2_fit$hyperparam_plot
pp_insample_plot(pred_data = ref_model2_fit$fitted_values)
```

For KGR models, we need to calculate gram matrices:

### Calculating gram matrix K from EPA data

Using the EPA air quality data, we can calculate the gram matrix K which will characterize the dependence structure of air quality (across 7 different pollutants and AQI) over time. This is done by calculating the squared difference between all of the observations at two different time points e.g. 64 observations for Jan 2014 compared with the 64 observations for Feb 2014. For the in sample analysis, the resulting matrix is 66x66 because we are holding out the last 6 months of observations.

```{r}
EPA_kernel = function(EPA_data = inla_dataset,unique_dates,rho_rbf,rho_periodic,sigma2){

  K_EPA = matrix(0,nrow=length(unique_dates),ncol=length(unique_dates))
  i = 1
  j = 1
  
  for(t1 in unique_dates){
    for (t2 in unique_dates){
      A = EPA_data %>% filter(date == t1)
      B = EPA_data %>% filter(date == t2)
      
      ABtest = c((A$AQI-B$AQI)^2,(A$lead-B$lead)^2,(A$co-B$co)^2,
                 (A$so2-B$so2)^2,(A$no2-B$no2)^2,(A$o3-B$o3)^2,
                 (A$pm10-B$pm10)^2,(A$pm25-B$pm25)^2) #58 counties * 8 measurements 
      
      # K_EPA[i,j] = exp(-sum(ABtest) / (2*rho_rbf)) * sigma2
      
      K_EPA[i,j] = exp(- (mean(ABtest))
                     / (2*rho_rbf)) * exp(- (2*sin(mean(abs(ABtest))*pi/12)^2)
                     / (rho_periodic)) * sigma2
      
      j = j+1
    }
    
    j = 1
    i = i+1
  }
  
  return(K_EPA)
}
```

```{r}
treatment_kernel = function(data = inla_dataset,unique_dates,rho_rbf,rho_periodic,sigma2){

  K_Z = matrix(0,nrow=length(unique_dates),ncol=length(unique_dates))
  i = 1
  j = 1
  
  for(t1 in unique_dates){
    for (t2 in unique_dates){
      A = data %>% filter(date == t1)
      B = data %>% filter(date == t2)
      
      ABtest = c(A$wildfire_pm25-B$wildfire_pm25)^2 
      
      # K_Z[i,j] = exp(-sum(ABtest) / (2*rho_rbf)) * sigma2
      
      K_Z[i,j] = exp(- (mean(ABtest))
                     / (2*rho_rbf)) * exp(- (2*sin(mean(abs(ABtest))*pi/12)^2)
                     / (rho_periodic)) * sigma2
      
      j = j+1
    }
    
    j = 1
    i = i+1
  }
  
  return(K_Z)
}


test = treatment_kernel(unique_dates = unique(inla_train_dataset$date),
                  rho_rbf=1,rho_periodic=1,sigma2=1)
```

### Calculating simple time kernel K

```{r}
time_kernel = function(unique_dates,rho_rbf,rho_periodic,sigma2){
  time_span = length(unique_dates)
  K_time = matrix(NA,nrow = time_span, ncol = time_span)
  
  for (i in 1:time_span){
    for (j in 1:time_span){
      # K_time[i,j] = exp(- (abs(i-j)^2) / (2*rho)) * sigma2
      
      K_time[i,j] = exp(- (abs(i-j)^2) / (2*rho_rbf)) * exp(- (2*sin(sum(abs(i-j))*pi/12)^2)
                     / (rho_periodic)) * sigma2
    }
  }
  
  return(K_time)
}

test2 = time_kernel(unique_dates = unique(inla_train_dataset$date),rho_rbf=1,
                                          rho_periodic=1,sigma2=1)
```

### Ensuring precision matrix is not computationally singular, so we jitter eigenvalues, using reciprocal condition number as constraint

```{r}
desingularize = function(covmatrix,threshold = 1e-2,increment = 0.01){
  
  if (any(is.na(covmatrix)) || any(is.infinite(covmatrix))) {
    # Add an initial small perturbation to the diagonal elements
    covmatrix = covmatrix + diag(0.1, nrow(covmatrix))  
    stop("Covariance matrix contains NaN or Inf values")
    }
  
  
  tracker = 0
  
  while (rcond(covmatrix) <= threshold){
    #Perform spectral decomposition
    ev = eigen(covmatrix)
    L = ev$values
    V = ev$vectors
    
    # #Add a little noise to eigenvalues to bring away from 0
    L = L + increment
    
    # #Add 0.01 to eigenvalues in bottom five percentile to bring away from 0
    # cutoff = quantile(abs(L),0.05)
    # L[which(abs(L) < cutoff)] = L[which(abs(L) < cutoff)] + 0.01
    
    #Calculate new precision matrix 
    covmatrix = V %*% diag(L) %*% t(V)
    
    tracker = tracker + 1
  }
  
  results_list = list(covmatrix,tracker)
  #sprintf("%s iterations of desingularizer applied",tracker)
  return(results_list)
}

# test = desingularize(K_time)
```

## KGR model 1

$\textbf{Y} | \boldsymbol{\Lambda}, \boldsymbol{\beta} , \textbf{F}, \textbf{Z}, \textbf{X}_{EPA}, \textbf{S} \sim Poisson(\boldsymbol{\Lambda})$

$\Lambda_{i,t} | \textbf{F}, \textbf{Z}, \textbf{X}_{EPA} = \exp(\beta_{c1} I \{ c=1 \} + \beta_{c2} I \{ c=2 \} + ... + \beta_{c5} I \{ c=5 \} + \beta_1 I \{t \, mod \, 12 = 1 \} + ... + \beta_{11} I \{t \, mod \, 12 = 11 \} + \beta_{12} \textbf{Z} + \boldsymbol{\beta}^T \textbf{X}_{EPA} + \textbf{F}_{i,t})$

$\textbf{F} | \textbf{S}, \rho_{rbf}, \rho_{p}, \sigma^2_{time} \sim \mathcal{GP}(\textbf{0},\textbf{K}_{time} \otimes \textbf{H}^2)$

$Cov(\textbf{F}_{i,t_1},\textbf{F}_{j,t_2}) = \left[ \textbf{K}_{time} \right]_{t_1,t_2} \left[ \textbf{H}^2 \right]_{i,j}$

```{r}
kgr_model1 = function(dataset, rho_time_rbf = 1, rho_time_periodic = 1, sigma2_time = 1, link=1){
  
  #Calculating gram matrix K_time
   K_time = time_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_time_rbf, 
                        rho_periodic = rho_time_periodic, sigma2 = sigma2_time)
  
  #Heatmap of resulting K 
  K_time_heatmap = matrix_heatmap(K_time,title = "K_time heatmap")
  
  #Calculate trace norm of gram matrix
  K_time_weight = norm((1/length(unique(dataset$date)))*K_time,type = "F")
  
  #Calculate proposed kernel
  covGP = kronecker(K_time/length(unique(dataset$date)),(H^2)/num_clus)
  
  #Need to ensure precision matrix is not computationally singular i.e det > 0
  covGP_jittered = desingularize(covGP,threshold = 1e-2,increment = 0.01)
  covGP = covGP_jittered[[1]]
  
  inv_covGP = solve(covGP)
  
  #Heatmap of resulting inv_covGP2 
  inv_covGP_heatmap = matrix_heatmap(inv_covGP,title = "")
  
  ###Fit INLA model 
  # kgr_formula1 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
  #   b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
  #   Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP)
  
  kgr_formula1_nonlinear = deaths ~ -1 + month + ns(b3_wildfire_pm25,df=3) + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP)
  
  model = inla(formula = kgr_formula1_nonlinear,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  # model = inla(inla.mdata(cbind(deaths,1), cbind(month, b3_wildfire_pm25, Intercept1, Intercept2, 
  #                                                Intercept3, Intercept4, Intercept5))
  #                         ~ -1 + month + b3_wildfire_pm25 + Intercept1 + Intercept2 + 
  #                           Intercept3 + Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP2),
  #              family = "0poisson",data = dataset,
  #               control.compute = list(dic=TRUE,waic=TRUE, return.marginals.predictor=TRUE),
  #               control.inla = list(strategy = "laplace"),
  #               control.predictor = list(compute = TRUE,link=link),
  #              control.family = list(link.simple = link.simple))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  marginal_fvs <- model$marginals.fitted.values
  
  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  kgr_model1_results = list(
    K_time_heatmap = K_time_heatmap,
    K_time_weight = K_time_weight/(K_time_weight + gfilter_weight),
    gfilter_weight = gfilter_weight/(K_time_weight + gfilter_weight),
    covmatrix = covGP,
    prec = inv_covGP,
    num_jitters = covGP_jittered[[2]],
    prec_heatmap = inv_covGP_heatmap,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model,
    marg_fitted_values = marginal_fvs
  )
  
  return(kgr_model1_results)
}

#Fit kgr_model1
kgr_model1_fit = kgr_model1(dataset = inla_train_dataset, rho_time_rbf = 536.519, rho_time_periodic = 5.707,sigma2_time = 2.773)

#Extract DIC and WAIC
kgr_model1_DIC = kgr_model1_fit$model_DIC
kgr_model1_WAIC = kgr_model1_fit$model_WAIC

#Get summaries of parameter estimates
kgr_model1_fit$model_summary
kgr_model1_fit$bri_hyperpar_summary
kgr_model1_fit$exp_effects
kgr_model1_fit$K_time_weight
kgr_model1_fit$gfilter_weight
kgr_model1_fit$num_jitters

#Show plots
kgr_model1_fit$K_time_heatmap
kgr_model1_fit$prec_heatmap
kgr_model1_fit$param_plot
kgr_model1_fit$hyperparam_plot
pp_insample_plot(pred_data = kgr_model1_fit$fitted_values)
```

## KGR model 2

$\textbf{Y} | \boldsymbol{\Lambda}, \boldsymbol{\beta} , \textbf{F}, \textbf{Z}, \textbf{X}_{EPA}, \textbf{S} \sim Poisson(\boldsymbol{\Lambda})$

$\Lambda_{i,t} | \textbf{F}, \textbf{Z}, \textbf{X}_{EPA} = \exp(\beta_{c1} I \{ c=1 \} + \beta_{c2} I \{ c=2 \} + ... + \beta_{c5} I \{ c=5 \} + \beta_1 I \{t \, mod \, 12 = 1 \} + ... + \beta_{11} I \{t \, mod \, 12 = 11 \} + \beta_{12} \textbf{Z} + \boldsymbol{\beta}^T \textbf{X}_{EPA} + \textbf{F}_{i,t})$

$\textbf{F} | \textbf{S}, \rho_{rbf}, \rho_{p}, \sigma^2_{Z} \sim \mathcal{GP}(\textbf{0},\textbf{K}_{Z} \otimes \textbf{H}^2)$

$Cov(\textbf{F}_{i,t_1},\textbf{F}_{j,t_2}) = \left[ \textbf{K}_{Z} \right]_{t_1,t_2} \left[ \textbf{H}^2 \right]_{i,j}$

```{r}
kgr_model2 = function(dataset, rho_Z_rbf = 1, rho_Z_periodic = 1, sigma2_Z = 1,link=1){
  
  #Calculate gram matrix K_Z
  K_Z = treatment_kernel(unique_dates = unique(dataset$date),
                     rho_rbf = rho_Z_rbf, rho_periodic = rho_Z_periodic, sigma2 = sigma2_Z)
  
  #Heatmap of resulting K 
  K_Z_heatmap = matrix_heatmap(K_Z,title = "K_Z Heatmap")
  
  #Calculate trace norm of gram matrix
  K_Z_weight = norm((1/60)*K_Z,type = "F")
  
  ###Load graph regression kernel 
  covGP2 = kronecker(K_Z/60,(H^2)/7)
  
  #Need to ensure precision matrix is not computationally singular i.e det > 0
  covGP_jittered = desingularize(covGP2,threshold = 1e-2,increment = 0.01)
  covGP2 = covGP_jittered[[1]]
  inv_covGP2 = solve(covGP2)
  # cov_Fnorm = norm(covGP2,type = "F")
  
  #Heatmap of resulting K 
  inv_covGP2_heatmap = matrix_heatmap(inv_covGP2,title = "")
  
  ###Fit INLA model 
  kgr_formula2 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP2)
  
  # kgr_formula2_nonlinear = deaths ~ -1 + month + ns(b3_wildfire_pm25,df=3) + b3_AQI + b3_lead + b3_co + 
  #   b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
  #   Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP2)
  
  
  model = inla(formula = kgr_formula2,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  marginal_fvs <- model$marginals.fitted.values

  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  kgr_model2_results = list(
    K_Z_heatmap = K_Z_heatmap,
    K_Z_weight = K_Z_weight/(K_Z_weight + gfilter_weight),
    gfilter_weight = gfilter_weight/(K_Z_weight + gfilter_weight),
    covmatrix = covGP2,
    prec = inv_covGP2,
    num_jitters = covGP_jittered[[2]],
    prec_heatmap = inv_covGP2_heatmap,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model,
    marg_fitted_values = marginal_fvs
  )
  
  return(kgr_model2_results)
}

#Fit kgr_model2
kgr_model2_fit = kgr_model2(dataset = inla_train_dataset,rho_Z_rbf = 0.0020604943,
                            rho_Z_periodic = 0.0024283451, sigma2_Z = 3.941096)

#Extract DIC and WAIC 
kgr_model2_DIC = kgr_model2_fit$model_DIC
kgr_model2_WAIC = kgr_model2_fit$model_WAIC

#Get summaries of parameter estimates
kgr_model2_fit$model_summary
kgr_model2_fit$bri_hyperpar_summary
kgr_model2_fit$exp_effects
kgr_model2_fit$K_Z_weight
kgr_model2_fit$gfilter_weight
kgr_model2_fit$num_jitters

#Show plots
kgr_model2_fit$prec_heatmap
kgr_model2_fit$K_Z_heatmap
kgr_model2_fit$param_plot
kgr_model2_fit$hyperparam_plot
pp_insample_plot(pred_data = kgr_model2_fit$fitted_values)
```

## KGR model 3

$\textbf{Y} | \boldsymbol{\Lambda}, \boldsymbol{\beta} , \textbf{F}, \textbf{Z}, \textbf{X}_{EPA}, \textbf{S} \sim Poisson(\boldsymbol{\Lambda})$

$\Lambda_{i,t} | \textbf{F}, \textbf{Z}, \textbf{X}_{EPA} = \exp(\beta_{c1} I \{ c=1 \} + \beta_{c2} I \{ c=2 \} + ... + \beta_{c5} I \{ c=5 \} + \beta_1 I \{t \, mod \, 12 = 1 \} + ... + \beta_{11} I \{t \, mod \, 12 = 11 \} + \beta_{12} \textbf{Z} + \boldsymbol{\beta}^T \textbf{X}_{EPA} + \textbf{F}_{i,t})$

$\textbf{F} | \textbf{S}, \rho^{time}_{rbf}, \rho^{time}_{p}, \rho^{EPA}_{rbf}, \rho^{EPA}_{p}, \sigma^2 \sim \mathcal{GP}(\textbf{0},(\textbf{K}_{time} \odot \textbf{K}_{EPA}) \otimes \textbf{H}^2)$

$Cov(\textbf{F}_{i,t_1},\textbf{F}_{j,t_2}) = \left[ \textbf{K}^{time} \right]_{t_1,t_2} \left[ \textbf{K}^{EPA} \right]_{t_1,t_2} \left[ \textbf{H}^2 \right]_{i,j}$

```{r}
kgr_model3 = function(dataset,rho_Z_rbf = 1,rho_Z_periodic = 1,rho_time_rbf = 1,rho_time_periodic = 1,sigma2 = 1,link=1){
  
  ###Calculating gram matrix K_EPA
  # K_EPA = EPA_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_EPA_rbf,
  #                    rho_periodic = rho_EPA_periodic,sigma2 = 1)

  K_Z = treatment_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_Z_rbf,
                     rho_periodic = rho_Z_periodic,sigma2 = 1)
  
  #Heatmap of resulting K 
  K_Z_heatmap = matrix_heatmap(K_Z,title = "K_Z heatmap")
  
  ###Calculating gram matrix K_time
 K_time = time_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_time_rbf, 
                      rho_periodic = rho_time_periodic, sigma2 = sigma2)
  
  #Heatmap of resulting K 
  K_time_heatmap = matrix_heatmap(K_time,title = "K_time heatmap")
  
  K_Z_norm = norm(K_Z,type = "F")
  K_time_norm = norm(K_time,type = "F")
  
  #Calculate trace norm of gram matrix
  gram = (K_Z*K_time)/sigma2
  K_weight = norm((1/length(unique(dataset$date)))*gram,type = "F")
  
  
  ###Load graph regression kernel 
  # covGP3 = kronecker(gram,H^2)
  covGP3 = kronecker(gram/length(unique(dataset$date)),(H^2)/num_clus)
  
  #Need to ensure precision matrix is not computationally singular i.e det > 0
  covGP_jittered = desingularize(covGP3,threshold = 1e-2,increment = 0.01)
  covGP3 = covGP_jittered[[1]]
  
  inv_covGP3 = solve(covGP3)
  
  #Heatmap of resulting K 
  inv_covGP3_heatmap = matrix_heatmap(inv_covGP3,title = "")
  
  ###Fit INLA model 
  # kgr_formula3 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
  #   b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
  #   Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP3)

  kgr_formula3_nonlinear = deaths ~ -1 + month + ns(b3_wildfire_pm25,df=3) + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP3)
  
  model = inla(formula = kgr_formula3_nonlinear,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  marginal_fvs <- model$marginals.fitted.values
  
  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  kgr_model3_results = list(
    K_Z_heatmap = K_Z_heatmap,
    K_time_heatmap = K_time_heatmap,
    K_Z_weight = K_Z_norm / (K_Z_norm + K_time_norm),
    K_time_weight = K_time_norm / (K_Z_norm + K_time_norm),
    K_weight = K_weight/(K_weight + gfilter_weight),
    gfilter_weight = gfilter_weight/(K_weight + gfilter_weight),
    covmatrix = covGP3,
    prec = inv_covGP3,
    num_jitters = covGP_jittered[[2]],
    prec_heatmap = inv_covGP3_heatmap,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model,
    marg_fitted_values = marginal_fvs
  )
  
  return(kgr_model3_results)
}

#Fit kgr_model3
# kgr_model3_fit = kgr_model3(dataset = inla_train_dataset, 
#                             rho_EPA_rbf = 280.607, 
#                             rho_EPA_periodic = 528.775,
#                             rho_time_rbf = 554.262, 
#                             rho_time_periodic = 543.737,
#                             sigma2 = 3.797)

kgr_model3_fit = kgr_model3(dataset = inla_train_dataset, 
                            rho_Z_rbf = 37801.48, 
                            rho_Z_periodic = 571512.56,
                            rho_time_rbf = 39.5276, 
                            rho_time_periodic = 380794.8,
                            sigma2 = 1.124888)

#Extract DIC and WAIC
kgr_model3_DIC = kgr_model3_fit$model_DIC
kgr_model3_WAIC = kgr_model3_fit$model_WAIC

#Get summaries of parameter estimates
kgr_model3_fit$model_summary
kgr_model3_fit$bri_hyperpar_summary
kgr_model3_fit$exp_effects
kgr_model3_fit$K_Z_weight
kgr_model3_fit$K_time_weight
kgr_model3_fit$K_weight
kgr_model3_fit$gfilter_weight
kgr_model3_fit$num_jitters

#Show plots
kgr_model3_fit$K_time_heatmap
kgr_model3_fit$K_Z_heatmap
kgr_model3_fit$prec_heatmap
kgr_model3_fit$param_plot
kgr_model3_fit$hyperparam_plot
pp_insample_plot(pred_data = kgr_model3_fit$fitted_values)
```

## KGR model 4

$\textbf{Y} | \boldsymbol{\Lambda}, \boldsymbol{\beta} , \textbf{F}, \textbf{Z}, \textbf{X}_{EPA}, \textbf{S} \sim Poisson(\boldsymbol{\Lambda})$

$\Lambda_{i,t} | \textbf{F}, \textbf{Z}, \textbf{X}_{EPA} = \exp(\beta_{c1} I \{ c=1 \} + \beta_{c2} I \{ c=2 \} + ... + \beta_{c5} I \{ c=5 \} + \beta_1 I \{t \, mod \, 12 = 1 \} + ... + \beta_{11} I \{t \, mod \, 12 = 11 \} + \beta_{12} \textbf{Z} + \boldsymbol{\beta}^T \textbf{X}_{EPA} + \textbf{F}_{i,t})$

$\textbf{F} | \textbf{S}, \rho^{time}_{rbf}, \rho^{time}_{p}, \rho^{Z}_{rbf}, \rho^{Z}_{p}, \sigma^2_{time}, \sigma^2_Z \sim \mathcal{GP}(\textbf{0},(\textbf{K}_{time} + \textbf{K}_{Z}) \otimes \textbf{H}^2)$

$Cov(\textbf{F}_{i,t_1},\textbf{F}_{j,t_2}) = (\left[ \textbf{K}^{time} \right]_{t_1,t_2} + \left[ \textbf{K}^{Z} \right]_{t_1,t_2}) \left[ \textbf{H}^2 \right]_{i,j}$

```{r}
kgr_model4 = function(dataset,rho_Z_rbf = 1,rho_Z_periodic = 1,rho_time_rbf = 1,rho_time_periodic = 1,sigma2_Z = 1,sigma2_time = 1,link=1){
  
  ###Calculating gram matrix K_EPA
  # K_EPA = EPA_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_EPA_rbf,
  #                    rho_periodic = rho_EPA_periodic,sigma2 = 1)

  K_Z = treatment_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_Z_rbf,
                     rho_periodic = rho_Z_periodic,sigma2 = sigma2_Z)
  
  #Heatmap of resulting K 
  K_Z_heatmap = matrix_heatmap(K_Z,title = "K_Z heatmap")
  
  ###Calculating gram matrix K_time
  K_time = time_kernel(unique_dates = unique(dataset$date),rho_rbf = rho_time_rbf, 
                      rho_periodic = rho_time_periodic, sigma2 = sigma2_time)
  
  #Heatmap of resulting K 
  K_time_heatmap = matrix_heatmap(K_time,title = "K_time heatmap")
  
  K_Z_norm = norm(K_Z,type = "F")
  K_time_norm = norm(K_time,type = "F")
  
  #Calculate trace norm of gram matrix
  gram = (K_Z + K_time)*0.5
  K_weight = norm((1/length(unique(dataset$date)))*gram,type = "F")
  
  
  ###Load graph regression kernel 
  # covGP4 = kronecker(gram,H^2)
  covGP4 = kronecker(gram/length(unique(dataset$date)),(H^2)/num_clus)
  
  #Need to ensure precision matrix is not computationally singular i.e det > 0
  covGP_jittered = desingularize(covGP4,threshold = 1e-2,increment = 0.01)
  covGP4 = covGP_jittered[[1]]
  
  inv_covGP4 = solve(covGP4)
  
  #Heatmap of resulting K 
  inv_covGP4_heatmap = matrix_heatmap(inv_covGP4,title = "")
  
  ###Fit INLA model 
  kgr_formula4 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP4)

  # kgr_formula4_nonlinear = deaths ~ -1 + month + ns(b3_wildfire_pm25,df=3) + b3_AQI + b3_lead + b3_co + 
  #   b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
  #   Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP4)
  
  model = inla(formula = kgr_formula4,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  marginal_fvs <- model$marginals.fitted.values
  
  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  kgr_model4_results = list(
    K_Z_heatmap = K_Z_heatmap,
    K_time_heatmap = K_time_heatmap,
    K_Z_weight = K_Z_norm / (K_Z_norm + K_time_norm),
    K_time_weight = K_time_norm / (K_Z_norm + K_time_norm),
    K_weight = K_weight/(K_weight + gfilter_weight),
    gfilter_weight = gfilter_weight/(K_weight + gfilter_weight),
    covmatrix = covGP4,
    prec = inv_covGP4,
    num_jitters = covGP_jittered[[2]],
    prec_heatmap = inv_covGP4_heatmap,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model,
    marg_fitted_values = marginal_fvs
  )
  
  return(kgr_model4_results)
}

#Fit kgr_model4
# kgr_model4_fit = kgr_model4(dataset = inla_train_dataset, 
#                             rho_EPA_rbf = 280.607, 
#                             rho_EPA_periodic = 528.775,
#                             rho_time_rbf = 554.262, 
#                             rho_time_periodic = 543.737,
#                             sigma2 = 3.797)

kgr_model4_fit = kgr_model4(dataset = inla_train_dataset, 
                            rho_Z_rbf = 0.0008842497,
                            rho_Z_periodic = 0.0012119621,
                            rho_time_rbf = 0.0037851329,
                            rho_time_periodic = 0.003501742,
                            sigma2_Z = 4.310359,
                            sigma2_time = 0.9296752, link=1)

#Extract DIC and WAIC
kgr_model4_DIC = kgr_model4_fit$model_DIC
kgr_model4_WAIC = kgr_model4_fit$model_WAIC

#Get summaries of parameter estimates
kgr_model4_fit$model_summary
kgr_model4_fit$bri_hyperpar_summary
kgr_model4_fit$exp_effects
kgr_model4_fit$K_Z_weight
kgr_model4_fit$K_time_weight
kgr_model4_fit$K_weight
kgr_model4_fit$gfilter_weight
kgr_model4_fit$num_jitters

#Show plots
kgr_model4_fit$K_time_heatmap
kgr_model4_fit$K_Z_heatmap
kgr_model4_fit$prec_heatmap
kgr_model4_fit$param_plot
kgr_model4_fit$hyperparam_plot
pp_insample_plot(pred_data = kgr_model4_fit$fitted_values)
```

## KGR model 5

### Set up AR(1), DL(3,6,12), and Interaction kernel structures

```{r}
EPA_county_list = list()

county_EPA_data = function(county){
  EPA_county_ts = matrix(nrow=length(unique(final_EPA_data$date)))

  for (j in 4:10){
    covariate_ts = final_EPA_data %>% filter(county_labels == county)
    EPA_county_ts = cbind(EPA_county_ts,covariate_ts[,j])
  }
  EPA_county_ts[,1] = covariate_ts$AQI
  colnames(EPA_county_ts) = c("AQI","Lead","CO","SO2","NO2","O3","PM10","PM2.5")
  
  return(EPA_county_ts)
}

for (i in 1:num_counties){
  EPA_county_list[[i]] = county_EPA_data(county_ids$county[i])
}
```


```{r}
wildfire_pm25_county_list = list()

county_wildfire_pm25_data = function(id){ 
  covariate_ts = b3_inla_dataset %>% filter(county_id == id)
  wildfire_county_ts = cbind(covariate_ts$wildfire_pm25,
  covariate_ts$b3_wildfire_pm25)
  
  colnames(wildfire_county_ts) = c("wildfire_pm25","b3_wildfire_pm25")
  
  return(wildfire_county_ts)
}

for (i in 1:num_counties) {
  wildfire_pm25_county_list[[i]] = county_wildfire_pm25_data(i)
}
```

```{r}
decompose_wildfire_ts = function(wildfire_ts) {
  S_scaled = wildfire_ts
  S_scaled = scale(wildfire_ts)
  colnames(S_scaled) = c("wildfire_pm25","b3_wildfire_pm25")
  S_scaled_ts = ts(S_scaled, frequency = 12)
  
  ts_decomposed <- lapply(colnames(S_scaled_ts), function(x) {stl(S_scaled_ts[, x], s.window = "periodic")})
  names(ts_decomposed) = colnames(S_scaled_ts)

  S_seasonal = ts_decomposed[[1]]$time.series[, "seasonal"]
  S_trend = ts_decomposed[[1]]$time.series[, "trend"]
  S_random = ts_decomposed[[1]]$time.series[, "remainder"]

  S_seasonal = cbind(S_seasonal,ts_decomposed[[2]]$time.series[, "seasonal"])
  S_trend = cbind(S_trend,ts_decomposed[[2]]$time.series[, "trend"])
  S_random = cbind(S_random,ts_decomposed[[2]]$time.series[, "remainder"])

  colnames(S_seasonal)= colnames(S_scaled_ts)
  colnames(S_trend)= colnames(S_scaled_ts)
  colnames(S_random)= colnames(S_scaled_ts)

  S_DL = S_seasonal + S_random
  S_DL = data.frame(S_DL)
 
  W = matrix(nrow=(nrow(S_random))^2)
  num_cols = ncol(S_random)
  col_num = 2
  
  for (i in 1:num_cols) {
    for (j in 1:num_cols) {
      interaction_col = kronecker(S_random[,i], S_random[,j])
      W = cbind(W, interaction_col)
      colnames(W)[col_num] = sprintf("%sx%s", colnames(S_scaled)[i], colnames(S_scaled)[j])
      col_num = col_num + 1
    }
  }
  W = W[,3]
  
  list = list(S_seasonal, S_trend, S_random, S_DL, W)
  names(list) = c("S_seasonal", "S_trend", "S_random", "S_DL", "W")
  return(list)
}

decomposed_wildfire_county_data = lapply(wildfire_pm25_county_list, decompose_wildfire_ts)
```

```{r}
S_random_all_wildfire = do.call(cbind, lapply(decomposed_wildfire_county_data, function(x) x$S_random))
S_DL_all_wildfire = do.call(cbind, lapply(decomposed_wildfire_county_data, function(x) x$S_DL))
W_all_wildfire = do.call(cbind, lapply(decomposed_wildfire_county_data, function(x) x$W))
```

AR(1) structure: 

```{r}
K_AR_periodic_county = list()

for (c in 1:num_counties){
  
  #Grab S_random data for county c
  county_data = decomposed_wildfire_county_data[[c]]
  S_random_county = county_data$S_random
  
  #Create a list to contain covariance matrix for each col (should I get rid of the for loop and just do wildfire_pm2.5)
  K_AR_periodic_list = list()
  
  rho_AR = 1
  sigma2_AR = 1
  
  time_span = nrow(S_random_county)
  
  #Calculate a AR 1 covariance matrix for each col and store in list
  for (i in 1:2){
    ts =  S_random_county[,i]
    K_covariate_periodic = matrix(nrow=time_span,ncol=time_span)
    
    for(j in 1:time_span){
      for (k in 1:time_span){
        if (abs(j-k) <= 1){
          K_covariate_periodic[j,k] = exp(- ((ts[j] - ts[k])^2) / (2*rho_AR)) * exp(- (2*sin((abs(ts[j] - ts[k]))*pi/12)^2)
                       / (rho_AR)) * sigma2_AR
        }
        else{
          K_covariate_periodic[j,k] = 0
          }
      }
    }
    K_AR_periodic_list[[i]] = K_covariate_periodic
  }

  names(K_AR_periodic_list) = colnames(S_random_county)
  
  #Add each pollutant's covariance matrix to get AR 1 matrix for each county
  K_AR_periodic = (1/2)*(K_AR_periodic_list[[1]] + K_AR_periodic_list[[2]])
  K_AR_periodic_weights = c(norm(K_AR_periodic_list[[1]],type = "F"),
  norm(K_AR_periodic_list[[2]],type = "F"))

  # K_AR_periodic = matrix(0,nrow=time_span,ncol=time_span)
  
  # K_AR_periodic_weights = rep(1,length(K_AR_periodic_list))
  
  # for(i in 1:length(K_AR_periodic_list)){
  #   K_AR_periodic = K_AR_periodic + ((1/2)*K_AR_periodic_list[[i]])
    
  #   K_AR_periodic_weights[i] = norm(K_AR_periodic_list[[i]],type = "F")
  # }
  
  K_AR_periodic_weights = K_AR_periodic_weights / sum(K_AR_periodic_weights)
  print(K_AR_periodic_weights)

  K_AR_periodic_county[[c]] = K_AR_periodic
}
```

Combining AR kernels for each county together: 

```{r}
K_AR_periodic = matrix(0,nrow=time_span,ncol=time_span)

K_AR_periodic_weights = rep(1,num_counties)

for(i in 1:num_counties){
  K_AR_periodic = K_AR_periodic + ((1/num_counties)*K_AR_periodic_county[[i]])
  
  K_AR_periodic_weights[i] = norm(K_AR_periodic_county[[i]],type = "F")
}

K_AR_periodic_weights = K_AR_periodic_weights / sum(K_AR_periodic_weights)

print(K_AR_periodic_weights)

#Heatmap of resulting K 
matrix_heatmap(K_AR_periodic,title = "Periodic AR 1 Covariance Structure")
```

DL(3,6,12) structure 

```{r}
K_DL_periodic_county = list()

for (c in 1:num_counties){
  
  #Grab S_DL data for county c
  county_data = decomposed_wildfire_county_data[[c]]
  S_DL_county = county_data$S_DL #has original and backshifted cols if needed
  S_DL_county = S_DL_county[,1] #only take original ts for now
  
  #Create a list to store covariance matrix for each DL 
  K_DL_periodic_list = list()
  
  dl_lags = c(3,6,12)
  tracker = 1
  
  for (i in dl_lags){
    
    K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
    
    rho_DL = 1
    sigma2_DL = 1
    
    #Calculate DL covariance matrix for specified lag   
    for(j in 1:length(S_DL_county)){
      for (k in 1:length(S_DL_county)){
        
        if ((abs(j-k) == 0) || (abs(j-k) == i)){
          K_DL_periodic[j,k] = exp(- (sum(S_DL_county[j] - S_DL_county[k])^2)
                               / (2*rho_DL)) * exp(- (2*sin(sum(abs(S_DL_county[j] - S_DL_county[k]))*pi/12)^2)
                               / (rho_DL)) * sigma2_DL
        } 
        else{
          K_DL_periodic[j,k] = 0
          }
      }
    }
    K_DL_periodic_list[[tracker]] = K_DL_periodic
    tracker = tracker+1
  }
  
  #Combine the 3 DL covariance matrices together
  K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
  
  K_DL_periodic_weights = rep(1,length(K_DL_periodic_list))
  
  for(i in 1:length(K_DL_periodic_list)){
    K_DL_periodic = K_DL_periodic + ((1/3)*K_DL_periodic_list[[i]])
    K_DL_periodic_weights[i] = norm(K_DL_periodic_list[[i]],type = "F")
  }
  
  K_DL_periodic_weights = K_DL_periodic_weights / sum(K_DL_periodic_weights)
  print(K_DL_periodic_weights)
  
  #Store DL(3,6,12) covariance matrix for each county 
  K_DL_periodic_county[[c]] = K_DL_periodic
}
```

Combining DL kernels for each county together: 

```{r}
K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
K_DL_periodic_weights = rep(1,num_counties)

for(i in 1:num_counties){
  K_DL_periodic = K_DL_periodic + ((1/num_counties)*K_DL_periodic_county[[i]])
  K_DL_periodic_weights[i] = norm(K_DL_periodic_county[[i]],type = "F")
}

K_DL_periodic_weights = K_DL_periodic_weights / sum(K_DL_periodic_weights)
print(K_DL_periodic_weights)

matrix_heatmap(K_DL_periodic,title = "Periodic DL (3,6,12) Covariance Structure")
```

Interaction structure: 

```{r}
K_Interaction_periodic_county = list()

for (c in 1:num_counties){
  
  #Grab interaction pair data for county c
  county_data = decomposed_wildfire_county_data[[c]]
  W_county = county_data$W
  time_span = length(W_county)

  K_interaction_periodic_list = list()
  
  #Create sequence of indices corresponding to comparisons for real time and one lag interaction effects
  idx = seq(1,time_span,by=70)
  
  #Calculate a kernel for each interaction pair  
  K_int0_periodic = matrix(nrow = sqrt(time_span),ncol = sqrt(time_span))
  
  rho_int = 1
  sigma2_int = 1
  
  for (i in 1:sqrt(time_span)){
    for (j in 1:sqrt(time_span)){       
      #Locally periodic kernels 
      K_int0_periodic[i,j] = exp(- ((W_county[idx[i]] - W_county[idx[j]])^2) / (2*rho_int)) * 
      exp(- (2*sin((abs(W_county[idx[i]] - W_county[idx[j]]))*pi/12)^2) / (rho_int)) * sigma2_int
    }
  }
  
  # matrix_heatmap(K_int0_periodic,title = "Periodic Interaction Covariance Structure")

  K_Interaction_periodic_county[[c]] = K_int0_periodic
  
  ###DON'T NEED IF WE ONLY HAVE ONE INTERACTION PAIR
  # #Combine kernels for each interaction pair together
  # K_interaction_periodic = matrix(0,nrow=60,ncol=60)
  
  # K_interaction_periodic_weights = rep(1,length(K_interaction_periodic_list))
  
  # for(i in 1:length(K_interaction_periodic_list)){
  #   K_interaction_periodic = K_interaction_periodic + ((1/length(K_interaction_periodic_list))*K_interaction_periodic_list[[i]])
    
  #   K_interaction_periodic_weights[i] = norm(K_interaction_periodic_list[[i]],type = "F")
  # }
  
  # K_interaction_periodic_weights = K_interaction_periodic_weights / sum(K_interaction_periodic_weights)
  # print(K_interaction_periodic_weights)

  # matrix_heatmap(K_interaction_periodic,title = "Periodic Interaction Covariance Structure")
  
  # #Store final interaction kernel (for all pairs) for each cluster 
  # K_Interaction_periodic_cluster[[c]] = K_interaction_periodic

}
```

Combining interaction kernels from each cluster together:

```{r}
K_interaction_periodic = matrix(0,nrow=sqrt(time_span),ncol=sqrt(time_span))
K_interaction_periodic_weights = rep(1,num_counties)

for(i in 1:num_counties){
  K_interaction_periodic = K_interaction_periodic + ((1/num_counties)*K_Interaction_periodic_county[[i]])

  K_interaction_periodic_weights[i] = norm(K_Interaction_periodic_county[[i]],type = "F")
}

K_interaction_periodic_weights = K_interaction_periodic_weights / sum(K_interaction_periodic_weights)
print(K_interaction_periodic_weights)

matrix_heatmap(K_interaction_periodic,title = "Periodic Interaction Covariance Structure")
```

### Fit the model

```{r,warning=FALSE}
kgr_model5 = function(dataset, rho_AR_rbf = 1, rho_AR_periodic = 1, rho_DL_rbf = 1, rho_DL_periodic = 1, rho_int_rbf = 1, rho_int_periodic = 1, sigma2_AR = 1, sigma2_DL = 1, sigma2_int = 1, link = 1){

  ###Calculating gram matrix K_AR
  K_AR_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab S_random data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    S_random_county = county_data$S_random
    
    #Create a list to contain covariance matrix for each col (should I get rid of the for loop and just do wildfire_pm2.5)
    K_AR_periodic_list = list()
    
    rho_AR = 1
    sigma2_AR = 1
    
    time_span = nrow(S_random_county)
    
    #Calculate a AR 1 covariance matrix for each col and store in list
    for (i in 1:2){
      ts =  S_random_county[,i]
      K_covariate_periodic = matrix(nrow=time_span,ncol=time_span)
      
      for(j in 1:time_span){
        for (k in 1:time_span){
          if (abs(j-k) <= 1){
            K_covariate_periodic[j,k] = exp(- ((ts[j] - ts[k])^2) / (2*rho_AR)) * exp(- (2*sin((abs(ts[j] - ts[k]))*pi/12)^2)
                        / (rho_AR)) * sigma2_AR
          }
          else{
            K_covariate_periodic[j,k] = 0
            }
        }
      }
      K_AR_periodic_list[[i]] = K_covariate_periodic
    }

    names(K_AR_periodic_list) = colnames(S_random_county)
    
    #Add each pollutant's covariance matrix to get AR 1 matrix for each county
    K_AR_periodic = (1/2)*(K_AR_periodic_list[[1]] + K_AR_periodic_list[[2]])
    K_AR_periodic_weights = c(norm(K_AR_periodic_list[[1]],type = "F"),
    norm(K_AR_periodic_list[[2]],type = "F"))

    # K_AR_periodic = matrix(0,nrow=time_span,ncol=time_span)
    
    # K_AR_periodic_weights = rep(1,length(K_AR_periodic_list))
    
    # for(i in 1:length(K_AR_periodic_list)){
    #   K_AR_periodic = K_AR_periodic + ((1/2)*K_AR_periodic_list[[i]])
      
    #   K_AR_periodic_weights[i] = norm(K_AR_periodic_list[[i]],type = "F")
    # }
    
    K_AR_periodic_weights = K_AR_periodic_weights / sum(K_AR_periodic_weights)
    K_AR_periodic_county[[c]] = K_AR_periodic
  }

  K_AR_periodic = matrix(0,nrow=time_span,ncol=time_span)

  K_AR_periodic_weights = rep(1,num_counties)

  for(i in 1:num_counties){
    K_AR_periodic = K_AR_periodic + ((1/num_counties)*K_AR_periodic_county[[i]])
    
    K_AR_periodic_weights[i] = norm(K_AR_periodic_county[[i]],type = "F")
  }

  K_AR_periodic_weights = K_AR_periodic_weights / sum(K_AR_periodic_weights)
  K_AR_norm = norm(K_AR_periodic,type = "F")
  K_AR_heatmap = matrix_heatmap(K_AR_periodic,title = "Periodic AR 1 Covariance Structure")

  ###Calculating gram matrix K_DL
  K_DL_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab S_DL data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    S_DL_county = county_data$S_DL #has original and backshifted cols if needed
    S_DL_county = S_DL_county[,1] #only take original ts for now
    
    #Create a list to store covariance matrix for each DL 
    K_DL_periodic_list = list()
    
    dl_lags = c(3,6,12)
    tracker = 1
    
    for (i in dl_lags){
      
      K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
      
      rho_DL = 1
      sigma2_DL = 1
      
      #Calculate DL covariance matrix for specified lag   
      for(j in 1:length(S_DL_county)){
        for (k in 1:length(S_DL_county)){
          
          if ((abs(j-k) == 0) || (abs(j-k) == i)){
            K_DL_periodic[j,k] = exp(- (sum(S_DL_county[j] - S_DL_county[k])^2)
                                / (2*rho_DL)) * exp(- (2*sin(sum(abs(S_DL_county[j] - S_DL_county[k]))*pi/12)^2)
                                / (rho_DL)) * sigma2_DL
          } 
          else{
            K_DL_periodic[j,k] = 0
            }
        }
      }
      K_DL_periodic_list[[tracker]] = K_DL_periodic
      tracker = tracker+1
    }
    
    #Combine the 3 DL covariance matrices together
    K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
    
    K_DL_periodic_weights = rep(1,length(K_DL_periodic_list))
    
    for(i in 1:length(K_DL_periodic_list)){
      K_DL_periodic = K_DL_periodic + ((1/3)*K_DL_periodic_list[[i]])
      K_DL_periodic_weights[i] = norm(K_DL_periodic_list[[i]],type = "F")
    }
    
    K_DL_periodic_weights = K_DL_periodic_weights / sum(K_DL_periodic_weights)
   
    #Store DL(3,6,12) covariance matrix for each county 
    K_DL_periodic_county[[c]] = K_DL_periodic
  }

  K_DL_periodic = matrix(0,nrow=time_span,ncol=time_span)
  K_DL_periodic_weights = rep(1,num_counties)

  for(i in 1:num_counties){
    K_DL_periodic = K_DL_periodic + ((1/num_counties)*K_DL_periodic_county[[i]])
    K_DL_periodic_weights[i] = norm(K_DL_periodic_county[[i]],type = "F")
  }

  K_DL_periodic_weights = K_DL_periodic_weights / sum(K_DL_periodic_weights)
  K_DL_norm = norm(K_DL_periodic,type = "F")
  K_DL_heatmap = matrix_heatmap(K_DL_periodic,title = "Periodic DL (3,6,12) Covariance Structure")
  
  ###Calculating gram matrix K_Interaction
  K_Interaction_periodic_county = list()

  for (c in 1:num_counties){
    
    #Grab interaction pair data for county c
    county_data = decomposed_wildfire_county_data[[c]]
    W_county = county_data$W
    time_sq = length(W_county)

    K_interaction_periodic_list = list()
    
    #Create sequence of indices corresponding to comparisons for real time and one lag interaction effects
    idx = seq(1,time_sq,by=70)
    
    #Calculate a kernel for each interaction pair  
    K_int0_periodic = matrix(nrow = time_span,ncol = time_span)
    
    rho_int = 1
    sigma2_int = 1
    
    for (i in 1:time_span){
      for (j in 1:time_span){       
        #Locally periodic kernels 
        K_int0_periodic[i,j] = exp(- ((W_county[idx[i]] - W_county[idx[j]])^2) / (2*rho_int)) * 
        exp(- (2*sin((abs(W_county[idx[i]] - W_county[idx[j]]))*pi/12)^2) / (rho_int)) * sigma2_int
      }
    }
    
    # matrix_heatmap(K_int0_periodic,title = "Periodic Interaction Covariance Structure")

    K_Interaction_periodic_county[[c]] = K_int0_periodic
  }

  matrix_heatmap(K_int0_periodic,title = "Periodic Interaction Covariance Structure")

  K_interaction_periodic = matrix(0,nrow=time_span,ncol=time_span)
  K_interaction_periodic_weights = rep(1,num_counties)

  for(i in 1:num_counties){
    K_interaction_periodic = K_interaction_periodic + ((1/num_counties)*K_Interaction_periodic_county[[i]])

    K_interaction_periodic_weights[i] = norm(K_Interaction_periodic_county[[i]],type = "F")
  }

  K_interaction_periodic_weights = K_interaction_periodic_weights / sum(K_interaction_periodic_weights)
  K_Int_norm = norm(K_interaction_periodic,type = "F")
  K_Interaction_heatmap = matrix_heatmap(K_interaction_periodic,title = "Periodic Interaction Covariance Structure")
  
  gram = (1/3)*(K_AR_periodic+K_DL_periodic+K_interaction_periodic)
  K_weight = norm((1/time_span)*gram,type = "F")
  
  ###Load graph regression kernel 
  # covGP5 = kronecker(gram,H^2)
  covGP5 = kronecker(gram/time_span,(H^2)/num_clus)
  
  #Need to ensure precision matrix is not computationally singular i.e det > 0
  covGP_jittered = desingularize(covGP5,threshold = 1e-2,increment = 0.01)
  covGP5 = covGP_jittered[[1]]
  
  inv_covGP5 = solve(covGP5)
  
  #Heatmap of resulting K 
  inv_covGP5_heatmap = matrix_heatmap(inv_covGP5,title = "")
  
  ###Fit INLA model 
  # kgr_formula5 = deaths ~ -1 + month + b3_wildfire_pm25 + b3_AQI + b3_lead + b3_co + 
  #   b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
  #   Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP3)

  kgr_formula5_nonlinear = deaths ~ -1 + month + ns(b3_wildfire_pm25,df=3) + b3_AQI + b3_lead + b3_co + 
    b3_so2 + b3_no2 + b3_o3 + b3_pm10 + Intercept1 + Intercept2 + Intercept3 + 
    Intercept4 + Intercept5 + f(idx,model = "generic0",Cmatrix = inv_covGP5)
  
  model = inla(formula = kgr_formula5_nonlinear,family = "poisson",data = dataset,
                  control.compute = list(dic=TRUE,waic=TRUE,
                                         return.marginals.predictor=TRUE),
                  control.inla = list(strategy = "laplace"),
                  control.predictor = list(compute = TRUE, link = link))
  
  ###Extract relevant information and store in the list
  model_summary <- model$summary.fixed
  bri_hyperpar_summary <- bri.hyperpar.summary(model)
  model_DIC <- model$dic$dic
  model_WAIC <- model$waic$waic
  preds_model <- model$summary.fitted.values
  preds_model <- cbind(dataset$county, dataset$date, dataset$county_id, dataset$b3_wildfire_pm25, preds_model)
  colnames(preds_model) <- c("county", "date", "county_id", "b3-wildfire-pm25", "mean", "sd", "0.025quant", "0.5quant", "0.975quant", "mode")
  marginal_fvs <- model$marginals.fitted.values
  
  #Exponentiating parameter to get better interpretation of estimates 
  multeff <- exp(model$summary.fixed$mean)
  names(multeff) <- model$names.fixed
  
  #Plot of each parameters' posterior density 
  mf <- melt(model$marginals.fixed)
  cf <- spread(mf,Var2,value)
  names(cf)[2] <- 'parameter'
  param_plot = ggplot(cf,aes(x=x,y=y)) + geom_line()+facet_wrap(~ parameter, 
             scales="free") + geom_vline(xintercept=0) + ylab("density")
  
  #Plot of precision of random effect (main hyperparameter of interest)
  sden <- data.frame(bri.hyper.sd(model$marginals.hyperpar[[1]]))
  hyperparam_plot = ggplot(sden,aes(x,y)) + geom_line() + ylab("density") + 
             xlab("linear predictor")
  
  #Store the results in the list
  kgr_model5_results = list(
    K_AR_heatmap = K_AR_heatmap, 
    K_DL_heatmap = K_DL_heatmap,
    K_Interaction_heatmap = K_Interaction_heatmap,
    K_AR_weight = K_AR_norm / (K_AR_norm + K_DL_norm + K_Int_norm),
    K_DL_weight = K_DL_norm / (K_AR_norm + K_DL_norm + K_Int_norm),
    K_Int_weight = K_Int_norm / (K_AR_norm + K_DL_norm + K_Int_norm),
    K_weight = K_weight/(K_weight + gfilter_weight),
    gfilter_weight = gfilter_weight/(K_weight + gfilter_weight),
    covmatrix = covGP5,
    prec = inv_covGP5,
    num_jitters = covGP_jittered[[2]],
    prec_heatmap = inv_covGP5_heatmap,
    model_summary = model_summary,
    bri_hyperpar_summary = bri_hyperpar_summary,
    exp_effects = multeff,
    param_plot = param_plot,
    hyperparam_plot = hyperparam_plot,
    model_DIC = model_DIC,
    model_WAIC = model_WAIC,
    fitted_values = preds_model,
    marg_fitted_values = marginal_fvs
  )
  
  return(kgr_model5_results)
}

#Fit kgr_model5
kgr_model5_fit = kgr_model5(dataset = inla_train_dataset, rho_AR_rbf = 1.019648e-04, 
                            rho_AR_periodic = 3.904021e-04, rho_DL_rbf = 6.926401e-04,
                            rho_DL_periodic = 1.378534e-05, rho_int_rbf = 1.494036e-04,
                            rho_int_periodic = 3.228066e-04, sigma2_AR = 5.567759e-01,
                            sigma2_DL = 5.955738e-01, sigma2_int = 2.712031, link=1)

#Extract DIC and WAIC 
kgr_model5_DIC = kgr_model5_fit$model_DIC
kgr_model5_WAIC = kgr_model5_fit$model_WAIC

#Get summaries of parameter estimates
kgr_model5_fit$model_summary
kgr_model5_fit$bri_hyperpar_summary
kgr_model5_fit$exp_effects
kgr_model5_fit$K_AR_weight
kgr_model5_fit$K_DL_weight
kgr_model5_fit$K_Int_weight
kgr_model5_fit$K_weight
kgr_model5_fit$gfilter_weight
kgr_model5_fit$num_jitters

#Show plots
kgr_model5_fit$K_AR_heatmap
kgr_model5_fit$K_DL_heatmap
kgr_model5_fit$K_Interaction_heatmap
kgr_model5_fit$prec_heatmap
kgr_model5_fit$param_plot
kgr_model5_fit$hyperparam_plot
pp_insample_plot(pred_data = kgr_model5_fit$fitted_values)
```
Fit some nonlinear models: 

```{r,eval=FALSE}
m.poly = inla(mpg ~ hp + I(hp^2) + I(hp^3), 
              data = test, control.predictor = list(compute = TRUE))
m.poly$summary.fixed
fvs.poly = m.poly$summary.fitted.values

ggplot(test,aes(x=hp,y=mpg)) + 
        geom_point() + geom_line(aes(y=fvs.poly$mean),color = "red") + 
        geom_ribbon(aes(ymin = fvs.poly$`0.025quant`,
                        ymax = fvs.poly$`0.975quant`),alpha = 0.3) 



knots = seq(50,350,length.out=3)
m.bsplines = inla(mpg ~ bs(hp,knots = knots),
                 data = test, control.predictor = list(compute = TRUE))
m.bsplines$summary.fixed
fvs.bsplines = m.bsplines$summary.fitted.values

ggplot(test,aes(x=hp,y=mpg)) + 
        geom_point() + geom_line(aes(y=fvs.bsplines$mean),color = "red") + 
        geom_ribbon(aes(ymin = fvs.bsplines$`0.025quant`,
                        ymax = fvs.bsplines$`0.975quant`),alpha = 0.3) 



m.nsplines = inla(mpg ~ ns(hp,df=3),
                 data = test, control.predictor = list(compute = TRUE))
m.nsplines$summary.fixed
fvs.nsplines = m.nsplines$summary.fitted.values

ggplot(test,aes(x=hp,y=mpg)) + 
        geom_point() + geom_line(aes(y=fvs.nsplines$mean),color = "red") + 
        geom_ribbon(aes(ymin = fvs.nsplines$`0.025quant`,
                        ymax = fvs.nsplines$`0.975quant`),alpha = 0.3) 
```

We can add splines to our log intensity function pretty easily 


# Fit models out of sample:

## BYM model

```{r}
#Run model 
ref_model2_outfit = ref_model2(dataset = inla_dataset_toimpute,a_prec_prior = 1,b_prec_prior = 5e-4,a_phi_prior = 1,b_phi_prior = 5e-4)

#Extract DIC and WAIC
ref_model2_DIC = ref_model2_outfit$model_DIC
ref_model2_WAIC = ref_model2_outfit$model_WAIC

#Get summaries of parameter estimates
ref_model2_outfit$model_summary %>% kable %>% kable_styling()
ref_model2_outfit$bri_hyperpar_summary %>% kable %>% kable_styling()
ref_model2_outfit$exp_effects

#Show plots
ref_model2_outfit$param_plot
ref_model2_outfit$hyperparam_plot
pp_outsample_plot(ref_data = inla_dataset,
                  pred_data = ref_model2_outfit$fitted_values)
```

## KGR model 1

```{r}
#Fit kgr_model1
kgr_model1_outfit = kgr_model1(dataset = inla_dataset_toimpute,rho_time_rbf = 536.519, rho_time_periodic = 5.707,sigma2_time = 2.773)

#Extract DIC and WAIC
kgr_model1_DIC = kgr_model1_outfit$model_DIC
kgr_model1_WAIC = kgr_model1_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model1_outfit$model_summary
kgr_model1_outfit$bri_hyperpar_summary
kgr_model1_outfit$exp_effects
kgr_model1_outfit$K_time_weight
kgr_model1_outfit$gfilter_weight

#Show plots
kgr_model1_outfit$K_time_heatmap
kgr_model1_outfit$param_plot
kgr_model1_outfit$hyperparam_plot
pp_outsample_plot(ref_data = inla_dataset,
                  pred_data = kgr_model1_outfit$fitted_values)
```

## KGR model 2

```{r}
#Fit kgr_model1
kgr_model2_outfit = kgr_model2(dataset = inla_dataset_toimpute,
                               rho_Z_rbf = 0.0020604943, rho_Z_periodic = 0.0024283451,
                               sigma2_Z = 3.941096, link=1)

#Extract DIC and WAIC
kgr_model2_DIC = kgr_model2_outfit$model_DIC
kgr_model2_WAIC = kgr_model2_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model2_outfit$model_summary
kgr_model2_outfit$bri_hyperpar_summary
kgr_model2_outfit$exp_effects
kgr_model2_outfit$K_time_weight
kgr_model2_outfit$gfilter_weight

#Show plots
kgr_model2_outfit$K_time_heatmap
kgr_model2_outfit$param_plot
kgr_model2_outfit$hyperparam_plot
pp_outsample_plot(ref_data = inla_dataset,
                  pred_data = kgr_model2_outfit$fitted_values)
```

## KGR model 3

```{r,warning=FALSE}
#Fit kgr_model3
kgr_model3_outfit = kgr_model3(dataset = inla_dataset_toimpute, 
                            rho_Z_rbf = 37801.48, 
                            rho_Z_periodic = 571512.56,
                            rho_time_rbf = 39.5276, 
                            rho_time_periodic = 380794.8,
                            sigma2 = 1.124888)

#Extract DIC and WAIC
kgr_model3_DIC = kgr_model3_outfit$model_DIC
kgr_model3_WAIC = kgr_model3_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model3_outfit$model_summary
kgr_model3_outfit$bri_hyperpar_summary
kgr_model3_outfit$exp_effects
kgr_model3_outfit$K_weight
kgr_model3_outfit$gfilter_weight

#Show plots
kgr_model3_outfit$K_time_heatmap
kgr_model3_outfit$K_Z_heatmap
kgr_model3_outfit$param_plot
kgr_model3_outfit$hyperparam_plot
pp_outsample_plot(ref_data = inla_dataset,
                  pred_data = kgr_model3_outfit$fitted_values)
```

## KGR model 4

```{r,warning=FALSE}
#Fit kgr_model3
kgr_model4_outfit = kgr_model4(dataset = inla_dataset_toimpute, 
                            rho_Z_rbf = 0.0008842497,
                            rho_Z_periodic = 0.0012119621,
                            rho_time_rbf = 0.0037851329,
                            rho_time_periodic = 0.003501742,
                            sigma2_Z = 4.310359,
                            sigma2_time = 0.9296752, link=1)

#Extract DIC and WAIC
kgr_model4_DIC = kgr_model4_outfit$model_DIC
kgr_model4_WAIC = kgr_model4_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model4_outfit$model_summary
kgr_model4_outfit$bri_hyperpar_summary
kgr_model4_outfit$exp_effects
kgr_model4_outfit$K_weight
kgr_model4_outfit$gfilter_weight

#Show plots
kgr_model4_outfit$K_time_heatmap
kgr_model4_outfit$K_Z_heatmap
kgr_model4_outfit$param_plot
kgr_model4_outfit$hyperparam_plot
pp_outsample_plot(ref_data = inla_dataset,
                  pred_data = kgr_model4_outfit$fitted_values)
```

## KGR model 5

```{r,warning=FALSE}
#Fit kgr_model5
kgr_model5_outfit = kgr_model5(dataset = inla_dataset_toimpute, rho_AR_rbf = 1.019648e-04, 
                            rho_AR_periodic = 3.904021e-04, rho_DL_rbf = 6.926401e-04,
                            rho_DL_periodic = 1.378534e-05, rho_int_rbf = 1.494036e-04,
                            rho_int_periodic = 3.228066e-04, sigma2_AR = 5.567759e-01,
                            sigma2_DL = 5.955738e-01, sigma2_int = 2.712031, link=1)

#Extract DIC and WAIC
kgr_model5_DIC = kgr_model5_outfit$model_DIC
kgr_model5_WAIC = kgr_model5_outfit$model_WAIC

#Get summaries of parameter estimates
kgr_model5_outfit$model_summary
kgr_model5_outfit$bri_hyperpar_summary
kgr_model5_outfit$exp_effects
kgr_model5_outfit$K_AR_weight
kgr_model5_outfit$K_DL_weight
kgr_model5_outfit$K_Int_weight
kgr_model5_outfit$K_weight
kgr_model5_outfit$gfilter_weight
kgr_model5_outfit$num_jitters

#Show plots
kgr_model5_outfit$K_AR_heatmap
kgr_model5_outfit$K_DL_heatmap
kgr_model5_outfit$K_Interaction_heatmap
kgr_model5_outfit$prec_heatmap
kgr_model5_outfit$param_plot
kgr_model5_outfit$hyperparam_plot
pp_outsample_plot(ref_data = inla_dataset,
                  pred_data = kgr_model5_outfit$fitted_values)
```

# Evaluating model fit

## Comparing out of sample MAE, MASE, and MAPE for different clusters between models

```{r}
print(degree_connectivity)
```

Mean absolute error (MAE) is a measure of the average size of the mistakes in a collection of predictions, without taking their direction into account

MAE = $\frac{1}{h_{max}} \sum_{h=1}^{h_{max}} | \hat \lambda_{t+h}^{obs} - \hat \lambda_{t+h}|$ where $\hat \lambda_{t+h}^{obs}$ is the average of the number of deaths observed for month $t+h$ over all years

```{r}
MAE_table = matrix(nrow=num_counties,ncol=6)

for (i in 1:num_counties){
  actual = inla_dataset %>% filter(county_id == i)  %>% data.frame()

  pm_1 = ref_model2_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()
  pm_2 = kgr_model1_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_3 = kgr_model2_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_4 = kgr_model3_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_5 = kgr_model4_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_6 = kgr_model5_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()

  actual_test = c()
  
  for (j in 1:12){
    actual_j = actual %>% filter(month == j) %>% select(deaths)
    est_lambda = mean(actual_j$deaths)
    actual_test = c(actual_test,as.numeric(est_lambda))
  }
  
  actual_test = c(actual_test[5:12],actual_test)

  pm_1_test = pm_1[50:69,]
  pm_2_test = pm_2[50:69,]
  pm_3_test = pm_3[50:69,]
  pm_4_test = pm_4[50:69,]
  pm_5_test = pm_5[50:69,]
  pm_6_test = pm_6[50:69,]

  actual_test_mean = mean(actual_test)

  mae1 = mean(abs(actual_test - pm_1_test$mean))
  mae2 = mean(abs(actual_test - pm_2_test$mean))
  mae3 = mean(abs(actual_test - pm_3_test$mean))
  mae4 = mean(abs(actual_test - pm_4_test$mean))
  mae5 = mean(abs(actual_test - pm_5_test$mean))
  mae6 = mean(abs(actual_test - pm_6_test$mean))
  
  MAE_table[i,] = c(mae1,mae2,mae3,mae4,mae5,mae6)
  # MAE_table[,i] = MAE_table[,i] / actual_test_mean
}

#Table 2: MAE on test dataset

MAE_table = data.frame(MAE_table)

rownames(MAE_table) = clusterlabels$counties
colnames(MAE_table) = c("BYM model","Proposed KGR model 1","Proposed KGR model 2","Proposed KGR model 3","Proposed KGR model 4","Proposed KGR model 5")

MAE_table
```

Mean absolute scaled error (MASE) is a measure of the accuracy of forecasts. It is the mean absolute error of the forecast values, divided by the mean absolute error of the in-sample one-step naive forecast.

MASE = $\frac{\frac{1}{n} \sum_{t=1}^n |\hat \lambda_{t+h}^{obs} - \hat \lambda_{t+h}|}{\frac{1}{n-1} \sum_{t=2}^n |\hat \lambda_{t+h}^{obs} - \hat \lambda_{t+h-1}^{obs}|}$

```{r}
MASE_table = matrix(nrow=num_counties,ncol=6)

for (i in 1:num_counties){
  actual = inla_dataset %>% filter(county_id == i)  %>% data.frame()

  pm_1 = ref_model2_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()
  pm_2 = kgr_model1_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_3 = kgr_model2_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_4 = kgr_model3_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_5 = kgr_model4_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_6 = kgr_model5_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  
  actual_test = c()
  
  for (j in 1:12){
    actual_j = actual %>% filter(month == j) %>% select(deaths)
    est_lambda = mean(actual_j$deaths)
    
    if(est_lambda == 0){
      est_lambda = est_lambda+0.0001
    }
    
    actual_test = c(actual_test,as.numeric(est_lambda))
  }
  
  actual_test = c(actual_test[5:12],actual_test)

  pm_1_test = pm_1$mean[50:69]
  pm_2_test = pm_2$mean[50:69]
  pm_3_test = pm_3$mean[50:69]
  pm_4_test = pm_4$mean[50:69]
  pm_5_test = pm_5$mean[50:69]
  pm_6_test = pm_6$mean[50:69]

  actual_test_mean = mean(actual_test)
  
  values1 = c()
  values2 = c()
  values3 = c()
  values4 = c()
  values5 = c()
  values6 = c()
  
  for (j in 2:length(actual_test)){
    error1 = (abs(actual_test[j] - pm_1_test[j])) / (abs(actual_test[j] - actual_test[j-1])) / (length(actual_test)-1)
    values1 = c(values1,error1)
    
    error2 = (abs(actual_test[j] - pm_2_test[j])) / (abs(actual_test[j] - actual_test[j-1])) / (length(actual_test)-1)
    values2 = c(values2,error2)
    
    error3 = (abs(actual_test[j] - pm_3_test[j])) / (abs(actual_test[j] - actual_test[j-1])) / (length(actual_test)-1)
    values3 = c(values3,error3)
    
    error4 = (abs(actual_test[j] - pm_4_test[j])) / (abs(actual_test[j] - actual_test[j-1])) / (length(actual_test)-1)
    values4 = c(values4,error4)
    
    error5 = (abs(actual_test[j] - pm_5_test[j])) / (abs(actual_test[j] - actual_test[j-1])) / (length(actual_test)-1)
    values5 = c(values5,error5)
    
    error6 = (abs(actual_test[j] - pm_6_test[j])) / (abs(actual_test[j] - actual_test[j-1])) / (length(actual_test)-1)
    values6 = c(values6,error6)
  }
  
  mase1 = mean(values1)
  mase2 = mean(values2)
  mase3 = mean(values3)
  mase4 = mean(values4)
  mase5 = mean(values5)
  mase6 = mean(values6)
  
  MASE_table[i,] = c(mase1,mase2,mase3,mase4,mase5,mase6)
  # MASE_table[i,] = MASE_table[,i] / actual_test_mean
}

#Table 3: MASE on test dataset

MASE_table = data.frame(MASE_table)

rownames(MASE_table) = clusterlabels$counties
colnames(MASE_table) = c("BYM model","Proposed KGR model 1","Proposed KGR model 2","Proposed KGR model 3","Proposed KGR model 4","Proposed KGR model 5")

MASE_table
```

Mean absolute percentage error (MAPE), also known as mean absolute percentage deviation (MAPD), is a measure of prediction accuracy of a forecasting method in statistics, expressing accuracy as a ratio

MAPE = $\frac{100}{n} \sum_{t=1}^n |\frac{\hat \lambda_{t+h}^{obs} - \hat \lambda_{t+h}}{\hat \lambda_{t+h}^{obs}}|$

```{r}
MAPE_table = matrix(nrow=num_counties,ncol=6)

for (i in 1:num_counties){
  actual = inla_dataset %>% filter(county_id == i)  %>% data.frame()

  pm_1 = ref_model2_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()
  pm_2 = kgr_model1_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_3 = kgr_model2_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_4 = kgr_model3_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_5 = kgr_model4_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()
  pm_6 = kgr_model5_outfit$fitted_values %>% filter(county_id == i) %>%
    select(mean,sd) %>% data.frame()

  actual_test = c()
  
  for (j in 1:12){
    actual_j = actual %>% filter(month == j) %>% select(deaths)
    est_lambda = mean(actual_j$deaths)
    
    if(est_lambda == 0){
      est_lambda = est_lambda+0.0001
    }
    
    actual_test = c(actual_test,as.numeric(est_lambda))
  }
  
  actual_test = c(actual_test[5:12],actual_test)
  
  pm_1_test = pm_1[50:69,]
  pm_2_test = pm_2[50:69,]
  pm_3_test = pm_3[50:69,]
  pm_4_test = pm_4[50:69,]
  pm_5_test = pm_5[50:69,]
  pm_6_test = pm_6[50:69,]

  actual_test_mean = mean(actual_test)

  mape1 = mean(abs((actual_test - pm_1_test$mean)/actual_test))
  mape2 = mean(abs((actual_test - pm_2_test$mean)/actual_test))
  mape3 = mean(abs((actual_test - pm_3_test$mean)/actual_test))
  mape4 = mean(abs((actual_test - pm_4_test$mean)/actual_test))
  mape5 = mean(abs((actual_test - pm_5_test$mean)/actual_test))
  mape6 = mean(abs((actual_test - pm_6_test$mean)/actual_test))
  
  MAPE_table[i,] = c(mape1,mape2,mape3,mape4,mape5,mape6)
  # MAPE_table[i,] = MAPE_table[,i] / actual_test_mean
}

#Table 2: RMSE on test dataset

MAPE_table = data.frame(MAPE_table)

rownames(MAPE_table) = clusterlabels$counties
colnames(MAPE_table) = c("BYM model","Proposed KGR model 1","Proposed KGR model 2","Proposed KGR model 3","Proposed KGR model 4","Proposed KGR model 5")

MAPE_table
```

```{r}
#Overall fit
test_RMSE_table = matrix(nrow=num_counties,ncol=6)

for (i in 1:num_counties){
  actual = inla_dataset %>% filter(county_id == i) %>% data.frame()

  pm_1 = ref_model2_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()
  pm_2 = kgr_model1_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()
  pm_3 = kgr_model2_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()
  pm_4 = kgr_model3_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()
  pm_5 = kgr_model4_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()
  pm_6 = kgr_model5_outfit$fitted_values %>% filter(county_id == i) %>% 
    select(mean,sd) %>% data.frame()

  actual_test = c()
  
  for (j in 1:12){
    actual_j = actual %>% filter(month == j) %>% select(deaths)
    est_lambda = mean(actual_j$deaths)
    
    actual_test = c(actual_test,as.numeric(est_lambda))
  }
  
  actual_test = c(actual_test[5:12],actual_test)
  
  pm_1_test = pm_1[50:69,]
  pm_2_test = pm_2[50:69,]
  pm_3_test = pm_3[50:69,]
  pm_4_test = pm_4[50:69,]
  pm_5_test = pm_5[50:69,]
  pm_6_test = pm_6[50:69,]

  actual_test_mean = mean(actual_test)

  test_rmse1 = sqrt(mean((actual_test - pm_1_test$mean)^2))
  test_rmse2 = sqrt(mean((actual_test - pm_2_test$mean)^2))
  test_rmse3 = sqrt(mean((actual_test - pm_3_test$mean)^2))
  test_rmse4 = sqrt(mean((actual_test - pm_4_test$mean)^2))
  test_rmse5 = sqrt(mean((actual_test - pm_5_test$mean)^2))
  test_rmse6 = sqrt(mean((actual_test - pm_6_test$mean)^2))
  
  test_RMSE_table[i,] = c(test_rmse1,test_rmse2,test_rmse3,test_rmse4,test_rmse5,test_rmse6)
  test_RMSE_table[i,] = test_RMSE_table[i,] / actual_test_mean
}

#Table 2: RMSE on test dataset

test_RMSE_table = data.frame(test_RMSE_table)

rownames(test_RMSE_table) = clusterlabels$counties
colnames(test_RMSE_table) = c("BYM model","Proposed KGR model 1","Proposed KGR model 2",
                              "Proposed KGR model 3","Proposed KGR model 4","Proposed KGR model 5")

test_RMSE_table
```

# Comparing coverage rates of credible intervals produced by each model

```{r}
coverage = rep(0,6)

true_values = inla_dataset$deaths
models_fvs = list(ref_model2_outfit$fitted_values,
                  kgr_model1_outfit$fitted_values,
                  kgr_model2_outfit$fitted_values,
                  kgr_model3_outfit$fitted_values,
                  kgr_model4_outfit$fitted_values,
                  kgr_model5_outfit$fitted_values)

# true_values = inla_test_dataset$deaths
# 
# r2_fvs = ref_model2_outfit$fitted_values %>% filter(date >= "2018-05-01")
# k1_fvs = kgr_model1_outfit$fitted_values %>% filter(date >= "2018-05-01")
# k3_fvs = kgr_model3_outfit$fitted_values %>% filter(date >= "2018-05-01")
# 
# models_fvs = list(r2_fvs,k1_fvs,k3_fvs)

for (i in 1:6){
  lci = models_fvs[[i]] %>% select('0.025quant')
  uci = models_fvs[[i]] %>% select('0.975quant')
  
  captured = (true_values >= lci$'0.025quant' & true_values <= uci$'0.975quant')
  coverage[i] = sum(captured)/length(captured)
}

coverage = data.frame(coverage)
colnames(coverage) = "95% coverage"
rownames(coverage) = c("BYM model","Proposed KGR model 1","Proposed KGR model 2",
                       "Proposed KGR model 3","Proposed KGR model 4","Proposed KGR model 5")

coverage
```

# What is DRF? DRF models the expected value of potential outcomes Y as a function of the treatment level and covariates

Use models fit above as DRF - DRF is a convex function $f: \mathcal{Z} x \mathcal{X} \rightarrow \mathcal{Y}$ s.t. $Y(Z) = Y(I,Z,\textbf{X}) = f(Z,\textbf{X}) + \textbf{F]$

$$
E[Y|Z = z_{i,t}, \textbf{X} = \textbf{x}_{i,t}] = E[Y(z_{i,t}) | Z = z_{i,t}, \textbf{X} = \textbf{x}_{i,t}] = E[Y(z_{i,t}) | \textbf{X} = \textbf{x}_{i,t}] = E[f(z_{i,t},\textbf{X}) + \textbf{F}_{i,t} | \textbf{X} = \textbf{x}_{i,t}] = f(z_{i,t},\textbf{x}_{i,t})
$$

Counterfactual prediction: change treatment level Z while holding covariates X fixed -\> estimate all potential outcomes for a given individual OR fix Z while varying X -\> estimate how treatment responses vary with individual level characteristics

# ADRF: average of DRF over all individuals in the population

$E[Y(z)] = E\{ E[Y(z) | \textbf{X}] \} = E[f(z,\textbf{X})] = \int_{\mathcal{X}} f(z,\textbf{x}) dP{\textbf{X}}(x) = \mu(z)$

Estimate by choosing a set of treatment values Z and averaging the predicted potential outcomes over the empirical distribution of X i.e. for each value z, estimate predicted value from DRF for every subject i in the sample and average those together

For each value of Z, predict Y for each county i using X_i -\> average those Ys together -\> one Ybar for each Z -\> ADRF(z)

Calculate ADRF for each month -\> 12 ADRFs -\> average across months -\> final ADRF

Create separate datasets to feed into INLA for each value of Z

```{r,eval=FALSE}
causal_dataset = inla_dataset

z_grid = seq(0,max(causal_dataset$wildfire_pm25),length.out = 20)

monthly_avg_dataset = causal_dataset %>% group_by(county,month) %>% summarize(aqi_avg = mean(b3_AQI),lead_avg = mean(b3_lead),
co_avg = mean(b3_co),so2_avg = mean(b3_so2),no2_avg = mean(b3_no2),o3_avg = mean(b3_o3),pm10_avg = mean(b3_pm10),
pm25_avg = mean(b3_pm25),treatment = mean(b3_wildfire_pm25))

ADRF_data_list = list()

for(i in 1:length(z_grid)){
  monthly_avg_dataset$treatment = z_grid[i]
  monthly_avg_dataset$deaths = NA
  monthly_avg_dataset$date  = "2018-05-01" #put a random date insample as placeholder
  ADRF_data_list[[i]] = monthly_avg_dataset
}
```

Calculate ADRF(z) for each month separately -\> 12 ADRFs -\> avg across months

Attach training data to potential outcome (ADRF) data

```{r,eval=FALSE}
ADRF_month = function(Month,model_num){
  ADRF_table = data.frame(month=NA,treatment=NA,potential_outcome=NA)

  for(i in 1:length(ADRF_data_list)){
    test = ADRF_data_list[[i]]
    month_test = test %>% filter(month == Month)

    inla_addon = inla_train_dataset[,c("county","month","b3_AQI","b3_lead","b3_co","b3_so2","b3_no2","b3_o3","b3_pm10","b3_pm25","b3_wildfire_pm25","deaths","date")]

    colnames(month_test) = colnames(inla_addon)
    ADRF_data = rbind(inla_addon,month_test)

    ADRF_data = ADRF_data %>% left_join(county_ids, by = "county")

    #Add multiple intercept columns, one for each cluster 
    ADRF_data <- ADRF_data %>%
      mutate(
        Intercept1 = ifelse(cluster == 1, 1, NA),
        Intercept2 = ifelse(cluster == 2, 1, NA),
        Intercept3 = ifelse(cluster == 3, 1, NA),
        Intercept4 = ifelse(cluster == 4, 1, NA),
        Intercept5 = ifelse(cluster == 5, 1, NA)
      ) 

    ADRF_data$idx = 1:nrow(ADRF_data)
    
    if(model_num == 1){
        model_ADRF = ref_model2(dataset = ADRF_data,a_prec_prior = 1,
                                b_prec_prior = 5e-4,a_phi_prior = 1,b_phi_prior = 5e-4)
    } else if(model_num == 2){
        model_ADRF = kgr_model1(dataset = ADRF_data,rho_time_rbf = 536.519, 
                                rho_time_periodic = 5.707,sigma2_time = 2.773)
    } else if(model_num == 3){
        model_ADRF = kgr_model2(dataset = ADRF_data, rho_Z_rbf = 0.0020604943,
                                rho_Z_periodic = 0.0024283451, sigma2_Z = 3.941096)
    } else if(model_num == 4){
        model_ADRF = kgr_model3(dataset = ADRF_data,
                            rho_Z_rbf = 37801.48, 
                            rho_Z_periodic = 571512.56,
                            rho_time_rbf = 39.5276, 
                            rho_time_periodic = 380794.8,
                            sigma2 = 1.124888)
    } else if(model_num == 5){
        model_ADRF = kgr_model4(dataset = ADRF_data,
                            rho_Z_rbf = 0.0008842497,
                            rho_Z_periodic = 0.0012119621,
                            rho_time_rbf = 0.0037851329,
                            rho_time_periodic = 0.003501742,
                            sigma2_Z = 4.310359,
                            sigma2_time = 0.9296752)
    } else if(model_num == 6){
        model_ADRF = kgr_model5(dataset = ADRF_data, rho_AR_rbf = 1.019648e-04, 
                        rho_AR_periodic = 3.904021e-04, rho_DL_rbf = 6.926401e-04,
                        rho_DL_periodic = 1.378534e-05, rho_int_rbf = 1.494036e-04,
                        rho_int_periodic = 3.228066e-04, sigma2_AR = 5.567759e-01,
                        sigma2_DL = 5.955738e-01, sigma2_int = 2.712031, link=1)
    }


    ADRF_preds = tail(model_ADRF$fitted_values,58)

    ADRF_pred = mean(ADRF_preds$mean)
    ADRF_response = data.frame(month=Month,treatment=z_grid[i],potential_outcome=ADRF_pred)

    ADRF_table = rbind(ADRF_table,ADRF_response)
  }
  
ADRF_table = ADRF_table[-1,]

  return(ADRF_table)
}

ADRF_Jan_ref2 = ADRF_month(Month=1,model_num = 1)
ADRF_Jan_kgr2 = ADRF_month(Month=1,model_num = 3)
ADRF_Jan_kgr3 = ADRF_month(Month=1,model_num = 4)
ADRF_Jan_kgr4 = ADRF_month(Month=1,model_num = 5)


# ADRF_Jan_kgr1_nonlinear = ADRF_month(Month=1,model_num=2)
# ADRF_Jan_kgr5_nonlinear = ADRF_month(Month=1,model_num=6)
```

Calculate some ADRFs: 

```{r}
ADRF_Jan_ref2 = readRDS("ADRF_Jan_ref2.rds")
ADRF_Jan_kgr1_nonlinear = readRDS("ADRF_Jan_kgr1_nonlinear.rds")
ADRF_Jan_kgr2 = readRDS("ADRF_Jan_kgr2.rds")
ADRF_Jan_kgr3 = readRDS("ADRF_Jan_kgr3.rds")
ADRF_Jan_kgr4 = readRDS("ADRF_Jan_kgr4.rds")
ADRF_Jan_kgr5_nonlinear = readRDS("ADRF_1_kgr5.rds")

ggplot(ADRF_Jan_ref2,aes(x=treatment,y=potential_outcome)) + geom_line()
ggplot(ADRF_Jan_kgr1_nonlinear,aes(x=treatment,y=potential_outcome)) + geom_line()
ggplot(ADRF_Jan_kgr2,aes(x=treatment,y=potential_outcome)) + geom_line()
ggplot(ADRF_Jan_kgr3,aes(x=treatment,y=potential_outcome)) + geom_line()
ggplot(ADRF_Jan_kgr4,aes(x=treatment,y=potential_outcome)) + geom_line()
ggplot(ADRF_Jan_kgr5_nonlinear,aes(x=treatment,y=potential_outcome)) + geom_line()
```
Calculate ADRF for every month

```{r,eval=FALSE}
for (i in 2:12){
  ref_model = ADRF_month(Month=i,model_num=1)
  kgr5 = ADRF_month(Month=i,model_num=6)
  
  file1 = sprintf("ADRF_%d_ref2.rds",i)
  file2 = sprintf("ADRF_%d_kgr5.rds",i)

  
  saveRDS(ref_model,file1)
  saveRDS(kgr5,file2)
}

# ADRF_Feb = test

# test = readRDS("ADRF_Jan_kgr1.rds")
# 
# plot = ggplot(ADRF_table,aes(x=treatment,y=potential_outcome)) + geom_line()
# plot
```

Average ADRF for each month into a single ADRF for each model 

```{r}
ADRF_overall = function(model_num,print=TRUE){
  potential_outcomes = rep(1,20)
  
  for (i in 1:12){
    if(model_num == 1){
      
      file = sprintf("ADRF_%d_ref2.rds",i) 
      ADRF_month = readRDS(file)
      potential_outcomes = cbind(potential_outcomes,ADRF_month$potential_outcome)
      
    } else if(model_num == 2){
        
      file = sprintf("ADRF_%d_kgr1.rds",i) 
      ADRF_month = readRDS(file)
      potential_outcomes = cbind(potential_outcomes,ADRF_month$potential_outcome)
      
    } else if(model_num == 3){
      
      file = sprintf("ADRF_%d_kgr2.rds",i) 
      ADRF_month = readRDS(file)
      potential_outcomes = cbind(potential_outcomes,ADRF_month$potential_outcome)  
      
    } else if(model_num == 4){
      
      file = sprintf("ADRF_%d_kgr3.rds",i) 
      ADRF_month = readRDS(file)
      potential_outcomes = cbind(potential_outcomes,ADRF_month$potential_outcome)  
      
    } else if(model_num == 5){
      
      file = sprintf("ADRF_%d_kgr4.rds",i) 
      ADRF_month = readRDS(file)
      potential_outcomes = cbind(potential_outcomes,ADRF_month$potential_outcome)  
      
    } else if(model_num == 6){
      
      file = sprintf("ADRF_%d_kgr5.rds",i) 
      ADRF_month = readRDS(file)
      potential_outcomes = cbind(potential_outcomes,ADRF_month$potential_outcome)  
      
    }
  }
  
  potential_outcomes = potential_outcomes[,-1]
  EYz = apply(potential_outcomes,1,mean)
  ADRF = data.frame(z_grid,EYz)
  colnames(ADRF) = c("treatment","potential_outcome")
  
  if(print == TRUE){
    ADRF_plot = ggplot(ADRF,aes(x=treatment,y=potential_outcome)) + geom_line()
    return(ADRF_plot)
  } else{
    return(ADRF)
  }
}

ADRF_overall(model_num=6)
```


```{r,eval=FALSE}
ADRF_Jan = readRDS("ADRF_1_kgr5.rds")
ADRF_Feb = readRDS("ADRF_2_kgr5.rds")
ADRF_Mar = readRDS("ADRF_3_kgr5.rds")
ADRF_Apr = readRDS("ADRF_4_kgr5.rds")
ADRF_May = readRDS("ADRF_5_kgr5.rds")
ADRF_Jun = readRDS("ADRF_6_kgr5.rds")
ADRF_Jul = readRDS("ADRF_7_kgr5.rds")
ADRF_Aug = readRDS("ADRF_8_kgr5.rds")
ADRF_Sep = readRDS("ADRF_9_kgr5.rds")
ADRF_Oct = readRDS("ADRF_10_kgr5.rds")
ADRF_Nov = readRDS("ADRF_11_kgr5.rds")
ADRF_Dec = readRDS("ADRF_12_kgr5.rds")


ADRF_kgr5 = ADRF_Jan$treatment
EYz = cbind(ADRF_Jan$potential_outcome,ADRF_Feb$potential_outcome,ADRF_Mar$potential_outcome,
            ADRF_Apr$potential_outcome,ADRF_May$potential_outcome,ADRF_Jun$potential_outcome,
            ADRF_Jul$potential_outcome,ADRF_Aug$potential_outcome,ADRF_Sep$potential_outcome,
            ADRF_Oct$potential_outcome,ADRF_Nov$potential_outcome,ADRF_Dec$potential_outcome)
EYz = apply(EYz,1,mean)
ADRF_kgr5 = data.frame(ADRF_kgr5,EYz)
colnames(ADRF_kgr5) = c("treatment","potential_outcome")

ggplot(ADRF_kgr5,aes(x=treatment,y=potential_outcome)) + geom_line()
```





ADRF function prototype:

```{r,eval=FALSE}
test = ADRF_data_list[[1]]
Jan_test = test %>% filter(month == 1)

inla_addon = inla_train_dataset[,c("county","month","b3_AQI","b3_lead","b3_co","b3_so2","b3_no2","b3_o3","b3_pm10","b3_pm25","b3_wildfire_pm25","deaths","date")]

colnames(Jan_test) = colnames(inla_addon)
ADRF_data = rbind(inla_addon,Jan_test)

ADRF_data = ADRF_data %>% left_join(county_ids, by = "county")

#Add multiple intercept columns, one for each cluster 
ADRF_data <- ADRF_data %>%
  mutate(
    Intercept1 = ifelse(cluster == 1, 1, NA),
    Intercept2 = ifelse(cluster == 2, 1, NA),
    Intercept3 = ifelse(cluster == 3, 1, NA),
    Intercept4 = ifelse(cluster == 4, 1, NA),
    Intercept5 = ifelse(cluster == 5, 1, NA)
  ) 

ADRF_data$idx = 1:nrow(ADRF_data)
```

Fit INLA model -\> avg fitted values for each county together -\> perform for each month

```{r,eval=FALSE}
ref_model2_ADRF = ref_model2(dataset = ADRF_data,a_prec_prior = 1,b_prec_prior = 5e-4,a_phi_prior = 1,b_phi_prior = 5e-4)

kgr_model1_ADRF = kgr_model1(dataset = ADRF_data,rho_time_rbf = 536.519, rho_time_periodic = 5.707,sigma2_time = 2.773)

kgr_model3_ADRF = kgr_model3(dataset = ADRF_data, 
                            rho_EPA_rbf = 37801.48, 
                            rho_EPA_periodic = 571512.56,
                            rho_time_rbf = 39.5276, 
                            rho_time_periodic = 380794.8,
                            sigma2 = 1.124888)


ADRF_preds = tail(kgr_model3_ADRF$fitted_values,58)
ADRF_pred = mean(ADRF_preds$mean)
ADRF_response = data.frame(month=1,treatment=0,potential_outcome=ADRF_pred)
```

# AGTE: evaluate ADRF at two levels of treatment for which we want to compare and take the difference

$\tau_{AGTE}(z,z_0) := E[Y(z) - Y(z_0)]$

Estimate by evaluating ADRF at two levels of treatment that we are interested in (if z=0 then we are looking at AGTET) -\> take the difference

Ex: ADRF(100) - ADRF(0)

$\tau_{AGTET}(Z,0) := E[Y(Z) - Y(0) | Z > 0]$

Holding covariate values constant and making same comparison gives us CAGTE -\> comparing potential outcomes for different individuals (values of X) but same treatment level allows us to estimate heterogeneity in treatment effects

$\tau_{CAGTE}(z,z_0,\textbf{x}) := E[Y(z) - Y(z_0) | \textbf{X} = \textbf{x}]$

CAGTE(z,0,X) = ADRF(z\|x) - ADRF(0\|x)

CAGTE1: Fix X to be mean of air pollutant confounders and a fixed month e.g. June

CAGTE2: Fix X to be mode of air pollutant confounders and a fixed month e.g. June
